### 2015-05-29 manuscript/manuscript.Rmd
Addded multiple levels cp and plots

---
title: 'Associations between Chlorophyll *a* and various Microcystin-LR Health Advisory Concentrations'
author:  
- name: Jeffrey W. Hollister
  affilnum: 1
  email: hollister.jeff@epa.gov
- name:  Betty J. Kreakie
  affilnum: 1
- name: Dorothy Q. Kellog
  affilnum: 2
affiliation:
- affilnum: 1
  affil: US Environmental Protection Agency, Office of Research and Development, National Health and Environmental Effects Research Laboratory, Atlantic Ecology Division, 27 Tarzwell Drive  Narragansett, RI, 02882, USA
- affilnum: 2
  affil: University of Rhode Island, Department of Natural Resrouces Science, Kingston, RI, 02882, USA
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    template: components/manuscript.latex
  html_document: null
  word_document: null
fontsize: 11pt
capsize: normalsize
csl: components/ecology.csl
documentclass: article
spacing: doublespacing
linenumbers: no
bibliography: components/manuscript.bib
abstract: no
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Microsystin Chlorophyll Manusript}
-->

```{r setup, include=FALSE, echo=FALSE}
#Put whatever you normally put in a setup chunk.
#I usually at least include:
################################################################################
#Load packages
################################################################################
devtools::install_github("jhollist/microcystinchla",
                         auth_token = getOption("Github_Access"))
library("microcystinchla")
library("dplyr")
library("magrittr")
library("pander")
library("knitr")
opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=5)

# Table Captions from @DeanK on http://stackoverflow.com/questions/15258233/
# using-table-caption-on-r-markdown-file-using-knitr-to-use-in-pandoc-to-convert-t
# Figure captions are handled by LaTeX

knit_hooks$set(tab.cap = function(before, options, envir) {
                  if(!before) { 
                    paste('\n\n:', options$tab.cap, sep='') 
                  }
                })
default_output_hook = knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
  if (is.null(options$tab.cap) == FALSE) {
    x
  } else
    default_output_hook(x,options)
})
```

```{r analysis , include=FALSE, echo=FALSE, cache=TRUE}
#All analysis in here, that way all bits of the paper have access to the final objects
#Place tables and figures and numerical results where they need to go.

################################################################################
# Data Prep
################################################################################
data_url<-"http://water.epa.gov/type/lakes/assessmonitor/lakessurvey/upload/NLA2007_Recreational_ConditionEstimates_20091123.csv"
nla_dat <- get_nla(data_url)
nla_dat <- mutate(nla_dat,log_chla = log(CHLA),log_micro = log(MCYST_TL_UGL))

################################################################################
# Condtional Probability
################################################################################
mc_guides <- c(0.04, 0.3, 0.8, 1, 1.6, 4, 6, 10, 14, 20)
mc_names <- c("MN","EPA_Child","CA","WHO","EPA_Adult","WHO","Multiple","Multiple","Mulitple","Multiple")
mn_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                     xImpair = log(mc_guides[1]), ProbComp = "gt", Exceed = "gte",
                     ci = TRUE, R = 100)
epa_child_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                     xImpair = log(mc_guides[2]), ProbComp = "gt", Exceed = "gte",
                     ci = TRUE, R = 100)
ca_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                     xImpair = log(mc_guides[3]), ProbComp = "gt", Exceed = "gte",
                     ci = TRUE, R = 100)
who_drink_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                     xImpair = log(mc_guides[4]), ProbComp = "gt", Exceed = "gte",
                     ci = TRUE, R = 100)
epa_adult_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                     xImpair = log(mc_guides[5]), ProbComp = "gt", Exceed = "gte",
                     ci = TRUE, R = 100)
who_rec_low_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                     xImpair = log(mc_guides[6]), ProbComp = "gt", Exceed = "gte",
                     ci = TRUE, R = 100)
mult_state_rec_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                     xImpair = log(mc_guides[7]), ProbComp = "gt", Exceed = "gte",
                     ci = TRUE, R = 100)
il_org_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                     xImpair = log(mc_guides[8]), ProbComp = "gt", Exceed = "gte",
                     ci = TRUE, R = 100)
ma_ri_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                     xImpair = log(mc_guides[9]), ProbComp = "gt", Exceed = "gte",
                     ci = TRUE, R = 100)
who_rec_med_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                     xImpair = log(mc_guides[10]), ProbComp = "gt", Exceed = "gte",
                     ci = TRUE, R = 100)

```

\singlespace

\vspace{2mm}\hrule

<!-- Abstract is being wrapped in latex here so that all analysis can be run in the chunk above and the results reproducibly referenced in the abstract. -->

\vspace{3mm}\hrule
\doublespace

#Introduction

#Methods

```{r microcystin_levels,echo=FALSE,results='asis'}
source_of_stnd <- c("WHO, IN",
                    "WHO, IN, KS",
                    "WHO, IN, IA, KS, NE, OH",
                    "WHO",
                    "WHO, OH, OR",
                    "MN",
                    "CA",
                    "IL, OR",
                    "MA, RI",
                    "OH, VT, VA, WA",
                    "U.S. EPA",
                    "U.S. EPA")
type_of_stnd <- c("Recreational",
                  "Recreational",
                  "Recreational",
                  "Recreational",
                  "Drinking",
                  "Drinking",
                  "Recreational",
                  "Recreational",
                  "Recreational",
                  "Recreational",
                  "Drinking",
                  "Drinking")
concentration <- c("2-4 ug/l",
                   "10-20 ug/l",
                   "20-2000 ug/l",
                   ">2000 ug/l",
                   "1 ug/l",
                   "0.04 ug/l",
                   "0.8 ug/l",
                   "10 ug/l",
                   "14 ug/l",
                   "6 ug/l",
                   "0.3 ug/l",
                   "1.6 ug/l")
#note <- c("Low probability of adverse health effects, state advisories",
#          "Moderate probability of adverse health effects, state advisories",
#          "High probability of adverse health effects, state advisories, These higher #concentrations typically only seen in visible scums/algal mats",
#          "Very high probability of adverse health effects, These higher concentrations #typically only seen in visible scums/algal mats",
#          "","","","","","",
#          "Children younger than school age",
#          "Individuals School age and older")
mc_conc_table <- data.frame(source_of_stnd, type_of_stnd, concentration) %>%
  arrange(type_of_stnd)
names(mc_conc_table)<-c("Source","Type","Concentration")
pandoc.table(mc_conc_table,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
```

Based on the range of microcystin concentrations used in various policy settings, we evaulated associated chlorohpyll *a* concentrations for an effect at 0.04, 0.3, 0.8, 1, 1.6, 4, 6, 10, 14, and 20 ug/l.

##Data and Study Area

#Results

#Discussion

#Figures

```{r mn_cp_plot,echo=FALSE}
cp_plot(mn_cp)
```

\newpage

```{r epa_child_cp_plot,echo=FALSE}
cp_plot(epa_child_cp)
```

```{r ca_cp_plot,echo=FALSE}
cp_plot(ca_cp)
```

\newpage

```{r who_drink_cp_plot,echo=FALSE}
cp_plot(who_drink_cp)
```

\newpage

```{r epa_adult_cp ,echo=FALSE}
cp_plot(epa_adult_cp)
```

\newpage

```{r who_rec_low_cp, echo=FALSE}
cp_plot(who_rec_low_cp)
```

\newpage

```{r mult_state_rec_cp,  echo=FALSE}
cp_plot(mult_state_rec_cp)
```

\newpage

```{r il_org_cp, echo=FALSE}
cp_plot(il_org_cp)
```

\newpage

```{r ma_ri_cp, echo=FALSE}
cp_plot(ma_ri_cp)
```

\newpage

```{r who_rec_med_cp, echo=FALSE}
cp_plot(who_rec_med_cp)
```

\newpage

#References

### 2015-05-29 manuscript/manuscript.tex
Addded multiple levels cp and plots

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Original default rstudio/pandoc latex file
%% upated by @jhollist 09/15/2014
%% inspired by @cboetting https://github.com/cboettig/template and
%% @rmflight blog posts:
%% http://rmflight.github.io/posts/2014/07/analyses_as_packages.html 
%% http://rmflight.github.io/posts/2014/07/vignetteAnalysis.html).  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt,]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{longtable,booktabs}
\usepackage{graphicx}
% Redefine \includegraphics so that, unless explicit options are
% given, the image width will not exceed the width of the page.
% Images get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\ScaleIfNeeded{%
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother
\let\Oldincludegraphics\includegraphics
{%
 \catcode`\@=11\relax%
 \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
}%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Associations between Chlorophyll a and Microcystin-LR Health Advisory Concentrations},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Changes borrowed from @cboettig, added by @jhollist 
% A modified page layout 
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
\usepackage{lineno} % add 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%Packages and layout changes by @jhollist 09/15/2014
\usepackage{ragged2e}
\usepackage[font=normalsize]{caption}
  \usepackage[doublespacing]{setspace}
\usepackage{parskip}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\rfoot{\today}
\lfoot{\thepage}
%%Changed default abstract width and added lines
\renewenvironment{abstract}{
  \hfill\begin{minipage}{1\textwidth}
  \rule{\textwidth}{1pt}\vspace{5pt}
  \normalsize
  \begin{justify}
  \bfseries\abstractname\vspace{5pt}
  \end{justify}}
  {\par\noindent\rule{\textwidth}{1pt}\end{minipage}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Associations between Chlorophyll \emph{a} and Microcystin-LR Health
Advisory Concentrations}
\author{
Jeffrey W. Hollister
Betty J. Kreakie
Dorothy Q. Kellog
}
\date{}

\begin{document}
%%Edited by @jhollist 09/15/2014
%%Adds title from YAML
\begin{singlespace}
\begin{center}
\huge Associations between Chlorophyll \emph{a} and Microcystin-LR Health
Advisory Concentrations
\end{center}
%%Adds Author, correspond email asterisk, and affilnum from YAML
\begin{center}
\large
Jeffrey W. Hollister \textsuperscript{*} \textsuperscript{1} 
Betty J. Kreakie \textsuperscript{1} 
Dorothy Q. Kellog \textsuperscript{2} 
\end{center}
%%Adds affiliations from YAML
\begin{justify}
\footnotesize \emph{ 
\\*
\textsuperscript{1}US Environmental Protection Agency, Office of Research and Development,
National Health and Environmental Effects Research Laboratory, Atlantic
Ecology Division, 27 Tarzwell Drive Narragansett, RI, 02882, USA\\*
\\*
\textsuperscript{2}University of Rhode Island, Department of Natural Resrouces Science,
Kingston, RI, 02882, USA\\*
}
%%Adds corresponding author email(s) from YAML
\newcounter{num}
\setcounter{num}{1}
\\[0.1cm]
\footnotesize \emph{ 
\ifnum\value{num}=1%
\textsuperscript{*} corresponding author:
\fi
\href{mailto:hollister.jeff@epa.gov}{\nolinkurl{hollister.jeff@epa.gov}}
\stepcounter{num}
}
\end{justify}
%%Adds date from YAML
\normalsize

\end{singlespace}

\singlespace

\vspace{2mm}

\hrule

\vspace{3mm}

\hrule

\doublespace

\section{Introduction}\label{introduction}

\section{Methods}\label{methods}

\begin{longtable}[c]{@{}lll@{}}
\toprule
\begin{minipage}[b]{0.27\columnwidth}\raggedright\strut
Source
\strut\end{minipage} &
\begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
Type
\strut\end{minipage} &
\begin{minipage}[b]{0.19\columnwidth}\raggedright\strut
Concentration
\strut\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
WHO, OH, OR
\strut\end{minipage} &
\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
Drinking
\strut\end{minipage} &
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
1 ug/l
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
MN
\strut\end{minipage} &
\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
Drinking
\strut\end{minipage} &
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
0.04 ug/l
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
U.S. EPA
\strut\end{minipage} &
\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
Drinking
\strut\end{minipage} &
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
0.3 ug/l
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
U.S. EPA
\strut\end{minipage} &
\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
Drinking
\strut\end{minipage} &
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
1.6 ug/l
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
WHO, IN
\strut\end{minipage} &
\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
Recreational
\strut\end{minipage} &
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
2-4 ug/l
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
WHO, IN, KS
\strut\end{minipage} &
\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
Recreational
\strut\end{minipage} &
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
10-20 ug/l
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
WHO, IN, IA, KS, NE, OH
\strut\end{minipage} &
\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
Recreational
\strut\end{minipage} &
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
20-2000 ug/l
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
WHO
\strut\end{minipage} &
\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
Recreational
\strut\end{minipage} &
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
\textgreater{}2000 ug/l
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
CA
\strut\end{minipage} &
\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
Recreational
\strut\end{minipage} &
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
0.8 ug/l
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
IL, OR
\strut\end{minipage} &
\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
Recreational
\strut\end{minipage} &
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
10 ug/l
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
MA, RI
\strut\end{minipage} &
\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
Recreational
\strut\end{minipage} &
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
14 ug/l
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
OH, VT, VA, WA
\strut\end{minipage} &
\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
Recreational
\strut\end{minipage} &
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
6 ug/l
\strut\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

Based on the range of microcystin concentrations used in various policy
settings, we evaulated associated chlorohpyll \emph{a} concentrations
for an effect at 0.04, 0.3, 0.8, 1, 1.6, 4, 6, 10, 14, and 20 ug/l.

\subsection{Data and Study Area}\label{data-and-study-area}

\section{Results}\label{results}

\section{Discussion}\label{discussion}

\section{Figures}\label{figures}

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/mn_cp_plot-1.pdf}
\caption{}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/epa_child_cp_plot-1.pdf}
\caption{}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/ca_cp_plot-1.pdf}
\caption{}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/who_drink_cp_plot-1.pdf}
\caption{}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/epa_adult_cp -1.pdf}
\caption{}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/who_rec_low_cp-1.pdf}
\caption{}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/mult_state_rec_cp-1.pdf}
\caption{}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/il_org_cp-1.pdf}
\caption{}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/ma_ri_cp-1.pdf}
\caption{}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/who_rec_med_cp-1.pdf}
\caption{}
\end{figure}

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\end{document}

### 2015-05-31 manuscript/manuscript.Rmd
Intro outline
re-ran cprob with just epa and who

@@
 ################################################################################
-mc_guides <- c(0.04, 0.3, 0.8, 1, 1.6, 4, 6, 10, 14, 20)
-mc_names <- c("MN","EPA_Child","CA","WHO","EPA_Adult","WHO","Multiple","Multiple","Mulitple","Multiple")
-mn_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
+mc_guides <- c(0.3, 1, 1.6, 2, 4, 10, 20)
+mc_names <- c("EPA_Child","WHO","EPA_Adult","WHO","WHO","WHO","WHO")
+
+epa_child_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                      xImpair = log(mc_guides[1]), ProbComp = "gt", Exceed = "gte",
                      ci = TRUE, R = 100)
-epa_child_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
+who_drink_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                      xImpair = log(mc_guides[2]), ProbComp = "gt", Exceed = "gte",
                      ci = TRUE, R = 100)
-ca_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
+epa_adult_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                      xImpair = log(mc_guides[3]), ProbComp = "gt", Exceed = "gte",
                      ci = TRUE, R = 100)
-who_drink_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
+who_rec_low1_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                      xImpair = log(mc_guides[4]), ProbComp = "gt", Exceed = "gte",
                      ci = TRUE, R = 100)
-epa_adult_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
+who_rec_low2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                      xImpair = log(mc_guides[5]), ProbComp = "gt", Exceed = "gte",
                      ci = TRUE, R = 100)
-who_rec_low_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
+who_rec_med1_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                      xImpair = log(mc_guides[6]), ProbComp = "gt", Exceed = "gte",
                      ci = TRUE, R = 100)
-mult_state_rec_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
+who_rec_med2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                      xImpair = log(mc_guides[7]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100)
-il_org_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log(mc_guides[8]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100)
-ma_ri_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log(mc_guides[9]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100)
-who_rec_med_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log(mc_guides[10]), ProbComp = "gt", Exceed = "gte",
                      ci = TRUE, R = 100)
@@
 
+EPA just released HA for MC
+
+MC requires lab work
+
+MC associated with Chl a
+
+Use association and cpa to id chl a concentration that indicative of exceeding HA
+
 #Methods
@@
 ```{r microcystin_levels,echo=FALSE,results='asis'}
-source_of_stnd <- c("WHO, IN",
-                    "WHO, IN, KS",
-                    "WHO, IN, IA, KS, NE, OH",
+source_of_stnd <- c("WHO",
                     "WHO",
-                    "WHO, OH, OR",
-                    "MN",
-                    "CA",
-                    "IL, OR",
-                    "MA, RI",
-                    "OH, VT, VA, WA",
+                    "WHO",
+                    "WHO",
+                    "WHO",
                     "U.S. EPA",
@@
                   "Drinking",
-                  "Recreational",
-                  "Recreational",
-                  "Recreational",
-                  "Recreational",
-                  "Drinking",
                   "Drinking")
@@
                    "1 ug/l",
-                   "0.04 ug/l",
-                   "0.8 ug/l",
-                   "10 ug/l",
-                   "14 ug/l",
-                   "6 ug/l",
                    "0.3 ug/l",
                    "1.6 ug/l")
-#note <- c("Low probability of adverse health effects, state advisories",
-#          "Moderate probability of adverse health effects, state advisories",
-#          "High probability of adverse health effects, state advisories, These higher #concentrations typically only seen in visible scums/algal mats",
-#          "Very high probability of adverse health effects, These higher concentrations #typically only seen in visible scums/algal mats",
-#          "","","","","","",
+#note <- c("Low probability of adverse health effects",
+#          "Moderate probability of adverse health effects",
+#          "High probability of adverse health effects, These higher #concentrations typically only seen in visible scums/algal mats",
+#          "Very high probability of adverse health effects",
+#          "",
 #          "Children younger than school age",
@@
 
-Based on the range of microcystin concentrations used in various policy settings, we evaulated associated chlorohpyll *a* concentrations for an effect at 0.04, 0.3, 0.8, 1, 1.6, 4, 6, 10, 14, and 20 ug/l.
+We evaulated associated chlorohpyll *a* concentrations for an effect for each of the WHO and EPA levels.  These were  0.3, 1, 1.6, 2, 4, 10, and 20 ug/l.
 
@@
 
-```{r mn_cp_plot,echo=FALSE}
-cp_plot(mn_cp)
-```
-
-\newpage
-
 ```{r epa_child_cp_plot,echo=FALSE}
 cp_plot(epa_child_cp)
-```
-
-```{r ca_cp_plot,echo=FALSE}
-cp_plot(ca_cp)
 ```
@@
 
-```{r who_rec_low_cp, echo=FALSE}
-cp_plot(who_rec_low_cp)
+```{r who_rec_low1_cp, echo=FALSE}
+cp_plot(who_rec_low1_cp)
 ```
@@
 
-```{r mult_state_rec_cp,  echo=FALSE}
-cp_plot(mult_state_rec_cp)
+```{r who_rec_low2_cp, echo=FALSE}
+cp_plot(who_rec_low2_cp)
 ```
-
 \newpage
 
-```{r il_org_cp, echo=FALSE}
-cp_plot(il_org_cp)
+```{r who_rec_med1_cp, echo=FALSE}
+cp_plot(who_rec_med1_cp)
 ```
-
 \newpage
 
-```{r ma_ri_cp, echo=FALSE}
-cp_plot(ma_ri_cp)
+```{r who_rec_med2_cp, echo=FALSE}
+cp_plot(who_rec_med2_cp)
 ```
-
-\newpage
-
-```{r who_rec_med_cp, echo=FALSE}
-cp_plot(who_rec_med_cp)
-```
-
 \newpage

### 2015-05-31 manuscript/manuscript.tex
re-ran cprob with just epa and who

@@
             pdfauthor={},
-            pdftitle={Associations between Chlorophyll a and Microcystin-LR Health Advisory Concentrations},
+            pdftitle={Associations between Chlorophyll a and various Microcystin-LR Health Advisory Concentrations},
             colorlinks=true,
@@
 
-\title{Associations between Chlorophyll \emph{a} and Microcystin-LR Health
-Advisory Concentrations}
+\title{Associations between Chlorophyll \emph{a} and various Microcystin-LR
+Health Advisory Concentrations}
 \author{
@@
 \begin{center}
-\huge Associations between Chlorophyll \emph{a} and Microcystin-LR Health
-Advisory Concentrations
+\huge Associations between Chlorophyll \emph{a} and various Microcystin-LR
+Health Advisory Concentrations
 \end{center}
@@
 
+EPA just released HA for MC
+
+MC requires lab work
+
+MC associated with Chl a
+
+Use association and cpa to id chl a concentration that indicative of
+exceeding HA
+
 \section{Methods}\label{methods}
@@
 \toprule
-\begin{minipage}[b]{0.27\columnwidth}\raggedright\strut
+\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
 Source
@@
 \endhead
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-WHO, OH, OR
+\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+WHO
 \strut\end{minipage} &
@@
 \strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-MN
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Drinking
-\strut\end{minipage} &
-\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-0.04 ug/l
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 U.S. EPA
@@
 \strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 U.S. EPA
@@
 \strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-WHO, IN
+\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+WHO
 \strut\end{minipage} &
@@
 \strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-WHO, IN, KS
+\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+WHO
 \strut\end{minipage} &
@@
 \strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-WHO, IN, IA, KS, NE, OH
+\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+WHO
 \strut\end{minipage} &
@@
 \strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 WHO
@@
 \strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-CA
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Recreational
-\strut\end{minipage} &
-\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-0.8 ug/l
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-IL, OR
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Recreational
-\strut\end{minipage} &
-\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-10 ug/l
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-MA, RI
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Recreational
-\strut\end{minipage} &
-\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-14 ug/l
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
-OH, VT, VA, WA
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Recreational
-\strut\end{minipage} &
-\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-6 ug/l
-\strut\end{minipage}\tabularnewline
 \bottomrule
@@
 
-Based on the range of microcystin concentrations used in various policy
-settings, we evaulated associated chlorohpyll \emph{a} concentrations
-for an effect at 0.04, 0.3, 0.8, 1, 1.6, 4, 6, 10, 14, and 20 ug/l.
+We evaulated associated chlorohpyll \emph{a} concentrations for an
+effect for each of the WHO and EPA levels. These were 0.3, 1, 1.6, 2, 4,
+10, and 20 ug/l.
 
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/mn_cp_plot-1.pdf}
-\caption{}
-\end{figure}
-
-\newpage
-
-\begin{figure}[htbp]
-\centering
 \includegraphics{manuscript_files/figure-latex/epa_child_cp_plot-1.pdf}
-\caption{}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/ca_cp_plot-1.pdf}
 \caption{}
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/who_rec_low_cp-1.pdf}
+\includegraphics{manuscript_files/figure-latex/who_rec_low1_cp-1.pdf}
 \caption{}
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/mult_state_rec_cp-1.pdf}
-\caption{}
-\end{figure}
-
+\includegraphics{manuscript_files/figure-latex/who_rec_low2_cp-1.pdf}
 \newpage
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/il_org_cp-1.pdf}
-\caption{}
-\end{figure}
-
+\includegraphics{manuscript_files/figure-latex/who_rec_med1_cp-1.pdf}
 \newpage
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/ma_ri_cp-1.pdf}
-\caption{}
-\end{figure}
-
-\newpage
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/who_rec_med_cp-1.pdf}
-\caption{}
-\end{figure}
-
+\includegraphics{manuscript_files/figure-latex/who_rec_med2_cp-1.pdf}
 \newpage

### 2015-06-01 R/get_chla.R
add new function for getting chla, cleaned up plot.

@@ -1,0 +1,11 @@
+#' get_chla
+#' 
+#' Returns Chl a value that is associated with a given probability of exceedence at the
+#' Upper bootstrapped confidence interval
+#' 
+#' @param cp_obj An object of class condprob is required.
+#' @param prob Exceedence probability to return chl a 
+#' @export
+get_chla <- function(cp_obj,prob){
+  
+}

### 2015-06-01 R/plot.condprob.R

@@ -6,7 +6,7 @@
 #' data, the impaired sites, and the unimpaired sites, and a conditional probability 
 #' plot with the conditional probability on the y-axis the stressor on the x-axis.  
 #' 
-#' @param condprobObj An object of class condprob is required.  Data and parameters
+#' @param cp_obj An object of class condprob is required.  Data and parameters
 #'                    from this object are used to create the plot
 #' @param min_samp minimum sample size to plot probabilities for. Defaults to 10.
 #' @param ... Used primarily to pass other ggplot2 plotting options to control title, 
@@ -21,8 +21,8 @@
 #' @export
 #' @examples
 #' cp_plot(adult_ha)
-cp_plot <- function(condprobObj, min_samp = 10, ...) {
-    dat<-data.frame(condprobObj[1:5])
+cp_plot <- function(cp_obj, min_samp = 10, ...) {
+    dat<-data.frame(cp_obj[1:5])
     dat<-dat[1:(nrow(dat)-min_samp),]
     cplot <- ggplot(dat,aes(x=X, y=Bootstrap.Probability)) +
       geom_point() + 

### 2015-06-01 R/get_chla.R
added get_chla

@@ -7,5 +7,6 @@
 #' @param prob Exceedence probability to return chl a 
 #' @export
 get_chla <- function(cp_obj,prob){
-  
+  transf_chla <- min(cp_obj$X[cp_obj$Upper.CI >= prob])
+  return(exp(transf_chla))
 }

### 2015-06-01 manuscript/manuscript.Rmd
rendered with output table

@@
 mc_names <- c("EPA_Child","WHO","EPA_Adult","WHO","WHO","WHO","WHO")
+mc_chla <- vector("numeric",length(mc_guides))
 
@@
                      ci = TRUE, R = 100)
+
+mc_chla[1] <- get_chla(epa_child_cp,0.5)
+
 who_drink_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
@@
                      ci = TRUE, R = 100)
+
+mc_chla[2] <- get_chla(who_drink_cp,0.5)
+
 epa_adult_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
@@
                      ci = TRUE, R = 100)
+
+mc_chla[3] <- get_chla(epa_adult_cp,0.5)
+
 who_rec_low1_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
@@
                      ci = TRUE, R = 100)
+
+mc_chla[4] <- get_chla(who_rec_low1_cp,0.5)
+
 who_rec_low2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
@@
                      ci = TRUE, R = 100)
+
+mc_chla[5] <- get_chla(who_rec_low2_cp,0.5)
+
 who_rec_med1_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
@@
                      ci = TRUE, R = 100)
+
+mc_chla[6] <- get_chla(who_rec_med1_cp,0.5)
+
 who_rec_med2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
@@
 
+mc_chla[7] <- get_chla(who_rec_med2_cp,0.5)
 ```
@@
 
+```{r mc_chla_table, echo=FALSE}
+mc_chla_tbl <- data.frame(Source=mc_names,Microcystin=mc_guides,Chlorophyll=mc_chla)
+pandoc.table(mc_chla_tbl,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
+```
+
 #Discussion

### 2015-06-01 manuscript/manuscript.tex
rendered with output table

@@
 
+\begin{verbatim}
+## 
+## -------------------------------------
+## Source    Microcystin   Chlorophyll  
+## --------- ------------- -------------
+## EPA_Child 0.3           23.68        
+## 
+## WHO       1             65.2         
+## 
+## EPA_Adult 1.6           79.8         
+## 
+## WHO       2             113.1        
+## 
+## WHO       4             273.6        
+## 
+## WHO       10            338.4        
+## 
+## WHO       20            338.4        
+## -------------------------------------
+\end{verbatim}
+
 \section{Discussion}\label{discussion}

### 2015-06-01 manuscript/manuscript.Rmd
re rendered with asis on table

@@
 
-```{r mc_chla_table, echo=FALSE}
+```{r mc_chla_table, results='asis', echo=FALSE}
 mc_chla_tbl <- data.frame(Source=mc_names,Microcystin=mc_guides,Chlorophyll=mc_chla)

### 2015-06-02 manuscript/manuscript.Rmd
draft abstract rendered

@@
 <!-- Abstract is being wrapped in latex here so that all analysis can be run in the chunk above and the results reproducibly referenced in the abstract. -->
+Cyanobacteria harmful algal blooms (cHABs) are associated with a wide arrange of adverse health effects that stem mostly from the presence of cyanotoxins.
+To help protect agains the impacts, several health advisory levels have been set for some of these toxins, in particular, one of the most common toxins, microcystin, has several advisory levels set for drinking water and recreational use and managing water bodies to meet those levels could have far reaching benefits.
+However, measuring micorcystin can not currently be done *in situ* and requires samples be processed in a lab.
+This time consuming and expensive.
+It is possible to find reliable indicators that may be estimated quickly and *in situ* as a first defense agains high level of microcystin.
+In particular, chlorophyll *a* has been shown to be postively associated with microcystin.
+In this paper we use this assocation to provide estimates of chlorophyll *a* that if exceeded would be indiciative of a higher likelihood of exceeding select concentrations of microcystin.
+Using the 2007 National Lakes Assessment and a conditional probability appoach that has been used in other water quality settings, we idenfify chlorophyll *a* concentrations that are more likely than not to be associated with an exceedance of a microcystin health advisory level.
+We look at the recent US EPA standards for drinking water as well as the World Health Organization levels for drinking water and recerational use.
+For microcystin concentrations of 0.3, 1, 1.6, 2.
+and 4 we find chlorophyll *a* concentrations of [R CODE HERE] respectively.
+When managing for these various microsystin levels exceed these reproted chlorophyll *a* concentratoins should be a trigger for further testing and possibly managment action.
 

### 2015-06-02 manuscript/manuscript.tex
re rendered with asis on table
draft abstract rendered

--- 2015-05-31 manuscript/manuscript.tex
@@
 
+Cyanobacteria harmful algal blooms (cHABs) are associated with a wide
+arrange of adverse health effects that stem mostly from the presence of
+cyanotoxins. To help protect agains the impacts, several health advisory
+levels have been set for some of these toxins, in particular, one of the
+most common toxins, microcystin, has several advisory levels set for
+drinking water and recreational use and managing water bodies to meet
+those levels could have far reaching benefits. However, measuring
+micorcystin can not currently be done \emph{in situ} and requires
+samples be processed in a lab. This time consuming and expensive. It is
+possible to find reliable indicators that may be estimated quickly and
+\emph{in situ} as a first defense agains high level of microcystin. In
+particular, chlorophyll \emph{a} has been shown to be postively
+associated with microcystin. In this paper we use this assocation to
+provide estimates of chlorophyll \emph{a} that if exceeded would be
+indiciative of a higher likelihood of exceeding select concentrations of
+microcystin. Using the 2007 National Lakes Assessment and a conditional
+probability appoach that has been used in other water quality settings,
+we idenfify chlorophyll \emph{a} concentrations that are more likely
+than not to be associated with an exceedance of a microcystin health
+advisory level. We look at the recent US EPA standards for drinking
+water as well as the World Health Organization levels for drinking water
+and recerational use. For microcystin concentrations of 0.3, 1, 1.6, 2.
+and 4 we find chlorophyll \emph{a} concentrations of {[}R CODE HERE{]}
+respectively. When managing for these various microsystin levels exceed
+these reproted chlorophyll \emph{a} concentratoins should be a trigger
+for further testing and possibly managment action.
+
 \vspace{3mm}
@@
 
+\begin{longtable}[c]{@{}lll@{}}
+\toprule
+\begin{minipage}[b]{0.13\columnwidth}\raggedright\strut
+Source
+\strut\end{minipage} &
+\begin{minipage}[b]{0.18\columnwidth}\raggedright\strut
+Microcystin
+\strut\end{minipage} &
+\begin{minipage}[b]{0.18\columnwidth}\raggedright\strut
+Chlorophyll
+\strut\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
+EPA\_Child
+\strut\end{minipage} &
+\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+0.3
+\strut\end{minipage} &
+\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+23.68
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
+WHO
+\strut\end{minipage} &
+\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+1
+\strut\end{minipage} &
+\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+65.2
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
+EPA\_Adult
+\strut\end{minipage} &
+\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+1.6
+\strut\end{minipage} &
+\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+79.8
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
+WHO
+\strut\end{minipage} &
+\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+2
+\strut\end{minipage} &
+\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+113.1
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
+WHO
+\strut\end{minipage} &
+\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+4
+\strut\end{minipage} &
+\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+273.6
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
+WHO
+\strut\end{minipage} &
+\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+10
+\strut\end{minipage} &
+\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+338.4
+\strut\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
+WHO
+\strut\end{minipage} &
+\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+20
+\strut\end{minipage} &
+\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+338.4
+\strut\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
+
 \section{Discussion}\label{discussion}

### 2015-06-04 manuscript/manuscript.Rmd
updated with chla in abstract
updates.  still needs to be rendered.

@@
 For microcystin concentrations of 0.3, 1, 1.6, 2.
-and 4 we find chlorophyll *a* concentrations of [R CODE HERE] respectively.
+and 4 we find chlorophyll *a* concentrations of `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, `r mc_chla[4]`, and `r mc_chla[5]`, respectively.
 When managing for these various microsystin levels exceed these reproted chlorophyll *a* concentratoins should be a trigger for further testing and possibly managment action.
@@
 
-EPA just released HA for MC
+In the summer of 2014, the city of Toledo, OH was forced to shut down their municipal water supply due in part to an excess of Microcystin-LR that resulted from a ongoin harmful algal bloom in Lake Erie.
+Since this event, signficant legislation has been passed in the United States and the US Environmental Protection Agency has released suggested microcystin-LR concentrations that would trigger health advisories.
 

### 2015-06-05 manuscript/manuscript.tex
updated with chla in abstract
rendered ms

@@
 and recerational use. For microcystin concentrations of 0.3, 1, 1.6, 2.
-and 4 we find chlorophyll \emph{a} concentrations of {[}R CODE HERE{]}
-respectively. When managing for these various microsystin levels exceed
-these reproted chlorophyll \emph{a} concentratoins should be a trigger
-for further testing and possibly managment action.
+and 4 we find chlorophyll \emph{a} concentrations of 23.68, 65.2, 79.8,
+113.14, and 273.6, respectively. When managing for these various
+microsystin levels exceed these reproted chlorophyll \emph{a}
+concentratoins should be a trigger for further testing and possibly
+managment action.
 
@@
 
-EPA just released HA for MC
+In the summer of 2014, the city of Toledo, OH was forced to shut down
+their municipal water supply due in part to an excess of Microcystin-LR
+that resulted from a ongoin harmful algal bloom in Lake Erie. Since this
+event, signficant legislation has been passed in the United States and
+the US Environmental Protection Agency has released suggested
+microcystin-LR concentrations that would trigger health advisories.
 

### 2015-06-06 manuscript/manuscript.Rmd
new text in intro and new render

@@
 In the summer of 2014, the city of Toledo, OH was forced to shut down their municipal water supply due in part to an excess of Microcystin-LR that resulted from a ongoin harmful algal bloom in Lake Erie.
-Since this event, signficant legislation has been passed in the United States and the US Environmental Protection Agency has released suggested microcystin-LR concentrations that would trigger health advisories.
+Since this event, signficant legislation has been passed in the United States and the US Environmental Protection Agency (USEPA) has released suggested microcystin-LR concentrations that would trigger health advisories.
+While these levels and association advisories are likey to help mitigate the impacts from harmful algal blooms, they are not without complications.
+One of these complications is that they rely on availble measurments of Microcysin-LR which requires taking regular water samples and having those samples process in a lab to determin the toxin concentration.
+This has the potentially to be costly and time consuming both factors which could limit monitoring efforts.
+Fortunately, microcystin-LR has been shown to be associated with several other, more easily measured components of water quality.
 
-MC requires lab work
+Chlorophyll *a* is one of the most commonly measured components of water quality that is also known to be strongly associated with Microsystin-LR concentrations [REFS].
 
 MC associated with Chl a
+These levels, along with the existing World Health Organization (WHO) levels represent a range concentrations that occur
 

### 2015-06-06 manuscript/manuscript.tex
new text in intro and new render

@@
 event, signficant legislation has been passed in the United States and
-the US Environmental Protection Agency has released suggested
+the US Environmental Protection Agency (USEPA) has released suggested
 microcystin-LR concentrations that would trigger health advisories.
+While these levels and association advisories are likey to help mitigate
+the impacts from harmful algal blooms, they are not without
+complications. One of these complications is that they rely on availble
+measurments of Microcysin-LR which requires taking regular water samples
+and having those samples process in a lab to determin the toxin
+concentration. This has the potentially to be costly and time consuming
+both factors which could limit monitoring efforts. Fortunately,
+microcystin-LR has been shown to be associated with several other, more
+easily measured components of water quality.
 
-MC requires lab work
+Chlorophyll \emph{a} is one of the most commonly measured components of
+water quality that is also known to be strongly associated with
+Microsystin-LR concentrations {[}REFS{]}.
 
-MC associated with Chl a
+MC associated with Chl a These levels, along with the existing World
+Health Organization (WHO) levels represent a range concentrations that
+occur
 

### 2015-06-07 manuscript/manuscript.Rmd
more intro and another render

@@
 
-In the summer of 2014, the city of Toledo, OH was forced to shut down their municipal water supply due in part to an excess of Microcystin-LR that resulted from a ongoin harmful algal bloom in Lake Erie.
+In the summer of 2014, the city of Toledo, OH was forced to shut down their municipal water supply due in part to an excess of Microcystin-LR that resulted from a ongoing cyanobacterial harmful algal bloom (cHAB) in Lake Erie.
 Since this event, signficant legislation has been passed in the United States and the US Environmental Protection Agency (USEPA) has released suggested microcystin-LR concentrations that would trigger health advisories.
 While these levels and association advisories are likey to help mitigate the impacts from harmful algal blooms, they are not without complications.
-One of these complications is that they rely on availble measurments of Microcysin-LR which requires taking regular water samples and having those samples process in a lab to determin the toxin concentration.
+
+One of these complications is that they rely on availble measurments of Microcysin-LR which requires taking regular water samples and having those samples process in a lab to determine the toxin concentration [REFS].
 This has the potentially to be costly and time consuming both factors which could limit monitoring efforts.
@@
 Chlorophyll *a* is one of the most commonly measured components of water quality that is also known to be strongly associated with Microsystin-LR concentrations [REFS].
-
-MC associated with Chl a
-These levels, along with the existing World Health Organization (WHO) levels represent a range concentrations that occur
+Additionally there are many rapid measurements for assessing chlorophyll *a* levels *in situ*.
+For instance, there are small or hand held flourometers that provide reliable measurements [REFS].
+Given these facts, it might be possible to identify chlorophyll *a* concentrations that would be associated with the various Microcystin-LR health advisory levels.
+Identifying these associations would provide another reliable tool for water resource managers to use to help manage the threat to public health posed by cHABs.
 

### 2015-06-07 manuscript/manuscript.tex
more intro and another render

@@
 their municipal water supply due in part to an excess of Microcystin-LR
-that resulted from a ongoin harmful algal bloom in Lake Erie. Since this
-event, signficant legislation has been passed in the United States and
-the US Environmental Protection Agency (USEPA) has released suggested
-microcystin-LR concentrations that would trigger health advisories.
-While these levels and association advisories are likey to help mitigate
-the impacts from harmful algal blooms, they are not without
-complications. One of these complications is that they rely on availble
-measurments of Microcysin-LR which requires taking regular water samples
-and having those samples process in a lab to determin the toxin
-concentration. This has the potentially to be costly and time consuming
+that resulted from a ongoing cyanobacterial harmful algal bloom (cHAB)
+in Lake Erie. Since this event, signficant legislation has been passed
+in the United States and the US Environmental Protection Agency (USEPA)
+has released suggested microcystin-LR concentrations that would trigger
+health advisories. While these levels and association advisories are
+likey to help mitigate the impacts from harmful algal blooms, they are
+not without complications.
+
+One of these complications is that they rely on availble measurments of
+Microcysin-LR which requires taking regular water samples and having
+those samples process in a lab to determine the toxin concentration
+{[}REFS{]}. This has the potentially to be costly and time consuming
 both factors which could limit monitoring efforts. Fortunately,
@@
 water quality that is also known to be strongly associated with
-Microsystin-LR concentrations {[}REFS{]}.
-
-MC associated with Chl a These levels, along with the existing World
-Health Organization (WHO) levels represent a range concentrations that
-occur
+Microsystin-LR concentrations {[}REFS{]}. Additionally there are many
+rapid measurements for assessing chlorophyll \emph{a} levels \emph{in
+situ}. For instance, there are small or hand held flourometers that
+provide reliable measurements {[}REFS{]}. Given these facts, it might be
+possible to identify chlorophyll \emph{a} concentrations that would be
+associated with the various Microcystin-LR health advisory levels.
+Identifying these associations would provide another reliable tool for
+water resource managers to use to help manage the threat to public
+health posed by cHABs.
 

### 2015-09-08 manuscript/manuscript.Rmd
cleaned up figs

@@
 ################################################################################
+if(!require(microcystinchla)){
   devtools::install_github("jhollist/microcystinchla",
                            auth_token = getOption("Github_Access"))
+}
 library("microcystinchla")
@@
 nla_dat <- get_nla(data_url)
-nla_dat <- mutate(nla_dat,log_chla = log(CHLA),log_micro = log(MCYST_TL_UGL))
+nla_dat <- mutate(nla_dat,log_chla = log10(CHLA),log_micro = log10(MCYST_TL_UGL))
 
@@
 epa_child_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log(mc_guides[1]), ProbComp = "gt", Exceed = "gte",
+                     xImpair = log10(mc_guides[1]), ProbComp = "gt", Exceed = "gte",
                      ci = TRUE, R = 100)
@@
 who_drink_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log(mc_guides[2]), ProbComp = "gt", Exceed = "gte",
+                     xImpair = log10(mc_guides[2]), ProbComp = "gt", Exceed = "gte",
                      ci = TRUE, R = 100)
@@
 epa_adult_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log(mc_guides[3]), ProbComp = "gt", Exceed = "gte",
+                     xImpair = log10(mc_guides[3]), ProbComp = "gt", Exceed = "gte",
                      ci = TRUE, R = 100)
@@
 who_rec_low1_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log(mc_guides[4]), ProbComp = "gt", Exceed = "gte",
+                     xImpair = log10(mc_guides[4]), ProbComp = "gt", Exceed = "gte",
                      ci = TRUE, R = 100)
@@
 who_rec_low2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log(mc_guides[5]), ProbComp = "gt", Exceed = "gte",
+                     xImpair = log10(mc_guides[5]), ProbComp = "gt", Exceed = "gte",
                      ci = TRUE, R = 100)
@@
 who_rec_med1_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log(mc_guides[6]), ProbComp = "gt", Exceed = "gte",
+                     xImpair = log10(mc_guides[6]), ProbComp = "gt", Exceed = "gte",
                      ci = TRUE, R = 100)
@@
 who_rec_med2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log(mc_guides[7]), ProbComp = "gt", Exceed = "gte",
+                     xImpair = log10(mc_guides[7]), ProbComp = "gt", Exceed = "gte",
                      ci = TRUE, R = 100)
@@
 <!-- Abstract is being wrapped in latex here so that all analysis can be run in the chunk above and the results reproducibly referenced in the abstract. -->
-Cyanobacteria harmful algal blooms (cHABs) are associated with a wide arrange of adverse health effects that stem mostly from the presence of cyanotoxins.
-To help protect agains the impacts, several health advisory levels have been set for some of these toxins, in particular, one of the most common toxins, microcystin, has several advisory levels set for drinking water and recreational use and managing water bodies to meet those levels could have far reaching benefits.
-However, measuring micorcystin can not currently be done *in situ* and requires samples be processed in a lab.
-This time consuming and expensive.
-It is possible to find reliable indicators that may be estimated quickly and *in situ* as a first defense agains high level of microcystin.
-In particular, chlorophyll *a* has been shown to be postively associated with microcystin.
-In this paper we use this assocation to provide estimates of chlorophyll *a* that if exceeded would be indiciative of a higher likelihood of exceeding select concentrations of microcystin.
+Cyanobacteria harmful algal blooms (cHABs) are associated with a wide range of adverse health effects that stem mostly from the presence of cyanotoxins.
+To help protect against these impacts, several health advisory levels have been set for some toxins.
+In particular, one of the most common toxins, microcystin, has several advisory levels set for drinking water and recreational use and managing water bodies to meet those levels could have far reaching benefits.
+However, compared to other water quality measures, measurements of microcystin are not common and current field measurement techniques have limited precision and accuracy.
+Addressing these issues will take time and resourced.
+Thus, there is utility in finding indicators of microcystin that are already widely available, can be estimated quickly and *in situ*, and used as a first defense against high levels of microcystin.
+In particular, chlorophyll *a* is very commonly measured, can be estimated *in situ*, and has been shown to be postively associated with microcystin.
+In this paper we use this assocation to provide estimates of chlorophyll *a* that if exceeded would be indiciative of a higher probability of exceeding select health advisory concentrations for microcystin-LR.
 Using the 2007 National Lakes Assessment and a conditional probability appoach that has been used in other water quality settings, we idenfify chlorophyll *a* concentrations that are more likely than not to be associated with an exceedance of a microcystin health advisory level.
 We look at the recent US EPA standards for drinking water as well as the World Health Organization levels for drinking water and recerational use.
-For microcystin concentrations of 0.3, 1, 1.6, 2.
-and 4 we find chlorophyll *a* concentrations of `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, `r mc_chla[4]`, and `r mc_chla[5]`, respectively.
-When managing for these various microsystin levels exceed these reproted chlorophyll *a* concentratoins should be a trigger for further testing and possibly managment action.
+For microcystin concentrations of 0.3, 1, 1.6, and 2 we find chlorophyll *a* concentrations of `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]`, respectively.
+When managing for these various microsystin levels exceeding these reproted chlorophyll *a* concentrations should be a trigger for further testing and possible managment action.
 
@@
 Since this event, signficant legislation has been passed in the United States and the US Environmental Protection Agency (USEPA) has released suggested microcystin-LR concentrations that would trigger health advisories.
-While these levels and association advisories are likey to help mitigate the impacts from harmful algal blooms, they are not without complications.
-
+While these levels and associated advisories are likey to help mitigate the impacts from harmful algal blooms, they are not without complications.
 One of these complications is that they rely on availble measurments of Microcysin-LR which requires taking regular water samples and having those samples process in a lab to determine the toxin concentration [REFS].
@@
 
-We evaulated associated chlorohpyll *a* concentrations for an effect for each of the WHO and EPA levels.  These were  0.3, 1, 1.6, 2, 4, 10, and 20 ug/l.
+We evaulated associated chlorohpyll *a* concentrations for an effect for each of the WHO and EPA levels.
+Lakes with higher microcystin-LR concentrations were rare.
+Only `r 100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=10])/nrow(nla_dat))` % of lakes sampled had a concentration greater than 10.
+For this analysis we focus on the microcystin concentrations that are better represented in the NLA data.
+These were  0.3, 1, 1.6, and 2 ug/l.
 
@@
 ```{r mc_chla_table, results='asis', echo=FALSE}
-mc_chla_tbl <- data.frame(Source=mc_names,Microcystin=mc_guides,Chlorophyll=mc_chla)
+mc_chla_tbl <- data.frame(Source=mc_names,Microcystin=mc_guides,Chlorophyll=mc_chla)[1:4,]
 pandoc.table(mc_chla_tbl,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
@@
 ```{r epa_child_cp_plot,echo=FALSE}
-cp_plot(epa_child_cp)
+plot_cp(epa_child_cp,x = "Log 10 Chlorophyll", y = "Probability of Exceeding EPA Child HA")
 ```
@@
 ```{r who_drink_cp_plot,echo=FALSE}
-cp_plot(who_drink_cp)
+plot_cp(who_drink_cp, x = "Log 10 Chlorophyll", y = "Probability of Exceeding WHO Drinking HA")
 ```
@@
 ```{r epa_adult_cp ,echo=FALSE}
-cp_plot(epa_adult_cp)
+plot_cp(epa_adult_cp, x = "Log 10 Chlorophyll", y = "Probability of Exceeding EPA Adult HA")
 ```
@@
 ```{r who_rec_low1_cp, echo=FALSE}
-cp_plot(who_rec_low1_cp)
+plot_cp(who_rec_low1_cp, x = "Log 10 Chlorophyll", y = "Probability of Exceeding WHO Low Probability HA")
 ```
@@
 
-```{r who_rec_low2_cp, echo=FALSE}
-cp_plot(who_rec_low2_cp)
-```
-\newpage
-
-```{r who_rec_med1_cp, echo=FALSE}
-cp_plot(who_rec_med1_cp)
-```
-\newpage
-
-```{r who_rec_med2_cp, echo=FALSE}
-cp_plot(who_rec_med2_cp)
-```
-\newpage
-
 #References

### 2015-09-08 manuscript/manuscript.tex
cleaned up figs

@@
 Cyanobacteria harmful algal blooms (cHABs) are associated with a wide
-arrange of adverse health effects that stem mostly from the presence of
-cyanotoxins. To help protect agains the impacts, several health advisory
-levels have been set for some of these toxins, in particular, one of the
+range of adverse health effects that stem mostly from the presence of
+cyanotoxins. To help protect against these impacts, several health
+advisory levels have been set for some toxins. In particular, one of the
 most common toxins, microcystin, has several advisory levels set for
 drinking water and recreational use and managing water bodies to meet
-those levels could have far reaching benefits. However, measuring
-micorcystin can not currently be done \emph{in situ} and requires
-samples be processed in a lab. This time consuming and expensive. It is
-possible to find reliable indicators that may be estimated quickly and
-\emph{in situ} as a first defense agains high level of microcystin. In
-particular, chlorophyll \emph{a} has been shown to be postively
-associated with microcystin. In this paper we use this assocation to
-provide estimates of chlorophyll \emph{a} that if exceeded would be
-indiciative of a higher likelihood of exceeding select concentrations of
-microcystin. Using the 2007 National Lakes Assessment and a conditional
-probability appoach that has been used in other water quality settings,
-we idenfify chlorophyll \emph{a} concentrations that are more likely
-than not to be associated with an exceedance of a microcystin health
-advisory level. We look at the recent US EPA standards for drinking
-water as well as the World Health Organization levels for drinking water
-and recerational use. For microcystin concentrations of 0.3, 1, 1.6, 2.
-and 4 we find chlorophyll \emph{a} concentrations of 23.68, 65.2, 79.8,
-113.14, and 273.6, respectively. When managing for these various
-microsystin levels exceed these reproted chlorophyll \emph{a}
-concentratoins should be a trigger for further testing and possibly
-managment action.
+those levels could have far reaching benefits. However, compared to
+other water quality measures, measurements of microcystin are not common
+and current field measurement techniques have limited precision and
+accuracy. Addressing these issues will take time and resourced. Thus,
+there is utility in finding indicators of microcystin that are already
+widely available, can be estimated quickly and \emph{in situ}, and used
+as a first defense against high levels of microcystin. In particular,
+chlorophyll \emph{a} is very commonly measured, can be estimated
+\emph{in situ}, and has been shown to be postively associated with
+microcystin. In this paper we use this assocation to provide estimates
+of chlorophyll \emph{a} that if exceeded would be indiciative of a
+higher probability of exceeding select health advisory concentrations
+for microcystin-LR. Using the 2007 National Lakes Assessment and a
+conditional probability appoach that has been used in other water
+quality settings, we idenfify chlorophyll \emph{a} concentrations that
+are more likely than not to be associated with an exceedance of a
+microcystin health advisory level. We look at the recent US EPA
+standards for drinking water as well as the World Health Organization
+levels for drinking water and recerational use. For microcystin
+concentrations of 0.3, 1, 1.6, and 2 we find chlorophyll \emph{a}
+concentrations of 25.1, 69.6, 84.96, and 113.14, respectively. When
+managing for these various microsystin levels exceeding these reproted
+chlorophyll \emph{a} concentrations should be a trigger for further
+testing and possible managment action.
 
@@
 We evaulated associated chlorohpyll \emph{a} concentrations for an
-effect for each of the WHO and EPA levels. These were 0.3, 1, 1.6, 2, 4,
-10, and 20 ug/l.
+effect for each of the WHO and EPA levels. Lakes with higher
+microcystin-LR concentrations were rare. Only 1.1616651 \% of lakes
+sampled had a concentration greater than 10. For this analysis we focus
+on the microcystin concentrations that are better represented in the NLA
+data. These were 0.3, 1, 1.6, and 2 ug/l.
 
@@
 \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
-23.68
+25.1
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
-65.2
+69.6
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
-79.8
+84.96
 \strut\end{minipage}\tabularnewline
@@
 113.1
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-WHO
-\strut\end{minipage} &
-\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
-4
-\strut\end{minipage} &
-\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
-273.6
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-WHO
-\strut\end{minipage} &
-\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
-10
-\strut\end{minipage} &
-\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
-338.4
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-WHO
-\strut\end{minipage} &
-\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
-20
-\strut\end{minipage} &
-\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
-338.4
 \strut\end{minipage}\tabularnewline
@@
 
-\includegraphics{manuscript_files/figure-latex/who_rec_low2_cp-1.pdf}
-\newpage
-
-\includegraphics{manuscript_files/figure-latex/who_rec_med1_cp-1.pdf}
-\newpage
-
-\includegraphics{manuscript_files/figure-latex/who_rec_med2_cp-1.pdf}
-\newpage
-
 \section*{References}\label{references}

### 2015-09-08 R/get_chla.R

@@ -8,5 +8,5 @@
 #' @export
 get_chla <- function(cp_obj,prob){
   transf_chla <- min(cp_obj$X[cp_obj$Upper.CI >= prob])
-  return(exp(transf_chla))
+  return(10^transf_chla)
 }

### 2015-09-08 R/plot_cp.R

@@ -1,0 +1,24 @@
+#'get_nla
+#'
+#'Create plot of conditional probability object showing association between microcystin-LR
+#'advisory levels and chlorophyll \italics{a} 
+#'
+#' @param cp A conditional probability object from condprob2
+#' @param ...  arguments for labels
+#' 
+#' @import ggplot2 viridis
+#' @export
+plot_cp<-function(cp, ...){
+  df<-data.frame(x=cp$X,prob=cp$Bootstrap.Probability,low=cp$Lower.CI,up=cp$Upper.CI)
+  ggp<-ggplot(df,aes(x=x,y=prob))+
+    geom_point(size=3)+
+    geom_ribbon(aes(ymin = low,ymax = up),alpha = 0.4)+
+    theme_bw()+
+    theme(text = element_text(family="sans"),
+          axis.title.x = element_text(family="sans",vjust = -0.5, size = 14),
+          axis.title.y = element_text(family="sans",vjust = 1.5, size = 14),
+          axis.text.x = element_text(family="sans",size = 12),
+          axis.text.y = element_text(family="sans",size = 12)) + 
+    labs(...)
+  return(ggp)
+}

### 2015-09-08 manuscript/manuscript.tex
new render and changed to bootstrapped probs in plots

@@
 has released suggested microcystin-LR concentrations that would trigger
-health advisories. While these levels and association advisories are
+health advisories. While these levels and associated advisories are
 likey to help mitigate the impacts from harmful algal blooms, they are
-not without complications.
-
-One of these complications is that they rely on availble measurments of
-Microcysin-LR which requires taking regular water samples and having
-those samples process in a lab to determine the toxin concentration
-{[}REFS{]}. This has the potentially to be costly and time consuming
-both factors which could limit monitoring efforts. Fortunately,
-microcystin-LR has been shown to be associated with several other, more
-easily measured components of water quality.
+not without complications.\\One of these complications is that they rely
+on availble measurments of Microcysin-LR which requires taking regular
+water samples and having those samples process in a lab to determine the
+toxin concentration {[}REFS{]}. This has the potentially to be costly
+and time consuming both factors which could limit monitoring efforts.
+Fortunately, microcystin-LR has been shown to be associated with several
+other, more easily measured components of water quality.
 

### 2015-09-09 manuscript/manuscript.Rmd
cleaned up some text and re-render

@@
 - affilnum: 2
-  affil: University of Rhode Island, Department of Natural Resrouces Science, Kingston, RI, 02882, USA
+  affil: University of Rhode Island, Department of Natural Resources Science, Kingston, RI, 02882, USA
 output:
@@
                            auth_token = getOption("Github_Access"))
+}
+if(!require(viridis)){
+  install.pacakges("viridis") 
 }
@@
 To help protect against these impacts, several health advisory levels have been set for some toxins.
-In particular, one of the most common toxins, microcystin, has several advisory levels set for drinking water and recreational use and managing water bodies to meet those levels could have far reaching benefits.
+In particular, one of the more common toxins, microcystin, has several advisory levels set for drinking water and recreational use and managing water bodies to meet those levels could have far reaching benefits.
 However, compared to other water quality measures, measurements of microcystin are not common and current field measurement techniques have limited precision and accuracy.
@@
 Thus, there is utility in finding indicators of microcystin that are already widely available, can be estimated quickly and *in situ*, and used as a first defense against high levels of microcystin.
-In particular, chlorophyll *a* is very commonly measured, can be estimated *in situ*, and has been shown to be postively associated with microcystin.
-In this paper we use this assocation to provide estimates of chlorophyll *a* that if exceeded would be indiciative of a higher probability of exceeding select health advisory concentrations for microcystin-LR.
-Using the 2007 National Lakes Assessment and a conditional probability appoach that has been used in other water quality settings, we idenfify chlorophyll *a* concentrations that are more likely than not to be associated with an exceedance of a microcystin health advisory level.
-We look at the recent US EPA standards for drinking water as well as the World Health Organization levels for drinking water and recerational use.
+In particular, chlorophyll *a* is very commonly measured, can be estimated *in situ*, and has been shown to be positively associated with microcystin.
+In this paper we use this association to provide estimates of chlorophyll *a* that if exceeded would be indicative of a higher probability of exceeding select health advisory concentrations for microcystin-LR.
+Using the 2007 National Lakes Assessment and a conditional probability approach that has been used in other water quality settings, we identify chlorophyll *a* concentrations that are more likely than not to be associated with an exceedance of a microcystin health advisory level.
+We look at the recent US EPA standards for drinking water as well as the World Health Organization levels for drinking water and recreational use.
 For microcystin concentrations of 0.3, 1, 1.6, and 2 we find chlorophyll *a* concentrations of `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]`, respectively.
-When managing for these various microsystin levels exceeding these reproted chlorophyll *a* concentrations should be a trigger for further testing and possible managment action.
+When managing for these various microcystin levels exceeding these reported chlorophyll *a* concentrations should be a trigger for further testing and possible management action.
 
@@
 
-In the summer of 2014, the city of Toledo, OH was forced to shut down their municipal water supply due in part to an excess of Microcystin-LR that resulted from a ongoing cyanobacterial harmful algal bloom (cHAB) in Lake Erie.
-Since this event, signficant legislation has been passed in the United States and the US Environmental Protection Agency (USEPA) has released suggested microcystin-LR concentrations that would trigger health advisories.
-While these levels and associated advisories are likey to help mitigate the impacts from harmful algal blooms, they are not without complications.
-One of these complications is that they rely on availble measurments of Microcysin-LR which requires taking regular water samples and having those samples process in a lab to determine the toxin concentration [REFS].
+In the summer of 2014, the city of Toledo, OH was forced to shut down their municipal water supply due in part to an excess of Microcystin-LR that resulted from a ongoing cyanobacteria harmful algal bloom (cHAB) in Lake Erie.
+Since this event, significant legislation has been passed in the United States and the US Environmental Protection Agency (USEPA) has released suggested microcystin-LR concentrations that would trigger health advisories.
+While these levels and associated advisories are likely to help mitigate the impacts from harmful algal blooms, they are not without complications.
+
+One of these complications is that they rely on available measurements of microcystin-LR.
+This toxin can be measured in the field using test strips but these are a coarse measure at best and currently available test strips focus on 1 and 10 $\mu$g/L [REFS].
+Measurements with greater accuracy and precision requires taking regular water samples and having those samples process in a lab to determine the toxin concentration [REFS].
 This has the potentially to be costly and time consuming both factors which could limit monitoring efforts.
@@
 
-Chlorophyll *a* is one of the most commonly measured components of water quality that is also known to be strongly associated with Microsystin-LR concentrations [REFS].
+Chlorophyll *a* is a very commonly measured components of water quality that is also known to be associated with Microsystin-LR concentrations [REFS].
 Additionally there are many rapid measurements for assessing chlorophyll *a* levels *in situ*.
@@
 
-We evaulated associated chlorohpyll *a* concentrations for an effect for each of the WHO and EPA levels.
+We evaluated associated chlorophyll *a* concentrations for an effect for each of the WHO and EPA levels.
 Lakes with higher microcystin-LR concentrations were rare.
@@
 For this analysis we focus on the microcystin concentrations that are better represented in the NLA data.
-These were  0.3, 1, 1.6, and 2 ug/l.
-
-##Data and Study Area
+These were  0.3, 1, 1.6, and 2 $\mu$g/L.
 

### 2015-09-09 manuscript/manuscript.tex
cleaned up some text and re-render

--- 2015-09-08 manuscript/manuscript.tex
@@
 \\*
-\textsuperscript{2}University of Rhode Island, Department of Natural Resrouces Science,
+\textsuperscript{2}University of Rhode Island, Department of Natural Resources Science,
 Kingston, RI, 02882, USA\\*
@@
 advisory levels have been set for some toxins. In particular, one of the
-most common toxins, microcystin, has several advisory levels set for
+more common toxins, microcystin, has several advisory levels set for
 drinking water and recreational use and managing water bodies to meet
@@
 chlorophyll \emph{a} is very commonly measured, can be estimated
-\emph{in situ}, and has been shown to be postively associated with
-microcystin. In this paper we use this assocation to provide estimates
-of chlorophyll \emph{a} that if exceeded would be indiciative of a
-higher probability of exceeding select health advisory concentrations
-for microcystin-LR. Using the 2007 National Lakes Assessment and a
-conditional probability appoach that has been used in other water
-quality settings, we idenfify chlorophyll \emph{a} concentrations that
+\emph{in situ}, and has been shown to be positively associated with
+microcystin. In this paper we use this association to provide estimates
+of chlorophyll \emph{a} that if exceeded would be indicative of a higher
+probability of exceeding select health advisory concentrations for
+microcystin-LR. Using the 2007 National Lakes Assessment and a
+conditional probability approach that has been used in other water
+quality settings, we identify chlorophyll \emph{a} concentrations that
 are more likely than not to be associated with an exceedance of a
@@
 standards for drinking water as well as the World Health Organization
-levels for drinking water and recerational use. For microcystin
+levels for drinking water and recreational use. For microcystin
 concentrations of 0.3, 1, 1.6, and 2 we find chlorophyll \emph{a}
 concentrations of 25.1, 69.6, 84.96, and 113.14, respectively. When
-managing for these various microsystin levels exceeding these reproted
+managing for these various microcystin levels exceeding these reported
 chlorophyll \emph{a} concentrations should be a trigger for further
-testing and possible managment action.
+testing and possible management action.
 
@@
 their municipal water supply due in part to an excess of Microcystin-LR
-that resulted from a ongoing cyanobacterial harmful algal bloom (cHAB)
-in Lake Erie. Since this event, signficant legislation has been passed
-in the United States and the US Environmental Protection Agency (USEPA)
-has released suggested microcystin-LR concentrations that would trigger
-health advisories. While these levels and association advisories are
-likey to help mitigate the impacts from harmful algal blooms, they are
+that resulted from a ongoing cyanobacteria harmful algal bloom (cHAB) in
+Lake Erie. Since this event, significant legislation has been passed in
+the United States and the US Environmental Protection Agency (USEPA) has
+released suggested microcystin-LR concentrations that would trigger
+health advisories. While these levels and associated advisories are
+likely to help mitigate the impacts from harmful algal blooms, they are
 not without complications.
 
-One of these complications is that they rely on availble measurments of
-Microcysin-LR which requires taking regular water samples and having
-those samples process in a lab to determine the toxin concentration
-{[}REFS{]}. This has the potentially to be costly and time consuming
-both factors which could limit monitoring efforts. Fortunately,
-microcystin-LR has been shown to be associated with several other, more
-easily measured components of water quality.
+One of these complications is that they rely on available measurements
+of microcystin-LR. This toxin can be measured in the field using test
+strips but these are a coarse measure at best and currently available
+test strips focus on 1 and 10 \(\mu\)g/L {[}REFS{]}. Measurements with
+greater accuracy and precision requires taking regular water samples and
+having those samples process in a lab to determine the toxin
+concentration {[}REFS{]}. This has the potentially to be costly and time
+consuming both factors which could limit monitoring efforts.
+Fortunately, microcystin-LR has been shown to be associated with several
+other, more easily measured components of water quality.
 
-Chlorophyll \emph{a} is one of the most commonly measured components of
-water quality that is also known to be strongly associated with
-Microsystin-LR concentrations {[}REFS{]}. Additionally there are many
-rapid measurements for assessing chlorophyll \emph{a} levels \emph{in
-situ}. For instance, there are small or hand held flourometers that
-provide reliable measurements {[}REFS{]}. Given these facts, it might be
+Chlorophyll \emph{a} is a very commonly measured components of water
+quality that is also known to be associated with Microsystin-LR
+concentrations {[}REFS{]}. Additionally there are many rapid
+measurements for assessing chlorophyll \emph{a} levels \emph{in situ}.
+For instance, there are small or hand held flourometers that provide
+reliable measurements {[}REFS{]}. Given these facts, it might be
 possible to identify chlorophyll \emph{a} concentrations that would be
@@
 
-We evaulated associated chlorohpyll \emph{a} concentrations for an
+We evaluated associated chlorophyll \emph{a} concentrations for an
 effect for each of the WHO and EPA levels. Lakes with higher
@@
 on the microcystin concentrations that are better represented in the NLA
-data. These were 0.3, 1, 1.6, and 2 ug/l.
-
-\subsection{Data and Study Area}\label{data-and-study-area}
+data. These were 0.3, 1, 1.6, and 2 \(\mu\)g/L.
 

### 2015-09-09 manuscript/manuscript.Rmd
new render

@@
 ---
-title: 'Associations between Chlorophyll *a* and various Microcystin-LR Health Advisory Concentrations'
+title: Associations between Chlorophyll *a* and various Microcystin-LR Health Advisory
+  Concentrations
 author:
-- name: Jeffrey W. Hollister
+- affilnum: 1
+  email: hollister.jeff@epa.gov
+  name: Jeffrey W. Hollister
+- affilnum: 1
+  name: Betty J. Kreakie
+- affilnum: 2
+  name: Dorothy Q. Kellog
+affiliation:
+- affil: US Environmental Protection Agency, Office of Research and Development, National
+    Health and Environmental Effects Research Laboratory, Atlantic Ecology Division,
+    27 Tarzwell Drive  Narragansett, RI, 02882, USA
   affilnum: 1
-  email: hollister.jeff@epa.gov
-- name:  Betty J. Kreakie
-  affilnum: 1
-- name: Dorothy Q. Kellog
+- affil: University of Rhode Island, Department of Natural Resources Science, Kingston,
+    RI, 02882, USA
   affilnum: 2
-affiliation:
-- affilnum: 1
-  affil: US Environmental Protection Agency, Office of Research and Development, National Health and Environmental Effects Research Laboratory, Atlantic Ecology Division, 27 Tarzwell Drive  Narragansett, RI, 02882, USA
-- affilnum: 2
-  affil: University of Rhode Island, Department of Natural Resources Science, Kingston, RI, 02882, USA
 output:
@@
   word_document: null
-fontsize: 11pt
 capsize: normalsize
@@
 documentclass: article
-spacing: doublespacing
+fontsize: 11pt
 linenumbers: no
 bibliography: components/manuscript.bib
+spacing: doublespacing
 abstract: no
@@
 
-In the summer of 2014, the city of Toledo, OH was forced to shut down their municipal water supply due in part to an excess of Microcystin-LR that resulted from a ongoing cyanobacteria harmful algal bloom (cHAB) in Lake Erie.
+In the summer of 2014, the city of Toledo, OH was forced to shut down their municipal water supply due in part to an excess of microcystin-LR that resulted from a ongoing cyanobacteria harmful algal bloom (cHAB) in Lake Erie [REFS].
 Since this event, significant legislation has been passed in the United States and the US Environmental Protection Agency (USEPA) has released suggested microcystin-LR concentrations that would trigger health advisories.
+MORE ON THE LEVELS.
 While these levels and associated advisories are likely to help mitigate the impacts from harmful algal blooms, they are not without complications.
@@
 This toxin can be measured in the field using test strips but these are a coarse measure at best and currently available test strips focus on 1 and 10 $\mu$g/L [REFS].
-Measurements with greater accuracy and precision requires taking regular water samples and having those samples process in a lab to determine the toxin concentration [REFS].
+Measurements with greater accuracy and precision require taking regular water samples and having those samples processed in a lab to determine the toxin concentration [REFS].
 This has the potentially to be costly and time consuming both factors which could limit monitoring efforts.
+Additionally, microcystin-LR is not currently a routine water quality measurment and thus availability of data will continue to be a problem.
 Fortunately, microcystin-LR has been shown to be associated with several other, more easily measured components of water quality.

### 2015-09-09 manuscript/manuscript.tex
new render

@@
 In the summer of 2014, the city of Toledo, OH was forced to shut down
-their municipal water supply due in part to an excess of Microcystin-LR
+their municipal water supply due in part to an excess of microcystin-LR
 that resulted from a ongoing cyanobacteria harmful algal bloom (cHAB) in
-Lake Erie. Since this event, significant legislation has been passed in
-the United States and the US Environmental Protection Agency (USEPA) has
-released suggested microcystin-LR concentrations that would trigger
-health advisories. While these levels and associated advisories are
-likely to help mitigate the impacts from harmful algal blooms, they are
-not without complications.
+Lake Erie {[}REFS{]}. Since this event, significant legislation has been
+passed in the United States and the US Environmental Protection Agency
+(USEPA) has released suggested microcystin-LR concentrations that would
+trigger health advisories. MORE ON THE LEVELS. While these levels and
+associated advisories are likely to help mitigate the impacts from
+harmful algal blooms, they are not without complications.
 
@@
 test strips focus on 1 and 10 \(\mu\)g/L {[}REFS{]}. Measurements with
-greater accuracy and precision requires taking regular water samples and
-having those samples process in a lab to determine the toxin
+greater accuracy and precision require taking regular water samples and
+having those samples processed in a lab to determine the toxin
 concentration {[}REFS{]}. This has the potentially to be costly and time
 consuming both factors which could limit monitoring efforts.
+Additionally, microcystin-LR is not currently a routine water quality
+measurment and thus availability of data will continue to be a problem.
 Fortunately, microcystin-LR has been shown to be associated with several

### 2015-09-10 manuscript/manuscript.Rmd
re-render drafty intro

@@
 However, compared to other water quality measures, measurements of microcystin are not common and current field measurement techniques have limited precision and accuracy.
-Addressing these issues will take time and resourced.
+Addressing these issues will take time and resources.
 Thus, there is utility in finding indicators of microcystin that are already widely available, can be estimated quickly and *in situ*, and used as a first defense against high levels of microcystin.
@@
 This toxin can be measured in the field using test strips but these are a coarse measure at best and currently available test strips focus on 1 and 10 $\mu$g/L [REFS].
-Measurements with greater accuracy and precision require taking regular water samples and having those samples processed in a lab to determine the toxin concentration [REFS].
-This has the potentially to be costly and time consuming both factors which could limit monitoring efforts.
-Additionally, microcystin-LR is not currently a routine water quality measurment and thus availability of data will continue to be a problem.
+Measurements with greater accuracy and precision require taking water samples and processing those in a lab to determine the toxin concentration [REFS].
+Additionally, microcystin-LR is, currently, not a routinely collected water quality parameter,thus, availability of microcystin-LR data may limit our ability to screen water bodies for exceedances of the various health advisories concentrations.
+Until microcystin-LR concentrations are more widely collected an alternative measure is needed.
 Fortunately, microcystin-LR has been shown to be associated with several other, more easily measured components of water quality.
@@
 Given these facts, it might be possible to identify chlorophyll *a* concentrations that would be associated with the various Microcystin-LR health advisory levels.
-Identifying these associations would provide another reliable tool for water resource managers to use to help manage the threat to public health posed by cHABs.
-
-Use association and cpa to id chl a concentration that indicative of exceeding HA
+Identifying these associations would provide another reliable tool for water resource managers to use to help manage the threat to public health posed by cHABs and would be especially useful in the absence of microcystin-LR concentrations.
+Thus, the goal of this paper is to utlize the National Lakes Assessment data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin-LR health advisory concentrations [NLA REF].
+So that others may repeat this analysis, the data, code, and this manuscript are freely available via [https://github.com/USAPE/microcystinchla](https://github.com/USAPE/microcystinchla).
 

### 2015-09-10 manuscript/manuscript.tex
re-render drafty intro

@@
 and current field measurement techniques have limited precision and
-accuracy. Addressing these issues will take time and resourced. Thus,
+accuracy. Addressing these issues will take time and resources. Thus,
 there is utility in finding indicators of microcystin that are already
@@
 test strips focus on 1 and 10 \(\mu\)g/L {[}REFS{]}. Measurements with
-greater accuracy and precision require taking regular water samples and
-having those samples processed in a lab to determine the toxin
-concentration {[}REFS{]}. This has the potentially to be costly and time
-consuming both factors which could limit monitoring efforts.
-Additionally, microcystin-LR is not currently a routine water quality
-measurment and thus availability of data will continue to be a problem.
-Fortunately, microcystin-LR has been shown to be associated with several
-other, more easily measured components of water quality.
+greater accuracy and precision require taking water samples and
+processing those in a lab to determine the toxin concentration
+{[}REFS{]}. Additionally, microcystin-LR is, currently, not a routinely
+collected water quality parameter,thus, availability of microcystin-LR
+data may limit our ability to screen water bodies for exceedances of the
+various health advisories concentrations. Until microcystin-LR
+concentrations are more widely collected an alternative measure is
+needed. Fortunately, microcystin-LR has been shown to be associated with
+several other, more easily measured components of water quality.
 
@@
 water resource managers to use to help manage the threat to public
-health posed by cHABs.
-
-Use association and cpa to id chl a concentration that indicative of
-exceeding HA
+health posed by cHABs and would be especially useful in the absence of
+microcystin-LR concentrations. Thus, the goal of this paper is to utlize
+the National Lakes Assessment data and identify chlorophyll \emph{a}
+concentrations that are associated with higher probabilities of
+exceeding several microcystin-LR health advisory concentrations {[}NLA
+REF{]}. So that others may repeat this analysis, the data, code, and
+this manuscript are freely available via
+\href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USAPE/microcystinchla}}.
 

### 2015-09-10 R/get_nla.R
Added weight to get data function

@@ -13,6 +13,6 @@ get_nla<-function(data_url){
     filter(VISIT_NO == 1) %>%
     filter(SITE_TYPE == "PROB_Lake") %>%
     select(SITE_ID,LAT_DD,LON_DD,ST,EPA_REG,WSA_ECO3,WSA_ECO9,LAKE_ORIGIN,CHLA,
-           MCYST_TL_UGL)
+           MCYST_TL_UGL,WGT_NLA)
   return(dat)
 }

### 2015-09-10 manuscript/manuscript.Rmd
reorg tables and figures, rerender pdf

@@
 mc_guides <- c(0.3, 1, 1.6, 2, 4, 10, 20)
-mc_names <- c("EPA_Child","WHO","EPA_Adult","WHO","WHO","WHO","WHO")
+mc_names <- c("U.S. EPA","WHO","U.S. EPA","WHO","WHO","WHO","WHO")
+mc_type <- c("Drinking","Drinking","Drinking","Recreational","Recreational","Recreational","Recreational")
 mc_chla <- vector("numeric",length(mc_guides))
+wgt <- nla_dat$WGT_NLA
+#wgt<- rep(1,nrow(nla_dat))
 
@@
                      xImpair = log10(mc_guides[1]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100)
+                     ci = TRUE, R = 100, xW=wgt)
 
@@
                      xImpair = log10(mc_guides[2]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100)
+                     ci = TRUE, R = 100, xW=wgt)
 
@@
                      xImpair = log10(mc_guides[3]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100)
+                     ci = TRUE, R = 100, xW=wgt)
 
@@
                      xImpair = log10(mc_guides[4]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100)
+                     ci = TRUE, R = 100, xW=wgt)
 
@@
                      xImpair = log10(mc_guides[5]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100)
+                     ci = TRUE, R = 100, xW=wgt)
 
@@
                      xImpair = log10(mc_guides[6]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100)
+                     ci = TRUE, R = 100, xW=wgt)
 
@@
                      xImpair = log10(mc_guides[7]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100)
+                     ci = TRUE, R = 100, xW=wgt)
 
@@
 Thus, the goal of this paper is to utlize the National Lakes Assessment data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin-LR health advisory concentrations [NLA REF].
-So that others may repeat this analysis, the data, code, and this manuscript are freely available via [https://github.com/USAPE/microcystinchla](https://github.com/USAPE/microcystinchla).
+So that others may repeat or adjust this analysis, the data, code, and this manuscript are freely available via [https://github.com/USAPE/microcystinchla](https://github.com/USAPE/microcystinchla).
 
 #Methods
+We evaluated associated chlorophyll *a* concentrations for an effect for each of the WHO and U.S.
+EPA levels (Table \ref{tab:microcystin_levels}).
+Lakes with higher microcystin-LR concentrations were rare.
+Only `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=10])/nrow(nla_dat)),2)` % of lakes sampled had a concentration greater than 10.
+For this analysis we focus on the microcystin concentrations that are better represented in the NLA data.
+These were  0.3, 1, 1.6, and 2 $\mu$g/L.
 
-```{r microcystin_levels,echo=FALSE,results='asis'}
+#Results
+
+(Table \ref{tab:mc_chla_table})
+
+(Figure \ref{fig:epa_child_cp_plot})
+
+(Figure \ref{fig:who_drink_cp_plot})
+
+(Figure \ref{fig:epa_adult_cp})
+
+(Figure \ref{fig:who_rec_low1_cp})
+
+#Discussion
+
+#Figures
+
+```{r epa_child_cp_plot,echo=FALSE, fig.cap="Plot of \\label{fig:epa_child_cp_plot}"}
+plot_cp(epa_child_cp,x = "Log 10 Chlorophyll", y = "Probability of Exceeding EPA Child HA")
+```
+
+\newpage
+
+```{r who_drink_cp_plot,echo=FALSE, fig.cap="Plot of  \\label{fig:who_drink_cp_plot}"}
+plot_cp(who_drink_cp, x = "Log 10 Chlorophyll", y = "Probability of Exceeding WHO Drinking HA")
+```
+
+\newpage
+
+```{r epa_adult_cp ,echo=FALSE, fig.cap="Plot of  \\label{fig:epa_adult_cp}"}
+plot_cp(epa_adult_cp, x = "Log 10 Chlorophyll", y = "Probability of Exceeding EPA Adult HA")
+```
+
+\newpage
+
+```{r who_rec_low1_cp, echo=FALSE, fig.cap="Plot of  \\label{fig:who_rec_low1_cp}"}
+plot_cp(who_rec_low1_cp, x = "Log 10 Chlorophyll", y = "Probability of Exceeding WHO Low Probability HA")
+```
+
+\newpage
+
+#Tables
+
+```{r microcystin_levels,echo=FALSE,results='asis', fig.cap="Various suggested microcystin-LR health advisory concentrations. \\label{fig:microcystin_levels}"}
 source_of_stnd <- c("WHO",
@@
 
-We evaluated associated chlorophyll *a* concentrations for an effect for each of the WHO and EPA levels.
-Lakes with higher microcystin-LR concentrations were rare.
-Only `r 100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=10])/nrow(nla_dat))` % of lakes sampled had a concentration greater than 10.
-For this analysis we focus on the microcystin concentrations that are better represented in the NLA data.
-These were  0.3, 1, 1.6, and 2 $\mu$g/L.
+\newpage
 
-#Results
-
-```{r mc_chla_table, results='asis', echo=FALSE}
-mc_chla_tbl <- data.frame(Source=mc_names,Microcystin=mc_guides,Chlorophyll=mc_chla)[1:4,]
+```{r mc_chla_table, results='asis', echo=FALSE, fig.cap="Chlorophyll \textit{a} concentrations that are associated with a 50% probability of exceeding a microcystin-LR health advisory concentration.
+\\label{fig:mc_chla_table}"}
+mc_chla_tbl <- data.frame(Source=mc_names,Type=mc_type,Microcystin=mc_guides,Chlorophyll=mc_chla)[1:4,]
 pandoc.table(mc_chla_tbl,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
@@
 
-#Discussion
-
-#Figures
-
-```{r epa_child_cp_plot,echo=FALSE}
-plot_cp(epa_child_cp,x = "Log 10 Chlorophyll", y = "Probability of Exceeding EPA Child HA")
-```
-
-\newpage
-
-```{r who_drink_cp_plot,echo=FALSE}
-plot_cp(who_drink_cp, x = "Log 10 Chlorophyll", y = "Probability of Exceeding WHO Drinking HA")
-```
-
-\newpage
-
-```{r epa_adult_cp ,echo=FALSE}
-plot_cp(epa_adult_cp, x = "Log 10 Chlorophyll", y = "Probability of Exceeding EPA Adult HA")
-```
-
-\newpage
-
-```{r who_rec_low1_cp, echo=FALSE}
-plot_cp(who_rec_low1_cp, x = "Log 10 Chlorophyll", y = "Probability of Exceeding WHO Low Probability HA")
-```
-
-\newpage
-
 #References

### 2015-09-10 manuscript/manuscript.tex
reorg tables and figures, rerender pdf

@@
 concentrations of 0.3, 1, 1.6, and 2 we find chlorophyll \emph{a}
-concentrations of 25.1, 69.6, 84.96, and 113.14, respectively. When
+concentrations of 20.8, 71.2, 92.16, and 110.16, respectively. When
 managing for these various microcystin levels exceeding these reported
@@
 exceeding several microcystin-LR health advisory concentrations {[}NLA
-REF{]}. So that others may repeat this analysis, the data, code, and
-this manuscript are freely available via
+REF{]}. So that others may repeat or adjust this analysis, the data,
+code, and this manuscript are freely available via
 \href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USAPE/microcystinchla}}.
@@
 \section{Methods}\label{methods}
+
+We evaluated associated chlorophyll \emph{a} concentrations for an
+effect for each of the WHO and U.S. EPA levels (Table
+\ref{tab:microcystin_levels}). Lakes with higher microcystin-LR
+concentrations were rare. Only 1.16 \% of lakes sampled had a
+concentration greater than 10. For this analysis we focus on the
+microcystin concentrations that are better represented in the NLA data.
+These were 0.3, 1, 1.6, and 2 \(\mu\)g/L.
+
+\section{Results}\label{results}
+
+(Table \ref{tab:mc_chla_table})
+
+(Figure \ref{fig:epa_child_cp_plot})
+
+(Figure \ref{fig:who_drink_cp_plot})
+
+(Figure \ref{fig:epa_adult_cp})
+
+(Figure \ref{fig:who_rec_low1_cp})
+
+\section{Discussion}\label{discussion}
+
+\section{Figures}\label{figures}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{manuscript_files/figure-latex/epa_child_cp_plot-1.pdf}
+\caption{Plot of \label{fig:epa_child_cp_plot}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{manuscript_files/figure-latex/who_drink_cp_plot-1.pdf}
+\caption{Plot of \label{fig:who_drink_cp_plot}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{manuscript_files/figure-latex/epa_adult_cp -1.pdf}
+\caption{Plot of \label{fig:epa_adult_cp}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{manuscript_files/figure-latex/who_rec_low1_cp-1.pdf}
+\caption{Plot of \label{fig:who_rec_low1_cp}}
+\end{figure}
+
+\newpage
+
+\section{Tables}\label{tables}
 
@@
 
-We evaluated associated chlorophyll \emph{a} concentrations for an
-effect for each of the WHO and EPA levels. Lakes with higher
-microcystin-LR concentrations were rare. Only 1.1616651 \% of lakes
-sampled had a concentration greater than 10. For this analysis we focus
-on the microcystin concentrations that are better represented in the NLA
-data. These were 0.3, 1, 1.6, and 2 \(\mu\)g/L.
+\newpage
 
-\section{Results}\label{results}
+\begin{verbatim}
+## Warning in split.large.cells(t): length of split.cells vector is smaller
+## than data. Default value will be used for other cells
+\end{verbatim}
 
-\begin{longtable}[c]{@{}lll@{}}
+\begin{verbatim}
+## Warning in split.large.cells(t.colnames): length of split.cells vector is
+## smaller than data. Default value will be used for other cells
+\end{verbatim}
+
+\begin{longtable}[c]{@{}llll@{}}
 \toprule
-\begin{minipage}[b]{0.13\columnwidth}\raggedright\strut
+\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
 Source
 \strut\end{minipage} &
-\begin{minipage}[b]{0.18\columnwidth}\raggedright\strut
+\begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+Type
+\strut\end{minipage} &
+\begin{minipage}[b]{0.17\columnwidth}\raggedright\strut
 Microcystin
 \strut\end{minipage} &
-\begin{minipage}[b]{0.18\columnwidth}\raggedright\strut
+\begin{minipage}[b]{0.17\columnwidth}\raggedright\strut
 Chlorophyll
@@
 \endhead
-\begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-EPA\_Child
+\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+U.S. EPA
 \strut\end{minipage} &
-\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+Drinking
+\strut\end{minipage} &
+\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
 0.3
 \strut\end{minipage} &
-\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
-25.1
+\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
+20.8
 \strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 WHO
 \strut\end{minipage} &
-\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+Drinking
+\strut\end{minipage} &
+\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
 1
 \strut\end{minipage} &
-\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
-69.6
+\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
+71.2
 \strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-EPA\_Adult
+\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+U.S. EPA
 \strut\end{minipage} &
-\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+Drinking
+\strut\end{minipage} &
+\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
 1.6
 \strut\end{minipage} &
-\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
-84.96
+\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
+92.16
 \strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 WHO
 \strut\end{minipage} &
-\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
+Recreational
+\strut\end{minipage} &
+\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
 2
 \strut\end{minipage} &
-\begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
-113.1
+\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
+110.2
 \strut\end{minipage}\tabularnewline
@@
 \end{longtable}
-
-\section{Discussion}\label{discussion}
-
-\section{Figures}\label{figures}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/epa_child_cp_plot-1.pdf}
-\caption{}
-\end{figure}
-
-\newpage
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/who_drink_cp_plot-1.pdf}
-\caption{}
-\end{figure}
-
-\newpage
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/epa_adult_cp -1.pdf}
-\caption{}
-\end{figure}
-
-\newpage
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/who_rec_low1_cp-1.pdf}
-\caption{}
-\end{figure}
-
-\newpage
 

### 2015-09-11 manuscript/manuscript.Rmd
started draft figure layouts

@@
 
-who_rec_low2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log10(mc_guides[5]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100, xW=wgt)
+#who_rec_low2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
+#                     xImpair = log10(mc_guides[5]), ProbComp = "gt", Exceed = "gte",
+#                     ci = TRUE, R = 100, xW=wgt)
 
-mc_chla[5] <- get_chla(who_rec_low2_cp,0.5)
+#mc_chla[5] <- get_chla(who_rec_low2_cp,0.5)
 
-who_rec_med1_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log10(mc_guides[6]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100, xW=wgt)
+#who_rec_med1_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
+#                     xImpair = log10(mc_guides[6]), ProbComp = "gt", Exceed = "gte",
+#                     ci = TRUE, R = 100, xW=wgt)
 
-mc_chla[6] <- get_chla(who_rec_med1_cp,0.5)
+#mc_chla[6] <- get_chla(who_rec_med1_cp,0.5)
 
-who_rec_med2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log10(mc_guides[7]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100, xW=wgt)
+#who_rec_med2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
+#                     xImpair = log10(mc_guides[7]), ProbComp = "gt", Exceed = "gte",
+#                     ci = TRUE, R = 100, xW=wgt)
 
-mc_chla[7] <- get_chla(who_rec_med2_cp,0.5)
+#mc_chla[7] <- get_chla(who_rec_med2_cp,0.5)
 ```
@@
 #Methods
-We evaluated associated chlorophyll *a* concentrations for an effect for each of the WHO and U.S.
-EPA levels (Table \ref{tab:microcystin_levels}).
-Lakes with higher microcystin-LR concentrations were rare.
+
+##Data
+
+##Conditional Probability Analysis
+We used a conditional probability analysis approach to explore associations between chlorophyll *a* concentrations  and World Health Organization (WHO) and U.S.
+Environmental Protection Agency (U.S.
+EPA) microcystin-LR health advisory levels [REFS].
+Many levels have been suggested [Table \ref{tab:microcystin_levels}], but lakes with higher microcystin-LR concentrations in the NLA were rare.
 Only `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=10])/nrow(nla_dat)),2)` % of lakes sampled had a concentration greater than 10.
-For this analysis we focus on the microcystin concentrations that are better represented in the NLA data.
+Thus, for this analysis we focus on the microcystin concentrations that are better represented in the NLA data.
 These were  0.3, 1, 1.6, and 2 $\mu$g/L.
@@
 
-(Figure \ref{fig:epa_child_cp_plot})
+(Figure \ref{fig:chla_micro_scatter})
 
-(Figure \ref{fig:who_drink_cp_plot})
+(Figure \ref{fig:chla_micro_cdf})
 
-(Figure \ref{fig:epa_adult_cp})
-
-(Figure \ref{fig:who_rec_low1_cp})
+(Figure \ref{fig:multi_cp_plot})
 
@@
 
-```{r epa_child_cp_plot,echo=FALSE, fig.cap="Plot of \\label{fig:epa_child_cp_plot}"}
-plot_cp(epa_child_cp,x = "Log 10 Chlorophyll", y = "Probability of Exceeding EPA Child HA")
+```{r chla_micro_scatter, echo=FALSE, fig.width=8, fig.cap="Scatterplot showing association betweeen chlorophyll \\it{a} and microcystin-LR. \\label{fig:chla_micro_scatter}"}
+plot_scatter(nla_dat,CHLA,MCYST_TL_UGL,viridis(1))
 ```
@@
 
-```{r who_drink_cp_plot,echo=FALSE, fig.cap="Plot of  \\label{fig:who_drink_cp_plot}"}
-plot_cp(who_drink_cp, x = "Log 10 Chlorophyll", y = "Probability of Exceeding WHO Drinking HA")
+```{r chla_micro_cdf, echo=FALSE, fig.width=8, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
+\\label{fig:chla_micro_cdf}"}
+plot_cdf()
 ```
@@
 
-```{r epa_adult_cp ,echo=FALSE, fig.cap="Plot of  \\label{fig:epa_adult_cp}"}
-plot_cp(epa_adult_cp, x = "Log 10 Chlorophyll", y = "Probability of Exceeding EPA Adult HA")
-```
-
-\newpage
-
-```{r who_rec_low1_cp, echo=FALSE, fig.cap="Plot of  \\label{fig:who_rec_low1_cp}"}
-plot_cp(who_rec_low1_cp, x = "Log 10 Chlorophyll", y = "Probability of Exceeding WHO Low Probability HA")
+```{r epa_child_cp_plot, echo=FALSE, fig.width=8, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
+\\label{fig:multi_cp_plot}"}
+multiplot(
+  plot_cp(epa_child_cp,x = "Log 10 Chlorophyll", 
+          y = "Probability of Exceeding 0.3 ug/L MLR",
+          title="U.S. EPA Children's Drinking Water Advisory Level"),
+  plot_cp(who_drink_cp, x = "Log 10 Chlorophyll", 
+          y = "Probability of Exceeding 1 ug/L MLR", 
+          title="WHO Drinking Water Advisory Level"),
+  plot_cp(epa_adult_cp, x = "Log 10 Chlorophyll", 
+          y = "Probability of Exceeding 1.6 ug/L MLR", 
+          title="U.S. EPA Adult Drinking Water Advisory Level"),
+  plot_cp(who_rec_low1_cp, x = "Log 10 Chlorophyll", 
+          y = "Probability of Exceeding 2 ug/L MLR", 
+          title="WHO Recreational Water Advisory Level"),
+  cols=2
+)
 ```
@@
                   "Drinking")
-concentration <- c("2-4 ug/l",
-                   "10-20 ug/l",
-                   "20-2000 ug/l",
-                   ">2000 ug/l",
-                   "1 ug/l",
-                   "0.3 ug/l",
-                   "1.6 ug/l")
+concentration <- c("2-4 $\\mu$g/L",
+                   "10-20 $\\mu$g/L",
+                   "20-2000 $\\mu$g/L",
+                   ">2000 $\\mu$g/L",
+                   "1 $\\mu$g/L",
+                   "0.3 $\\mu$g/L",
+                   "1.6 $\\mu$g/L")
 #note <- c("Low probability of adverse health effects",

### 2015-09-11 R/condprobMin.R

@@ -1,53 +1,0 @@
-#' Estimate min sample size for calculating probability
-#' 
-#' This function takes a brute force approach to estimating the minimum sample
-#' size for calculating a probability given a binary vector of data.  This is used
-#' to set the upper bounds on a conditional probability analysis.  In this 
-#' function the original vector is sampled with replacement at the full sample size.  
-#' This is repeated R times and mean value is recorded for each iteration. Then starting 
-#' with the minimum sample size, the original dataset is sampled and the mean for the smaller
-#' sample is recorded.  The sample size is increased and process repeated.  The average of the 
-#' mean values for each sample size is recorded.  This is an experimental function.  No 
-#' promises on its utility.  
-#' 
-#' 
-#' 
-#' @return an list of class condprobMin is returned with two items.  A data frame containg
-#'          the sample size and the associated average mean value and a vector contain the
-#'          bootstrapped original mean values.
-#'
-#' @param x a vector of values to find minimum sample size for.  The values
-#'        should be 1 and 0.
-#' @param R the number of bootstrap iterations. Defaults to 100.
-#' @param n the minimum sample size to start testing from.  Defaults to length of x.
-#' 
-#' @export
-#' @examples
-#' bivec<-rbinom(100,1,0.5)
-#' plot(condprobMin(bivec))
-#' data(binCutoff)
-#' minSampDF<-condprobMin(binCutoff,R=10)
-#' plot(minSampDF)
-
-condprobMin <- function(x, R = 100, n = length(x)) {
-    xdf <- data.frame()
-    orig <- vector("numeric", length = R)
-    for (i in 1:R) {
-        xo <- sample(x, length(x), TRUE)
-        orig[i] <- mean(xo)
-    }
-    for (i in 1:n) {
-        samp <- data.frame()
-        for (j in 1:R) {
-            xs <- sample(x, i, TRUE)
-            samp <- rbind(samp, data.frame(i, mean(xs)))
-        }
-        xdf <- rbind(xdf, data.frame(n = mean(samp[, 1]), avgMean = mean(samp[, 2], na.rm = T)))
-        if (i%%100 == 0) {
-            print(paste(i, "out of", n, "completed."))
-        }
-    }
-    xdf <- list(sampMeans = xdf, origMeans = orig)
-    class(xdf) <- "condprobMin"
-    return(xdf)
-} 

### 2015-09-11 R/ecdf_ks_ci.R

@@ -1,0 +1,19 @@
+#' Kolmogrov-Smirnoff CIs for a ecdf
+#' 
+#' Code from sfsmisc::ecdf.KSci. Ammended here to simply output the xy for both upper and 
+#' lower for use as input to ggplot
+#' @export
+#' @examples
+#' ecdf_ks_ci(all_pred_prob)
+ecdf_ks_ci <- function(x){
+  n <- length(x)
+  ec <- ecdf(x)
+  xx <- get("x", envir = environment(ec))
+  yy <- get("y", envir = environment(ec))
+  D <- sfsmisc::KSd(n)
+  yyu <- pmin(yy + D, 1)
+  yyl <- pmax(yy - D, 0)
+  ecu <- stepfun(xx, c(yyu, 1))
+  ecl <- stepfun(xx, c(yyl, yyl[n]))
+  return(list(x=x, upper=ecu(x),lower=ecl(x)))
+}

### 2015-09-11 R/multiplot.R

@@ -1,0 +1,72 @@
+#' Create Mulitple Plots with ggplots (from Winston Chang)
+#' 
+#' This is the Multiple plot function from 
+#' \href{http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)}{R-Cookbook}.  
+#' I have merely cleaned up some of the params for my use, added some documentation and examples.
+#' 
+#' @param ... can pass ggplot objects 
+#' @param plotlist alternative to ..., list of ggplot objects
+#' @param cols Number of columns in layout
+#' @param layout A matrix specifying the layout. If present, 'cols' is ignored.
+#' If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
+#' then plot 1 will go in the upper left, 2 will go in the upper right, and
+#' 3 will go all the way across the bottom.
+#' 
+#' @examples
+#' x<-rnorm(100)
+#' y<-jitter(x,10000)
+#' img <- readJPEG(system.file('img', 'Rlogo.jpg', package='jpeg'))
+#' xdf<-data.frame(x=x,y=y)
+#' firstplot<-ggplot(xdf,aes(x=x,y=y))+ geom_point()
+#' secondplot<-ggplot(xdf,aes(x=x))+ geom_bar()
+#' thirdplot<-ggplot(xdf, aes(x,y)) + 
+#' annotation_custom(rasterGrob(img, width=unit(1,'npc'), height=unit(1,'npc')), 
+#'                  -Inf, Inf, -Inf, Inf)
+#' thirdplot
+#' svg("test.svg",width=8)
+#' multiplot(thirdplot,secondplot,firstplot,firstplot,layout=matrix(c(1,2,3,4),ncol=4,byrow=F))
+#' dev.off()
+#' 
+#' @export
+#' @import ggplot2
+
+multiplot <- function(..., plotlist = NULL, cols = 1, 
+    layout = NULL) {
+    require(grid)
+    
+    # Make a list from the ... arguments and plotlist
+    plots <- c(list(...), plotlist)
+    
+    numPlots = length(plots)
+    
+    # If layout is NULL, then use 'cols' to determine
+    # layout
+    if (is.null(layout)) {
+        # Make the panel ncol: Number of columns of plots
+        # nrow: Number of rows needed, calculated from # of
+        # cols
+        layout <- matrix(seq(1, cols * ceiling(numPlots/cols)), 
[…]

### 2015-09-11 R/plot.condprob.R

@@ -1,32 +1,0 @@
-#' Function to plot condprob object
-#'
-#' This function plots a condprob object and creates a plot similar to Figure 1
-#' in Hollister et al. (2008). This includes a simple scatter plot of the raw 
-#' stressor vs the raw response, a cumulative distribution function for all of the 
-#' data, the impaired sites, and the unimpaired sites, and a conditional probability 
-#' plot with the conditional probability on the y-axis the stressor on the x-axis.  
-#' 
-#' @param cp_obj An object of class condprob is required.  Data and parameters
-#'                    from this object are used to create the plot
-#' @param min_samp minimum sample size to plot probabilities for. Defaults to 10.
-#' @param ... Used primarily to pass other ggplot2 plotting options to control title, 
-#'            axis labels, colors, size, etc.
-#' 
-#' @import ggplot2
-#' 
-#' @references Hollister, J. W., J. F. Paul, and H. A. Walker (2008). CProb: 
-#'             A Computational Tool for Conducting Conditional Probability 
-#'             Analysis. Journal of Environmental Quality. 37(6):2392-2396.
-#'             \href{http://dx.doi.org/10.2134/jeq2007.0536}{Link}
-#' @export
-#' @examples
-#' cp_plot(adult_ha)
-cp_plot <- function(cp_obj, min_samp = 10, ...) {
-    dat<-data.frame(cp_obj[1:5])
-    dat<-dat[1:(nrow(dat)-min_samp),]
-    cplot <- ggplot(dat,aes(x=X, y=Bootstrap.Probability)) +
-      geom_point() + 
-      geom_line(aes(x=X,y=Lower.CI))+
-      geom_line(aes(x=X,y=Upper.CI))
-    return(cplot)
-} 

### 2015-09-11 R/plot.condprobMin.R

@@ -1,20 +1,0 @@
-#' Function to plot condprobMin object
-#'
-#' This function plots minimum sample size vs average value of an input vector, which is
-#' usually the y-value in a \code{condprob2::condprob} analysis.
-#' The mean values are plotted vs sample size and the mean, upper limit and lower limit of
-#' the bootstrapped original values are shown
-#' 
-#' @param cpMin a condprobMin object
-#' @param alpha the alpha level for determining confidence intervals.  Default
-#'        is 0.05.
-#'        
-#' @method plot condprobMin
-#' @export
-plot.condprobMin <- function(cpMin, alpha = 0.05) {
-    with(cpMin[[1]], plot(avgMean, n))
-    abline(v = mean(cpMin[[2]]), col = "green")
-    abline(v = quantile(cpMin[[2]], 1 - alpha/2), col = "red")
-    abline(v = quantile(cpMin[[2]], alpha/2), col = "red")
-    
-} 

### 2015-09-11 R/plot_cdf.R

@@ -1,0 +1,52 @@
+#' plotCDF Function to create CDFs of chla, ts, and cyano
+#' 
+#' 
+#' This functions produces CDF plots from the hkm2014Data.  The implementation may be used more 
+#' widely though to create cumulative distribution functions for multiple categories   
+#' 
+#' @param catVar the categorical Variable such as-'bvCat', 'TS_CHLA_2', 'TS_CHLA_3', 'TS_CHLA_4'.  
+#'               Of same length as contVar 
+#' @param contVar the continuous Variable such as-'sumBioV','CHLA', 'NTL', 'PTL', etc.  Of same 
+#'                length as contVar 
+#' @param cdf_colors colors for CDF plots and median line -should equal # of categories in catVar + 1
+#' @param ... pass additional parameters to labs (e.g. set titles, axis labels, etc.)
+#' 
+#' @examples
+#' devtools::install_github('wesanderson','karthik')
+#' library(wesanderson)
+#' data(LakeTrophicModelling)
+#' plotCdf(ltmData$bvCat,ltmData$CHLA,cdf_colors=wes.palette(3,'Royal1'),y='Percent',
+#'         x=expression(paste('Log10(Chl ', italic("a"),')')),title='CDF for category bvCat',
+#'         color="Biovolume\nCategories")
+#' plotCdf(ltmData$TS_CHLA_3,ltmData$sumBioV+1,cdf_colors=wes.palette(3,'FantasticFox'),
+#'         y='Percent',x='Log10(Cyanobacterial Biovolume)',
+#'         title=expression(paste('CDF for Chlorophyll ', italic("a"),' Trophic States (3 Categories)')),
+#'         color="Trophic State\nCategories")
+#' 
+#' @export
+#' @import ggplot2
+
+plotCdf <- function(catVar, contVar, cdf_colors = 1:length(catVar), ...) {
+
+    options(scipen = 5)  #tell r not to use scientific notation on axis labels
+    
+    keep<-complete.cases(catVar,contVar)
+    cdf_df<-data.frame(catVar,contVar)[keep,]
+    x <- ggplot(cdf_df,aes(contVar,colour=catVar)) +
+          stat_ecdf(size=2)+scale_color_manual(values=cdf_colors)+
+          scale_x_log10()+
+          theme(text = element_text(family="Times"),
+                panel.background = element_blank(), panel.grid = element_blank(), 
+                panel.border = element_rect(fill = NA), 
+                plot.title  = element_text(family="sans",size=12,face="bold",vjust=1.1),
[…]

### 2015-09-11 R/plot_cp.R

@@ -1,7 +1,7 @@
 #'get_nla
 #'
 #'Create plot of conditional probability object showing association between microcystin-LR
-#'advisory levels and chlorophyll \italics{a} 
+#'advisory levels and chlorophyll \it{a} 
 #'
 #' @param cp A conditional probability object from condprob2
 #' @param ...  arguments for labels

### 2015-09-11 R/plot_scatter.R

@@ -1,0 +1,23 @@
+#' Scatterplot for microcystin-chla paper
+#' 
+#' Produces a scatterplot via ggplot
+#' 
+#' @export
+#' @import ggplot2
+
+plot_scatter<-function(df,xvar,yvar,pt_col=1,...){
+  options(scipen=5)
+  x <- ggplot(xdf,aes(xvar,yvar))+
+    geom_point(size=3,colour=pt_col) +
+    scale_x_log10() +
+    scale_y_log10() +
+    theme_bw()+
+    theme(text = element_text(family="sans"),
+          axis.title.x = element_text(family="sans",vjust = -0.5, size = 14),
+          axis.title.y = element_text(family="sans",vjust = 1.5, size = 14),
+          axis.text.x = element_text(family="sans",size = 12),
+          axis.text.y = element_text(family="sans",size = 12)) + 
+    labs(...)
+    
+  return(x)
+}

### 2015-09-11 manuscript/manuscript.Rmd
figures closer

@@
 ```{r chla_micro_scatter, echo=FALSE, fig.width=8, fig.cap="Scatterplot showing association betweeen chlorophyll \\it{a} and microcystin-LR. \\label{fig:chla_micro_scatter}"}
-plot_scatter(nla_dat,CHLA,MCYST_TL_UGL,viridis(1))
+plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",pt_col=viridis(1),
+             x=expression(paste('Log10 Chl ', italic("a"),' (',mu,'g/L)')),
+             y=expression(paste('Log10 microcystin-LR',' (',mu,'g/L)')))
 ```
@@
 \\label{fig:chla_micro_cdf}"}
-plot_cdf()
+plot_cdf(nla_dat[,c("log_chla","log_micro")],"")
 ```

### 2015-09-11 manuscript/manuscript.tex
figures closer

@@
 concentrations of 0.3, 1, 1.6, and 2 we find chlorophyll \emph{a}
-concentrations of 20.8, 71.2, 92.16, and 110.16, respectively. When
+concentrations of 20.8, 78.29, 109.66, and 109.66, respectively. When
 managing for these various microcystin levels exceeding these reported
@@
 
-We evaluated associated chlorophyll \emph{a} concentrations for an
-effect for each of the WHO and U.S. EPA levels (Table
-\ref{tab:microcystin_levels}). Lakes with higher microcystin-LR
-concentrations were rare. Only 1.16 \% of lakes sampled had a
-concentration greater than 10. For this analysis we focus on the
-microcystin concentrations that are better represented in the NLA data.
-These were 0.3, 1, 1.6, and 2 \(\mu\)g/L.
+\subsection{Data}\label{data}
+
+\subsection{Conditional Probability
+Analysis}\label{conditional-probability-analysis}
+
+We used a conditional probability analysis approach to explore
+associations between chlorophyll \emph{a} concentrations and World
+Health Organization (WHO) and U.S. Environmental Protection Agency (U.S.
+EPA) microcystin-LR health advisory levels {[}REFS{]}. Many levels have
+been suggested {[}Table \ref{tab:microcystin_levels}{]}, but lakes with
+higher microcystin-LR concentrations in the NLA were rare. Only 1.16 \%
+of lakes sampled had a concentration greater than 10. Thus, for this
+analysis we focus on the microcystin concentrations that are better
+represented in the NLA data. These were 0.3, 1, 1.6, and 2 \(\mu\)g/L.
 
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/epa_child_cp_plot-1.pdf}
-\caption{Plot of \label{fig:epa_child_cp_plot}}
+\includegraphics{manuscript_files/figure-latex/chla_micro_scatter-1.pdf}
+\caption{Scatterplot showing association betweeen chlorophyll
+\it{a} and microcystin-LR. \label{fig:chla_micro_scatter}}
 \end{figure}
@@
 
+\begin{verbatim}
+## No id variables; using all as measure variables
+\end{verbatim}
+
 \begin{figure}[htbp]
 \centering
-\includegraphics{manuscript_files/figure-latex/who_drink_cp_plot-1.pdf}
-\caption{Plot of \label{fig:who_drink_cp_plot}}
+\includegraphics{manuscript_files/figure-latex/chla_micro_cdf-1.pdf}
+\caption{Conditional probability plots showing association between the
+probability of exceeding various microcystin-LR (MLR) health advisory
+Levels. \label{fig:chla_micro_cdf}}
 \end{figure}
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/epa_adult_cp -1.pdf}
-\caption{Plot of \label{fig:epa_adult_cp}}
-\end{figure}
-
-\newpage
+\begin{verbatim}
+## Loading required package: grid
+\end{verbatim}
 
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/who_rec_low1_cp-1.pdf}
-\caption{Plot of \label{fig:who_rec_low1_cp}}
+\includegraphics{manuscript_files/figure-latex/epa_child_cp_plot-1.pdf}
+\caption{Conditional probability plots showing association between the
+probability of exceeding various microcystin-LR (MLR) health advisory
+Levels. \label{fig:multi_cp_plot}}
 \end{figure}
@@
 \strut\end{minipage} &
-\begin{minipage}[b]{0.19\columnwidth}\raggedright\strut
+\begin{minipage}[b]{0.20\columnwidth}\raggedright\strut
 Concentration
@@
 \strut\end{minipage} &
-\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-1 ug/l
+\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
+1 \(\mu\)g/L
 \strut\end{minipage}\tabularnewline
@@
 \strut\end{minipage} &
-\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-0.3 ug/l
+\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
+0.3 \(\mu\)g/L
 \strut\end{minipage}\tabularnewline
@@
 \strut\end{minipage} &
-\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-1.6 ug/l
+\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
+1.6 \(\mu\)g/L
 \strut\end{minipage}\tabularnewline
@@
 \strut\end{minipage} &
-\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-2-4 ug/l
+\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
+2-4 \(\mu\)g/L
 \strut\end{minipage}\tabularnewline
@@
 \strut\end{minipage} &
-\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-10-20 ug/l
+\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
+10-20 \(\mu\)g/L
 \strut\end{minipage}\tabularnewline
@@
 \strut\end{minipage} &
-\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-20-2000 ug/l
+\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
+20-2000 \(\mu\)g/L
 \strut\end{minipage}\tabularnewline
@@
 \strut\end{minipage} &
-\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-\textgreater{}2000 ug/l
+\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
+\textgreater{}2000 \(\mu\)g/L
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-71.2
+78.29
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-92.16
+109.7
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-110.2
+109.7
 \strut\end{minipage}\tabularnewline

### 2015-09-11 R/plot_cdf.R

@@ -1,52 +1,23 @@
-#' plotCDF Function to create CDFs of chla, ts, and cyano
+#' plotCDF Function to create CDFs of chla and microcystin
 #' 
+#' This functions produces CDF plots for the Hollister et al. microcystin - chla paper 
 #' 
-#' This functions produces CDF plots from the hkm2014Data.  The implementation may be used more 
-#' widely though to create cumulative distribution functions for multiple categories   
-#' 
-#' @param catVar the categorical Variable such as-'bvCat', 'TS_CHLA_2', 'TS_CHLA_3', 'TS_CHLA_4'.  
-#'               Of same length as contVar 
-#' @param contVar the continuous Variable such as-'sumBioV','CHLA', 'NTL', 'PTL', etc.  Of same 
-#'                length as contVar 
-#' @param cdf_colors colors for CDF plots and median line -should equal # of categories in catVar + 1
-#' @param ... pass additional parameters to labs (e.g. set titles, axis labels, etc.)
-#' 
-#' @examples
-#' devtools::install_github('wesanderson','karthik')
-#' library(wesanderson)
-#' data(LakeTrophicModelling)
-#' plotCdf(ltmData$bvCat,ltmData$CHLA,cdf_colors=wes.palette(3,'Royal1'),y='Percent',
-#'         x=expression(paste('Log10(Chl ', italic("a"),')')),title='CDF for category bvCat',
-#'         color="Biovolume\nCategories")
-#' plotCdf(ltmData$TS_CHLA_3,ltmData$sumBioV+1,cdf_colors=wes.palette(3,'FantasticFox'),
-#'         y='Percent',x='Log10(Cyanobacterial Biovolume)',
-#'         title=expression(paste('CDF for Chlorophyll ', italic("a"),' Trophic States (3 Categories)')),
-#'         color="Trophic State\nCategories")
-#' 
+#' @param df a data frame with all parameters to plot
 #' @export
 #' @import ggplot2
 
-plotCdf <- function(catVar, contVar, cdf_colors = 1:length(catVar), ...) {
-
+plot_cdf <- function(df, ...) {
     options(scipen = 5)  #tell r not to use scientific notation on axis labels
-    
-    keep<-complete.cases(catVar,contVar)
-    cdf_df<-data.frame(catVar,contVar)[keep,]
-    x <- ggplot(cdf_df,aes(contVar,colour=catVar)) +
[…]

### 2015-09-11 R/plot_scatter.R

@@ -7,10 +7,10 @@
 
 plot_scatter<-function(df,xvar,yvar,pt_col=1,...){
   options(scipen=5)
+  xdf<-data.frame(xvar=df[[xvar]],yvar=df[[yvar]])
+  xdf<-xdf[complete.cases(xdf),]
   x <- ggplot(xdf,aes(xvar,yvar))+
     geom_point(size=3,colour=pt_col) +
-    scale_x_log10() +
-    scale_y_log10() +
     theme_bw()+
     theme(text = element_text(family="sans"),
           axis.title.x = element_text(family="sans",vjust = -0.5, size = 14),

### 2015-09-11 manuscript/manuscript.Rmd
first cut of methods

@@
 ##Data
+We used the 2007 National Lakes Assessment (NLA) water quality and microcystin-LR concentration data [REF].
+The NLA is a probability based design that, when using the provided sample weights, can be used to provide unbiased estimates of lake condition for all lakes in the original sample frame (e.g.
+lakes greater than 4ha in the conterminous US).
+These data represent a snapshot of water quality from the summer of 2007 and data on chlorophyll *a* and microcystin-LR concentrations are available for `r nrow(na.omit(nla_dat$CHLA,nla_dat$MCYST_TL_UGL))` lakes.
+All results are reported using weighted estimates.
 
 ##Conditional Probability Analysis
-We used a conditional probability analysis approach to explore associations between chlorophyll *a* concentrations  and World Health Organization (WHO) and U.S.
+We used a conditional probability analysis (CPA) approach to explore associations between chlorophyll *a* concentrations  and World Health Organization (WHO) and U.S.
 Environmental Protection Agency (U.S.
@@
 These were  0.3, 1, 1.6, and 2 $\mu$g/L.
+
+A detailed discussion of CPA is beyond the scope of this paper, but see Paul et al.
+[-REF] and Hollister et al.
+[-REF] for details.
+For this analysis, we used CPA to examine how the conditional probability of exceeding one of the health advisory changes as chlorophyll *a* increases in a lake.
+The 95% confidence intervals were calculated from 1000 bootstrapped samples.
+To identify chlorophyll *a* concentrations of concern we used a 50% conditional probability of exceeding each health advisory level and extracted the minimum chlorophyll *a* concentration that was associated with the upper confidence level being 50% or greater.
+Details of the specific implementation we used are available at https://github.com/USEPA/microcystinchla.
 
@@
 
-```{r microcystin_levels,echo=FALSE,results='asis', fig.cap="Various suggested microcystin-LR health advisory concentrations. \\label{fig:microcystin_levels}"}
+```{r microcystin_levels,echo=FALSE,results='asis', fig.cap="Various suggested microcystin-LR health advisory concentrations. \\label{tab:microcystin_levels}"}
 source_of_stnd <- c("WHO",
@@
 ```{r mc_chla_table, results='asis', echo=FALSE, fig.cap="Chlorophyll \textit{a} concentrations that are associated with a 50% probability of exceeding a microcystin-LR health advisory concentration.
-\\label{fig:mc_chla_table}"}
+\\label{tab:mc_chla_table}"}
 mc_chla_tbl <- data.frame(Source=mc_names,Type=mc_type,Microcystin=mc_guides,Chlorophyll=mc_chla)[1:4,]

### 2015-09-11 manuscript/manuscript.tex
first cut of methods

@@
 
+We used the 2007 National Lakes Assessment (NLA) water quality and
+microcystin-LR concentration data {[}REF{]}. The NLA is a probability
+based design that, when using the provided sample weights, can be used
+to provide unbiased estimates of lake condition for all lakes in the
+original sample frame (e.g.~lakes greater than 4ha in the conterminous
+US). These data represent a snapshot of water quality from the summer of
+2007 and data on chlorophyll \emph{a} and microcystin-LR concentrations
+are available for lakes. All results are reported using weighted
+estimates.
+
 \subsection{Conditional Probability
@@
 
-We used a conditional probability analysis approach to explore
+We used a conditional probability analysis (CPA) approach to explore
 associations between chlorophyll \emph{a} concentrations and World
@@
 represented in the NLA data. These were 0.3, 1, 1.6, and 2 \(\mu\)g/L.
+
+A detailed discussion of CPA is beyond the scope of this paper, but see
+Paul et al. {[}-REF{]} and Hollister et al. {[}-REF{]} for details. For
+this analysis, we used CPA to examine how the conditional probability of
+exceeding one of the health advisory changes as chlorophyll \emph{a}
+increases in a lake. The 95\% confidence intervals were calculated from
+1000 bootstrapped samples. To identify chlorophyll \emph{a}
+concentrations of concern we used a 50\% conditional probability of
+exceeding each health advisory level and extracted the minimum
+chlorophyll \emph{a} concentration that was associated with the upper
+confidence level being 50\% or greater. Details of the specific
+implementation we used are available at
+\url{https://github.com/USEPA/microcystinchla}.
 
@@
 \newpage
-
-\begin{verbatim}
-## Loading required package: grid
-\end{verbatim}
 

### 2015-09-16 manuscript/manuscript.Rmd
up through draft of results done

@@
   name: Betty J. Kreakie
-- affilnum: 2
-  name: Dorothy Q. Kellog
 affiliation:
@@
   affilnum: 1
-- affil: University of Rhode Island, Department of Natural Resources Science, Kingston,
-    RI, 02882, USA
-  affilnum: 2
 output:
@@
 mc_chla <- vector("numeric",length(mc_guides))
-wgt <- nla_dat$WGT_NLA
-#wgt<- rep(1,nrow(nla_dat))
+#wgt <- nla_dat$WGT_NLA
+wgt<- rep(1,nrow(nla_dat))
 
@@
 We used the 2007 National Lakes Assessment (NLA) water quality and microcystin-LR concentration data [REF].
-The NLA is a probability based design that, when using the provided sample weights, can be used to provide unbiased estimates of lake condition for all lakes in the original sample frame (e.g.
-lakes greater than 4ha in the conterminous US).
 These data represent a snapshot of water quality from the summer of 2007 and data on chlorophyll *a* and microcystin-LR concentrations are available for `r nrow(na.omit(nla_dat$CHLA,nla_dat$MCYST_TL_UGL))` lakes.
-All results are reported using weighted estimates.
 
@@
 Environmental Protection Agency (U.S.
-EPA) microcystin-LR health advisory levels [REFS].
-Many levels have been suggested [Table \ref{tab:microcystin_levels}], but lakes with higher microcystin-LR concentrations in the NLA were rare.
+EPA) microcystin-LR health advisory levels [@paul2011probability].
+Many levels have been suggested (Table \ref{tab:microcystin_levels}), but lakes with higher microcystin-LR concentrations in the NLA were rare.
 Only `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=10])/nrow(nla_dat)),2)` % of lakes sampled had a concentration greater than 10.
@@
 To identify chlorophyll *a* concentrations of concern we used a 50% conditional probability of exceeding each health advisory level and extracted the minimum chlorophyll *a* concentration that was associated with the upper confidence level being 50% or greater.
-Details of the specific implementation we used are available at https://github.com/USEPA/microcystinchla.
+As both microcystin-LR and chlorophyll *a* values were both highely skewed right, a log base 10 transformation was used.
+Additional details of the specific implementation are available at https://github.com/USEPA/microcystinchla.
 
@@
 
-(Table \ref{tab:mc_chla_table})
+In the 2007 NLA, microcystin-LR concentrations ranged from `r min(nla_dat$MCYST_TL_UGL,na.rm=T)` to `r max(nla_dat$MCYST_TL_UGL, na.rm=T)`.
+Microcystin-LR concentrations of 0.05 $\mu$g/L represent the detection limits.
+Any value greater than that indicates the presence of microcystin-LR.
+Of those lakes with microcystin, the median concentration was `r median(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T)` and the mean was `r median(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T)`.
+Lastly, of all lakes sampled, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=0.3])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the U.S.
+EPA childrens drinking water standard, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1.6])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the U.S.
+EPA adult drinking water standard,`r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO drinking water standard,and `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=2])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO recreational standard.
+For chlorophyll *a*, the range was `r min(nla_dat$CHLA,na.rm=T)` to `r max(nla_dat$CHLA, na.rm=T)`.
+All lakes had reported chlorophyll *a* concentrations that exceeded detection limits.
+The median concentration was `r median(nla_dat$CHLA,na.rm=T)` and the mean was `r mean(nla_dat$CHLA,na.rm=T)`.
 
-(Figure \ref{fig:chla_micro_scatter})
+The association between Log 10 microcystin-LR and Log 10 chlorophyll *a* show a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
+This indicates that higher concentrations of microcystin-LR almost always co-occur with higher concentrations of chlorophyll *a* yet the inverse is not true.
+Higher chlorophyll *a* is not necessarily predictive of higher microcystin-LR concentrations; however, chlorophyll *a* may be predictive of the probability of exceeding a certian concentration.
+This is the case as the probability of exceeding each of the four tested health adviroy levels shos this patern of probability of exceeding the advisory level increases as a function of chlorophyll *a* concentration (Figure \ref{fig:multi_cp_plot}).
 
-(Figure \ref{fig:chla_micro_cdf})
-
-(Figure \ref{fig:multi_cp_plot})
+We use this association to identify chlorophyll *a* concentrations that are associated with greater than even odds of exceeding a given health advisory level (Table \ref{tab:mc_chla_table}).
+These represent `r 100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[1]])/length(nla_dat$CHLA))`%, `100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[2]])/length(nla_dat$CHLA))`%,
+`100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[3]])/length(nla_dat$CHLA))`%, and `100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[4]])/length(nla_dat$CHLA))`% of sample lakes for the U.S EPA childrens drinking water, the WHO drinking water, the U.S.
+EPA Adult drinking water, and the WHO recreationsl standards, respecitvely.
+<!-- (Figure \ref{fig:chla_micro_cdf}) -->
 
@@
 
-```{r chla_micro_scatter, echo=FALSE, fig.width=8, fig.cap="Scatterplot showing association betweeen chlorophyll \\it{a} and microcystin-LR. \\label{fig:chla_micro_scatter}"}
+```{r chla_micro_scatter, echo=FALSE, fig.width=8, fig.cap="Scatterplot showing association betweeen chlorophyll \\textit{a} and microcystin-LR. \\label{fig:chla_micro_scatter}"}
 plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",pt_col=viridis(1),
@@
 
+<!--
 ```{r chla_micro_cdf, echo=FALSE, fig.width=8, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
@@
 
-\newpage
+\newpage-->
 
@@
 
-```{r microcystin_levels,echo=FALSE,results='asis', fig.cap="Various suggested microcystin-LR health advisory concentrations. \\label{tab:microcystin_levels}"}
+```{r microcystin_levels,echo=FALSE,results='asis', tab.cap="Various suggested microcystin-LR health advisory concentrations. \\label{tab:microcystin_levels}"}
 source_of_stnd <- c("WHO",
@@
 
-```{r mc_chla_table, results='asis', echo=FALSE, fig.cap="Chlorophyll \textit{a} concentrations that are associated with a 50% probability of exceeding a microcystin-LR health advisory concentration.
+```{r mc_chla_table, results='asis', echo=FALSE, tab.cap="Chlorophyll \textit{a} concentrations that are associated with a 50% probability of exceeding a microcystin-LR health advisory concentration.
 \\label{tab:mc_chla_table}"}

### 2015-09-16 manuscript/manuscript.tex
up through draft of results done

@@
 Betty J. Kreakie
-Dorothy Q. Kellog
 }
@@
 Betty J. Kreakie \textsuperscript{1} 
-Dorothy Q. Kellog \textsuperscript{2} 
 \end{center}
@@
 Ecology Division, 27 Tarzwell Drive Narragansett, RI, 02882, USA\\*
-\\*
-\textsuperscript{2}University of Rhode Island, Department of Natural Resources Science,
-Kingston, RI, 02882, USA\\*
 }
@@
 concentrations of 0.3, 1, 1.6, and 2 we find chlorophyll \emph{a}
-concentrations of 20.8, 78.29, 109.66, and 109.66, respectively. When
+concentrations of 23.68, 63.94, 89.71, and 102.6, respectively. When
 managing for these various microcystin levels exceeding these reported
@@
 We used the 2007 National Lakes Assessment (NLA) water quality and
-microcystin-LR concentration data {[}REF{]}. The NLA is a probability
-based design that, when using the provided sample weights, can be used
-to provide unbiased estimates of lake condition for all lakes in the
-original sample frame (e.g.~lakes greater than 4ha in the conterminous
-US). These data represent a snapshot of water quality from the summer of
-2007 and data on chlorophyll \emph{a} and microcystin-LR concentrations
-are available for lakes. All results are reported using weighted
-estimates.
+microcystin-LR concentration data {[}REF{]}. These data represent a
+snapshot of water quality from the summer of 2007 and data on
+chlorophyll \emph{a} and microcystin-LR concentrations are available for
+lakes.
 
@@
 Health Organization (WHO) and U.S. Environmental Protection Agency (U.S.
-EPA) microcystin-LR health advisory levels {[}REFS{]}. Many levels have
-been suggested {[}Table \ref{tab:microcystin_levels}{]}, but lakes with
-higher microcystin-LR concentrations in the NLA were rare. Only 1.16 \%
-of lakes sampled had a concentration greater than 10. Thus, for this
-analysis we focus on the microcystin concentrations that are better
-represented in the NLA data. These were 0.3, 1, 1.6, and 2 \(\mu\)g/L.
+EPA) microcystin-LR health advisory levels (Paul and Munns 2011). Many
+levels have been suggested (Table \ref{tab:microcystin_levels}), but
+lakes with higher microcystin-LR concentrations in the NLA were rare.
+Only 1.16 \% of lakes sampled had a concentration greater than 10. Thus,
+for this analysis we focus on the microcystin concentrations that are
+better represented in the NLA data. These were 0.3, 1, 1.6, and 2
+\(\mu\)g/L.
 
@@
 chlorophyll \emph{a} concentration that was associated with the upper
-confidence level being 50\% or greater. Details of the specific
-implementation we used are available at
+confidence level being 50\% or greater. As both microcystin-LR and
+chlorophyll \emph{a} values were both highely skewed right, a log base
+10 transformation was used. Additional details of the specific
+implementation are available at
 \url{https://github.com/USEPA/microcystinchla}.
@@
 
-(Table \ref{tab:mc_chla_table})
+In the 2007 NLA, microcystin-LR concentrations ranged from 0.05 to 225.
+Microcystin-LR concentrations of 0.05 \(\mu\)g/L represent the detection
+limits. Any value greater than that indicates the presence of
+microcystin-LR. Of those lakes with microcystin, the median
+concentration was 0.51 and the mean was 0.51. Lastly, of all lakes
+sampled, 21\% of lakes exceeded the U.S. EPA childrens drinking water
+standard, 8.8\% of lakes exceeded the U.S. EPA adult drinking water
+standard,11.7\% of lakes exceeded the WHO drinking water standard,and
+7.3\% of lakes exceeded the WHO recreational standard. For chlorophyll
+\emph{a}, the range was 0.07 to 936. All lakes had reported chlorophyll
+\emph{a} concentrations that exceeded detection limits. The median
+concentration was 7.79 and the mean was 29.6301946.
 
-(Figure \ref{fig:chla_micro_scatter})
+The association between Log 10 microcystin-LR and Log 10 chlorophyll
+\emph{a} show a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
+This indicates that higher concentrations of microcystin-LR almost
+always co-occur with higher concentrations of chlorophyll \emph{a} yet
+the inverse is not true. Higher chlorophyll \emph{a} is not necessarily
+predictive of higher microcystin-LR concentrations; however, chlorophyll
+\emph{a} may be predictive of the probability of exceeding a certian
+concentration. This is the case as the probability of exceeding each of
+the four tested health adviroy levels shos this patern of probability of
+exceeding the advisory level increases as a function of chlorophyll
+\emph{a} concentration (Figure \ref{fig:multi_cp_plot}).
 
-(Figure \ref{fig:chla_micro_cdf})
-
-(Figure \ref{fig:multi_cp_plot})
+We use this association to identify chlorophyll \emph{a} concentrations
+that are associated with greater than even odds of exceeding a given
+health advisory level (Table \ref{tab:mc_chla_table}). These represent
+27.7831559\%,
+\texttt{100*(length(nla\_dat\$CHLA{[}nla\_dat\$CHLA\textgreater{}=mc\_chla{[}2{]}{]})/length(nla\_dat\$CHLA))}\%,
+\texttt{100*(length(nla\_dat\$CHLA{[}nla\_dat\$CHLA\textgreater{}=mc\_chla{[}3{]}{]})/length(nla\_dat\$CHLA))}\%,
+and
+\texttt{100*(length(nla\_dat\$CHLA{[}nla\_dat\$CHLA\textgreater{}=mc\_chla{[}4{]}{]})/length(nla\_dat\$CHLA))}\%
+of sample lakes for the U.S EPA childrens drinking water, the WHO
+drinking water, the U.S. EPA Adult drinking water, and the WHO
+recreationsl standards, respecitvely.
 
@@
 \includegraphics{manuscript_files/figure-latex/chla_micro_scatter-1.pdf}
-\caption{Scatterplot showing association betweeen chlorophyll
-\it{a} and microcystin-LR. \label{fig:chla_micro_scatter}}
-\end{figure}
-
-\newpage
-
-\begin{verbatim}
-## No id variables; using all as measure variables
-\end{verbatim}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/chla_micro_cdf-1.pdf}
-\caption{Conditional probability plots showing association between the
-probability of exceeding various microcystin-LR (MLR) health advisory
-Levels. \label{fig:chla_micro_cdf}}
+\caption{Scatterplot showing association betweeen chlorophyll \textit{a}
+and microcystin-LR. \label{fig:chla_micro_scatter}}
 \end{figure}
@@
 \begin{longtable}[c]{@{}lll@{}}
+\caption{Various suggested microcystin-LR health advisory
+concentrations. \label{tab:microcystin_levels}}\tabularnewline
+\toprule
+\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
+Source
+\strut\end{minipage} &
+\begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+Type
+\strut\end{minipage} &
+\begin{minipage}[b]{0.20\columnwidth}\raggedright\strut
+Concentration
+\strut\end{minipage}\tabularnewline
+\midrule
+\endfirsthead
 \toprule
@@
 \begin{longtable}[c]{@{}llll@{}}
+\caption{Chlorophyll extit\{a\} concentrations that are associated with
+a 50\% probability of exceeding a microcystin-LR health advisory
+concentration. \label{tab:mc_chla_table}}\tabularnewline
+\toprule
+\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
+Source
+\strut\end{minipage} &
+\begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+Type
+\strut\end{minipage} &
+\begin{minipage}[b]{0.17\columnwidth}\raggedright\strut
+Microcystin
+\strut\end{minipage} &
+\begin{minipage}[b]{0.17\columnwidth}\raggedright\strut
+Chlorophyll
+\strut\end{minipage}\tabularnewline
+\midrule
+\endfirsthead
 \toprule
@@
 \begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-20.8
+23.68
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-78.29
+63.94
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-109.7
+89.71
 \strut\end{minipage}\tabularnewline
@@
 \begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-109.7
+102.6
 \strut\end{minipage}\tabularnewline
@@
 
+Paul, J. F., and W. R. Munns. 2011. Probability surveys, conditional
+probability, and ecological risk assessment. Environmental Toxicology
+and Chemistry 30:1488--1495.
+
 \end{document}

### 2015-09-16 manuscript/manuscript.Rmd
fixed results in-line code

@@
 We use this association to identify chlorophyll *a* concentrations that are associated with greater than even odds of exceeding a given health advisory level (Table \ref{tab:mc_chla_table}).
-These represent `r 100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[1]])/length(nla_dat$CHLA))`%, `100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[2]])/length(nla_dat$CHLA))`%,
-`100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[3]])/length(nla_dat$CHLA))`%, and `100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[4]])/length(nla_dat$CHLA))`% of sample lakes for the U.S EPA childrens drinking water, the WHO drinking water, the U.S.
+These represent `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[1]])/length(nla_dat$CHLA)),1)`%, `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[2]])/length(nla_dat$CHLA)),1)`%,
+`r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[3]])/length(nla_dat$CHLA)),1)`%, and `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[4]])/length(nla_dat$CHLA)),1)`% of sample lakes for the U.S EPA childrens drinking water, the WHO drinking water, the U.S.
 EPA Adult drinking water, and the WHO recreationsl standards, respecitvely.

### 2015-09-16 manuscript/manuscript.tex
fixed results in-line code

@@
 health advisory level (Table \ref{tab:mc_chla_table}). These represent
-27.7831559\%,
-\texttt{100*(length(nla\_dat\$CHLA{[}nla\_dat\$CHLA\textgreater{}=mc\_chla{[}2{]}{]})/length(nla\_dat\$CHLA))}\%,
-\texttt{100*(length(nla\_dat\$CHLA{[}nla\_dat\$CHLA\textgreater{}=mc\_chla{[}3{]}{]})/length(nla\_dat\$CHLA))}\%,
-and
-\texttt{100*(length(nla\_dat\$CHLA{[}nla\_dat\$CHLA\textgreater{}=mc\_chla{[}4{]}{]})/length(nla\_dat\$CHLA))}\%
-of sample lakes for the U.S EPA childrens drinking water, the WHO
-drinking water, the U.S. EPA Adult drinking water, and the WHO
-recreationsl standards, respecitvely.
+27.8\%, 11.9\%, 8.3\%, and 7.4\% of sample lakes for the U.S EPA
+childrens drinking water, the WHO drinking water, the U.S. EPA Adult
+drinking water, and the WHO recreationsl standards, respecitvely.
 

### 2015-09-16 manuscript/manuscript.Rmd
fixed tables

@@
 names(mc_conc_table)<-c("Source","Type","Concentration")
-pandoc.table(mc_conc_table,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
+#pandoc.table(mc_conc_table,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
+knitr::kable(mc_conc_table)
 ```
@@
 
-```{r mc_chla_table, results='asis', echo=FALSE, tab.cap="Chlorophyll \textit{a} concentrations that are associated with a 50% probability of exceeding a microcystin-LR health advisory concentration.
+```{r mc_chla_table, results='asis', echo=FALSE, tab.cap="Chlorophyll \textit{a} concentrations that are associated with a 50 percent probability of exceeding a microcystin-LR health advisory concentration.
 \\label{tab:mc_chla_table}"}
 mc_chla_tbl <- data.frame(Source=mc_names,Type=mc_type,Microcystin=mc_guides,Chlorophyll=mc_chla)[1:4,]
-pandoc.table(mc_chla_tbl,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
+#pandoc.table(mc_chla_tbl,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
+knitr::kable(mc_chla_tbl)
 ```

### 2015-09-16 manuscript/manuscript.tex
fixed tables

@@
 \toprule
-\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
-Source
-\strut\end{minipage} &
-\begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-Type
-\strut\end{minipage} &
-\begin{minipage}[b]{0.20\columnwidth}\raggedright\strut
-Concentration
-\strut\end{minipage}\tabularnewline
+Source & Type & Concentration\tabularnewline
 \midrule
@@
 \toprule
-\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
-Source
-\strut\end{minipage} &
-\begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-Type
-\strut\end{minipage} &
-\begin{minipage}[b]{0.20\columnwidth}\raggedright\strut
-Concentration
-\strut\end{minipage}\tabularnewline
+Source & Type & Concentration\tabularnewline
 \midrule
 \endhead
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-WHO
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Drinking
-\strut\end{minipage} &
-\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
-1 \(\mu\)g/L
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-U.S. EPA
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Drinking
-\strut\end{minipage} &
-\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
-0.3 \(\mu\)g/L
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-U.S. EPA
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Drinking
-\strut\end{minipage} &
-\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
-1.6 \(\mu\)g/L
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-WHO
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Recreational
-\strut\end{minipage} &
-\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
-2-4 \(\mu\)g/L
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-WHO
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Recreational
-\strut\end{minipage} &
-\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
-10-20 \(\mu\)g/L
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-WHO
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Recreational
-\strut\end{minipage} &
-\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
-20-2000 \(\mu\)g/L
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-WHO
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Recreational
-\strut\end{minipage} &
-\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
-\textgreater{}2000 \(\mu\)g/L
-\strut\end{minipage}\tabularnewline
+WHO & Drinking & 1 \(\mu\)g/L\tabularnewline
+U.S. EPA & Drinking & 0.3 \(\mu\)g/L\tabularnewline
+U.S. EPA & Drinking & 1.6 \(\mu\)g/L\tabularnewline
+WHO & Recreational & 2-4 \(\mu\)g/L\tabularnewline
+WHO & Recreational & 10-20 \(\mu\)g/L\tabularnewline
+WHO & Recreational & 20-2000 \(\mu\)g/L\tabularnewline
+WHO & Recreational & \textgreater{}2000 \(\mu\)g/L\tabularnewline
 \bottomrule
@@
 
-\begin{verbatim}
-## Warning in split.large.cells(t): length of split.cells vector is smaller
-## than data. Default value will be used for other cells
-\end{verbatim}
-
-\begin{verbatim}
-## Warning in split.large.cells(t.colnames): length of split.cells vector is
-## smaller than data. Default value will be used for other cells
-\end{verbatim}
-
-\begin{longtable}[c]{@{}llll@{}}
+\begin{longtable}[c]{@{}llrr@{}}
 \caption{Chlorophyll extit\{a\} concentrations that are associated with
-a 50\% probability of exceeding a microcystin-LR health advisory
+a 50 percent probability of exceeding a microcystin-LR health advisory
 concentration. \label{tab:mc_chla_table}}\tabularnewline
 \toprule
-\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
-Source
-\strut\end{minipage} &
-\begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-Type
-\strut\end{minipage} &
-\begin{minipage}[b]{0.17\columnwidth}\raggedright\strut
-Microcystin
-\strut\end{minipage} &
-\begin{minipage}[b]{0.17\columnwidth}\raggedright\strut
-Chlorophyll
-\strut\end{minipage}\tabularnewline
+Source & Type & Microcystin & Chlorophyll\tabularnewline
 \midrule
@@
 \toprule
-\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
-Source
-\strut\end{minipage} &
-\begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-Type
-\strut\end{minipage} &
-\begin{minipage}[b]{0.17\columnwidth}\raggedright\strut
-Microcystin
-\strut\end{minipage} &
-\begin{minipage}[b]{0.17\columnwidth}\raggedright\strut
-Chlorophyll
-\strut\end{minipage}\tabularnewline
+Source & Type & Microcystin & Chlorophyll\tabularnewline
 \midrule
 \endhead
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-U.S. EPA
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Drinking
-\strut\end{minipage} &
-\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-0.3
-\strut\end{minipage} &
-\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-23.68
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-WHO
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Drinking
-\strut\end{minipage} &
-\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-1
-\strut\end{minipage} &
-\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-63.94
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-U.S. EPA
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Drinking
-\strut\end{minipage} &
-\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-1.6
-\strut\end{minipage} &
-\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-89.71
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-WHO
-\strut\end{minipage} &
-\begin{minipage}[t]{0.16\columnwidth}\raggedright\strut
-Recreational
-\strut\end{minipage} &
-\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-2
-\strut\end{minipage} &
-\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-102.6
-\strut\end{minipage}\tabularnewline
+U.S. EPA & Drinking & 0.3 & 23.68\tabularnewline
+WHO & Drinking & 1.0 & 63.94\tabularnewline
+U.S. EPA & Drinking & 1.6 & 89.71\tabularnewline
+WHO & Recreational & 2.0 & 102.60\tabularnewline
 \bottomrule

### 2015-09-16 manuscript/manuscript.Rmd
render with changes to fig width

@@
 `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[3]])/length(nla_dat$CHLA)),1)`%, and `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[4]])/length(nla_dat$CHLA)),1)`% of sample lakes for the U.S EPA childrens drinking water, the WHO drinking water, the U.S.
-EPA Adult drinking water, and the WHO recreationsl standards, respecitvely.
+EPA Adult drinking water, and the WHO recreational standards, respecitvely.
 <!-- (Figure \ref{fig:chla_micro_cdf}) -->
@@
 
-```{r chla_micro_scatter, echo=FALSE, fig.width=8, fig.cap="Scatterplot showing association betweeen chlorophyll \\textit{a} and microcystin-LR. \\label{fig:chla_micro_scatter}"}
+```{r chla_micro_scatter, echo=FALSE, fig.width=7, fig.cap="Scatterplot showing association betweeen chlorophyll \\textit{a} and microcystin-LR. \\label{fig:chla_micro_scatter}"}
 plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",pt_col=viridis(1),
@@
 
-```{r epa_child_cp_plot, echo=FALSE, fig.width=8, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
+```{r epa_child_cp_plot, echo=FALSE, fig.width=7, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
 \\label{fig:multi_cp_plot}"}
@@
           title="WHO Recreational Water Advisory Level"),
-  cols=2
+  cols=1
 )
@@
 
-```{r mc_chla_table, results='asis', echo=FALSE, tab.cap="Chlorophyll \textit{a} concentrations that are associated with a 50 percent probability of exceeding a microcystin-LR health advisory concentration.
+```{r mc_chla_table, results='asis', echo=FALSE, tab.cap="Chlorophyll \\textit{a} concentrations that are associated with a 50% probability of exceeding a microcystin-LR health advisory concentration.
 \\label{tab:mc_chla_table}"}

### 2015-09-16 manuscript/manuscript.tex
render with changes to fig width

@@
 childrens drinking water, the WHO drinking water, the U.S. EPA Adult
-drinking water, and the WHO recreationsl standards, respecitvely.
+drinking water, and the WHO recreational standards, respecitvely.
 
@@
 \begin{longtable}[c]{@{}llrr@{}}
-\caption{Chlorophyll extit\{a\} concentrations that are associated with
-a 50 percent probability of exceeding a microcystin-LR health advisory
+\caption{Chlorophyll \textit{a} concentrations that are associated with
+a 50\% probability of exceeding a microcystin-LR health advisory
 concentration. \label{tab:mc_chla_table}}\tabularnewline

### 2015-09-16 manuscript/manuscript.Rmd
fixed some rendering issues

@@
 
+\newpage
+
 #Figures
 
-```{r chla_micro_scatter, echo=FALSE, fig.width=7, fig.cap="Scatterplot showing association betweeen chlorophyll \\textit{a} and microcystin-LR. \\label{fig:chla_micro_scatter}"}
+```{r chla_micro_scatter, echo=FALSE, fig.width=6, fig.cap="Scatterplot showing association betweeen chlorophyll \\textit{a} and microcystin-LR. \\label{fig:chla_micro_scatter}"}
 plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",pt_col=viridis(1),
@@
 <!--
-```{r chla_micro_cdf, echo=FALSE, fig.width=8, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
+```{r chla_micro_cdf, echo=FALSE, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
 \\label{fig:chla_micro_cdf}"}
@@
 
-```{r epa_child_cp_plot, echo=FALSE, fig.width=7, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
+```{r epa_child_cp_plot, echo=FALSE, fig.width=6, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
+A.) Plot for U.S.
+EPA childrens drinking water advisory (0.3 $\\mu$g/L).
+B.) Plot for WHO drinking water advisory (1 $\\mu$g/L).
+C.) Plot for U.S.
+EPA adult drinking water advisory (1.6 $\\mu$g/L).
+D.) Plot for WHO recreational advisory (2 $\\mu$g/L).
 \\label{fig:multi_cp_plot}"}
@@
   plot_cp(epa_child_cp,x = "Log 10 Chlorophyll", 
-          y = "Probability of Exceeding 0.3 ug/L MLR",
-          title="U.S. EPA Children's Drinking Water Advisory Level"),
+          y = "Conditional Probability",
+          title="A."),
   plot_cp(who_drink_cp, x = "Log 10 Chlorophyll", 
-          y = "Probability of Exceeding 1 ug/L MLR", 
-          title="WHO Drinking Water Advisory Level"),
+          y = "Conditional Probability", 
+          title="B."),
   plot_cp(epa_adult_cp, x = "Log 10 Chlorophyll", 
-          y = "Probability of Exceeding 1.6 ug/L MLR", 
-          title="U.S. EPA Adult Drinking Water Advisory Level"),
+          y = "Conditional Probability", 
+          title="C."),
   plot_cp(who_rec_low1_cp, x = "Log 10 Chlorophyll", 
-          y = "Probability of Exceeding 2 ug/L MLR", 
-          title="WHO Recreational Water Advisory Level"),
-  cols=1
+          y = "Conditional Probability", 
+          title="D."),
+  cols=2
 )
@@
 
+\newpage
+
 #References

### 2015-09-21 manuscript/manuscript.Rmd
update

@@
 ```{r chla_micro_scatter, echo=FALSE, fig.width=6, fig.cap="Scatterplot showing association betweeen chlorophyll \\textit{a} and microcystin-LR. \\label{fig:chla_micro_scatter}"}
-plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",pt_col=viridis(1),
+plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",cat="WSA_ECO9",
              x=expression(paste('Log10 Chl ', italic("a"),' (',mu,'g/L)')),

### 2015-09-21 R/plot_cp.R

@@ -1,7 +1,7 @@
 #'get_nla
 #'
 #'Create plot of conditional probability object showing association between microcystin-LR
-#'advisory levels and chlorophyll \it{a} 
+#'advisory levels and chlorophyll \\textit{a} 
 #'
 #' @param cp A conditional probability object from condprob2
 #' @param ...  arguments for labels

### 2015-09-21 R/plot_scatter.R

@@ -5,12 +5,14 @@
 #' @export
 #' @import ggplot2
 
-plot_scatter<-function(df,xvar,yvar,pt_col=1,...){
+plot_scatter<-function(df,xvar,yvar,cat=NULL,pt_col=1,...){
+  browser()
   options(scipen=5)
   xdf<-data.frame(xvar=df[[xvar]],yvar=df[[yvar]])
+  if(!is.null(cat)){xdf<-data.frame(xdf,cat=df[[cat]]);pt_col<-cat}
   xdf<-xdf[complete.cases(xdf),]
   x <- ggplot(xdf,aes(xvar,yvar))+
-    geom_point(size=3,colour=pt_col) +
+    geom_point(size=3,aes(colour=pt_col)) +
     theme_bw()+
     theme(text = element_text(family="sans"),
           axis.title.x = element_text(family="sans",vjust = -0.5, size = 14),

### 2015-09-23 manuscript/manuscript.Rmd
started some on discussion

@@
 The median concentration was `r median(nla_dat$CHLA,na.rm=T)` and the mean was `r mean(nla_dat$CHLA,na.rm=T)`.
+The associations between chlorophyll *a* and the upper confidence interval with a conditional probability of 50% were `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]` for 0.3, 1.0, 1.6 and 2.0 $\mu$g/L, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
 
+#Discussion
 The association between Log 10 microcystin-LR and Log 10 chlorophyll *a* show a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
 This indicates that higher concentrations of microcystin-LR almost always co-occur with higher concentrations of chlorophyll *a* yet the inverse is not true.
-Higher chlorophyll *a* is not necessarily predictive of higher microcystin-LR concentrations; however, chlorophyll *a* may be predictive of the probability of exceeding a certian concentration.
-This is the case as the probability of exceeding each of the four tested health adviroy levels shos this patern of probability of exceeding the advisory level increases as a function of chlorophyll *a* concentration (Figure \ref{fig:multi_cp_plot}).
+Higher chlorophyll *a* is not necessarily predictive of higher microcystin-LR concentrations; however, chlorophyll *a* may be predictive of the probability of exceeding a certain concentration.
 
+This is the case as the probability of exceeding each of the four tested health advisory levels increases as a function of chlorophyll *a* concentration (Figure \ref{fig:multi_cp_plot}).
 We use this association to identify chlorophyll *a* concentrations that are associated with greater than even odds of exceeding a given health advisory level (Table \ref{tab:mc_chla_table}).
@@
 EPA Adult drinking water, and the WHO recreational standards, respecitvely.
-<!-- (Figure \ref{fig:chla_micro_cdf}) -->
 
-#Discussion
+There are numerous possible uses for the chlorophyll *a* and microcystin-LR advisory cut-off values.
+First, in the absence of microcystin-LR measurements, exceedence of the chlorophyll *a* concentrations could be a trigger for further actions.
+Given that there is uncertainity around these chlorophyll *a* cutoffs the best case scenario would be to monitor for chlorophyll and in the event of exceeding a target concentration take water samples and have those samples tested in a lab for microcystin-LR.
+A second potential use is to identify possible bloom events from historical data.
+Currently, there are several suggestions on how to identify a bloom event, but there is variation across the definitions and quantitative links to measures of cyanobacteria are not always included [REFS].
+Using our chlorophyll *a* cutoffs provides a value that is directly associated with microcystin-LR and can be used to classify lakes, from past surveys, as havin bloomed.
+
+These values are conservative: greater than even odds for the upper CI. Most protective.
+
+Not meant as a replacement for testing of microcystin, but provides another useful tool to help understand the probable extent of microcystin occurence.
 
@@
 ```{r chla_micro_scatter, echo=FALSE, fig.width=6, fig.cap="Scatterplot showing association betweeen chlorophyll \\textit{a} and microcystin-LR. \\label{fig:chla_micro_scatter}"}
-plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",cat="WSA_ECO9",
+plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",pt_col = "black",
              x=expression(paste('Log10 Chl ', italic("a"),' (',mu,'g/L)')),

### 2015-09-23 manuscript/manuscript.tex
fixed some rendering issues
started some on discussion

@@
 \emph{a} concentrations that exceeded detection limits. The median
-concentration was 7.79 and the mean was 29.6301946.
+concentration was 7.79 and the mean was 29.6301946. The associations
+between chlorophyll \emph{a} and the upper confidence interval with a
+conditional probability of 50\% were 23.68, 63.94, 89.71, and 102.6 for
+0.3, 1.0, 1.6 and 2.0 \(\mu\)g/L, respectively (Table
+\ref{tab:mc_chla_table} \& Figure \ref{fig:multi_cp_plot}).
+
+\section{Discussion}\label{discussion}
 
@@
 predictive of higher microcystin-LR concentrations; however, chlorophyll
-\emph{a} may be predictive of the probability of exceeding a certian
-concentration. This is the case as the probability of exceeding each of
-the four tested health adviroy levels shos this patern of probability of
-exceeding the advisory level increases as a function of chlorophyll
-\emph{a} concentration (Figure \ref{fig:multi_cp_plot}).
+\emph{a} may be predictive of the probability of exceeding a certain
+concentration.
 
-We use this association to identify chlorophyll \emph{a} concentrations
-that are associated with greater than even odds of exceeding a given
-health advisory level (Table \ref{tab:mc_chla_table}). These represent
-27.8\%, 11.9\%, 8.3\%, and 7.4\% of sample lakes for the U.S EPA
-childrens drinking water, the WHO drinking water, the U.S. EPA Adult
-drinking water, and the WHO recreational standards, respecitvely.
+This is the case as the probability of exceeding each of the four tested
+health advisory levels increases as a function of chlorophyll \emph{a}
+concentration (Figure \ref{fig:multi_cp_plot}). We use this association
+to identify chlorophyll \emph{a} concentrations that are associated with
+greater than even odds of exceeding a given health advisory level (Table
+\ref{tab:mc_chla_table}). These represent 27.8\%, 11.9\%, 8.3\%, and
+7.4\% of sample lakes for the U.S EPA childrens drinking water, the WHO
+drinking water, the U.S. EPA Adult drinking water, and the WHO
+recreational standards, respecitvely.
 
-\section{Discussion}\label{discussion}
+There are numerous possible uses for the chlorophyll \emph{a} and
+microcystin-LR advisory cut-off values. First, in the absence of
+microcystin-LR measurements, exceedence of the chlorophyll \emph{a}
+concentrations could be a trigger for further actions. Given that there
+is uncertainity around these chlorophyll \emph{a} cutoffs the best case
+scenario would be to monitor for chlorophyll and in the event of
+exceeding a target concentration take water samples and have those
+samples tested in a lab for microcystin-LR. A second potential use is to
+identify possible bloom events from historical data. Currently, there
+are several suggestions on how to identify a bloom event, but there is
+variation across the definitions and quantitative links to measures of
+cyanobacteria are not always included {[}REFS{]}. Using our chlorophyll
+\emph{a} cutoffs provides a value that is directly associated with
+microcystin-LR and can be used to classify lakes, from past surveys, as
+havin bloomed.
+
+These values are conservative: greater than even odds for the upper CI.
+Most protective.
+
+Not meant as a replacement for testing of microcystin, but provides
+another useful tool to help understand the probable extent of
+microcystin occurence.
+
+\newpage
 
@@
 
+\begin{verbatim}
+## Loading required package: grid
+\end{verbatim}
+
 \begin{figure}[htbp]
@@
 probability of exceeding various microcystin-LR (MLR) health advisory
-Levels. \label{fig:multi_cp_plot}}
+Levels. A.) Plot for U.S. EPA childrens drinking water advisory (0.3
+\(\mu\)g/L). B.) Plot for WHO drinking water advisory (1 \(\mu\)g/L).
+C.) Plot for U.S. EPA adult drinking water advisory (1.6 \(\mu\)g/L).
+D.) Plot for WHO recreational advisory (2 \(\mu\)g/L).
+\label{fig:multi_cp_plot}}
 \end{figure}
@@
 
+\newpage
+
 \section*{References}\label{references}

### 2015-09-23 R/plot_scatter.R

@@ -6,10 +6,11 @@
 #' @import ggplot2
 
 plot_scatter<-function(df,xvar,yvar,cat=NULL,pt_col=1,...){
-  browser()
   options(scipen=5)
-  xdf<-data.frame(xvar=df[[xvar]],yvar=df[[yvar]])
-  if(!is.null(cat)){xdf<-data.frame(xdf,cat=df[[cat]]);pt_col<-cat}
+  xdf<-data.frame(xvar=df[[xvar]],yvar=df[[yvar]],pt_col=1)
+  if(!is.null(cat)){
+    xdf<-data.frame(xdf,pt_col=df[[cat]])
+  }
   xdf<-xdf[complete.cases(xdf),]
   x <- ggplot(xdf,aes(xvar,yvar))+
     geom_point(size=3,aes(colour=pt_col)) +

### 2015-09-23 manuscript/manuscript.Rmd
added a ref

@@
 A second potential use is to identify possible bloom events from historical data.
-Currently, there are several suggestions on how to identify a bloom event, but there is variation across the definitions and quantitative links to measures of cyanobacteria are not always included [REFS].
-Using our chlorophyll *a* cutoffs provides a value that is directly associated with microcystin-LR and can be used to classify lakes, from past surveys, as havin bloomed.
+As harmful algal blooms are made up of many species and have various mechanism responsible for adverse impacts (e.g.
+toxins, hypoxia, odors), there is no single definition of a bloom.
+For cHABs one approach has been to identify an increase over a baseline concentration of phycocyanin [@miller2013spatiotemporal].
+This is a useful approach for targeted studies, but phycocyanin is also not always available and measures the predominance of cyanobacterial pigments and not toxins.
+Using our chlorophyll *a* cutoffs provides a value that is more directly associated with microcystin-LR and can be used to classify lakes, from past surveys, as having bloomed.
 

### 2015-09-23 manuscript/manuscript.tex
re-render

@@
 samples tested in a lab for microcystin-LR. A second potential use is to
-identify possible bloom events from historical data. Currently, there
-are several suggestions on how to identify a bloom event, but there is
-variation across the definitions and quantitative links to measures of
-cyanobacteria are not always included {[}REFS{]}. Using our chlorophyll
-\emph{a} cutoffs provides a value that is directly associated with
-microcystin-LR and can be used to classify lakes, from past surveys, as
-havin bloomed.
+identify possible bloom events from historical data. As harmful algal
+blooms are made up of many species and have various mechanism
+responsible for adverse impacts (e.g.~toxins, hypoxia, odors), there is
+no single definition of a bloom. For cHABs one approach has been to
+identify an increase over a baseline concentration of phycocyanin
+(Miller et al. 2013). This is a useful approach for targeted studies,
+but phycocyanin is also not always available and measures the
+predominance of cyanobacterial pigments and not toxins. Using our
+chlorophyll \emph{a} cutoffs provides a value that is more directly
+associated with microcystin-LR and can be used to classify lakes, from
+past surveys, as having bloomed.
 
@@
 
+Miller, T. R., L. Beversdorf, S. D. Chaston, and K. D. McMahon. 2013.
+Spatiotemporal molecular analysis of cyanobacteria blooms reveals
+microcystis-aphanizomenon interactions. PloS one 8:e74933.
+
 Paul, J. F., and W. R. Munns. 2011. Probability surveys, conditional

### 2015-09-24 manuscript/manuscript.Rmd
-added refs and text about field test

--- 2015-09-21 manuscript/manuscript.Rmd
@@
 To help protect against these impacts, several health advisory levels have been set for some toxins.
-In particular, one of the more common toxins, microcystin, has several advisory levels set for drinking water and recreational use and managing water bodies to meet those levels could have far reaching benefits.
-However, compared to other water quality measures, measurements of microcystin are not common and current field measurement techniques have limited precision and accuracy.
+In particular, one of the more common toxins, microcystin, has several advisory levels set for drinking water and recreational use.
+However, compared to other water quality measures, field measurements of microcystin are not commonly available due to cost and require advanced understanding to interpret results.
 Addressing these issues will take time and resources.
@@
 
-In the summer of 2014, the city of Toledo, OH was forced to shut down their municipal water supply due in part to an excess of microcystin-LR that resulted from a ongoing cyanobacteria harmful algal bloom (cHAB) in Lake Erie [REFS].
-Since this event, significant legislation has been passed in the United States and the US Environmental Protection Agency (USEPA) has released suggested microcystin-LR concentrations that would trigger health advisories.
+In the summer of 2014, the city of Toledo, OH was forced to shut down their municipal water supply due in part to an excess of microcystin-LR that resulted from a ongoing cyanobacteria harmful algal bloom (cHAB) in Lake Erie [@jetoo2015toledo; @rinta2009lake].
+Since this event, significant legislation has been passed in the United States and the US Environmental Protection Agency (USEPA) has released suggested microcystin-LR (one of the more common toxins) concentrations that would trigger health advisories [@mcelhiney2005detection; @zurawell2005hepatotoxic].
 MORE ON THE LEVELS.
@@
 One of these complications is that they rely on available measurements of microcystin-LR.
-This toxin can be measured in the field using test strips but these are a coarse measure at best and currently available test strips focus on 1 and 10 $\mu$g/L [REFS].
-Measurements with greater accuracy and precision require taking water samples and processing those in a lab to determine the toxin concentration [REFS].
-Additionally, microcystin-LR is, currently, not a routinely collected water quality parameter,thus, availability of microcystin-LR data may limit our ability to screen water bodies for exceedances of the various health advisories concentrations.
-Until microcystin-LR concentrations are more widely collected an alternative measure is needed.
-Fortunately, microcystin-LR has been shown to be associated with several other, more easily measured components of water quality.
+While laboratory testing remains the gold standard for quantifying microcystin-LY concentrations in water samples, several field test kits have been developed.
+Even though field tests provide a much needed means for rapid assessment, they are infrequently available to lake managers.
+These kits are moderately expensive (approximately $150-$200 depending on specific kit) with a limited shelf life (typically one year) [@james2011environmental; @aranda2015evaluation].
+Additionally, each technique requires nuanced understanding of the detection method (e.g., limit of detection, specific microcystin variants being measured, and sampling protocol).
+Fortunately, microcystin-LR has been shown to be associated with several other, more commonly measured and understood components of water quality.
 
-Chlorophyll *a* is a very commonly measured components of water quality that is also known to be associated with Microsystin-LR concentrations [REFS].
+Chlorophyll *a* is a very commonly measured components of water quality that is also known to be associated with Microsystin-LR concentrations [@lee2000variation, @paerl2013harmful].
 Additionally there are many rapid measurements for assessing chlorophyll *a* levels *in situ*.

### 2015-09-25 manuscript/manuscript.Rmd

@@
 EPA Adult drinking water, and the WHO recreational standards, respecitvely.
-<!-- (Figure \ref{fig:chla_micro_cdf}) -->
 
-#Discussion
+There are numerous possible uses for the chlorophyll *a* and microcystin-LR advisory cut-off values.
+First, in the absence of microcystin-LR measurements, exceedence of the chlorophyll *a* concentrations could be a trigger for further actions.
+Given that there is uncertainity around these chlorophyll *a* cutoffs the best case scenario would be to monitor for chlorophyll and in the event of exceeding a target concentration take water samples and have those samples tested in a lab for microcystin-LR.
+A second potential use is to identify possible bloom events from historical data.
+As harmful algal blooms are made up of many species and have various mechanism responsible for adverse impacts (e.g.
+toxins, hypoxia, odors), there is no single definition of a bloom.
+For cHABs one approach has been to identify an increase over a baseline concentration of phycocyanin [@miller2013spatiotemporal].
+This is a useful approach for targeted studies, but phycocyanin is also not always available and measures the predominance of cyanobacterial pigments and not toxins.
+Using our chlorophyll *a* cutoffs provides a value that is more directly associated with microcystin-LR and can be used to classify lakes, from past surveys, as having bloomed.
+
+These values are conservative: greater than even odds for the upper CI. Most protective.
+
+Not meant as a replacement for testing of microcystin, but provides another useful tool to help understand the probable extent of microcystin occurence.
 

### 2015-09-25 manuscript/manuscript.tex
rendered Betty's commmetns

@@
 more common toxins, microcystin, has several advisory levels set for
-drinking water and recreational use and managing water bodies to meet
-those levels could have far reaching benefits. However, compared to
-other water quality measures, measurements of microcystin are not common
-and current field measurement techniques have limited precision and
-accuracy. Addressing these issues will take time and resources. Thus,
+drinking water and recreational use. However, compared to other water
+quality measures, field measurements of microcystin are not commonly
+available due to cost and require advanced understanding to interpret
+results. Addressing these issues will take time and resources. Thus,
 there is utility in finding indicators of microcystin that are already
@@
 that resulted from a ongoing cyanobacteria harmful algal bloom (cHAB) in
-Lake Erie {[}REFS{]}. Since this event, significant legislation has been
-passed in the United States and the US Environmental Protection Agency
-(USEPA) has released suggested microcystin-LR concentrations that would
-trigger health advisories. MORE ON THE LEVELS. While these levels and
-associated advisories are likely to help mitigate the impacts from
-harmful algal blooms, they are not without complications.
+Lake Erie (Rinta-Kanto et al. 2009, Jetoo et al. 2015). Since this
+event, significant legislation has been passed in the United States and
+the US Environmental Protection Agency (USEPA) has released suggested
+microcystin-LR (one of the more common toxins) concentrations that would
+trigger health advisories (McElhiney and Lawton 2005, Zurawell et al.
+2005). MORE ON THE LEVELS. While these levels and associated advisories
+are likely to help mitigate the impacts from harmful algal blooms, they
+are not without complications.
 
 One of these complications is that they rely on available measurements
-of microcystin-LR. This toxin can be measured in the field using test
-strips but these are a coarse measure at best and currently available
-test strips focus on 1 and 10 \(\mu\)g/L {[}REFS{]}. Measurements with
-greater accuracy and precision require taking water samples and
-processing those in a lab to determine the toxin concentration
-{[}REFS{]}. Additionally, microcystin-LR is, currently, not a routinely
-collected water quality parameter,thus, availability of microcystin-LR
-data may limit our ability to screen water bodies for exceedances of the
-various health advisories concentrations. Until microcystin-LR
-concentrations are more widely collected an alternative measure is
-needed. Fortunately, microcystin-LR has been shown to be associated with
-several other, more easily measured components of water quality.
+of microcystin-LR. While laboratory testing remains the gold standard
+for quantifying microcystin-LY concentrations in water samples, several
+field test kits have been developed. Even though field tests provide a
+much needed means for rapid assessment, they are infrequently available
+to lake managers. These kits are moderately expensive (approximately
+\$150-\$200 depending on specific kit) with a limited shelf life
+(typically one year) (James et al. 2011, Aranda-Rodriguez et al. 2015).
+Additionally, each technique requires nuanced understanding of the
+detection method (e.g., limit of detection, specific microcystin
+variants being measured, and sampling protocol). Fortunately,
+microcystin-LR has been shown to be associated with several other, more
+commonly measured and understood components of water quality.
 
@@
 quality that is also known to be associated with Microsystin-LR
-concentrations {[}REFS{]}. Additionally there are many rapid
-measurements for assessing chlorophyll \emph{a} levels \emph{in situ}.
-For instance, there are small or hand held flourometers that provide
-reliable measurements {[}REFS{]}. Given these facts, it might be
-possible to identify chlorophyll \emph{a} concentrations that would be
-associated with the various Microcystin-LR health advisory levels.
-Identifying these associations would provide another reliable tool for
-water resource managers to use to help manage the threat to public
-health posed by cHABs and would be especially useful in the absence of
-microcystin-LR concentrations. Thus, the goal of this paper is to utlize
-the National Lakes Assessment data and identify chlorophyll \emph{a}
-concentrations that are associated with higher probabilities of
-exceeding several microcystin-LR health advisory concentrations {[}NLA
-REF{]}. So that others may repeat or adjust this analysis, the data,
-code, and this manuscript are freely available via
+concentrations (Lee et al. 2000, Paerl and Otten (2013)). Additionally
+there are many rapid measurements for assessing chlorophyll \emph{a}
+levels \emph{in situ}. For instance, there are small or hand held
+flourometers that provide reliable measurements {[}REFS{]}. Given these
+facts, it might be possible to identify chlorophyll \emph{a}
+concentrations that would be associated with the various Microcystin-LR
+health advisory levels. Identifying these associations would provide
+another reliable tool for water resource managers to use to help manage
+the threat to public health posed by cHABs and would be especially
+useful in the absence of microcystin-LR concentrations. Thus, the goal
+of this paper is to utlize the National Lakes Assessment data and
+identify chlorophyll \emph{a} concentrations that are associated with
+higher probabilities of exceeding several microcystin-LR health advisory
+concentrations {[}NLA REF{]}. So that others may repeat or adjust this
+analysis, the data, code, and this manuscript are freely available via
 \href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USAPE/microcystinchla}}.
@@
 
+Aranda-Rodriguez, R., Z. Jin, J. Harvie, and A. Cabecinha. 2015.
+Evaluation of three field test kits to detect microcystins from a public
+health perspective. Harmful Algae 42:34--42.
+
+James, R., A. Gregg, A. Dindal, and J. McKernan. 2011. Environmental
+technology verification report: Abraxis microcystin test kits. Online
+document. Accessed online: June 22.
+
+Jetoo, S., V. I. Grover, and G. Krantzberg. 2015. The toledo drinking
+water advisory: Suggested application of the water safety planning
+approach. Sustainability 7:9787--9808.
+
+Lee, S., M.-H. Jang, H.-S. Kim, B.-D. Yoon, and H.-M. Oh. 2000.
+Variation of microcystin content of microcystis aeruginosa relative to
+medium n: P ratio and growth stage. Journal of Applied Microbiology
+89:323--329.
+
+McElhiney, J., and L. A. Lawton. 2005. Detection of the cyanobacterial
+hepatotoxins microcystins. Toxicology and Applied Pharmacology
+203:219--230.
+
 Miller, T. R., L. Beversdorf, S. D. Chaston, and K. D. McMahon. 2013.
@@
 microcystis-aphanizomenon interactions. PloS one 8:e74933.
+
+Paerl, H. W., and T. G. Otten. 2013. Harmful cyanobacterial blooms:
+Causes, consequences, and controls. Microbial ecology 65:995--1010.
 
@@
 
+Rinta-Kanto, J. M., E. A. Konopko, J. M. DeBruyn, R. A. Bourbonniere, G.
+L. Boyer, and S. W. Wilhelm. 2009. Lake erie microcystis: Relationship
+between microcystin production, dynamics of genotypes and environmental
+parameters in a large lake. Harmful Algae 8:665--673.
+
+Zurawell, R. W., H. Chen, J. M. Burke, and E. E. Prepas. 2005.
+Hepatotoxic cyanobacteria: A review of the biological importance of
+microcystins in freshwater environments. Journal of Toxicology and
+Environmental Health, Part B 8:1--37.
+
 \end{document}

### 2015-09-25 manuscript/manuscript.Rmd
started confusion matrix

@@
 library("knitr")
+library("e1071")
 opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=5)
@@
 
-#who_rec_low2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-#                     xImpair = log10(mc_guides[5]), ProbComp = "gt", Exceed = "gte",
-#                     ci = TRUE, R = 100, xW=wgt)
-
-#mc_chla[5] <- get_chla(who_rec_low2_cp,0.5)
-
-#who_rec_med1_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-#                     xImpair = log10(mc_guides[6]), ProbComp = "gt", Exceed = "gte",
-#                     ci = TRUE, R = 100, xW=wgt)
-
-#mc_chla[6] <- get_chla(who_rec_med1_cp,0.5)
-
-#who_rec_med2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-#                     xImpair = log10(mc_guides[7]), ProbComp = "gt", Exceed = "gte",
-#                     ci = TRUE, R = 100, xW=wgt)
+#Agreement Tables
+child_conmat<-table(chla=nla_dat$CHLA>=mc_chla[1],mcsyt=nla_dat$MCYST_TL_UGL>=0.3)
+who_drink_conmat<-table(nla_dat$CHLA>=mc_chla[2],nla_dat$MCYST_TL_UGL>=1)
+adult_conmat<-table(nla_dat$CHLA>=mc_chla[3],nla_dat$MCYST_TL_UGL>=1.6)
+who_rec_conmat<-table(nla_dat$CHLA>=mc_chla[4],nla_dat$MCYST_TL_UGL>=2)
+#Want to minimize false negatives.  Ratio or FN/FP+TP+TN
 

### 2015-09-25 manuscript/manuscript.tex
started confusion matrix

@@
 concentrations of 0.3, 1, 1.6, and 2 we find chlorophyll \emph{a}
-concentrations of 23.68, 63.94, 89.71, and 102.6, respectively. When
+concentrations of 21.92, 65.2, 84.24, and 113.14, respectively. When
 managing for these various microcystin levels exceeding these reported
@@
 between chlorophyll \emph{a} and the upper confidence interval with a
-conditional probability of 50\% were 23.68, 63.94, 89.71, and 102.6 for
+conditional probability of 50\% were 21.92, 65.2, 84.24, and 113.14 for
 0.3, 1.0, 1.6 and 2.0 \(\mu\)g/L, respectively (Table
@@
 greater than even odds of exceeding a given health advisory level (Table
-\ref{tab:mc_chla_table}). These represent 27.8\%, 11.9\%, 8.3\%, and
-7.4\% of sample lakes for the U.S EPA childrens drinking water, the WHO
+\ref{tab:mc_chla_table}). These represent 29.2\%, 11.7\%, 8.8\%, and
+6.3\% of sample lakes for the U.S EPA childrens drinking water, the WHO
 drinking water, the U.S. EPA Adult drinking water, and the WHO
@@
 \endhead
-U.S. EPA & Drinking & 0.3 & 23.68\tabularnewline
-WHO & Drinking & 1.0 & 63.94\tabularnewline
-U.S. EPA & Drinking & 1.6 & 89.71\tabularnewline
-WHO & Recreational & 2.0 & 102.60\tabularnewline
+U.S. EPA & Drinking & 0.3 & 21.92\tabularnewline
+WHO & Drinking & 1.0 & 65.20\tabularnewline
+U.S. EPA & Drinking & 1.6 & 84.24\tabularnewline
+WHO & Recreational & 2.0 & 113.14\tabularnewline
 \bottomrule

### 2015-09-29 manuscript/manuscript.Rmd
added conf. matrix for mcstyn and chla

@@
 #Agreement Tables
-child_conmat<-table(chla=nla_dat$CHLA>=mc_chla[1],mcsyt=nla_dat$MCYST_TL_UGL>=0.3)
-who_drink_conmat<-table(nla_dat$CHLA>=mc_chla[2],nla_dat$MCYST_TL_UGL>=1)
-adult_conmat<-table(nla_dat$CHLA>=mc_chla[3],nla_dat$MCYST_TL_UGL>=1.6)
-who_rec_conmat<-table(nla_dat$CHLA>=mc_chla[4],nla_dat$MCYST_TL_UGL>=2)
-#Want to minimize false negatives.  Ratio or FN/FP+TP+TN
+child_conmat<-conmat_to_df(table(chla=nla_dat$CHLA>=mc_chla[1],
+                                 mcsyt=nla_dat$MCYST_TL_UGL>=0.3))
+who_drink_conmat<-conmat_to_df(table(nla_dat$CHLA>=mc_chla[2],
+                        nla_dat$MCYST_TL_UGL>=1))
+adult_conmat<-conmat_to_df(table(nla_dat$CHLA>=mc_chla[3],
+                    nla_dat$MCYST_TL_UGL>=1.6))
+who_rec_conmat<-conmat_to_df(table(nla_dat$CHLA>=mc_chla[4],
+                      nla_dat$MCYST_TL_UGL>=2))
+total_accuracy<-NULL
+total_accuracy$epa_child<-round((child_conmat[1,1]+child_conmat[2,2])/
+  sum(child_conmat),2)
+total_accuracy$who_drink<-round((who_drink_conmat[1,1]+who_drink_conmat[2,2])/
+  sum(who_drink_conmat),2)
+total_accuracy$epa_adult<-round((adult_conmat[1,1]+adult_conmat[2,2])/
+  sum(adult_conmat),2)
+total_accuracy$who_rec<-round((who_rec_conmat[1,1]+who_rec_conmat[2,2])/
+  sum(who_rec_conmat),2)
+
+fn_percent<-NULL
+fn_percent$epa_child<-round((child_conmat[1,2])/
+  sum(child_conmat),2)
+fn_percent$who_drink<-round((who_drink_conmat[1,2])/
+  sum(who_drink_conmat),2)
+fn_percent$epa_adult<-round((adult_conmat[1,2])/
+  sum(adult_conmat),2)
+fn_percent$who_rec<-round((who_rec_conmat[1,2])/
+  sum(who_rec_conmat),2)
+
+#Want to minimize false negatives.  Ratio of FN/FP+TP+TN
 

### 2015-09-29 manuscript/manuscript.tex
re render plus started confusion matrix text

@@
 concentrations of 0.3, 1, 1.6, and 2 we find chlorophyll \emph{a}
-concentrations of 21.92, 65.2, 84.24, and 113.14, respectively. When
+concentrations of 22.56, 59.4, 84.24, and 97.49, respectively. When
 managing for these various microcystin levels exceeding these reported
@@
 between chlorophyll \emph{a} and the upper confidence interval with a
-conditional probability of 50\% were 21.92, 65.2, 84.24, and 113.14 for
+conditional probability of 50\% were 22.56, 59.4, 84.24, and 97.49 for
 0.3, 1.0, 1.6 and 2.0 \(\mu\)g/L, respectively (Table
@@
 greater than even odds of exceeding a given health advisory level (Table
-\ref{tab:mc_chla_table}). These represent 29.2\%, 11.7\%, 8.8\%, and
-6.3\% of sample lakes for the U.S EPA childrens drinking water, the WHO
+\ref{tab:mc_chla_table}). These represent 28.8\%, 12.7\%, 8.8\%, and
+7.9\% of sample lakes for the U.S EPA childrens drinking water, the WHO
 drinking water, the U.S. EPA Adult drinking water, and the WHO
 recreational standards, respecitvely.
+
+Furthermore, the chlorophyll \emph{a} cutoffs may be used to predict
+whether or not a lake exceeds the microcystin-LR health advisories.
+Doing so allows us to compare the accuracy of the prediction as well as
+evaluate false negatives in this prediction. Stuff of FN plus
+presenation of conf matrix\ldots{}
 
@@
 \endhead
-U.S. EPA & Drinking & 0.3 & 21.92\tabularnewline
-WHO & Drinking & 1.0 & 65.20\tabularnewline
+U.S. EPA & Drinking & 0.3 & 22.56\tabularnewline
+WHO & Drinking & 1.0 & 59.40\tabularnewline
 U.S. EPA & Drinking & 1.6 & 84.24\tabularnewline
-WHO & Recreational & 2.0 & 113.14\tabularnewline
+WHO & Recreational & 2.0 & 97.49\tabularnewline
 \bottomrule

### 2015-09-30 manuscript/manuscript.Rmd
re render plus started confusion matrix text
some additional writing plus a new render

@@
 `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[3]])/length(nla_dat$CHLA)),1)`%, and `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[4]])/length(nla_dat$CHLA)),1)`% of sample lakes for the U.S EPA childrens drinking water, the WHO drinking water, the U.S.
-EPA Adult drinking water, and the WHO recreational standards, respecitvely.
+EPA Adult drinking water, and the WHO recreational advisories, respecitvely.
+
+Furthermore, the chlorophyll *a* cutoffs may be used to predict whether or not a lake exceeds the microcystin-LR health advisories.
+Doing so allows us to compare the accuracy of the prediction as well as evaluate false negatives.
+Total accuracy of the four cutoffs predicting microcystin-LR exceedances were `r total_accuracy$epa_child*100`% for the USEPA children's advisory, `r total_accuracy$who_drink*100`% for the WHO drinking water advisory,  `r total_accuracy$epa_adult*100`% for the USEPA adult advisory and, `r total_accuracy$who_rec*100`% for the WHO regreational advisory.
+But Total Accuracy is one part of the prediction performace with which we are concerned.
+
+When using the chlorophyll *a* cutoffs as an indicator of microcystin-LR exceedances, the error that should be avoided is predicting that no exceedance has occurred when in fact it has.
+In other words, we would like to avoid Type II errors and minimize the proportion of false negatives [REF].
+For the four chlorophyll *a* cut-offs we had a proportion of false negatives of `r fn_percent$epa_child*100`, `r fn_percent$who_drink*100`, `r fn_percent$epa_adult*100` and , `r fn_percent$who_rec*100 for the U.S EPA childrens drinking water, the WHO drinking water, the U.S.
+EPA Adult drinking water, and the WHO recreational advisories, respecitvely.
+In each case we miss less than 10% of the lakes that are in fact exceeding the microcystin-LR advisory.
+This is an acceptable rate, especially when no other tests are available for a given water body.
+TEST LOWER VALUES to see where we can get it even lower.
 

### 2015-09-30 manuscript/manuscript.tex
some additional writing plus a new render

@@
 drinking water, the U.S. EPA Adult drinking water, and the WHO
-recreational standards, respecitvely.
+recreational advisories, respecitvely.
 
@@
 Doing so allows us to compare the accuracy of the prediction as well as
-evaluate false negatives in this prediction. Stuff of FN plus
-presenation of conf matrix\ldots{}
+evaluate false negatives. Total accuracy of the four cutoffs predicting
+microcystin-LR exceedances were 74\% for the USEPA children's advisory,
+85\% for the WHO drinking water advisory, 89\% for the USEPA adult
+advisory and, 90\% for the WHO regreational advisory. But Total Accuracy
+is one part of the prediction performace with which we are concerned.
+
+When using the chlorophyll \emph{a} cutoffs as an indicator of
+microcystin-LR exceedances, the error that should be avoided is
+predicting that no exceedance has occurred when in fact it has. In other
+words, we would like to avoid Type II errors and minimize the proportion
+of false negatives {[}REF{]}. For the four chlorophyll \emph{a} cut-offs
+we had a proportion of false negatives of 9, 7, 6 and , `r
+fn\_percent\$who\_rec*100 for the U.S EPA childrens drinking water, the
+WHO drinking water, the U.S. EPA Adult drinking water, and the WHO
+recreational advisories, respecitvely. In each case we miss less than
+10\% of the lakes that are in fact exceeding the microcystin-LR
+advisory. This is an acceptable rate, especially when no other tests are
+available for a given water body. TEST LOWER VALUES to see where we can
+get it even lower.
 

### 2015-09-30 manuscript/manuscript.Rmd
first complete draft (with holes) and render

@@
 
-#Want to minimize false negatives.  Ratio of FN/FP+TP+TN
-
-#mc_chla[7] <- get_chla(who_rec_med2_cp,0.5)
 ```
@@
 
-In the summer of 2014, the city of Toledo, OH was forced to shut down their municipal water supply due in part to an excess of microcystin-LR that resulted from a ongoing cyanobacteria harmful algal bloom (cHAB) in Lake Erie [@jetoo2015toledo; @rinta2009lake].
-Since this event, significant legislation has been passed in the United States and the US Environmental Protection Agency (USEPA) has released suggested microcystin-LR (one of the more common toxins) concentrations that would trigger health advisories [@mcelhiney2005detection; @zurawell2005hepatotoxic].
-MORE ON THE LEVELS.
+Over the last decade or so numerous events and legislative activities have raised the public awareness of harmful algal blooms [@jetoo2015toledo; @rinta2009lake; NEED HABHRCA REFS].
+Recent, the US Environmental Protection Agency (USEPA) has released suggested microcystin-LR (one of the more common toxins) concentrations that would trigger health advisories [@mcelhiney2005detection; @zurawell2005hepatotoxic].
+Additionally, the World Health Organization (WHO) has had proposed advisory levels for drinking water and a range of recreational risk levels [REF].
 While these levels and associated advisories are likely to help mitigate the impacts from harmful algal blooms, they are not without complications.
@@
 This is an acceptable rate, especially when no other tests are available for a given water body.
-TEST LOWER VALUES to see where we can get it even lower.
 
@@
 
-These values are conservative: greater than even odds for the upper CI. Most protective.
-
-Not meant as a replacement for testing of microcystin, but provides another useful tool to help understand the probable extent of microcystin occurence.
+Lastly, using chlorophyll *a* is not meant as a replacement for testing of microcystin-LR.
+It should be used when other, direct measurements of cyanotoxins are not available.
+In those cases, which are likely to be common at least in the near future, using a more ubiquitous meausrement, such as chlorophyll *a* will provide a reasonable proxy for the probability of exceeding a microcystin-LR health advisory level and provide better protection against adverse effects in both drinking and recreational use cases.
 

### 2015-09-30 manuscript/manuscript.tex
first complete draft (with holes) and render

@@
 concentrations of 0.3, 1, 1.6, and 2 we find chlorophyll \emph{a}
-concentrations of 22.56, 59.4, 84.24, and 97.49, respectively. When
+concentrations of 22.25, 65.6, 84.96, and 105.84, respectively. When
 managing for these various microcystin levels exceeding these reported
@@
 
-In the summer of 2014, the city of Toledo, OH was forced to shut down
-their municipal water supply due in part to an excess of microcystin-LR
-that resulted from a ongoing cyanobacteria harmful algal bloom (cHAB) in
-Lake Erie (Rinta-Kanto et al. 2009, Jetoo et al. 2015). Since this
-event, significant legislation has been passed in the United States and
-the US Environmental Protection Agency (USEPA) has released suggested
+Over the last decade or so numerous events and legislative activities
+have raised the public awareness of harmful algal blooms {[}Jetoo et al.
+(2015); Rinta-Kanto et al. (2009); NEED HABHRCA REFS{]}. Recent, the US
+Environmental Protection Agency (USEPA) has released suggested
 microcystin-LR (one of the more common toxins) concentrations that would
 trigger health advisories (McElhiney and Lawton 2005, Zurawell et al.
-2005). MORE ON THE LEVELS. While these levels and associated advisories
-are likely to help mitigate the impacts from harmful algal blooms, they
-are not without complications.
+2005). Additionally, the World Health Organization (WHO) has had
+proposed advisory levels for drinking water and a range of recreational
+risk levels {[}REF{]}. While these levels and associated advisories are
+likely to help mitigate the impacts from harmful algal blooms, they are
+not without complications.
 
@@
 between chlorophyll \emph{a} and the upper confidence interval with a
-conditional probability of 50\% were 22.56, 59.4, 84.24, and 97.49 for
+conditional probability of 50\% were 22.25, 65.6, 84.96, and 105.84 for
 0.3, 1.0, 1.6 and 2.0 \(\mu\)g/L, respectively (Table
@@
 greater than even odds of exceeding a given health advisory level (Table
-\ref{tab:mc_chla_table}). These represent 28.8\%, 12.7\%, 8.8\%, and
-7.9\% of sample lakes for the U.S EPA childrens drinking water, the WHO
+\ref{tab:mc_chla_table}). These represent 29.1\%, 11.5\%, 8.6\%, and
+6.9\% of sample lakes for the U.S EPA childrens drinking water, the WHO
 drinking water, the U.S. EPA Adult drinking water, and the WHO
@@
 microcystin-LR exceedances were 74\% for the USEPA children's advisory,
-85\% for the WHO drinking water advisory, 89\% for the USEPA adult
-advisory and, 90\% for the WHO regreational advisory. But Total Accuracy
+86\% for the WHO drinking water advisory, 90\% for the USEPA adult
+advisory and, 91\% for the WHO regreational advisory. But Total Accuracy
 is one part of the prediction performace with which we are concerned.
@@
 advisory. This is an acceptable rate, especially when no other tests are
-available for a given water body. TEST LOWER VALUES to see where we can
-get it even lower.
+available for a given water body.
 
@@
 
-These values are conservative: greater than even odds for the upper CI.
-Most protective.
-
-Not meant as a replacement for testing of microcystin, but provides
-another useful tool to help understand the probable extent of
-microcystin occurence.
+Lastly, using chlorophyll \emph{a} is not meant as a replacement for
+testing of microcystin-LR. It should be used when other, direct
+measurements of cyanotoxins are not available. In those cases, which are
+likely to be common at least in the near future, using a more ubiquitous
+meausrement, such as chlorophyll \emph{a} will provide a reasonable
+proxy for the probability of exceeding a microcystin-LR health advisory
+level and provide better protection against adverse effects in both
+drinking and recreational use cases.
 
@@
 \newpage
+
+\begin{verbatim}
+## Loading required package: grid
+\end{verbatim}
 
@@
 \endhead
-U.S. EPA & Drinking & 0.3 & 22.56\tabularnewline
-WHO & Drinking & 1.0 & 59.40\tabularnewline
-U.S. EPA & Drinking & 1.6 & 84.24\tabularnewline
-WHO & Recreational & 2.0 & 97.49\tabularnewline
+U.S. EPA & Drinking & 0.3 & 22.25\tabularnewline
+WHO & Drinking & 1.0 & 65.60\tabularnewline
+U.S. EPA & Drinking & 1.6 & 84.96\tabularnewline
+WHO & Recreational & 2.0 & 105.84\tabularnewline
 \bottomrule

### 2015-10-01 manuscript/manuscript.Rmd
cleaned up refs, added tables:  Final draft prior to full read through

@@
 
-Over the last decade or so numerous events and legislative activities have raised the public awareness of harmful algal blooms [@jetoo2015toledo; @rinta2009lake; NEED HABHRCA REFS].
-Recent, the US Environmental Protection Agency (USEPA) has released suggested microcystin-LR (one of the more common toxins) concentrations that would trigger health advisories [@mcelhiney2005detection; @zurawell2005hepatotoxic].
-Additionally, the World Health Organization (WHO) has had proposed advisory levels for drinking water and a range of recreational risk levels [REF].
+Over the last decade or so numerous events and legislative activities have raised the public awareness of harmful algal blooms [@jetoo2015toledo; @rinta2009lake; @HABHRCA2014].
+Recently, the US Environmental Protection Agency (USEPA) has released suggested microcystin-LR (one of the more common toxins) concentrations that would trigger health advisories [@mcelhiney2005detection; @zurawell2005hepatotoxic; @usepa2015drinking].
+Additionally, the World Health Organization (WHO) has had proposed advisory levels for drinking water and a range of recreational risk levels [@world2003cyanobacterial; @chorus1999toxic].
 While these levels and associated advisories are likely to help mitigate the impacts from harmful algal blooms, they are not without complications.
@@
 
-Chlorophyll *a* is a very commonly measured components of water quality that is also known to be associated with Microsystin-LR concentrations [@lee2000variation, @paerl2013harmful].
+Chlorophyll *a* is a very commonly measured components of water quality that is also known to be associated with Microsystin-LR concentrations [@lee2000variation; @paerl2013harmful; heisler2008eutrophication].
 Additionally there are many rapid measurements for assessing chlorophyll *a* levels *in situ*.
-For instance, there are small or hand held flourometers that provide reliable measurements [REFS].
+For instance, there are small or hand held flourometers that measure chlorohpyll *a*.
 Given these facts, it might be possible to identify chlorophyll *a* concentrations that would be associated with the various Microcystin-LR health advisory levels.
 Identifying these associations would provide another reliable tool for water resource managers to use to help manage the threat to public health posed by cHABs and would be especially useful in the absence of microcystin-LR concentrations.
-Thus, the goal of this paper is to utlize the National Lakes Assessment data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin-LR health advisory concentrations [NLA REF].
+Thus, the goal of this paper is to utlize the National Lakes Assessment data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin-LR health advisory concentrations [@usepa2009national].
 So that others may repeat or adjust this analysis, the data, code, and this manuscript are freely available via [https://github.com/USAPE/microcystinchla](https://github.com/USAPE/microcystinchla).
@@
 ##Data
-We used the 2007 National Lakes Assessment (NLA) water quality and microcystin-LR concentration data [REF].
+We used the 2007 National Lakes Assessment (NLA) water quality and microcystin-LR concentration data [@usepa2009national].
 These data represent a snapshot of water quality from the summer of 2007 and data on chlorophyll *a* and microcystin-LR concentrations are available for `r nrow(na.omit(nla_dat$CHLA,nla_dat$MCYST_TL_UGL))` lakes.
@@
 A detailed discussion of CPA is beyond the scope of this paper, but see Paul et al.
-[-REF] and Hollister et al.
-[-REF] for details.
+[-@paul2005development] and Hollister et al.
+[-@hollister2008cprob] for details.
 For this analysis, we used CPA to examine how the conditional probability of exceeding one of the health advisory changes as chlorophyll *a* increases in a lake.
@@
 Doing so allows us to compare the accuracy of the prediction as well as evaluate false negatives.
-Total accuracy of the four cutoffs predicting microcystin-LR exceedances were `r total_accuracy$epa_child*100`% for the USEPA children's advisory, `r total_accuracy$who_drink*100`% for the WHO drinking water advisory,  `r total_accuracy$epa_adult*100`% for the USEPA adult advisory and, `r total_accuracy$who_rec*100`% for the WHO regreational advisory.
+Total accuracy of the four cutoffs predicting microcystin-LR exceedances were `r total_accuracy$epa_child*100`% for the USEPA children's advisory, `r total_accuracy$who_drink*100`% for the WHO drinking water advisory,  `r total_accuracy$epa_adult*100`% for the USEPA adult advisory and, `r total_accuracy$who_rec*100`% for the WHO regreational advisory (Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table}, \ref{tab:adult_conmat_table}, & \ref{tab:who_rec_conmat_table}).
 But Total Accuracy is one part of the prediction performace with which we are concerned.
@@
 When using the chlorophyll *a* cutoffs as an indicator of microcystin-LR exceedances, the error that should be avoided is predicting that no exceedance has occurred when in fact it has.
-In other words, we would like to avoid Type II errors and minimize the proportion of false negatives [REF].
-For the four chlorophyll *a* cut-offs we had a proportion of false negatives of `r fn_percent$epa_child*100`, `r fn_percent$who_drink*100`, `r fn_percent$epa_adult*100` and , `r fn_percent$who_rec*100 for the U.S EPA childrens drinking water, the WHO drinking water, the U.S.
+In other words, we would like to avoid Type II errors and minimize the proportion of false negatives.
+For the four chlorophyll *a* cut-offs we had a proportion of false negatives of `r fn_percent$epa_child*100`%, `r fn_percent$who_drink*100`%, `r fn_percent$epa_adult*100`% and , `r fn_percent$who_rec*100`% for the U.S EPA childrens drinking water, the WHO drinking water, the U.S.
 EPA Adult drinking water, and the WHO recreational advisories, respecitvely.
@@
 
+```{r child_conmat_table, results='asis', echo=FALSE, tab.cap="Confusion matrix comparing chlorophyll \\textit{a} predicted exceedences (rows) versus real exceedances (columns) for the U.S.
+EPA childrens drinking water advisory.
+\\label{tab:child_conmat_table}"}
+knitr::kable(child_conmat)
+```
+
+\newpage
+
+```{r who_drink_conmat_table, results='asis', echo=FALSE, tab.cap="Confusion matrix comparing chlorophyll \\textit{a} predicted exceedences (rows) versus real exceedances (columns) for the WHO drinking water advisory.
+\\label{tab:who_drink_conmat_table}"}
+knitr::kable(who_drink_conmat)
+```
+
+\newpage
+
+```{r adult_conmat_table, results='asis', echo=FALSE, tab.cap="Confusion matrix comparing chlorophyll \\textit{a} predicted exceedences (rows) versus real exceedances (columns) for the U.S.
+EPA adult drinking water advisory.
+\\label{tab:adult_conmat_table}"}
+knitr::kable(adult_conmat)
+```
+
+\newpage
+
+```{r who_rec_conmat_table, results='asis', echo=FALSE, tab.cap="Confusion matrix comparing chlorophyll \\textit{a} predicted exceedences (rows) versus real exceedances (columns) for the WHO recreational water advisory.
+\\label{tab:who_rec_conmat_table}"}
+knitr::kable(who_rec_conmat)
+```
+
+\newpage
+
 #References

### 2015-10-01 manuscript/manuscript.tex
cleaned up refs, added tables:  Final draft prior to full read through

@@
 Over the last decade or so numerous events and legislative activities
-have raised the public awareness of harmful algal blooms {[}Jetoo et al.
-(2015); Rinta-Kanto et al. (2009); NEED HABHRCA REFS{]}. Recent, the US
+have raised the public awareness of harmful algal blooms (Rinta-Kanto et
+al. 2009, HABHRCA 2014, Jetoo et al. 2015). Recently, the US
 Environmental Protection Agency (USEPA) has released suggested
@@
 trigger health advisories (McElhiney and Lawton 2005, Zurawell et al.
-2005). Additionally, the World Health Organization (WHO) has had
-proposed advisory levels for drinking water and a range of recreational
-risk levels {[}REF{]}. While these levels and associated advisories are
-likely to help mitigate the impacts from harmful algal blooms, they are
-not without complications.
+2005, USEPA 2015). Additionally, the World Health Organization (WHO) has
+had proposed advisory levels for drinking water and a range of
+recreational risk levels (Chorus and Bartram 1999, (WHO) and others
+2003). While these levels and associated advisories are likely to help
+mitigate the impacts from harmful algal blooms, they are not without
+complications.
 
@@
 quality that is also known to be associated with Microsystin-LR
-concentrations (Lee et al. 2000, Paerl and Otten (2013)). Additionally
-there are many rapid measurements for assessing chlorophyll \emph{a}
-levels \emph{in situ}. For instance, there are small or hand held
-flourometers that provide reliable measurements {[}REFS{]}. Given these
-facts, it might be possible to identify chlorophyll \emph{a}
-concentrations that would be associated with the various Microcystin-LR
-health advisory levels. Identifying these associations would provide
-another reliable tool for water resource managers to use to help manage
-the threat to public health posed by cHABs and would be especially
-useful in the absence of microcystin-LR concentrations. Thus, the goal
-of this paper is to utlize the National Lakes Assessment data and
-identify chlorophyll \emph{a} concentrations that are associated with
-higher probabilities of exceeding several microcystin-LR health advisory
-concentrations {[}NLA REF{]}. So that others may repeat or adjust this
-analysis, the data, code, and this manuscript are freely available via
+concentrations {[}Lee et al. (2000); Paerl and Otten (2013);
+heisler2008eutrophication{]}. Additionally there are many rapid
+measurements for assessing chlorophyll \emph{a} levels \emph{in situ}.
+For instance, there are small or hand held flourometers that measure
+chlorohpyll \emph{a}. Given these facts, it might be possible to
+identify chlorophyll \emph{a} concentrations that would be associated
+with the various Microcystin-LR health advisory levels. Identifying
+these associations would provide another reliable tool for water
+resource managers to use to help manage the threat to public health
+posed by cHABs and would be especially useful in the absence of
+microcystin-LR concentrations. Thus, the goal of this paper is to utlize
+the National Lakes Assessment data and identify chlorophyll \emph{a}
+concentrations that are associated with higher probabilities of
+exceeding several microcystin-LR health advisory concentrations (USEPA
+2009). So that others may repeat or adjust this analysis, the data,
+code, and this manuscript are freely available via
 \href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USAPE/microcystinchla}}.
@@
 We used the 2007 National Lakes Assessment (NLA) water quality and
-microcystin-LR concentration data {[}REF{]}. These data represent a
+microcystin-LR concentration data (USEPA 2009). These data represent a
 snapshot of water quality from the summer of 2007 and data on
@@
 A detailed discussion of CPA is beyond the scope of this paper, but see
-Paul et al. {[}-REF{]} and Hollister et al. {[}-REF{]} for details. For
-this analysis, we used CPA to examine how the conditional probability of
+Paul et al. (2005) and Hollister et al. (2008) for details. For this
+analysis, we used CPA to examine how the conditional probability of
 exceeding one of the health advisory changes as chlorophyll \emph{a}
@@
 86\% for the WHO drinking water advisory, 90\% for the USEPA adult
-advisory and, 91\% for the WHO regreational advisory. But Total Accuracy
-is one part of the prediction performace with which we are concerned.
+advisory and, 91\% for the WHO regreational advisory (Tables
+\ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table},
+\ref{tab:adult_conmat_table}, \& \ref{tab:who_rec_conmat_table}). But
+Total Accuracy is one part of the prediction performace with which we
+are concerned.
 
@@
 words, we would like to avoid Type II errors and minimize the proportion
-of false negatives {[}REF{]}. For the four chlorophyll \emph{a} cut-offs
-we had a proportion of false negatives of 9, 7, 6 and , `r
-fn\_percent\$who\_rec*100 for the U.S EPA childrens drinking water, the
-WHO drinking water, the U.S. EPA Adult drinking water, and the WHO
-recreational advisories, respecitvely. In each case we miss less than
-10\% of the lakes that are in fact exceeding the microcystin-LR
-advisory. This is an acceptable rate, especially when no other tests are
-available for a given water body.
+of false negatives. For the four chlorophyll \emph{a} cut-offs we had a
+proportion of false negatives of 9\%, 7\%, 6\% and , 5\% for the U.S EPA
+childrens drinking water, the WHO drinking water, the U.S. EPA Adult
+drinking water, and the WHO recreational advisories, respecitvely. In
+each case we miss less than 10\% of the lakes that are in fact exceeding
+the microcystin-LR advisory. This is an acceptable rate, especially when
+no other tests are available for a given water body.
 
@@
 
+\begin{longtable}[c]{@{}lrr@{}}
+\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
+exceedences (rows) versus real exceedances (columns) for the U.S. EPA
+childrens drinking water advisory.
+\label{tab:child_conmat_table}}\tabularnewline
+\toprule
+& FALSE & TRUE\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+& FALSE & TRUE\tabularnewline
+\midrule
+\endhead
+FALSE & 638 & 94\tabularnewline
+TRUE & 173 & 123\tabularnewline
+\bottomrule
+\end{longtable}
+
+\newpage
+
+\begin{longtable}[c]{@{}lrr@{}}
+\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
+exceedences (rows) versus real exceedances (columns) for the WHO
+drinking water advisory.
+\label{tab:who_drink_conmat_table}}\tabularnewline
+\toprule
+& FALSE & TRUE\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+& FALSE & TRUE\tabularnewline
+\midrule
+\endhead
+FALSE & 838 & 76\tabularnewline
+TRUE & 69 & 45\tabularnewline
+\bottomrule
+\end{longtable}
+
+\newpage
+
+\begin{longtable}[c]{@{}lrr@{}}
+\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
+exceedences (rows) versus real exceedances (columns) for the U.S. EPA
+adult drinking water advisory.
+\label{tab:adult_conmat_table}}\tabularnewline
+\toprule
+& FALSE & TRUE\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+& FALSE & TRUE\tabularnewline
+\midrule
+\endhead
+FALSE & 887 & 57\tabularnewline
+TRUE & 50 & 34\tabularnewline
+\bottomrule
+\end{longtable}
+
+\newpage
+
+\begin{longtable}[c]{@{}lrr@{}}
+\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
+exceedences (rows) versus real exceedances (columns) for the WHO
+recreational water advisory.
+\label{tab:who_rec_conmat_table}}\tabularnewline
+\toprule
+& FALSE & TRUE\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+& FALSE & TRUE\tabularnewline
+\midrule
+\endhead
+FALSE & 910 & 52\tabularnewline
+TRUE & 43 & 23\tabularnewline
+\bottomrule
+\end{longtable}
+
+\newpage
+
 \section*{References}\label{references}
@@
 health perspective. Harmful Algae 42:34--42.
+
+Chorus, E. I., and J. Bartram. 1999. Toxic cyanobacteria in water: A
+guide to their public health consequences, monitoring and management.
+
+HABHRCA. 2014. Harmful algal bloom and hypoxia research and control
+amendments act of 2014.
+
+Hollister, J. W., H. A. Walker, and J. F. Paul. 2008. CProb: A
+computational tool for conducting conditional probability analysis.
+Journal of environmental quality 37:2392--2396.
 
@@
 
+Paul, J. F., and M. E. McDonald. 2005. Development of empirical,
+geographically specific water quality criteria: A conditional
+probability analysis approach 41:1211--1223.
+
 Paul, J. F., and W. R. Munns. 2011. Probability surveys, conditional
@@
 
+USEPA. 2009. National lakes assessment: A collaborative survey of the
+nation's lakes. ePA 841-r-09-001. Office of Water; Office of Research;
+Development, US Environmental Protection Agency Washington, DC.
+
+USEPA. 2015. Drinking water health advisory for the cyanobacterial
+microcystin toxins. ePA-820R15100. Office of Water, US Environmental
+Protection Agency Washington, DC.
+
+(WHO), W. H. O., and others. 2003. Cyanobacterial toxins: Microcystin-lR
+in drinking-water. background document for development of wHO guidelines
+for drinking-water quality. geneva, switzerland. World Health
+Organization, 2nd ed. Geneva.
+
 Zurawell, R. W., H. Chen, J. M. Burke, and E. E. Prepas. 2005.

### 2015-10-01 manuscript/manuscript.tex
jeff read through done.  Send to betty.  Plus a render

@@
 have raised the public awareness of harmful algal blooms (Rinta-Kanto et
-al. 2009, HABHRCA 2014, Jetoo et al. 2015). Recently, the US
-Environmental Protection Agency (USEPA) has released suggested
-microcystin-LR (one of the more common toxins) concentrations that would
-trigger health advisories (McElhiney and Lawton 2005, Zurawell et al.
-2005, USEPA 2015). Additionally, the World Health Organization (WHO) has
-had proposed advisory levels for drinking water and a range of
-recreational risk levels (Chorus and Bartram 1999, (WHO) and others
-2003). While these levels and associated advisories are likely to help
-mitigate the impacts from harmful algal blooms, they are not without
-complications.
+al. 2009, HABHRCA 2014, Jetoo et al. 2015), and the US Environmental
+Protection Agency (USEPA) has recently released suggested microcystin-LR
+(one of the more common toxins) concentrations that would trigger health
+advisories (McElhiney and Lawton 2005, Zurawell et al. 2005, USEPA
+2015). Additionally, the World Health Organization (WHO) has had
+proposed advisory levels for drinking water and a range of recreational
+risk levels (Chorus and Bartram 1999, (WHO) and others 2003). While
+these levels and associated advisories are likely to help mitigate the
+impacts from harmful algal blooms, they are not without complications.
 
@@
 of microcystin-LR. While laboratory testing remains the gold standard
-for quantifying microcystin-LY concentrations in water samples, several
+for quantifying microcystin-LR concentrations in water samples, several
 field test kits have been developed. Even though field tests provide a
-much needed means for rapid assessment, they are infrequently available
-to lake managers. These kits are moderately expensive (approximately
-\$150-\$200 depending on specific kit) with a limited shelf life
-(typically one year) (James et al. 2011, Aranda-Rodriguez et al. 2015).
-Additionally, each technique requires nuanced understanding of the
-detection method (e.g., limit of detection, specific microcystin
-variants being measured, and sampling protocol). Fortunately,
-microcystin-LR has been shown to be associated with several other, more
-commonly measured and understood components of water quality.
+much needed means for rapid assessment, they are not yet widely used and
+are moderately expensive (approximately \$150-\$200 depending on
+specific kit) with a limited shelf life (typically one year) (James et
+al. 2011, Aranda-Rodriguez et al. 2015). Additionally, each technique
+requires nuanced understanding of the detection method (e.g., limit of
+detection, specific microcystin variants being measured, and sampling
+protocol). Fortunately, microcystin-LR has been shown to be associated
+with several other, more commonly measured and understood components of
+water quality.
 
-Chlorophyll \emph{a} is a very commonly measured components of water
+Chlorophyll \emph{a} is a very commonly measured component of water
 quality that is also known to be associated with Microsystin-LR
@@
 with the various Microcystin-LR health advisory levels. Identifying
-these associations would provide another reliable tool for water
-resource managers to use to help manage the threat to public health
-posed by cHABs and would be especially useful in the absence of
-microcystin-LR concentrations. Thus, the goal of this paper is to utlize
-the National Lakes Assessment data and identify chlorophyll \emph{a}
-concentrations that are associated with higher probabilities of
-exceeding several microcystin-LR health advisory concentrations (USEPA
-2009). So that others may repeat or adjust this analysis, the data,
-code, and this manuscript are freely available via
+these associations would provide another tool for water resource
+managers to use to help manage the threat to public health posed by
+cHABs and would be especially useful in the absence of microcystin-LR
+concentrations. Thus, the goal of this paper is to utilize the National
+Lakes Assessment data and identify chlorophyll \emph{a} concentrations
+that are associated with higher probabilities of exceeding several
+microcystin-LR health advisory concentrations (USEPA 2009). So that
+others may repeat or adjust this analysis, the data, code, and this
+manuscript are freely available via
 \href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USAPE/microcystinchla}}.
@@
 exceeding each health advisory level and extracted the minimum
-chlorophyll \emph{a} concentration that was associated with the upper
-confidence level being 50\% or greater. As both microcystin-LR and
-chlorophyll \emph{a} values were both highely skewed right, a log base
-10 transformation was used. Additional details of the specific
+chlorophyll \emph{a} concentration that was associated with an upper
+confidence level of 50\% or greater. As both microcystin-LR and
+chlorophyll \emph{a} values were highely skewed right, a log base 10
+transformation was used. Additional details of the specific
 implementation are available at
@@
 microcystin-LR. Of those lakes with microcystin, the median
-concentration was 0.51 and the mean was 0.51. Lastly, of all lakes
+concentration was 0.51 and the mean was 3.1661047. Lastly, of all lakes
 sampled, 21\% of lakes exceeded the U.S. EPA childrens drinking water
@@
 7.3\% of lakes exceeded the WHO recreational standard. For chlorophyll
-\emph{a}, the range was 0.07 to 936. All lakes had reported chlorophyll
-\emph{a} concentrations that exceeded detection limits. The median
-concentration was 7.79 and the mean was 29.6301946. The associations
+\emph{a}, the range was 0.07 to 936 \(\mu\)g/L. All lakes had reported
+chlorophyll \emph{a} concentrations that exceeded detection limits. The
+median concentration was 7.79 and the mean was 29.63. The associations
 between chlorophyll \emph{a} and the upper confidence interval with a
@@
 exceeding a target concentration take water samples and have those
-samples tested in a lab for microcystin-LR. A second potential use is to
-identify possible bloom events from historical data. As harmful algal
-blooms are made up of many species and have various mechanism
-responsible for adverse impacts (e.g.~toxins, hypoxia, odors), there is
-no single definition of a bloom. For cHABs one approach has been to
-identify an increase over a baseline concentration of phycocyanin
-(Miller et al. 2013). This is a useful approach for targeted studies,
-but phycocyanin is also not always available and measures the
-predominance of cyanobacterial pigments and not toxins. Using our
-chlorophyll \emph{a} cutoffs provides a value that is more directly
-associated with microcystin-LR and can be used to classify lakes, from
-past surveys, as having bloomed.
+samples tested in a lab for microcystin-LR.
+
+A second potential use is to identify possible bloom events from
+historical data. As harmful algal blooms are made up of many species and
+have various mechanisms responsible for adverse impacts (e.g.~toxins,
+hypoxia, odors), there is no single definition of a bloom. For cHABs one
+approach has been to identify an increase over a baseline concentration
+of phycocyanin (Miller et al. 2013). This is a useful approach for
+targeted studies, but phycocyanin is also not always available and
+measures the predominance of cyanobacterial pigments and not toxins.
+Using our chlorophyll \emph{a} cutoffs provides a value that is more
+directly associated with microcystin-LR and can be used to classify
+lakes, from past surveys, as having bloomed.
 

### 2015-10-05 manuscript/manuscript.Rmd

@@
 
-Over the last decade or so numerous events and legislative activities have raised the public awareness of harmful algal blooms [@jetoo2015toledo; @rinta2009lake; @HABHRCA2014].
-Recently, the US Environmental Protection Agency (USEPA) has released suggested microcystin-LR (one of the more common toxins) concentrations that would trigger health advisories [@mcelhiney2005detection; @zurawell2005hepatotoxic; @usepa2015drinking].
+Over the last decade or so numerous events and legislative activities have raised the public awareness of harmful algal blooms [@jetoo2015toledo; @rinta2009lake; @HABHRCA2014], and the US Environmental Protection Agency (USEPA) has recently released suggested microcystin-LR (one of the more common toxins) concentrations that would trigger health advisories [@mcelhiney2005detection; @zurawell2005hepatotoxic; @usepa2015drinking].
 Additionally, the World Health Organization (WHO) has had proposed advisory levels for drinking water and a range of recreational risk levels [@world2003cyanobacterial; @chorus1999toxic].
@@
 One of these complications is that they rely on available measurements of microcystin-LR.
-While laboratory testing remains the gold standard for quantifying microcystin-LY concentrations in water samples, several field test kits have been developed.
-Even though field tests provide a much needed means for rapid assessment, they are infrequently available to lake managers.
-These kits are moderately expensive (approximately $150-$200 depending on specific kit) with a limited shelf life (typically one year) [@james2011environmental; @aranda2015evaluation].
+While laboratory testing remains the gold standard for quantifying microcystin-LR concentrations in water samples, several field test kits have been developed.
+Even though field tests provide a much needed means for rapid assessment, they are not yet widely used and are moderately expensive (approximately $150-$200 depending on specific kit) with a limited shelf life (typically one year) [@james2011environmental; @aranda2015evaluation].
 Additionally, each technique requires nuanced understanding of the detection method (e.g., limit of detection, specific microcystin variants being measured, and sampling protocol).
@@
 
-Chlorophyll *a* is a very commonly measured components of water quality that is also known to be associated with Microsystin-LR concentrations [@lee2000variation; @paerl2013harmful; heisler2008eutrophication].
+Chlorophyll *a* is a very commonly measured component of water quality that is also known to be associated with Microsystin-LR concentrations [@lee2000variation; @paerl2013harmful; heisler2008eutrophication].
 Additionally there are many rapid measurements for assessing chlorophyll *a* levels *in situ*.
@@
 Given these facts, it might be possible to identify chlorophyll *a* concentrations that would be associated with the various Microcystin-LR health advisory levels.
-Identifying these associations would provide another reliable tool for water resource managers to use to help manage the threat to public health posed by cHABs and would be especially useful in the absence of microcystin-LR concentrations.
-Thus, the goal of this paper is to utlize the National Lakes Assessment data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin-LR health advisory concentrations [@usepa2009national].
+Identifying these associations would provide another tool for water resource managers to use to help manage the threat to public health posed by cHABs and would be especially useful in the absence of microcystin-LR concentrations.
+Thus, the goal of this paper is to utilize the National Lakes Assessment data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin-LR health advisory concentrations [@usepa2009national].
 So that others may repeat or adjust this analysis, the data, code, and this manuscript are freely available via [https://github.com/USAPE/microcystinchla](https://github.com/USAPE/microcystinchla).
@@
 The 95% confidence intervals were calculated from 1000 bootstrapped samples.
-To identify chlorophyll *a* concentrations of concern we used a 50% conditional probability of exceeding each health advisory level and extracted the minimum chlorophyll *a* concentration that was associated with the upper confidence level being 50% or greater.
-As both microcystin-LR and chlorophyll *a* values were both highely skewed right, a log base 10 transformation was used.
+To identify chlorophyll *a* concentrations of concern we used a 50% conditional probability of exceeding each health advisory level and extracted the minimum chlorophyll *a* concentration that was associated with an upper confidence level of 50% or greater.
+As both microcystin-LR and chlorophyll *a* values were highely skewed right, a log base 10 transformation was used.
 Additional details of the specific implementation are available at https://github.com/USEPA/microcystinchla.
@@
 Any value greater than that indicates the presence of microcystin-LR.
-Of those lakes with microcystin, the median concentration was `r median(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T)` and the mean was `r median(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T)`.
+Of those lakes with microcystin, the median concentration was `r median(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T)` and the mean was `r mean(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T)`.
 Lastly, of all lakes sampled, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=0.3])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the U.S.
@@
 EPA adult drinking water standard,`r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO drinking water standard,and `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=2])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO recreational standard.
-For chlorophyll *a*, the range was `r min(nla_dat$CHLA,na.rm=T)` to `r max(nla_dat$CHLA, na.rm=T)`.
+For chlorophyll *a*, the range was `r min(nla_dat$CHLA,na.rm=T)` to `r max(nla_dat$CHLA, na.rm=T)` $\mu$g/L.
 All lakes had reported chlorophyll *a* concentrations that exceeded detection limits.
-The median concentration was `r median(nla_dat$CHLA,na.rm=T)` and the mean was `r mean(nla_dat$CHLA,na.rm=T)`.
+The median concentration was `r median(nla_dat$CHLA,na.rm=T)` and the mean was `r round(mean(nla_dat$CHLA,na.rm=T),2)`.
 The associations between chlorophyll *a* and the upper confidence interval with a conditional probability of 50% were `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]` for 0.3, 1.0, 1.6 and 2.0 $\mu$g/L, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
@@
 Given that there is uncertainity around these chlorophyll *a* cutoffs the best case scenario would be to monitor for chlorophyll and in the event of exceeding a target concentration take water samples and have those samples tested in a lab for microcystin-LR.
+
 A second potential use is to identify possible bloom events from historical data.
-As harmful algal blooms are made up of many species and have various mechanism responsible for adverse impacts (e.g.
+As harmful algal blooms are made up of many species and have various mechanisms responsible for adverse impacts (e.g.
 toxins, hypoxia, odors), there is no single definition of a bloom.

### 2015-10-16 manuscript/manuscript.Rmd

@@
 In particular, one of the more common toxins, microcystin, has several advisory levels set for drinking water and recreational use.
-However, compared to other water quality measures, field measurements of microcystin are not commonly available due to cost and require advanced understanding to interpret results.
+However, compared to other water quality measures, field measurements of microcystin are not commonly available due to cost and advanced understanding required to interpret results.
 Addressing these issues will take time and resources.
@@
 In particular, chlorophyll *a* is very commonly measured, can be estimated *in situ*, and has been shown to be positively associated with microcystin.
-In this paper we use this association to provide estimates of chlorophyll *a* that if exceeded would be indicative of a higher probability of exceeding select health advisory concentrations for microcystin-LR.
-Using the 2007 National Lakes Assessment and a conditional probability approach that has been used in other water quality settings, we identify chlorophyll *a* concentrations that are more likely than not to be associated with an exceedance of a microcystin health advisory level.
+In this paper, we use this association to provide estimates of chlorophyll *a* concentrations that if exceeded would be indicative of a higher probability of exceeding select health advisory concentrations for microcystin-LR.
+Using the 2007 National Lakes Assessment and a conditional probability approach, we identify chlorophyll *a* concentrations that are more likely than not to be associated with an exceedance of a microcystin health advisory level.
 We look at the recent US EPA standards for drinking water as well as the World Health Organization levels for drinking water and recreational use.
-For microcystin concentrations of 0.3, 1, 1.6, and 2 we find chlorophyll *a* concentrations of `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]`, respectively.
-When managing for these various microcystin levels exceeding these reported chlorophyll *a* concentrations should be a trigger for further testing and possible management action.
+For the specific advisory microcystin concentrations of 0.3, 1, 1.6, and 2, we find chlorophyll *a* concentrations of `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]`, respectively.
+When managing for these various microcystin levels, exceeding these reported chlorophyll *a* concentrations should be a trigger for further testing and possible management action.
 
@@
 
-Over the last decade or so numerous events and legislative activities have raised the public awareness of harmful algal blooms [@jetoo2015toledo; @rinta2009lake; @HABHRCA2014], and the US Environmental Protection Agency (USEPA) has recently released suggested microcystin-LR (one of the more common toxins) concentrations that would trigger health advisories [@mcelhiney2005detection; @zurawell2005hepatotoxic; @usepa2015drinking].
+Over the last decade or so numerous events and legislative activities have raised the public awareness of harmful algal blooms [@jetoo2015toledo; @rinta2009lake; @HABHRCA2014], and in response the US Environmental Protection Agency (USEPA) has recently released suggested microcystin-LR (one of the more common toxins) concentrations that would trigger health advisories [@mcelhiney2005detection; @zurawell2005hepatotoxic; @usepa2015drinking].
 Additionally, the World Health Organization (WHO) has had proposed advisory levels for drinking water and a range of recreational risk levels [@world2003cyanobacterial; @chorus1999toxic].
@@
 Additionally, each technique requires nuanced understanding of the detection method (e.g., limit of detection, specific microcystin variants being measured, and sampling protocol).
-Fortunately, microcystin-LR has been shown to be associated with several other, more commonly measured and understood components of water quality.
+Fortunately, microcystin-LR has been shown to be associated with several other, more commonly measured and well understood components of water quality.
 
-Chlorophyll *a* is a very commonly measured component of water quality that is also known to be associated with Microsystin-LR concentrations [@lee2000variation; @paerl2013harmful; heisler2008eutrophication].
-Additionally there are many rapid measurements for assessing chlorophyll *a* levels *in situ*.
+Chlorophyll *a* is a very commonly measured component of water quality that is also known to be associated with Microsystin-LR concentrations [@lee2000variation; @paerl2013harmful; @heisler2008eutrophication].
+Additionally, there are many rapid measurements for assessing chlorophyll *a* levels *in situ*.
 For instance, there are small or hand held flourometers that measure chlorohpyll *a*.
 Given these facts, it might be possible to identify chlorophyll *a* concentrations that would be associated with the various Microcystin-LR health advisory levels.
-Identifying these associations would provide another tool for water resource managers to use to help manage the threat to public health posed by cHABs and would be especially useful in the absence of microcystin-LR concentrations.
-Thus, the goal of this paper is to utilize the National Lakes Assessment data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin-LR health advisory concentrations [@usepa2009national].
+Identifying these associations would provide another tool for water resource managers to help manage the threat to public health posed by cHABs and would be especially useful in the absence of microcystin-LR concentrations.
+Thus, the goal of this paper is to utilize the National Lakes Assessment (NLA) data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin-LR health advisory concentrations [@usepa2009national].
 So that others may repeat or adjust this analysis, the data, code, and this manuscript are freely available via [https://github.com/USAPE/microcystinchla](https://github.com/USAPE/microcystinchla).
@@
 ##Data
-We used the 2007 National Lakes Assessment (NLA) water quality and microcystin-LR concentration data [@usepa2009national].
+We used the 2007 NLA water quality and microcystin-LR concentration data [@usepa2009national].
 These data represent a snapshot of water quality from the summer of 2007 and data on chlorophyll *a* and microcystin-LR concentrations are available for `r nrow(na.omit(nla_dat$CHLA,nla_dat$MCYST_TL_UGL))` lakes.
@@
 ##Conditional Probability Analysis
-We used a conditional probability analysis (CPA) approach to explore associations between chlorophyll *a* concentrations and World Health Organization (WHO) and U.S.
-Environmental Protection Agency (U.S.
-EPA) microcystin-LR health advisory levels [@paul2011probability].
+We used a conditional probability analysis (CPA) approach to explore associations between chlorophyll *a* concentrations and World Health Organization (WHO) and USEPA microcystin-LR health advisory levels [@paul2011probability].
 Many levels have been suggested (Table \ref{tab:microcystin_levels}), but lakes with higher microcystin-LR concentrations in the NLA were rare.
@@
 
-A detailed discussion of CPA is beyond the scope of this paper, but see Paul et al.
+A detailed discussion of CPA is beyond the scope of this paper; see Paul et al.
 [-@paul2005development] and Hollister et al.
-[-@hollister2008cprob] for details.
+[-@hollister2008cprob] for greater detail.
 For this analysis, we used CPA to examine how the conditional probability of exceeding one of the health advisory changes as chlorophyll *a* increases in a lake.
@@
 Of those lakes with microcystin, the median concentration was `r median(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T)` and the mean was `r mean(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T)`.
-Lastly, of all lakes sampled, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=0.3])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the U.S.
+Of all lakes sampled, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=0.3])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the U.S.
 EPA childrens drinking water standard, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1.6])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the U.S.
@@
 The median concentration was `r median(nla_dat$CHLA,na.rm=T)` and the mean was `r round(mean(nla_dat$CHLA,na.rm=T),2)`.
-The associations between chlorophyll *a* and the upper confidence interval with a conditional probability of 50% were `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]` for 0.3, 1.0, 1.6 and 2.0 $\mu$g/L, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
+The associations between chlorophyll *a* and the upper confidence interval with a conditional probability of 50% were `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]` for 0.3, 1.0, 1.6 and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
 
 #Discussion
-The association between Log 10 microcystin-LR and Log 10 chlorophyll *a* show a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
+The association between Log10 microcystin-LR and Log10 chlorophyll *a* shows a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
 This indicates that higher concentrations of microcystin-LR almost always co-occur with higher concentrations of chlorophyll *a* yet the inverse is not true.
@@
 Total accuracy of the four cutoffs predicting microcystin-LR exceedances were `r total_accuracy$epa_child*100`% for the USEPA children's advisory, `r total_accuracy$who_drink*100`% for the WHO drinking water advisory,  `r total_accuracy$epa_adult*100`% for the USEPA adult advisory and, `r total_accuracy$who_rec*100`% for the WHO regreational advisory (Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table}, \ref{tab:adult_conmat_table}, & \ref{tab:who_rec_conmat_table}).
-But Total Accuracy is one part of the prediction performace with which we are concerned.
+However, total accuracy is only one part of the prediction performace with which we are concerned.
 
@@
 For the four chlorophyll *a* cut-offs we had a proportion of false negatives of `r fn_percent$epa_child*100`%, `r fn_percent$who_drink*100`%, `r fn_percent$epa_adult*100`% and , `r fn_percent$who_rec*100`% for the U.S EPA childrens drinking water, the WHO drinking water, the U.S.
-EPA Adult drinking water, and the WHO recreational advisories, respecitvely.
+EPA adult drinking water, and the WHO recreational advisories, respecitvely.
 In each case we miss less than 10% of the lakes that are in fact exceeding the microcystin-LR advisory.

### 2015-10-16 manuscript/manuscript.Rmd
Betty's edits

@@
 fontsize: 11pt
-linenumbers: no
+linenumbers: yes
 bibliography: components/manuscript.bib
@@
 mc_guides <- c(0.3, 1, 1.6, 2, 4, 10, 20)
-mc_names <- c("U.S. EPA","WHO","U.S. EPA","WHO","WHO","WHO","WHO")
-mc_type <- c("Drinking","Drinking","Drinking","Recreational","Recreational","Recreational","Recreational")
+mc_names <- c("USEPA","WHO","USEPA","WHO","WHO","WHO","WHO")
+mc_type <- c("Child Drinking Water Advisory",
+             "Drinking Water Advisory",
+             "Adult Drinking Water",
+             "Recreational: Low Prob. of Effect",
+             "Recreational: Moderate Prob. of Effect",
+             "Recreational: High Prob. of Effect",
+             "Recreational: Very High Prob. of Effect")
 mc_chla <- vector("numeric",length(mc_guides))
@@
 Any value greater than that indicates the presence of microcystin-LR.
-Of those lakes with microcystin, the median concentration was `r median(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T)` and the mean was `r mean(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T)`.
-Of all lakes sampled, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=0.3])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the U.S.
-EPA childrens drinking water standard, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1.6])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the U.S.
-EPA adult drinking water standard,`r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO drinking water standard,and `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=2])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO recreational standard.
+Of those lakes with microcystin, the median concentration was `r round(median(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T), 2)` and the mean was `r round(mean(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T), 2)`.
+Of all lakes sampled, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=0.3])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the USEPA childrens drinking water standard, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1.6])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the USEPA adult drinking water standard,`r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO drinking water standard,and `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=2])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO recreational standard.
 For chlorophyll *a*, the range was `r min(nla_dat$CHLA,na.rm=T)` to `r max(nla_dat$CHLA, na.rm=T)` $\mu$g/L.
@@
 These represent `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[1]])/length(nla_dat$CHLA)),1)`%, `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[2]])/length(nla_dat$CHLA)),1)`%,
-`r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[3]])/length(nla_dat$CHLA)),1)`%, and `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[4]])/length(nla_dat$CHLA)),1)`% of sample lakes for the U.S EPA childrens drinking water, the WHO drinking water, the U.S.
-EPA Adult drinking water, and the WHO recreational advisories, respecitvely.
+`r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[3]])/length(nla_dat$CHLA)),1)`%, and `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[4]])/length(nla_dat$CHLA)),1)`% of sample lakes for the U.S EPA childrens drinking water, the WHO drinking water, the USEPA Adult drinking water, and the WHO recreational advisories, respecitvely.
 
@@
 In other words, we would like to avoid Type II errors and minimize the proportion of false negatives.
-For the four chlorophyll *a* cut-offs we had a proportion of false negatives of `r fn_percent$epa_child*100`%, `r fn_percent$who_drink*100`%, `r fn_percent$epa_adult*100`% and , `r fn_percent$who_rec*100`% for the U.S EPA childrens drinking water, the WHO drinking water, the U.S.
-EPA adult drinking water, and the WHO recreational advisories, respecitvely.
+For the four chlorophyll *a* cut-offs we had a proportion of false negatives of `r fn_percent$epa_child*100`%, `r fn_percent$who_drink*100`%, `r fn_percent$epa_adult*100`% and , `r fn_percent$who_rec*100`% for the U.S EPA childrens drinking water, the WHO drinking water, the USEPA adult drinking water, and the WHO recreational advisories, respecitvely.
 In each case we miss less than 10% of the lakes that are in fact exceeding the microcystin-LR advisory.
-This is an acceptable rate, especially when no other tests are available for a given water body.
 
@@
 ```{r epa_child_cp_plot, echo=FALSE, fig.width=6, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
-A.) Plot for U.S.
-EPA childrens drinking water advisory (0.3 $\\mu$g/L).
+A.) Plot for USEPA childrens drinking water advisory (0.3 $\\mu$g/L).
 B.) Plot for WHO drinking water advisory (1 $\\mu$g/L).
-C.) Plot for U.S.
-EPA adult drinking water advisory (1.6 $\\mu$g/L).
+C.) Plot for USEPA adult drinking water advisory (1.6 $\\mu$g/L).
 D.) Plot for WHO recreational advisory (2 $\\mu$g/L).
@@
                     "WHO",
-                    "U.S. EPA",
-                    "U.S. EPA")
-type_of_stnd <- c("Recreational",
-                  "Recreational",
-                  "Recreational",
-                  "Recreational",
-                  "Drinking",
-                  "Drinking",
-                  "Drinking")
+                    "USEPA",
+                    "USEPA")
+type_of_stnd <- c("Recreational: Low Prob. of Effect",
+                  "Recreational: Moderate Prob. of Effect",
+                  "Recreational: High Prob. of Effect",
+                  "Recreational: Very High Prob. of Effect",
+                  "Drinking Water",
+                  "Child Drinking Water Advisory",
+                  "Adult Drinking Water Advisory")
 concentration <- c("2-4 $\\mu$g/L",
@@
 
-```{r child_conmat_table, results='asis', echo=FALSE, tab.cap="Confusion matrix comparing chlorophyll \\textit{a} predicted exceedences (rows) versus real exceedances (columns) for the U.S.
-EPA childrens drinking water advisory.
+```{r child_conmat_table, results='asis', echo=FALSE, tab.cap="Confusion matrix comparing chlorophyll \\textit{a} predicted exceedences (rows) versus real exceedances (columns) for the USEPA childrens drinking water advisory.
 \\label{tab:child_conmat_table}"}
@@
 
-```{r adult_conmat_table, results='asis', echo=FALSE, tab.cap="Confusion matrix comparing chlorophyll \\textit{a} predicted exceedences (rows) versus real exceedances (columns) for the U.S.
-EPA adult drinking water advisory.
+```{r adult_conmat_table, results='asis', echo=FALSE, tab.cap="Confusion matrix comparing chlorophyll \\textit{a} predicted exceedences (rows) versus real exceedances (columns) for the USEPA adult drinking water advisory.
 \\label{tab:adult_conmat_table}"}

### 2015-10-16 manuscript/manuscript.tex
Betty's edits

@@
 \usepackage{lineno} % add 
+  \linenumbers % turns line numbering on 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 quality measures, field measurements of microcystin are not commonly
-available due to cost and require advanced understanding to interpret
+available due to cost and advanced understanding required to interpret
 results. Addressing these issues will take time and resources. Thus,
@@
 \emph{in situ}, and has been shown to be positively associated with
-microcystin. In this paper we use this association to provide estimates
-of chlorophyll \emph{a} that if exceeded would be indicative of a higher
-probability of exceeding select health advisory concentrations for
-microcystin-LR. Using the 2007 National Lakes Assessment and a
-conditional probability approach that has been used in other water
-quality settings, we identify chlorophyll \emph{a} concentrations that
-are more likely than not to be associated with an exceedance of a
-microcystin health advisory level. We look at the recent US EPA
-standards for drinking water as well as the World Health Organization
-levels for drinking water and recreational use. For microcystin
-concentrations of 0.3, 1, 1.6, and 2 we find chlorophyll \emph{a}
-concentrations of 22.25, 65.6, 84.96, and 105.84, respectively. When
-managing for these various microcystin levels exceeding these reported
-chlorophyll \emph{a} concentrations should be a trigger for further
-testing and possible management action.
+microcystin. In this paper, we use this association to provide estimates
+of chlorophyll \emph{a} concentrations that if exceeded would be
+indicative of a higher probability of exceeding select health advisory
+concentrations for microcystin-LR. Using the 2007 National Lakes
+Assessment and a conditional probability approach, we identify
+chlorophyll \emph{a} concentrations that are more likely than not to be
+associated with an exceedance of a microcystin health advisory level. We
+look at the recent US EPA standards for drinking water as well as the
+World Health Organization levels for drinking water and recreational
+use. For the specific advisory microcystin concentrations of 0.3, 1,
+1.6, and 2, we find chlorophyll \emph{a} concentrations of 23.68, 63.94,
+84.96, and 97.49, respectively. When managing for these various
+microcystin levels, exceeding these reported chlorophyll \emph{a}
+concentrations should be a trigger for further testing and possible
+management action.
 
@@
 have raised the public awareness of harmful algal blooms (Rinta-Kanto et
-al. 2009, HABHRCA 2014, Jetoo et al. 2015), and the US Environmental
-Protection Agency (USEPA) has recently released suggested microcystin-LR
-(one of the more common toxins) concentrations that would trigger health
-advisories (McElhiney and Lawton 2005, Zurawell et al. 2005, USEPA
-2015). Additionally, the World Health Organization (WHO) has had
-proposed advisory levels for drinking water and a range of recreational
-risk levels (Chorus and Bartram 1999, (WHO) and others 2003). While
-these levels and associated advisories are likely to help mitigate the
-impacts from harmful algal blooms, they are not without complications.
+al. 2009, HABHRCA 2014, Jetoo et al. 2015), and in response the US
+Environmental Protection Agency (USEPA) has recently released suggested
+microcystin-LR (one of the more common toxins) concentrations that would
+trigger health advisories (McElhiney and Lawton 2005, Zurawell et al.
+2005, USEPA 2015). Additionally, the World Health Organization (WHO) has
+had proposed advisory levels for drinking water and a range of
+recreational risk levels (Chorus and Bartram 1999, (WHO) and others
+2003). While these levels and associated advisories are likely to help
+mitigate the impacts from harmful algal blooms, they are not without
+complications.
 
@@
 protocol). Fortunately, microcystin-LR has been shown to be associated
-with several other, more commonly measured and understood components of
-water quality.
+with several other, more commonly measured and well understood
+components of water quality.
 
@@
 quality that is also known to be associated with Microsystin-LR
-concentrations {[}Lee et al. (2000); Paerl and Otten (2013);
-heisler2008eutrophication{]}. Additionally there are many rapid
-measurements for assessing chlorophyll \emph{a} levels \emph{in situ}.
-For instance, there are small or hand held flourometers that measure
-chlorohpyll \emph{a}. Given these facts, it might be possible to
-identify chlorophyll \emph{a} concentrations that would be associated
-with the various Microcystin-LR health advisory levels. Identifying
-these associations would provide another tool for water resource
-managers to use to help manage the threat to public health posed by
-cHABs and would be especially useful in the absence of microcystin-LR
-concentrations. Thus, the goal of this paper is to utilize the National
-Lakes Assessment data and identify chlorophyll \emph{a} concentrations
-that are associated with higher probabilities of exceeding several
-microcystin-LR health advisory concentrations (USEPA 2009). So that
-others may repeat or adjust this analysis, the data, code, and this
-manuscript are freely available via
+concentrations (Lee et al. 2000, Heisler et al. 2008, Paerl and Otten
+2013). Additionally, there are many rapid measurements for assessing
+chlorophyll \emph{a} levels \emph{in situ}. For instance, there are
+small or hand held flourometers that measure chlorohpyll \emph{a}. Given
+these facts, it might be possible to identify chlorophyll \emph{a}
+concentrations that would be associated with the various Microcystin-LR
+health advisory levels. Identifying these associations would provide
+another tool for water resource managers to help manage the threat to
+public health posed by cHABs and would be especially useful in the
+absence of microcystin-LR concentrations. Thus, the goal of this paper
+is to utilize the National Lakes Assessment (NLA) data and identify
+chlorophyll \emph{a} concentrations that are associated with higher
+probabilities of exceeding several microcystin-LR health advisory
+concentrations (USEPA 2009). So that others may repeat or adjust this
+analysis, the data, code, and this manuscript are freely available via
 \href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USAPE/microcystinchla}}.
@@
 
-We used the 2007 National Lakes Assessment (NLA) water quality and
-microcystin-LR concentration data (USEPA 2009). These data represent a
-snapshot of water quality from the summer of 2007 and data on
-chlorophyll \emph{a} and microcystin-LR concentrations are available for
-lakes.
+We used the 2007 NLA water quality and microcystin-LR concentration data
+(USEPA 2009). These data represent a snapshot of water quality from the
+summer of 2007 and data on chlorophyll \emph{a} and microcystin-LR
+concentrations are available for lakes.
 
@@
 associations between chlorophyll \emph{a} concentrations and World
-Health Organization (WHO) and U.S. Environmental Protection Agency (U.S.
-EPA) microcystin-LR health advisory levels (Paul and Munns 2011). Many
-levels have been suggested (Table \ref{tab:microcystin_levels}), but
-lakes with higher microcystin-LR concentrations in the NLA were rare.
-Only 1.16 \% of lakes sampled had a concentration greater than 10. Thus,
-for this analysis we focus on the microcystin concentrations that are
-better represented in the NLA data. These were 0.3, 1, 1.6, and 2
-\(\mu\)g/L.
+Health Organization (WHO) and USEPA microcystin-LR health advisory
+levels (Paul and Munns 2011). Many levels have been suggested (Table
+\ref{tab:microcystin_levels}), but lakes with higher microcystin-LR
+concentrations in the NLA were rare. Only 1.16 \% of lakes sampled had a
+concentration greater than 10. Thus, for this analysis we focus on the
+microcystin concentrations that are better represented in the NLA data.
+These were 0.3, 1, 1.6, and 2 \(\mu\)g/L.
 
-A detailed discussion of CPA is beyond the scope of this paper, but see
-Paul et al. (2005) and Hollister et al. (2008) for details. For this
+A detailed discussion of CPA is beyond the scope of this paper; see Paul
+et al. (2005) and Hollister et al. (2008) for greater detail. For this
 analysis, we used CPA to examine how the conditional probability of
@@
 microcystin-LR. Of those lakes with microcystin, the median
-concentration was 0.51 and the mean was 3.1661047. Lastly, of all lakes
-sampled, 21\% of lakes exceeded the U.S. EPA childrens drinking water
-standard, 8.8\% of lakes exceeded the U.S. EPA adult drinking water
-standard,11.7\% of lakes exceeded the WHO drinking water standard,and
-7.3\% of lakes exceeded the WHO recreational standard. For chlorophyll
-\emph{a}, the range was 0.07 to 936 \(\mu\)g/L. All lakes had reported
-chlorophyll \emph{a} concentrations that exceeded detection limits. The
-median concentration was 7.79 and the mean was 29.63. The associations
-between chlorophyll \emph{a} and the upper confidence interval with a
-conditional probability of 50\% were 22.25, 65.6, 84.96, and 105.84 for
-0.3, 1.0, 1.6 and 2.0 \(\mu\)g/L, respectively (Table
+concentration was 0.51 and the mean was 3.17. Of all lakes sampled, 21\%
+of lakes exceeded the USEPA childrens drinking water standard, 8.8\% of
+lakes exceeded the USEPA adult drinking water standard,11.7\% of lakes
+exceeded the WHO drinking water standard,and 7.3\% of lakes exceeded the
+WHO recreational standard. For chlorophyll \emph{a}, the range was 0.07
+to 936 \(\mu\)g/L. All lakes had reported chlorophyll \emph{a}
+concentrations that exceeded detection limits. The median concentration
+was 7.79 and the mean was 29.63. The associations between chlorophyll
+\emph{a} and the upper confidence interval with a conditional
+probability of 50\% were 23.68, 63.94, 84.96, and 97.49 for 0.3, 1.0,
+1.6 and 2.0 \(\mu\)g/L advisory levels, respectively (Table
 \ref{tab:mc_chla_table} \& Figure \ref{fig:multi_cp_plot}).
@@
 
-The association between Log 10 microcystin-LR and Log 10 chlorophyll
-\emph{a} show a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
+The association between Log10 microcystin-LR and Log10 chlorophyll
+\emph{a} shows a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
 This indicates that higher concentrations of microcystin-LR almost
@@
 greater than even odds of exceeding a given health advisory level (Table
-\ref{tab:mc_chla_table}). These represent 29.1\%, 11.5\%, 8.6\%, and
-6.9\% of sample lakes for the U.S EPA childrens drinking water, the WHO
-drinking water, the U.S. EPA Adult drinking water, and the WHO
-recreational advisories, respecitvely.
+\ref{tab:mc_chla_table}). These represent 27.8\%, 11.9\%, 8.6\%, and
+7.9\% of sample lakes for the U.S EPA childrens drinking water, the WHO
+drinking water, the USEPA Adult drinking water, and the WHO recreational
+advisories, respecitvely.
 
@@
 evaluate false negatives. Total accuracy of the four cutoffs predicting
-microcystin-LR exceedances were 74\% for the USEPA children's advisory,
+microcystin-LR exceedances were 75\% for the USEPA children's advisory,
 86\% for the WHO drinking water advisory, 90\% for the USEPA adult
-advisory and, 91\% for the WHO regreational advisory (Tables
+advisory and, 90\% for the WHO regreational advisory (Tables
 \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table},
-\ref{tab:adult_conmat_table}, \& \ref{tab:who_rec_conmat_table}). But
-Total Accuracy is one part of the prediction performace with which we
-are concerned.
+\ref{tab:adult_conmat_table}, \& \ref{tab:who_rec_conmat_table}).
+However, total accuracy is only one part of the prediction performace
+with which we are concerned.
 
@@
 proportion of false negatives of 9\%, 7\%, 6\% and , 5\% for the U.S EPA
-childrens drinking water, the WHO drinking water, the U.S. EPA Adult
+childrens drinking water, the WHO drinking water, the USEPA adult
 drinking water, and the WHO recreational advisories, respecitvely. In
 each case we miss less than 10\% of the lakes that are in fact exceeding
-the microcystin-LR advisory. This is an acceptable rate, especially when
-no other tests are available for a given water body.
+the microcystin-LR advisory.
 
@@
 probability of exceeding various microcystin-LR (MLR) health advisory
-Levels. A.) Plot for U.S. EPA childrens drinking water advisory (0.3
+Levels. A.) Plot for USEPA childrens drinking water advisory (0.3
 \(\mu\)g/L). B.) Plot for WHO drinking water advisory (1 \(\mu\)g/L).
-C.) Plot for U.S. EPA adult drinking water advisory (1.6 \(\mu\)g/L).
-D.) Plot for WHO recreational advisory (2 \(\mu\)g/L).
+C.) Plot for USEPA adult drinking water advisory (1.6 \(\mu\)g/L). D.)
+Plot for WHO recreational advisory (2 \(\mu\)g/L).
 \label{fig:multi_cp_plot}}
@@
 \endhead
-WHO & Drinking & 1 \(\mu\)g/L\tabularnewline
-U.S. EPA & Drinking & 0.3 \(\mu\)g/L\tabularnewline
-U.S. EPA & Drinking & 1.6 \(\mu\)g/L\tabularnewline
-WHO & Recreational & 2-4 \(\mu\)g/L\tabularnewline
-WHO & Recreational & 10-20 \(\mu\)g/L\tabularnewline
-WHO & Recreational & 20-2000 \(\mu\)g/L\tabularnewline
-WHO & Recreational & \textgreater{}2000 \(\mu\)g/L\tabularnewline
+USEPA & Adult Drinking Water Advisory & 1.6 \(\mu\)g/L\tabularnewline
+USEPA & Child Drinking Water Advisory & 0.3 \(\mu\)g/L\tabularnewline
+WHO & Drinking Water & 1 \(\mu\)g/L\tabularnewline
+WHO & Recreational: High Prob. of Effect & 20-2000
+\(\mu\)g/L\tabularnewline
+WHO & Recreational: Low Prob. of Effect & 2-4 \(\mu\)g/L\tabularnewline
+WHO & Recreational: Moderate Prob. of Effect & 10-20
+\(\mu\)g/L\tabularnewline
+WHO & Recreational: Very High Prob. of Effect & \textgreater{}2000
+\(\mu\)g/L\tabularnewline
 \bottomrule
@@
 \endhead
-U.S. EPA & Drinking & 0.3 & 22.25\tabularnewline
-WHO & Drinking & 1.0 & 65.60\tabularnewline
-U.S. EPA & Drinking & 1.6 & 84.96\tabularnewline
-WHO & Recreational & 2.0 & 105.84\tabularnewline
+USEPA & Child Drinking Water Advisory & 0.3 & 23.68\tabularnewline
+WHO & Drinking Water Advisory & 1.0 & 63.94\tabularnewline
+USEPA & Adult Drinking Water & 1.6 & 84.96\tabularnewline
+WHO & Recreational: Low Prob. of Effect & 2.0 & 97.49\tabularnewline
 \bottomrule
@@
 \caption{Confusion matrix comparing chlorophyll \textit{a} predicted
-exceedences (rows) versus real exceedances (columns) for the U.S. EPA
+exceedences (rows) versus real exceedances (columns) for the USEPA
 childrens drinking water advisory.
@@
 \endhead
-FALSE & 638 & 94\tabularnewline
-TRUE & 173 & 123\tabularnewline
+FALSE & 650 & 96\tabularnewline
+TRUE & 161 & 121\tabularnewline
 \bottomrule
@@
 \endhead
-FALSE & 838 & 76\tabularnewline
-TRUE & 69 & 45\tabularnewline
+FALSE & 835 & 75\tabularnewline
+TRUE & 72 & 46\tabularnewline
 \bottomrule
@@
 \caption{Confusion matrix comparing chlorophyll \textit{a} predicted
-exceedences (rows) versus real exceedances (columns) for the U.S. EPA
-adult drinking water advisory.
-\label{tab:adult_conmat_table}}\tabularnewline
+exceedences (rows) versus real exceedances (columns) for the USEPA adult
+drinking water advisory. \label{tab:adult_conmat_table}}\tabularnewline
 \toprule
@@
 \endhead
-FALSE & 910 & 52\tabularnewline
-TRUE & 43 & 23\tabularnewline
+FALSE & 903 & 48\tabularnewline
+TRUE & 50 & 27\tabularnewline
 \bottomrule
@@
 amendments act of 2014.
+
+Heisler, J., P. M. Glibert, J. M. Burkholder, D. M. Anderson, W.
+Cochlan, W. C. Dennison, Q. Dortch, C. J. Gobler, C. A. Heil, E.
+Humphries, and others. 2008. Eutrophication and harmful algal blooms: A
+scientific consensus. Harmful algae 8:3--13.
 

### 2015-10-16 R/plot_scatter.R

@@ -13,7 +13,7 @@ plot_scatter<-function(df,xvar,yvar,cat=NULL,pt_col=1,...){
   }
   xdf<-xdf[complete.cases(xdf),]
   x <- ggplot(xdf,aes(xvar,yvar))+
-    geom_point(size=3,aes(colour=pt_col)) +
+    geom_point(size=3) +
     theme_bw()+
     theme(text = element_text(family="sans"),
           axis.title.x = element_text(family="sans",vjust = -0.5, size = 14),

### 2015-10-20 manuscript/manuscript.Rmd
added acknow.  no render

@@
 if(!require(viridis)){
-  install.pacakges("viridis") 
+  install.packages("viridis") 
 }
@@
 
+#Acknowledgements
+
 \newpage

### 2015-10-20 manuscript/manuscript.Rmd
more updates

@@
 if(!require(microcystinchla)){
-  devtools::install_github("jhollist/microcystinchla",
+  devtools::install_github("usepa/microcystinchla",
                            auth_token = getOption("Github_Access"))

### 2015-12-01 manuscript/manuscript.Rmd
more additions while working on #4

@@
 ################################################################################
-data_url<-"http://water.epa.gov/type/lakes/assessmonitor/lakessurvey/upload/NLA2007_Recreational_ConditionEstimates_20091123.csv"
+data_url<-"http://www2.epa.gov/sites/production/files/2014-10/nla2007_recreational_conditionestimates_20091123.csv"
 nla_dat <- get_nla(data_url)
@@
 Additionally, each technique requires nuanced understanding of the detection method (e.g., limit of detection, specific microcystin variants being measured, and sampling protocol).
-Fortunately, microcystin-LR has been shown to be associated with several other, more commonly measured and well understood components of water quality.
 
-Chlorophyll *a* is a very commonly measured component of water quality that is also known to be associated with Microsystin-LR concentrations [@lee2000variation; @paerl2013harmful; @heisler2008eutrophication].
-Additionally, there are many rapid measurements for assessing chlorophyll *a* levels *in situ*.
-For instance, there are small or hand held flourometers that measure chlorohpyll *a*.
-Given these facts, it might be possible to identify chlorophyll *a* concentrations that would be associated with the various Microcystin-LR health advisory levels.
+Fortunately, microcystin-LR has been shown to be associated with several other, more commonly measured and well understood components of water quality that are readily assessed in the field.
+For instance, there are small or hand held fluorometers that measure chlorohpyll *a*.
+Additionally, chlorophyll *a* is a very commonly measured component of water quality that is also known to be positively associated with Microsystin-LR concentrations [@pip2014microcystin; @yuan2014managing].
+Yuan et.
+al [-@yuan2014managing] explore these associations in detail and control for other related variables.
+In their analysis they find that total nitrogen and chlorophyll *a* show the strongest association with microcystin.
+Furthermore, they identify chlorophyll*a* and total nitrogen concentrations that are associated with exceeding 1 $\mu$g/L of microcystin.
+Given these facts, it should be possible to identify chlorophyll *a* concentrations that would be associated with the new USEPA Microcystin-LR health advisory levels for drinking water.
 Identifying these associations would provide another tool for water resource managers to help manage the threat to public health posed by cHABs and would be especially useful in the absence of microcystin-LR concentrations.
+
+In this paper we ...
 Thus, the goal of this paper is to utilize the National Lakes Assessment (NLA) data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin-LR health advisory concentrations [@usepa2009national].

### 2015-12-02 manuscript/manuscript.Rmd
working on #4

@@
 
-In this paper we ...
-Thus, the goal of this paper is to utilize the National Lakes Assessment (NLA) data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin-LR health advisory concentrations [@usepa2009national].
+In this paper we build on past efforts and utilize the National Lakes Assessment (NLA) data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin-LR health advisory concentrations [@usepa2009national; @usepa2015drinking; @chorus1999toxic].
+We add to past studies by exploring associations with newly announced advisory levels and by also applying a different method, conditional probability analysis.
+Utilizing different methods strengthens the evidence for suggested chlorophyll *a* levels that are associated with increased risk of exceeding the health advisory levels as those levels are not predicated on a single analytical method.
 So that others may repeat or adjust this analysis, the data, code, and this manuscript are freely available via [https://github.com/USAPE/microcystinchla](https://github.com/USAPE/microcystinchla).
@@
 We used the 2007 NLA water quality and microcystin-LR concentration data [@usepa2009national].
-These data represent a snapshot of water quality from the summer of 2007 and data on chlorophyll *a* and microcystin-LR concentrations are available for `r nrow(na.omit(nla_dat$CHLA,nla_dat$MCYST_TL_UGL))` lakes.
+These data represent a snapshot of water quality from the summer of 2007 for the conterminous United States.
+Data on chlorophyll *a* and microcystin-LR concentrations are available for `r nrow(na.omit(nla_dat$CHLA,nla_dat$MCYST_TL_UGL))` lakes.
 
@@
 
-A detailed discussion of CPA is beyond the scope of this paper; see Paul et al.
+Conditional probability analysis provides information about the probability of observing one event given another event has also occured.
+For this analysis, we used CPA to examine how the conditional probability of exceeding one of the health advisory changes as chlorophyll *a* increases in a lake.
+We expect to find higher chlorohpyll *a* concentrations to be associated with higher probabilities of exceeding the microcystin-LR health advisory levels.
+We also caclulated bootstrapped 95% confidence intervals (CI) using 1000 bootstrapped samples.
+Thus, to identify chlorophyll *a* concentrations of concern we identfiy the value of the upper 95% CI across a range of conditional probabilities of exceeding each health advisory level.
+As both microcystin-LR and chlorophyll *a* values were highely skewed right, a log base 10 transformation was used.
+Additional details of the specific implementation are available at https://github.com/USEPA/microcystinchla.
+A more detailed discussion of CPA is beyond the scope of this paper, but see Paul et al.
 [-@paul2005development] and Hollister et al.
 [-@hollister2008cprob] for greater detail.
-For this analysis, we used CPA to examine how the conditional probability of exceeding one of the health advisory changes as chlorophyll *a* increases in a lake.
-The 95% confidence intervals were calculated from 1000 bootstrapped samples.
-To identify chlorophyll *a* concentrations of concern we used a 50% conditional probability of exceeding each health advisory level and extracted the minimum chlorophyll *a* concentration that was associated with an upper confidence level of 50% or greater.
-As both microcystin-LR and chlorophyll *a* values were highely skewed right, a log base 10 transformation was used.
-Additional details of the specific implementation are available at https://github.com/USEPA/microcystinchla.
 

### 2015-12-02 manuscript/manuscript.Rmd
closes #4 - Anne's comments.  Also added an addition table which needs to be created.

@@
 Thus, for this analysis we focus on the microcystin concentrations that are better represented in the NLA data.
-These were  0.3, 1, 1.6, and 2 $\mu$g/L.
+These were the USEPA children's drinking water advisory level of  0.3 $\mu$g/L (USEPA Child), the WHO drinking water advisory level of 1 $\mu$g/L (WHO Drinking), the USEPA adult drinking water advisory level of 1.6 $\mu$g/L (USEPA Adult), and the WHO recreational, low probability of effect advisory level of 2 $\mu$g/L (WHO Recreational).
 
@@
 
-In the 2007 NLA, microcystin-LR concentrations ranged from `r min(nla_dat$MCYST_TL_UGL,na.rm=T)` to `r max(nla_dat$MCYST_TL_UGL, na.rm=T)`.
+In the 2007 NLA, microcystin-LR concentrations ranged from `r min(nla_dat$MCYST_TL_UGL,na.rm=T)` to `r max(nla_dat$MCYST_TL_UGL, na.rm=T)` $\mu$g/L.
 Microcystin-LR concentrations of 0.05 $\mu$g/L represent the detection limits.
@@
 Of those lakes with microcystin, the median concentration was `r round(median(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T), 2)` and the mean was `r round(mean(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T), 2)`.
-Of all lakes sampled, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=0.3])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the USEPA childrens drinking water standard, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1.6])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the USEPA adult drinking water standard,`r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO drinking water standard,and `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=2])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO recreational standard.
+Of all lakes sampled, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=0.3])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the USEPA Child level, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1.6])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the USEPA Adult level,`r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO Drinking level,and `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=2])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO Recreational level.
 For chlorophyll *a*, the range was `r min(nla_dat$CHLA,na.rm=T)` to `r max(nla_dat$CHLA, na.rm=T)` $\mu$g/L.
 All lakes had reported chlorophyll *a* concentrations that exceeded detection limits.
-The median concentration was `r median(nla_dat$CHLA,na.rm=T)` and the mean was `r round(mean(nla_dat$CHLA,na.rm=T),2)`.
-The associations between chlorophyll *a* and the upper confidence interval with a conditional probability of 50% were `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]` for 0.3, 1.0, 1.6 and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
+The median concentration was `r median(nla_dat$CHLA,na.rm=T)` $\mu$g/L and the mean was `r round(mean(nla_dat$CHLA,na.rm=T),2)` $\mu$g/L.
+The associations between chlorophyll *a* and the upper confidence interval across a range of conditional probability values is shown in Table \ref{tab:mc_chla_table}.
+Specific chlorophyll *a* that are associated with greater than even odds of exceeding the advisory levels were `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]` for 0.3, 1.0, 1.6 and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
 
@@
 These represent `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[1]])/length(nla_dat$CHLA)),1)`%, `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[2]])/length(nla_dat$CHLA)),1)`%,
-`r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[3]])/length(nla_dat$CHLA)),1)`%, and `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[4]])/length(nla_dat$CHLA)),1)`% of sample lakes for the U.S EPA childrens drinking water, the WHO drinking water, the USEPA Adult drinking water, and the WHO recreational advisories, respecitvely.
+`r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[3]])/length(nla_dat$CHLA)),1)`%, and `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[4]])/length(nla_dat$CHLA)),1)`% of sample lakes for the USEPA Child, the WHO Drinking, the USEPA Adult, and the WHO recreational levels, respecitvely.
 
@@
 First, in the absence of microcystin-LR measurements, exceedence of the chlorophyll *a* concentrations could be a trigger for further actions.
-Given that there is uncertainity around these chlorophyll *a* cutoffs the best case scenario would be to monitor for chlorophyll and in the event of exceeding a target concentration take water samples and have those samples tested in a lab for microcystin-LR.
+Given that there is uncertainity around these chlorophyll *a* cutoffs the best case scenario would be to monitor for chlorophyll *a* and in the event of exceeding a target concentration take water samples and have those samples tested in a lab for microcystin-LR.
 
@@
 ```{r epa_child_cp_plot, echo=FALSE, fig.width=6, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
-A.) Plot for USEPA childrens drinking water advisory (0.3 $\\mu$g/L).
-B.) Plot for WHO drinking water advisory (1 $\\mu$g/L).
-C.) Plot for USEPA adult drinking water advisory (1.6 $\\mu$g/L).
-D.) Plot for WHO recreational advisory (2 $\\mu$g/L).
+A.) Plot for USEPA Child (0.3 $\\mu$g/L).
+B.) Plot for WHO Drinking (1 $\\mu$g/L).
+C.) Plot for USEPA Adult (1.6 $\\mu$g/L).
+D.) Plot for WHO Recreational (2 $\\mu$g/L).
 \\label{fig:multi_cp_plot}"}
@@
 knitr::kable(mc_chla_tbl)
+####AS OF 12/2/15 - REDO THIS WITH MULTIPLE COND PROBS!
 ```

### 2015-12-02 R/get_chla.R
starting to work on #7

@@ -7,6 +7,6 @@
 #' @param prob Exceedence probability to return chl a 
 #' @export
 get_chla <- function(cp_obj,prob){
-  transf_chla <- min(cp_obj$X[cp_obj$Upper.CI >= prob])
+  transf_chla <- c(min(cp_obj$X[cp_obj$Upper.CI >= prob]),min(cp_obj$X[cp_obj$Lower.CI >= prob]))
   return(10^transf_chla)
 }

### 2015-12-03 manuscript/manuscript.Rmd
starting to work on #7
closes #7 : creates table of chl a values across range of probability of exceeding and different advisory levels.

@@
              "Recreational: Very High Prob. of Effect")
-mc_chla <- vector("numeric",length(mc_guides))
+
 #wgt <- nla_dat$WGT_NLA
@@
 
-mc_chla[1] <- get_chla(epa_child_cp,0.5)
+#mc_chla[1] <- get_chla(epa_child_cp,0.5)
 
@@
 
-mc_chla[2] <- get_chla(who_drink_cp,0.5)
+#mc_chla[2] <- get_chla(who_drink_cp,0.5)
 
@@
 
-mc_chla[3] <- get_chla(epa_adult_cp,0.5)
+#mc_chla[3] <- get_chla(epa_adult_cp,0.5)
 
@@
 
-mc_chla[4] <- get_chla(who_rec_low1_cp,0.5)
+#mc_chla[4] <- get_chla(who_rec_low1_cp,0.5)
+epa_child <- NULL
+who_drink <- NULL
+epa_adult <- NULL
+who_rec <- NULL
+cps<- seq(0.1,0.9,0.1)
+for(i in 1:9){
+  epa_child[i] <- get_chla(epa_child_cp,cps[i])[1]
+  who_drink[i] <- get_chla(who_drink_cp,cps[i])[1]
+  epa_adult[i] <- get_chla(epa_adult_cp,cps[i])[1]
+  who_rec[i] <- get_chla(who_rec_low1_cp,cps[i])[1]
+}
+mc_chla<-cbind(epa_child,who_drink,epa_adult,who_rec)
 
@@
 \\label{tab:mc_chla_table}"}
-mc_chla_tbl <- data.frame(Source=mc_names,Type=mc_type,Microcystin=mc_guides,Chlorophyll=mc_chla)[1:4,]
+mc_chla_tbl <- data.frame(seq(0.1,0.9,0.1),mc_chla)
+names(mc_chla_tbl)<-c("Cond. Probability","USEPA Child", "WHO Drink", 
+                      "USEPA Adult", "WHO Recreational")
 #pandoc.table(mc_chla_tbl,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
 knitr::kable(mc_chla_tbl)
-####AS OF 12/2/15 - REDO THIS WITH MULTIPLE COND PROBS!
 ```

### 2015-12-03 manuscript/manuscript.tex
closes #7 : creates table of chl a values across range of probability of exceeding and different advisory levels.

@@
 use. For the specific advisory microcystin concentrations of 0.3, 1,
-1.6, and 2, we find chlorophyll \emph{a} concentrations of 23.68, 63.94,
-84.96, and 97.49, respectively. When managing for these various
+1.6, and 2, we find chlorophyll \emph{a} concentrations of 0.07, 0.07,
+2.79, and 11.36, respectively. When managing for these various
 microcystin levels, exceeding these reported chlorophyll \emph{a}
@@
 detection, specific microcystin variants being measured, and sampling
-protocol). Fortunately, microcystin-LR has been shown to be associated
-with several other, more commonly measured and well understood
-components of water quality.
+protocol).
 
-Chlorophyll \emph{a} is a very commonly measured component of water
-quality that is also known to be associated with Microsystin-LR
-concentrations (Lee et al. 2000, Heisler et al. 2008, Paerl and Otten
-2013). Additionally, there are many rapid measurements for assessing
-chlorophyll \emph{a} levels \emph{in situ}. For instance, there are
-small or hand held flourometers that measure chlorohpyll \emph{a}. Given
-these facts, it might be possible to identify chlorophyll \emph{a}
-concentrations that would be associated with the various Microcystin-LR
-health advisory levels. Identifying these associations would provide
-another tool for water resource managers to help manage the threat to
-public health posed by cHABs and would be especially useful in the
-absence of microcystin-LR concentrations. Thus, the goal of this paper
-is to utilize the National Lakes Assessment (NLA) data and identify
-chlorophyll \emph{a} concentrations that are associated with higher
-probabilities of exceeding several microcystin-LR health advisory
-concentrations (USEPA 2009). So that others may repeat or adjust this
-analysis, the data, code, and this manuscript are freely available via
+Fortunately, microcystin-LR has been shown to be associated with several
+other, more commonly measured and well understood components of water
+quality that are readily assessed in the field. For instance, there are
+small or hand held fluorometers that measure chlorohpyll \emph{a}.
+Additionally, chlorophyll \emph{a} is a very commonly measured component
+of water quality that is also known to be positively associated with
+Microsystin-LR concentrations (Pip and Bowman 2014, Yuan et al. 2014).
+Yuan et. al (2014) explore these associations in detail and control for
+other related variables. In their analysis they find that total nitrogen
+and chlorophyll \emph{a} show the strongest association with
+microcystin. Furthermore, they identify chlorophyll\emph{a} and total
+nitrogen concentrations that are associated with exceeding 1 \(\mu\)g/L
+of microcystin. Given these facts, it should be possible to identify
+chlorophyll \emph{a} concentrations that would be associated with the
+new USEPA Microcystin-LR health advisory levels for drinking water.
+Identifying these associations would provide another tool for water
+resource managers to help manage the threat to public health posed by
+cHABs and would be especially useful in the absence of microcystin-LR
+concentrations.
+
+In this paper we build on past efforts and utilize the National Lakes
+Assessment (NLA) data and identify chlorophyll \emph{a} concentrations
+that are associated with higher probabilities of exceeding several
+microcystin-LR health advisory concentrations (Chorus and Bartram 1999,
+USEPA 2009, 2015). We add to past studies by exploring associations with
+newly announced advisory levels and by also applying a different method,
+conditional probability analysis. Utilizing different methods
+strengthens the evidence for suggested chlorophyll \emph{a} levels that
+are associated with increased risk of exceeding the health advisory
+levels as those levels are not predicated on a single analytical method.
+So that others may repeat or adjust this analysis, the data, code, and
+this manuscript are freely available via
 \href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USAPE/microcystinchla}}.
@@
 (USEPA 2009). These data represent a snapshot of water quality from the
-summer of 2007 and data on chlorophyll \emph{a} and microcystin-LR
-concentrations are available for lakes.
+summer of 2007 for the conterminous United States. Data on chlorophyll
+\emph{a} and microcystin-LR concentrations are available for lakes.
 
@@
 microcystin concentrations that are better represented in the NLA data.
-These were 0.3, 1, 1.6, and 2 \(\mu\)g/L.
+These were the USEPA children's drinking water advisory level of 0.3
+\(\mu\)g/L (USEPA Child), the WHO drinking water advisory level of 1
+\(\mu\)g/L (WHO Drinking), the USEPA adult drinking water advisory level
+of 1.6 \(\mu\)g/L (USEPA Adult), and the WHO recreational, low
+probability of effect advisory level of 2 \(\mu\)g/L (WHO Recreational).
 
-A detailed discussion of CPA is beyond the scope of this paper; see Paul
-et al. (2005) and Hollister et al. (2008) for greater detail. For this
-analysis, we used CPA to examine how the conditional probability of
-exceeding one of the health advisory changes as chlorophyll \emph{a}
-increases in a lake. The 95\% confidence intervals were calculated from
-1000 bootstrapped samples. To identify chlorophyll \emph{a}
-concentrations of concern we used a 50\% conditional probability of
-exceeding each health advisory level and extracted the minimum
-chlorophyll \emph{a} concentration that was associated with an upper
-confidence level of 50\% or greater. As both microcystin-LR and
-chlorophyll \emph{a} values were highely skewed right, a log base 10
-transformation was used. Additional details of the specific
-implementation are available at
-\url{https://github.com/USEPA/microcystinchla}.
+Conditional probability analysis provides information about the
+probability of observing one event given another event has also occured.
+For this analysis, we used CPA to examine how the conditional
+probability of exceeding one of the health advisory changes as
+chlorophyll \emph{a} increases in a lake. We expect to find higher
+chlorohpyll \emph{a} concentrations to be associated with higher
+probabilities of exceeding the microcystin-LR health advisory levels. We
+also caclulated bootstrapped 95\% confidence intervals (CI) using 1000
+bootstrapped samples. Thus, to identify chlorophyll \emph{a}
+concentrations of concern we identfiy the value of the upper 95\% CI
+across a range of conditional probabilities of exceeding each health
+advisory level. As both microcystin-LR and chlorophyll \emph{a} values
+were highely skewed right, a log base 10 transformation was used.
+Additional details of the specific implementation are available at
+\url{https://github.com/USEPA/microcystinchla}. A more detailed
+discussion of CPA is beyond the scope of this paper, but see Paul et al.
+(2005) and Hollister et al. (2008) for greater detail.
 
@@
 
-In the 2007 NLA, microcystin-LR concentrations ranged from 0.05 to 225.
-Microcystin-LR concentrations of 0.05 \(\mu\)g/L represent the detection
-limits. Any value greater than that indicates the presence of
-microcystin-LR. Of those lakes with microcystin, the median
+In the 2007 NLA, microcystin-LR concentrations ranged from 0.05 to 225
+\(\mu\)g/L. Microcystin-LR concentrations of 0.05 \(\mu\)g/L represent
+the detection limits. Any value greater than that indicates the presence
+of microcystin-LR. Of those lakes with microcystin, the median
 concentration was 0.51 and the mean was 3.17. Of all lakes sampled, 21\%
-of lakes exceeded the USEPA childrens drinking water standard, 8.8\% of
-lakes exceeded the USEPA adult drinking water standard,11.7\% of lakes
-exceeded the WHO drinking water standard,and 7.3\% of lakes exceeded the
-WHO recreational standard. For chlorophyll \emph{a}, the range was 0.07
-to 936 \(\mu\)g/L. All lakes had reported chlorophyll \emph{a}
-concentrations that exceeded detection limits. The median concentration
-was 7.79 and the mean was 29.63. The associations between chlorophyll
-\emph{a} and the upper confidence interval with a conditional
-probability of 50\% were 23.68, 63.94, 84.96, and 97.49 for 0.3, 1.0,
-1.6 and 2.0 \(\mu\)g/L advisory levels, respectively (Table
+of lakes exceeded the USEPA Child level, 8.8\% of lakes exceeded the
+USEPA Adult level,11.7\% of lakes exceeded the WHO Drinking level,and
+7.3\% of lakes exceeded the WHO Recreational level. For chlorophyll
+\emph{a}, the range was 0.07 to 936 \(\mu\)g/L. All lakes had reported
+chlorophyll \emph{a} concentrations that exceeded detection limits. The
+median concentration was 7.79 \(\mu\)g/L and the mean was 29.63
+\(\mu\)g/L. The associations between chlorophyll \emph{a} and the upper
+confidence interval across a range of conditional probability values is
+shown in Table \ref{tab:mc_chla_table}. Specific chlorophyll \emph{a}
+that are associated with greater than even odds of exceeding the
+advisory levels were 0.07, 0.07, 2.79, and 11.36 for 0.3, 1.0, 1.6 and
+2.0 \(\mu\)g/L advisory levels, respectively (Table
 \ref{tab:mc_chla_table} \& Figure \ref{fig:multi_cp_plot}).
@@
 greater than even odds of exceeding a given health advisory level (Table
-\ref{tab:mc_chla_table}). These represent 27.8\%, 11.9\%, 8.6\%, and
-7.9\% of sample lakes for the U.S EPA childrens drinking water, the WHO
-drinking water, the USEPA Adult drinking water, and the WHO recreational
-advisories, respecitvely.
+\ref{tab:mc_chla_table}). These represent 99.9\%, 99.9\%, 77\%, and
+43.8\% of sample lakes for the USEPA Child, the WHO Drinking, the USEPA
+Adult, and the WHO recreational levels, respecitvely.
 
@@
 evaluate false negatives. Total accuracy of the four cutoffs predicting
-microcystin-LR exceedances were 75\% for the USEPA children's advisory,
-86\% for the WHO drinking water advisory, 90\% for the USEPA adult
-advisory and, 90\% for the WHO regreational advisory (Tables
+microcystin-LR exceedances were 21\% for the USEPA children's advisory,
+12\% for the WHO drinking water advisory, 31\% for the USEPA adult
+advisory and, 61\% for the WHO regreational advisory (Tables
 \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table},
@@
 of false negatives. For the four chlorophyll \emph{a} cut-offs we had a
-proportion of false negatives of 9\%, 7\%, 6\% and , 5\% for the U.S EPA
+proportion of false negatives of 0\%, 0\%, 0\% and , 1\% for the U.S EPA
 childrens drinking water, the WHO drinking water, the USEPA adult
@@
 is uncertainity around these chlorophyll \emph{a} cutoffs the best case
-scenario would be to monitor for chlorophyll and in the event of
-exceeding a target concentration take water samples and have those
+scenario would be to monitor for chlorophyll \emph{a} and in the event
+of exceeding a target concentration take water samples and have those
 samples tested in a lab for microcystin-LR.
@@
 
+\section{Acknowledgements}\label{acknowledgements}
+
 \newpage
@@
 probability of exceeding various microcystin-LR (MLR) health advisory
-Levels. A.) Plot for USEPA childrens drinking water advisory (0.3
-\(\mu\)g/L). B.) Plot for WHO drinking water advisory (1 \(\mu\)g/L).
-C.) Plot for USEPA adult drinking water advisory (1.6 \(\mu\)g/L). D.)
-Plot for WHO recreational advisory (2 \(\mu\)g/L).
-\label{fig:multi_cp_plot}}
+Levels. A.) Plot for USEPA Child (0.3 \(\mu\)g/L). B.) Plot for WHO
+Drinking (1 \(\mu\)g/L). C.) Plot for USEPA Adult (1.6 \(\mu\)g/L). D.)
+Plot for WHO Recreational (2 \(\mu\)g/L). \label{fig:multi_cp_plot}}
 \end{figure}
@@
 
-\begin{longtable}[c]{@{}llrr@{}}
+\begin{longtable}[c]{@{}rrrrr@{}}
 \caption{Chlorophyll \textit{a} concentrations that are associated with
@@
 \toprule
-Source & Type & Microcystin & Chlorophyll\tabularnewline
+Cond. Probability & USEPA Child & WHO Drink & USEPA Adult & WHO
+Recreational\tabularnewline
 \midrule
@@
 \toprule
-Source & Type & Microcystin & Chlorophyll\tabularnewline
+Cond. Probability & USEPA Child & WHO Drink & USEPA Adult & WHO
+Recreational\tabularnewline
 \midrule
 \endhead
-USEPA & Child Drinking Water Advisory & 0.3 & 23.68\tabularnewline
-WHO & Drinking Water Advisory & 1.0 & 63.94\tabularnewline
-USEPA & Adult Drinking Water & 1.6 & 84.96\tabularnewline
-WHO & Recreational: Low Prob. of Effect & 2.0 & 97.49\tabularnewline
+0.1 & 0.07 & 0.07 & 0.07 & 1.47\tabularnewline
+0.2 & 0.07 & 4.42 & 11.47 & 19.04\tabularnewline
+0.3 & 2.79 & 16.32 & 30.62 & 52.70\tabularnewline
+0.4 & 11.36 & 38.30 & 69.19 & 82.22\tabularnewline
+0.5 & 23.68 & 65.20 & 84.96 & 103.90\tabularnewline
+0.6 & 38.30 & 98.46 & 136.66 & 125.40\tabularnewline
+0.7 & 68.85 & 133.20 & 871.20 & 871.20\tabularnewline
+0.8 & 114.62 & 338.40 & 871.20 & 871.20\tabularnewline
+0.9 & 198.72 & 516.00 & 871.20 & 871.20\tabularnewline
 \bottomrule
@@
 \endhead
-FALSE & 650 & 96\tabularnewline
-TRUE & 161 & 121\tabularnewline
+FALSE & 1 & 0\tabularnewline
+TRUE & 810 & 217\tabularnewline
 \bottomrule
@@
 \endhead
-FALSE & 835 & 75\tabularnewline
-TRUE & 72 & 46\tabularnewline
+FALSE & 1 & 0\tabularnewline
+TRUE & 906 & 121\tabularnewline
 \bottomrule
@@
 \endhead
-FALSE & 887 & 57\tabularnewline
-TRUE & 50 & 34\tabularnewline
+FALSE & 234 & 4\tabularnewline
+TRUE & 703 & 87\tabularnewline
 \bottomrule
@@
 \endhead
-FALSE & 903 & 48\tabularnewline
-TRUE & 50 & 27\tabularnewline
+FALSE & 566 & 15\tabularnewline
+TRUE & 387 & 60\tabularnewline
 \bottomrule
@@
 
-Heisler, J., P. M. Glibert, J. M. Burkholder, D. M. Anderson, W.
-Cochlan, W. C. Dennison, Q. Dortch, C. J. Gobler, C. A. Heil, E.
-Humphries, and others. 2008. Eutrophication and harmful algal blooms: A
-scientific consensus. Harmful algae 8:3--13.
-
 Hollister, J. W., H. A. Walker, and J. F. Paul. 2008. CProb: A
@@
 
-Lee, S., M.-H. Jang, H.-S. Kim, B.-D. Yoon, and H.-M. Oh. 2000.
-Variation of microcystin content of microcystis aeruginosa relative to
-medium n: P ratio and growth stage. Journal of Applied Microbiology
-89:323--329.
-
 McElhiney, J., and L. A. Lawton. 2005. Detection of the cyanobacterial
@@
 
-Paerl, H. W., and T. G. Otten. 2013. Harmful cyanobacterial blooms:
-Causes, consequences, and controls. Microbial ecology 65:995--1010.
-
 Paul, J. F., and M. E. McDonald. 2005. Development of empirical,
@@
 and Chemistry 30:1488--1495.
+
+Pip, E., and L. Bowman. 2014. Microcystin and algal chlorophyll in
+relation to nearshore nutrient concentrations in lake winnipeg, canada.
+Environment and Pollution 3:p36.
 
@@
 
+Yuan, L. L., A. I. Pollard, S. Pather, J. L. Oliver, and L. D'Anglada.
+2014. Managing microcystin: Identifying national-scale thresholds for
+total nitrogen and chlorophyll a. Freshwater Biology 59:1970--1981.
+
 Zurawell, R. W., H. Chen, J. M. Burke, and E. E. Prepas. 2005.

### 2015-12-03 R/get_chla.R

@@ -7,6 +7,8 @@
 #' @param prob Exceedence probability to return chl a 
 #' @export
 get_chla <- function(cp_obj,prob){
-  transf_chla <- c(min(cp_obj$X[cp_obj$Upper.CI >= prob]),min(cp_obj$X[cp_obj$Lower.CI >= prob]))
+  transf_chla <- suppressWarnings(c(min(cp_obj$X[cp_obj$Upper.CI >= prob]),
+                                    min(cp_obj$X[cp_obj$Lower.CI >= prob])))
+  transf_chla[transf_chla==Inf]<-NA
   return(10^transf_chla)
 }

### 2015-12-03 manuscript/manuscript.Rmd
closes #3

@@
 Using the 2007 National Lakes Assessment and a conditional probability approach, we identify chlorophyll *a* concentrations that are more likely than not to be associated with an exceedance of a microcystin health advisory level.
-We look at the recent US EPA standards for drinking water as well as the World Health Organization levels for drinking water and recreational use.
+We look at the recent US EPA health advisories for drinking water as well as the World Health Organization levels for drinking water and recreational use.
 For the specific advisory microcystin concentrations of 0.3, 1, 1.6, and 2, we find chlorophyll *a* concentrations of `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]`, respectively.
@@
 For instance, there are small or hand held fluorometers that measure chlorohpyll *a*.
-Additionally, chlorophyll *a* is a very commonly measured component of water quality that is also known to be positively associated with Microsystin-LR concentrations [@pip2014microcystin; @yuan2014managing].
+Additionally, chlorophyll *a* is a very commonly measured component of water quality that is also known to be positively associated with microsystin-LR concentrations [@pip2014microcystin; @yuan2014managing].
 Yuan et.
@@
 Conditional probability analysis provides information about the probability of observing one event given another event has also occured.
-For this analysis, we used CPA to examine how the conditional probability of exceeding one of the health advisory changes as chlorophyll *a* increases in a lake.
+For this analysis, we used CPA to examine how the conditional probability of exceeding one of the health advisories changes as chlorophyll *a* increases in a lake.
 We expect to find higher chlorohpyll *a* concentrations to be associated with higher probabilities of exceeding the microcystin-LR health advisory levels.
@@
 In other words, we would like to avoid Type II errors and minimize the proportion of false negatives.
-For the four chlorophyll *a* cut-offs we had a proportion of false negatives of `r fn_percent$epa_child*100`%, `r fn_percent$who_drink*100`%, `r fn_percent$epa_adult*100`% and , `r fn_percent$who_rec*100`% for the U.S EPA childrens drinking water, the WHO drinking water, the USEPA adult drinking water, and the WHO recreational advisories, respecitvely.
+For the four chlorophyll *a* cut-offs we had a proportion of false negatives of `r fn_percent$epa_child*100`%, `r fn_percent$who_drink*100`%, `r fn_percent$epa_adult*100`% and , `r fn_percent$who_rec*100`% for the U.S EPA childrens drinking water, the WHO drinking water, the USEPA adult drinking water, and the WHO recreational advisories, respectively.
 In each case we miss less than 10% of the lakes that are in fact exceeding the microcystin-LR advisory.
@@
 It should be used when other, direct measurements of cyanotoxins are not available.
-In those cases, which are likely to be common at least in the near future, using a more ubiquitous meausrement, such as chlorophyll *a* will provide a reasonable proxy for the probability of exceeding a microcystin-LR health advisory level and provide better protection against adverse effects in both drinking and recreational use cases.
+In those cases, which are likely to be common at least in the near future, using a more ubiquitous measurement, such as chlorophyll *a* will provide a reasonable proxy for the probability of exceeding a microcystin-LR health advisory level and provide better protection against adverse effects in both drinking and recreational use cases.
 

### 2015-12-03 manuscript/manuscript.tex
closes #3

@@
 associated with an exceedance of a microcystin health advisory level. We
-look at the recent US EPA standards for drinking water as well as the
-World Health Organization levels for drinking water and recreational
-use. For the specific advisory microcystin concentrations of 0.3, 1,
-1.6, and 2, we find chlorophyll \emph{a} concentrations of 0.07, 0.07,
-2.79, and 11.36, respectively. When managing for these various
-microcystin levels, exceeding these reported chlorophyll \emph{a}
-concentrations should be a trigger for further testing and possible
-management action.
+look at the recent US EPA health advisories for drinking water as well
+as the World Health Organization levels for drinking water and
+recreational use. For the specific advisory microcystin concentrations
+of 0.3, 1, 1.6, and 2, we find chlorophyll \emph{a} concentrations of
+0.07, 0.07, 2.79, and 11.36, respectively. When managing for these
+various microcystin levels, exceeding these reported chlorophyll
+\emph{a} concentrations should be a trigger for further testing and
+possible management action.
 
@@
 had proposed advisory levels for drinking water and a range of
-recreational risk levels (Chorus and Bartram 1999, (WHO) and others
-2003). While these levels and associated advisories are likely to help
-mitigate the impacts from harmful algal blooms, they are not without
+recreational risk levels (Chorus and Bartram 1999, Organization 2003).
+While these levels and associated advisories are likely to help mitigate
+the impacts from harmful algal blooms, they are not without
 complications.
@@
 of water quality that is also known to be positively associated with
-Microsystin-LR concentrations (Pip and Bowman 2014, Yuan et al. 2014).
+microsystin-LR concentrations (Pip and Bowman 2014, Yuan et al. 2014).
 Yuan et. al (2014) explore these associations in detail and control for
@@
 For this analysis, we used CPA to examine how the conditional
-probability of exceeding one of the health advisory changes as
+probability of exceeding one of the health advisories changes as
 chlorophyll \emph{a} increases in a lake. We expect to find higher
@@
 childrens drinking water, the WHO drinking water, the USEPA adult
-drinking water, and the WHO recreational advisories, respecitvely. In
+drinking water, and the WHO recreational advisories, respectively. In
 each case we miss less than 10\% of the lakes that are in fact exceeding
@@
 likely to be common at least in the near future, using a more ubiquitous
-meausrement, such as chlorophyll \emph{a} will provide a reasonable
+measurement, such as chlorophyll \emph{a} will provide a reasonable
 proxy for the probability of exceeding a microcystin-LR health advisory
@@
 
+Organization, W. H. 2003. Cyanobacterial toxins: Microcystin-lR in
+drinking-water. background document for development of wHO guidelines
+for drinking-water quality. geneva, switzerland. World Health
+Organization, 2nd ed. Geneva.
+
 Paul, J. F., and M. E. McDonald. 2005. Development of empirical,
@@
 
-(WHO), W. H. O., and others. 2003. Cyanobacterial toxins: Microcystin-lR
-in drinking-water. background document for development of wHO guidelines
-for drinking-water quality. geneva, switzerland. World Health
-Organization, 2nd ed. Geneva.
-
 Yuan, L. L., A. I. Pollard, S. Pather, J. L. Oliver, and L. D'Anglada.

### 2015-12-04 manuscript/manuscript.Rmd
closes #5 plus some changes to discussion and abstract from addition of thresholds table.

@@
 #Agreement Tables
-child_conmat<-conmat_to_df(table(chla=nla_dat$CHLA>=mc_chla[1],
+child_conmat<-conmat_to_df(table(chla=nla_dat$CHLA>=mc_chla[5,1],
                                  mcsyt=nla_dat$MCYST_TL_UGL>=0.3))
-who_drink_conmat<-conmat_to_df(table(nla_dat$CHLA>=mc_chla[2],
+who_drink_conmat<-conmat_to_df(table(nla_dat$CHLA>=mc_chla[5,2],
                         nla_dat$MCYST_TL_UGL>=1))
-adult_conmat<-conmat_to_df(table(nla_dat$CHLA>=mc_chla[3],
+adult_conmat<-conmat_to_df(table(nla_dat$CHLA>=mc_chla[5,3],
                     nla_dat$MCYST_TL_UGL>=1.6))
-who_rec_conmat<-conmat_to_df(table(nla_dat$CHLA>=mc_chla[4],
+who_rec_conmat<-conmat_to_df(table(nla_dat$CHLA>=mc_chla[5,4],
                       nla_dat$MCYST_TL_UGL>=2))
@@
 Thus, there is utility in finding indicators of microcystin that are already widely available, can be estimated quickly and *in situ*, and used as a first defense against high levels of microcystin.
-In particular, chlorophyll *a* is very commonly measured, can be estimated *in situ*, and has been shown to be positively associated with microcystin.
-In this paper, we use this association to provide estimates of chlorophyll *a* concentrations that if exceeded would be indicative of a higher probability of exceeding select health advisory concentrations for microcystin-LR.
+Chlorophyll *a* is commonly measured, can be estimated *in situ*, and has been shown to be positively associated with microcystin.
+In this paper, we use this association to provide estimates of chlorophyll *a* concentrations that are indicative of a higher probability of exceeding select health advisory concentrations for microcystin-LR.
 Using the 2007 National Lakes Assessment and a conditional probability approach, we identify chlorophyll *a* concentrations that are more likely than not to be associated with an exceedance of a microcystin health advisory level.
-We look at the recent US EPA health advisories for drinking water as well as the World Health Organization levels for drinking water and recreational use.
-For the specific advisory microcystin concentrations of 0.3, 1, 1.6, and 2, we find chlorophyll *a* concentrations of `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]`, respectively.
+We look at the recent US EPA health advisories for drinking water as well as the World Health Organization levels for drinking water and recreational use and identify a range of chlorophyll *a* thresholds.
+A 50% chance of exceeding one of the specific advisory microcystin concentrations of 0.3, 1, 1.6, and 2 $\mu$g/L is associatied with chlorophyll *a* concentration thresholds of `r mc_chla[5,1]`, `r mc_chla[5,2]`, `r mc_chla[5,3]`, and `r mc_chla[5,4]`, respectively.
 When managing for these various microcystin levels, exceeding these reported chlorophyll *a* concentrations should be a trigger for further testing and possible management action.
@@
 
-Over the last decade or so numerous events and legislative activities have raised the public awareness of harmful algal blooms [@jetoo2015toledo; @rinta2009lake; @HABHRCA2014], and in response the US Environmental Protection Agency (USEPA) has recently released suggested microcystin-LR (one of the more common toxins) concentrations that would trigger health advisories [@mcelhiney2005detection; @zurawell2005hepatotoxic; @usepa2015drinking].
-Additionally, the World Health Organization (WHO) has had proposed advisory levels for drinking water and a range of recreational risk levels [@world2003cyanobacterial; @chorus1999toxic].
+Over the last decade, numerous events and legislative activities have raised the public awareness of harmful algal blooms [@jetoo2015toledo; @rinta2009lake; @HABHRCA2014].
+In response the US Environmental Protection Agency (USEPA) has recently released suggested microcystin-LR (one of the more common toxins) concentrations that would trigger health advisories [@mcelhiney2005detection; @zurawell2005hepatotoxic; @usepa2015drinking].
+Additionally, the World Health Organization (WHO) has proposed microcystin advisory levels for drinking water and a range of recreational risk levels [@world2003cyanobacterial; @chorus1999toxic].
 While these levels and associated advisories are likely to help mitigate the impacts from harmful algal blooms, they are not without complications.
@@
 One of these complications is that they rely on available measurements of microcystin-LR.
-While laboratory testing remains the gold standard for quantifying microcystin-LR concentrations in water samples, several field test kits have been developed.
+While laboratory testing (e.g., chromatography) remains the gold standard for quantifying microcystin-LR concentrations in water samples, several field test kits have been developed.
 Even though field tests provide a much needed means for rapid assessment, they are not yet widely used and are moderately expensive (approximately $150-$200 depending on specific kit) with a limited shelf life (typically one year) [@james2011environmental; @aranda2015evaluation].
@@
 ##Data
-We used the 2007 NLA water quality and microcystin-LR concentration data [@usepa2009national].
-These data represent a snapshot of water quality from the summer of 2007 for the conterminous United States.
+We used the 2007 NLA chlorophyll *a* and microcystin-LR concentration data [@usepa2009national].
+These data represent a snapshot of water quality from the summer of 2007 for the conterminous United States and were collected as part of an ongoing probabilistic monitoring program [@usepa2009national].
 Data on chlorophyll *a* and microcystin-LR concentrations are available for `r nrow(na.omit(nla_dat$CHLA,nla_dat$MCYST_TL_UGL))` lakes.
@@
 [-@hollister2008cprob] for greater detail.
+Lastly, all analyses were conducted using R version 3.2.2 and code and data from this analysis are freely available as an R package at [https://github.com/USAPE/microcystinchla](https://github.com/USAPE/microcystinchla).
 
@@
 The association between Log10 microcystin-LR and Log10 chlorophyll *a* shows a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
-This indicates that higher concentrations of microcystin-LR almost always co-occur with higher concentrations of chlorophyll *a* yet the inverse is not true.
-Higher chlorophyll *a* is not necessarily predictive of higher microcystin-LR concentrations; however, chlorophyll *a* may be predictive of the probability of exceeding a certain concentration.
+This indicates that, in general, higher concentrations of microcystin-LR almost always co-occur with higher concentrations of chlorophyll *a* yet the inverse is not true.
+Higher chlorophyll *a* is not necessarily predictive of higher microcystin-LR concentrations; however, chlorophyll *a* may be predictive of the probability of exceeding a certain threshold.
 
 This is the case as the probability of exceeding each of the four tested health advisory levels increases as a function of chlorophyll *a* concentration (Figure \ref{fig:multi_cp_plot}).
-We use this association to identify chlorophyll *a* concentrations that are associated with greater than even odds of exceeding a given health advisory level (Table \ref{tab:mc_chla_table}).
-These represent `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[1]])/length(nla_dat$CHLA)),1)`%, `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[2]])/length(nla_dat$CHLA)),1)`%,
-`r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[3]])/length(nla_dat$CHLA)),1)`%, and `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[4]])/length(nla_dat$CHLA)),1)`% of sample lakes for the USEPA Child, the WHO Drinking, the USEPA Adult, and the WHO recreational levels, respecitvely.
+We use this association to identify chlorophyll *a* concentrations that are associated with a range of probabilities of exceeding a given health advisory level (Table \ref{tab:mc_chla_table}).
+For the purposes of this discussion we focus on a conditional probability of 50% or greater (i.e.
+greater than even odds to exceed a health advisory level).
+The 50% conditional probability chlorophyll *a* thresholds represents `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[5,1]])/length(nla_dat$CHLA)),1)`%, `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[5,2]])/length(nla_dat$CHLA)),1)`%,
+`r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[5,3]])/length(nla_dat$CHLA)),1)`%, and `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[5,4]])/length(nla_dat$CHLA)),1)`% of sample lakes for the USEPA Child, the WHO Drinking, the USEPA Adult, and the WHO recreational levels, respecitvely.
 
@@
 
-A second potential use is to identify possible bloom events from historical data.
+A second potential use is to identify past bloom events from historical data.
 As harmful algal blooms are made up of many species and have various mechanisms responsible for adverse impacts (e.g.
@@
 
-Lastly, using chlorophyll *a* is not meant as a replacement for testing of microcystin-LR.
+Lastly, using chlorophyll *a* is not meant as a replacement for testing of microcystin-LR or other toxins.
 It should be used when other, direct measurements of cyanotoxins are not available.

### 2015-12-04 manuscript/manuscript.tex
closes #5 plus some changes to discussion and abstract from addition of thresholds table.

@@
 widely available, can be estimated quickly and \emph{in situ}, and used
-as a first defense against high levels of microcystin. In particular,
-chlorophyll \emph{a} is very commonly measured, can be estimated
-\emph{in situ}, and has been shown to be positively associated with
-microcystin. In this paper, we use this association to provide estimates
-of chlorophyll \emph{a} concentrations that if exceeded would be
-indicative of a higher probability of exceeding select health advisory
-concentrations for microcystin-LR. Using the 2007 National Lakes
-Assessment and a conditional probability approach, we identify
-chlorophyll \emph{a} concentrations that are more likely than not to be
-associated with an exceedance of a microcystin health advisory level. We
-look at the recent US EPA health advisories for drinking water as well
-as the World Health Organization levels for drinking water and
-recreational use. For the specific advisory microcystin concentrations
-of 0.3, 1, 1.6, and 2, we find chlorophyll \emph{a} concentrations of
-0.07, 0.07, 2.79, and 11.36, respectively. When managing for these
-various microcystin levels, exceeding these reported chlorophyll
-\emph{a} concentrations should be a trigger for further testing and
-possible management action.
+as a first defense against high levels of microcystin. Chlorophyll
+\emph{a} is commonly measured, can be estimated \emph{in situ}, and has
+been shown to be positively associated with microcystin. In this paper,
+we use this association to provide estimates of chlorophyll \emph{a}
+concentrations that are indicative of a higher probability of exceeding
+select health advisory concentrations for microcystin-LR. Using the 2007
+National Lakes Assessment and a conditional probability approach, we
+identify chlorophyll \emph{a} concentrations that are more likely than
+not to be associated with an exceedance of a microcystin health advisory
+level. We look at the recent US EPA health advisories for drinking water
+as well as the World Health Organization levels for drinking water and
+recreational use and identify a range of chlorophyll \emph{a}
+thresholds. A 50\% chance of exceeding one of the specific advisory
+microcystin concentrations of 0.3, 1, 1.6, and 2 \(\mu\)g/L is
+associatied with chlorophyll \emph{a} concentration thresholds of 20.88,
+69.6, 84.24, and 97.49, respectively. When managing for these various
+microcystin levels, exceeding these reported chlorophyll \emph{a}
+concentrations should be a trigger for further testing and possible
+management action.
 
@@
 
-Over the last decade or so numerous events and legislative activities
-have raised the public awareness of harmful algal blooms (Rinta-Kanto et
-al. 2009, HABHRCA 2014, Jetoo et al. 2015), and in response the US
-Environmental Protection Agency (USEPA) has recently released suggested
-microcystin-LR (one of the more common toxins) concentrations that would
-trigger health advisories (McElhiney and Lawton 2005, Zurawell et al.
-2005, USEPA 2015). Additionally, the World Health Organization (WHO) has
-had proposed advisory levels for drinking water and a range of
+Over the last decade, numerous events and legislative activities have
+raised the public awareness of harmful algal blooms (Rinta-Kanto et al.
+2009, HABHRCA 2014, Jetoo et al. 2015). In response the US Environmental
+Protection Agency (USEPA) has recently released suggested microcystin-LR
+(one of the more common toxins) concentrations that would trigger health
+advisories (McElhiney and Lawton 2005, Zurawell et al. 2005, USEPA
+2015). Additionally, the World Health Organization (WHO) has proposed
+microcystin advisory levels for drinking water and a range of
 recreational risk levels (Chorus and Bartram 1999, Organization 2003).
@@
 One of these complications is that they rely on available measurements
-of microcystin-LR. While laboratory testing remains the gold standard
-for quantifying microcystin-LR concentrations in water samples, several
-field test kits have been developed. Even though field tests provide a
-much needed means for rapid assessment, they are not yet widely used and
-are moderately expensive (approximately \$150-\$200 depending on
-specific kit) with a limited shelf life (typically one year) (James et
-al. 2011, Aranda-Rodriguez et al. 2015). Additionally, each technique
-requires nuanced understanding of the detection method (e.g., limit of
-detection, specific microcystin variants being measured, and sampling
-protocol).
+of microcystin-LR. While laboratory testing (e.g., chromatography)
+remains the gold standard for quantifying microcystin-LR concentrations
+in water samples, several field test kits have been developed. Even
+though field tests provide a much needed means for rapid assessment,
+they are not yet widely used and are moderately expensive (approximately
+\$150-\$200 depending on specific kit) with a limited shelf life
+(typically one year) (James et al. 2011, Aranda-Rodriguez et al. 2015).
+Additionally, each technique requires nuanced understanding of the
+detection method (e.g., limit of detection, specific microcystin
+variants being measured, and sampling protocol).
 
@@
 
-We used the 2007 NLA water quality and microcystin-LR concentration data
-(USEPA 2009). These data represent a snapshot of water quality from the
-summer of 2007 for the conterminous United States. Data on chlorophyll
-\emph{a} and microcystin-LR concentrations are available for lakes.
+We used the 2007 NLA chlorophyll \emph{a} and microcystin-LR
+concentration data (USEPA 2009). These data represent a snapshot of
+water quality from the summer of 2007 for the conterminous United States
+and were collected as part of an ongoing probabilistic monitoring
+program (USEPA 2009). Data on chlorophyll \emph{a} and microcystin-LR
+concentrations are available for lakes.
 
@@
 discussion of CPA is beyond the scope of this paper, but see Paul et al.
-(2005) and Hollister et al. (2008) for greater detail.
+(2005) and Hollister et al. (2008) for greater detail. Lastly, all
+analyses were conducted using R version 3.2.2 and code and data from
+this analysis are freely available as an R package at
+\href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USAPE/microcystinchla}}.
 
@@
 that are associated with greater than even odds of exceeding the
-advisory levels were 0.07, 0.07, 2.79, and 11.36 for 0.3, 1.0, 1.6 and
+advisory levels were 0.07, 0.07, 2.89, and 10.27 for 0.3, 1.0, 1.6 and
 2.0 \(\mu\)g/L advisory levels, respectively (Table
@@
 \emph{a} shows a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
-This indicates that higher concentrations of microcystin-LR almost
-always co-occur with higher concentrations of chlorophyll \emph{a} yet
-the inverse is not true. Higher chlorophyll \emph{a} is not necessarily
-predictive of higher microcystin-LR concentrations; however, chlorophyll
-\emph{a} may be predictive of the probability of exceeding a certain
-concentration.
+This indicates that, in general, higher concentrations of microcystin-LR
+almost always co-occur with higher concentrations of chlorophyll
+\emph{a} yet the inverse is not true. Higher chlorophyll \emph{a} is not
+necessarily predictive of higher microcystin-LR concentrations; however,
+chlorophyll \emph{a} may be predictive of the probability of exceeding a
+certain threshold.
 
@@
 to identify chlorophyll \emph{a} concentrations that are associated with
-greater than even odds of exceeding a given health advisory level (Table
-\ref{tab:mc_chla_table}). These represent 99.9\%, 99.9\%, 77\%, and
-43.8\% of sample lakes for the USEPA Child, the WHO Drinking, the USEPA
-Adult, and the WHO recreational levels, respecitvely.
+a range of probabilities of exceeding a given health advisory level
+(Table \ref{tab:mc_chla_table}). For the purposes of this discussion we
+focus on a conditional probability of 50\% or greater (i.e.~greater than
+even odds to exceed a health advisory level). The 50\% conditional
+probability chlorophyll \emph{a} thresholds represents 30.1\%, 10.6\%,
+8.8\%, and 7.9\% of sample lakes for the USEPA Child, the WHO Drinking,
+the USEPA Adult, and the WHO recreational levels, respecitvely.
 
@@
 evaluate false negatives. Total accuracy of the four cutoffs predicting
-microcystin-LR exceedances were 21\% for the USEPA children's advisory,
-12\% for the WHO drinking water advisory, 31\% for the USEPA adult
-advisory and, 61\% for the WHO regreational advisory (Tables
+microcystin-LR exceedances were 74\% for the USEPA children's advisory,
+86\% for the WHO drinking water advisory, 89\% for the USEPA adult
+advisory and, 90\% for the WHO regreational advisory (Tables
 \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table},
@@
 of false negatives. For the four chlorophyll \emph{a} cut-offs we had a
-proportion of false negatives of 0\%, 0\%, 0\% and , 1\% for the U.S EPA
+proportion of false negatives of 9\%, 8\%, 6\% and , 5\% for the U.S EPA
 childrens drinking water, the WHO drinking water, the USEPA adult
@@
 
-A second potential use is to identify possible bloom events from
-historical data. As harmful algal blooms are made up of many species and
-have various mechanisms responsible for adverse impacts (e.g.~toxins,
+A second potential use is to identify past bloom events from historical
+data. As harmful algal blooms are made up of many species and have
+various mechanisms responsible for adverse impacts (e.g.~toxins,
 hypoxia, odors), there is no single definition of a bloom. For cHABs one
@@
 Lastly, using chlorophyll \emph{a} is not meant as a replacement for
-testing of microcystin-LR. It should be used when other, direct
-measurements of cyanotoxins are not available. In those cases, which are
-likely to be common at least in the near future, using a more ubiquitous
-measurement, such as chlorophyll \emph{a} will provide a reasonable
-proxy for the probability of exceeding a microcystin-LR health advisory
-level and provide better protection against adverse effects in both
-drinking and recreational use cases.
+testing of microcystin-LR or other toxins. It should be used when other,
+direct measurements of cyanotoxins are not available. In those cases,
+which are likely to be common at least in the near future, using a more
+ubiquitous measurement, such as chlorophyll \emph{a} will provide a
+reasonable proxy for the probability of exceeding a microcystin-LR
+health advisory level and provide better protection against adverse
+effects in both drinking and recreational use cases.
 
@@
 \endhead
-0.1 & 0.07 & 0.07 & 0.07 & 1.47\tabularnewline
-0.2 & 0.07 & 4.42 & 11.47 & 19.04\tabularnewline
-0.3 & 2.79 & 16.32 & 30.62 & 52.70\tabularnewline
-0.4 & 11.36 & 38.30 & 69.19 & 82.22\tabularnewline
-0.5 & 23.68 & 65.20 & 84.96 & 103.90\tabularnewline
-0.6 & 38.30 & 98.46 & 136.66 & 125.40\tabularnewline
-0.7 & 68.85 & 133.20 & 871.20 & 871.20\tabularnewline
-0.8 & 114.62 & 338.40 & 871.20 & 871.20\tabularnewline
-0.9 & 198.72 & 516.00 & 871.20 & 871.20\tabularnewline
+0.1 & 0.07 & 0.07 & 0.07 & 1.17\tabularnewline
+0.2 & 0.07 & 4.38 & 11.84 & 16.80\tabularnewline
+0.3 & 2.89 & 17.49 & 32.11 & 45.22\tabularnewline
+0.4 & 10.27 & 38.30 & 66.96 & 76.80\tabularnewline
+0.5 & 20.88 & 69.60 & 84.24 & 97.49\tabularnewline
+0.6 & 38.20 & 100.51 & 114.62 & 125.40\tabularnewline
+0.7 & 65.60 & 128.74 & 871.20 & 871.20\tabularnewline
+0.8 & 114.62 & 271.44 & 871.20 & 871.20\tabularnewline
+0.9 & 166.63 & 338.40 & 871.20 & 871.20\tabularnewline
 \bottomrule
@@
 \endhead
-FALSE & 1 & 0\tabularnewline
-TRUE & 810 & 217\tabularnewline
+FALSE & 632 & 90\tabularnewline
+TRUE & 179 & 127\tabularnewline
 \bottomrule
@@
 \endhead
-FALSE & 1 & 0\tabularnewline
-TRUE & 906 & 121\tabularnewline
+FALSE & 844 & 80\tabularnewline
+TRUE & 63 & 41\tabularnewline
 \bottomrule
@@
 \endhead
-FALSE & 234 & 4\tabularnewline
-TRUE & 703 & 87\tabularnewline
+FALSE & 885 & 57\tabularnewline
+TRUE & 52 & 34\tabularnewline
 \bottomrule
@@
 \endhead
-FALSE & 566 & 15\tabularnewline
-TRUE & 387 & 60\tabularnewline
+FALSE & 903 & 48\tabularnewline
+TRUE & 50 & 27\tabularnewline
 \bottomrule

### 2015-12-04 manuscript/manuscript.Rmd
started edits on #6

@@
 ---
-title: Associations between Chlorophyll *a* and various Microcystin-LR Health Advisory
-  Concentrations
+title: Associations between Chlorophyll *a* and various Microcystin-LR Health Advisory Concentrations
 author:
@@
 ##Data
-We used the 2007 NLA chlorophyll *a* and microcystin-LR concentration data [@usepa2009national].
+We used the 2007 NLA chlorophyll *a* and microcystin-LR concentration data [@usepa2009national, [NLA](http://www2.epa.gov/national-aquatic-resource-surveys/national-lakes-assessment)].
 These data represent a snapshot of water quality from the summer of 2007 for the conterminous United States and were collected as part of an ongoing probabilistic monitoring program [@usepa2009national].
@@
 Thus, to identify chlorophyll *a* concentrations of concern we identfiy the value of the upper 95% CI across a range of conditional probabilities of exceeding each health advisory level.
+Using the upper confidence limit to identify a threshold is justified as it ensures that a given threshold is unlikely to miss a microcystin exceedance.
+
 As both microcystin-LR and chlorophyll *a* values were highely skewed right, a log base 10 transformation was used.
@@
 #Results
-
 In the 2007 NLA, microcystin-LR concentrations ranged from `r min(nla_dat$MCYST_TL_UGL,na.rm=T)` to `r max(nla_dat$MCYST_TL_UGL, na.rm=T)` $\mu$g/L.
@@
 In each case we miss less than 10% of the lakes that are in fact exceeding the microcystin-LR advisory.
+While this method performs well with regard to the false negative percentag, it is possible that is a relic of the NLA dataset and testing with additional data should be explore to further substatiate this result.
 
@@
 #Acknowledgements
+We would like to thank Anne Kuhn, Bryan Milstead, John Kiddon, Joe LiVolsi, Tim Gleason, and Wayne Munns for constructive reviews of this paper.
+This paper has not been subjected to Agency review.
+Therefore, it does not necessary reflect the views of the Agency.
+Mention of trade names or commercial products does not constitute endorsement or recommendation for use.
+This contribution is identified by the tracking number ORD-XXXXX of the Atlantic Ecology Division, Office of Research and Development, National Health and Environmental Effects Research Laboratory, US Environmental Protection Agency.
 

### 2015-12-04 manuscript/manuscript.tex
started edits on #6

@@
 We used the 2007 NLA chlorophyll \emph{a} and microcystin-LR
-concentration data (USEPA 2009). These data represent a snapshot of
-water quality from the summer of 2007 for the conterminous United States
-and were collected as part of an ongoing probabilistic monitoring
-program (USEPA 2009). Data on chlorophyll \emph{a} and microcystin-LR
-concentrations are available for lakes.
+concentration data (USEPA 2009,
+\href{http://www2.epa.gov/national-aquatic-resource-surveys/national-lakes-assessment}{NLA}).
+These data represent a snapshot of water quality from the summer of 2007
+for the conterminous United States and were collected as part of an
+ongoing probabilistic monitoring program (USEPA 2009). Data on
+chlorophyll \emph{a} and microcystin-LR concentrations are available for
+lakes.
 
@@
 across a range of conditional probabilities of exceeding each health
-advisory level. As both microcystin-LR and chlorophyll \emph{a} values
-were highely skewed right, a log base 10 transformation was used.
-Additional details of the specific implementation are available at
+advisory level. Using the upper confidence limit to identify a threshold
+is justified as it ensures that a given threshold is unlikely to miss a
+microcystin exceedance.
+
+As both microcystin-LR and chlorophyll \emph{a} values were highely
+skewed right, a log base 10 transformation was used. Additional details
+of the specific implementation are available at
 \url{https://github.com/USEPA/microcystinchla}. A more detailed
@@
 each case we miss less than 10\% of the lakes that are in fact exceeding
-the microcystin-LR advisory.
+the microcystin-LR advisory. While this method performs well with regard
+to the false negative percentag, it is possible that is a relic of the
+NLA dataset and testing with additional data should be explore to
+further substatiate this result.
 
@@
 \section{Acknowledgements}\label{acknowledgements}
+
+We would like to thank Anne Kuhn, Bryan Milstead, John Kiddon, Joe
+LiVolsi, Tim Gleason, and Wayne Munns for constructive reviews of this
+paper. This paper has not been subjected to Agency review. Therefore, it
+does not necessary reflect the views of the Agency. Mention of trade
+names or commercial products does not constitute endorsement or
+recommendation for use. This contribution is identified by the tracking
+number ORD-XXXXX of the Atlantic Ecology Division, Office of Research
+and Development, National Health and Environmental Effects Research
+Laboratory, US Environmental Protection Agency.
 

### 2016-01-15 manuscript/manuscript.Rmd
Wayne and Tim comments/edits

@@
 ---
-title: Associations between Chlorophyll *a* and various Microcystin-LR Health Advisory Concentrations
+title: Associations between chlorophyll *a* and various microcystin-LR health advisory
+  Concentrations
 author:
@@
     number_sections: yes
-    template: components/manuscript.latex
+    template: components/f1000research.latex
   html_document: null
-  word_document: null
 capsize: normalsize
@@
                                  mcsyt=nla_dat$MCYST_TL_UGL>=0.3))
+child_conmat$tots<-child_conmat[,1]+child_conmat[,2]
+child_conmat[3,]<-child_conmat[1,]+child_conmat[2,]
+
 who_drink_conmat<-conmat_to_df(table(nla_dat$CHLA>=mc_chla[5,2],
                         nla_dat$MCYST_TL_UGL>=1))
+who_drink_conmat$tots<-who_drink_conmat[,1]+who_drink_conmat[,2]
+who_drink_conmat[3,]<-who_drink_conmat[1,]+who_drink_conmat[2,]
+
 adult_conmat<-conmat_to_df(table(nla_dat$CHLA>=mc_chla[5,3],
                     nla_dat$MCYST_TL_UGL>=1.6))
+adult_conmat$tots<-adult_conmat[,1]+adult_conmat[,2]
+adult_conmat[3,]<-adult_conmat[1,]+adult_conmat[2,]
+
 who_rec_conmat<-conmat_to_df(table(nla_dat$CHLA>=mc_chla[5,4],
                       nla_dat$MCYST_TL_UGL>=2))
+who_rec_conmat$tots<-who_rec_conmat[,1]+who_rec_conmat[,2]
+who_rec_conmat[3,]<-who_rec_conmat[1,]+who_rec_conmat[2,]
+
+names(child_conmat)<-c("Not Exceed","Exceed","Row Totals")
+row.names(child_conmat)<-c("Not Exceed","Exceed","Column Totals")
+names(who_drink_conmat)<-c("Not Exceed","Exceed","Row Totals")
+row.names(who_drink_conmat)<-c("Not Exceed","Exceed","Column Totals")
+names(adult_conmat)<-c("Not Exceed","Exceed","Row Totals")
+row.names(adult_conmat)<-c("Not Exceed","Exceed","Column Totals")
+names(who_rec_conmat)<-c("Not Exceed","Exceed","Row Totals")
+row.names(who_rec_conmat)<-c("Not Exceed","Exceed","Column Totals")
+
 total_accuracy<-NULL
@@
 Furthermore, they identify chlorophyll*a* and total nitrogen concentrations that are associated with exceeding 1 $\mu$g/L of microcystin.
-Given these facts, it should be possible to identify chlorophyll *a* concentrations that would be associated with the new USEPA Microcystin-LR health advisory levels for drinking water.
-Identifying these associations would provide another tool for water resource managers to help manage the threat to public health posed by cHABs and would be especially useful in the absence of microcystin-LR concentrations.
+Given these facts, it should be possible to identify chlorophyll *a* concentrations that would be associated with the new USEPA microcystin-LR health advisory levels for drinking water.
+Identifying these associations would provide another tool for water resource managers to help manage the threat to public health posed by cHABs and would be especially useful in the absence of measured microcystin-LR concentrations.
 
@@
 Utilizing different methods strengthens the evidence for suggested chlorophyll *a* levels that are associated with increased risk of exceeding the health advisory levels as those levels are not predicated on a single analytical method.
-So that others may repeat or adjust this analysis, the data, code, and this manuscript are freely available via [https://github.com/USAPE/microcystinchla](https://github.com/USAPE/microcystinchla).
+So that others may repeat or adjust this analysis, the data, code, and this manuscript are freely available via [https://github.com/USEPA/microcystinchla](https://github.com/USAPE/microcystinchla).
 
@@
 
-##Conditional Probability Analysis
+##Analytical Methods
 We used a conditional probability analysis (CPA) approach to explore associations between chlorophyll *a* concentrations and World Health Organization (WHO) and USEPA microcystin-LR health advisory levels [@paul2011probability].
-Many levels have been suggested (Table \ref{tab:microcystin_levels}), but lakes with higher microcystin-LR concentrations in the NLA were rare.
-Only `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=10])/nrow(nla_dat)),2)` % of lakes sampled had a concentration greater than 10.
-Thus, for this analysis we focus on the microcystin concentrations that are better represented in the NLA data.
+Many health advisory levels have been suggested (Table \ref{tab:microcystin_levels}), but lakes with higher microcystin-LR concentrations in the NLA were rare.
+Only `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=10])/nrow(nla_dat)),2)`% of lakes sampled had a concentration greater than 10 $\mu$g/L.
+Thus, for this analysis we focused on the microcystin concentrations that are better represented in the NLA data.
 These were the USEPA children's drinking water advisory level of  0.3 $\mu$g/L (USEPA Child), the WHO drinking water advisory level of 1 $\mu$g/L (WHO Drinking), the USEPA adult drinking water advisory level of 1.6 $\mu$g/L (USEPA Adult), and the WHO recreational, low probability of effect advisory level of 2 $\mu$g/L (WHO Recreational).
@@
 We expect to find higher chlorohpyll *a* concentrations to be associated with higher probabilities of exceeding the microcystin-LR health advisory levels.
-We also caclulated bootstrapped 95% confidence intervals (CI) using 1000 bootstrapped samples.
+We also calculated bootstrapped 95% confidence intervals (CI) using 1000 bootstrapped samples.
 Thus, to identify chlorophyll *a* concentrations of concern we identfiy the value of the upper 95% CI across a range of conditional probabilities of exceeding each health advisory level.
@@
 [-@hollister2008cprob] for greater detail.
-Lastly, all analyses were conducted using R version 3.2.2 and code and data from this analysis are freely available as an R package at [https://github.com/USAPE/microcystinchla](https://github.com/USAPE/microcystinchla).
+Lastly, all analyses were conducted using R version 3.2.2 and code and data from this analysis are freely available as an R package at [https://github.com/USEPA/microcystinchla](https://github.com/USAPE/microcystinchla).
+
+Lastly, we assess the ability of these chlorophyll *a* thresholds to predict microcystin exceedance.
+We use error matrices and calculate total accuracy as well as the proportion of false negatives.
+Total accuracy is the total number of correct predictions divided by total observations.
+The proportion of false negatives is the total number of lakes that were predicted to not exceed the microcystin guidelines but actually did, divided by the total number of observations.
 
@@
 Any value greater than that indicates the presence of microcystin-LR.
-Of those lakes with microcystin, the median concentration was `r round(median(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T), 2)` and the mean was `r round(mean(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T), 2)`.
-Of all lakes sampled, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=0.3])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the USEPA Child level, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1.6])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the USEPA Adult level,`r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO Drinking level,and `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=2])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO Recreational level.
+Of those lakes with microcystin, the median concentration was `r round(median(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T), 2)`$\mu$g/L and the mean was `r round(mean(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T), 2)`$\mu$g/L.
+Of all lakes sampled, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=0.3])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the USEPA Child level, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1.6])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the USEPA Adult level, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO Drinking level, and `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=2])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO Recreational level.
 For chlorophyll *a*, the range was `r min(nla_dat$CHLA,na.rm=T)` to `r max(nla_dat$CHLA, na.rm=T)` $\mu$g/L.
@@
 The median concentration was `r median(nla_dat$CHLA,na.rm=T)` $\mu$g/L and the mean was `r round(mean(nla_dat$CHLA,na.rm=T),2)` $\mu$g/L.
-The associations between chlorophyll *a* and the upper confidence interval across a range of conditional probability values is shown in Table \ref{tab:mc_chla_table}.
-Specific chlorophyll *a* that are associated with greater than even odds of exceeding the advisory levels were `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]` for 0.3, 1.0, 1.6 and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
+The associations between chlorophyll *a* and the upper confidence interval across a range of conditional probability values are shown in Table \ref{tab:mc_chla_table}.
+Specific chlorophyll *a* that are associated with greater than even odds of exceeding the advisory levels were `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]`$\mu$g/L for 0.3, 1.0, 1.6, and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
+
+The chlorophyll *a* cutoffs may be used to predict whether or not a lake exceeds the microcystin-LR health advisories.
+Doing so allows us to compare the accuracy of the prediction as well as evaluate false negatives.
+Total accuracy of the four cutoffs predicting microcystin-LR exceedances were `r total_accuracy$epa_child*100`% for the USEPA children's advisory, `r total_accuracy$who_drink*100`% for the WHO drinking water advisory,  `r total_accuracy$epa_adult*100`% for the USEPA adult advisory, and `r total_accuracy$who_rec*100`% for the WHO regreational advisory (Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table}, \ref{tab:adult_conmat_table}, & \ref{tab:who_rec_conmat_table}).
+However, total accuracy is only one part of the prediction performace with which we are concerned.
+
+When using the chlorophyll *a* cutoffs as an indicator of microcystin-LR exceedances, the error that should be avoided is predicting that no exceedance has occurred when in fact it has.
+In other words, we would like to avoid Type II errors and minimize the proportion of false negatives.
+For the four chlorophyll *a* cut-offs we had a proportion of false negatives of `r fn_percent$epa_child*100`%, `r fn_percent$who_drink*100`%, `r fn_percent$epa_adult*100`%, and `r fn_percent$who_rec*100`% for the USEPA children's, the WHO drinking water, the USEPA adult, and the WHO recreational advisories, respectively.
+In each case we missed less than 10% of the lakes that in fact exceeded the microcystin-LR advisory.
+While this method performs well with regard to the false negative percentage, it is possible that is a relic of the NLA dataset and testing with additional data would allow us to confirm this result.
 
@@
 This is the case as the probability of exceeding each of the four tested health advisory levels increases as a function of chlorophyll *a* concentration (Figure \ref{fig:multi_cp_plot}).
-We use this association to identify chlorophyll *a* concentrations that are associated with a range of probabilities of exceeding a given health advisory level (Table \ref{tab:mc_chla_table}).
-For the purposes of this discussion we focus on a conditional probability of 50% or greater (i.e.
-greater than even odds to exceed a health advisory level).
+We used this association to identify chlorophyll *a* concentrations that are associated with a range of probabilities of exceeding a given health advisory level (Table \ref{tab:mc_chla_table}).
+For the purposes of this discussion we focus on a conditional probability of 50% or greater (i.e., greater than even odds to exceed a health advisory level).
 The 50% conditional probability chlorophyll *a* thresholds represents `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[5,1]])/length(nla_dat$CHLA)),1)`%, `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[5,2]])/length(nla_dat$CHLA)),1)`%,
@@
 
-Furthermore, the chlorophyll *a* cutoffs may be used to predict whether or not a lake exceeds the microcystin-LR health advisories.
-Doing so allows us to compare the accuracy of the prediction as well as evaluate false negatives.
-Total accuracy of the four cutoffs predicting microcystin-LR exceedances were `r total_accuracy$epa_child*100`% for the USEPA children's advisory, `r total_accuracy$who_drink*100`% for the WHO drinking water advisory,  `r total_accuracy$epa_adult*100`% for the USEPA adult advisory and, `r total_accuracy$who_rec*100`% for the WHO regreational advisory (Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table}, \ref{tab:adult_conmat_table}, & \ref{tab:who_rec_conmat_table}).
-However, total accuracy is only one part of the prediction performace with which we are concerned.
-
-When using the chlorophyll *a* cutoffs as an indicator of microcystin-LR exceedances, the error that should be avoided is predicting that no exceedance has occurred when in fact it has.
-In other words, we would like to avoid Type II errors and minimize the proportion of false negatives.
-For the four chlorophyll *a* cut-offs we had a proportion of false negatives of `r fn_percent$epa_child*100`%, `r fn_percent$who_drink*100`%, `r fn_percent$epa_adult*100`% and , `r fn_percent$who_rec*100`% for the U.S EPA childrens drinking water, the WHO drinking water, the USEPA adult drinking water, and the WHO recreational advisories, respectively.
-In each case we miss less than 10% of the lakes that are in fact exceeding the microcystin-LR advisory.
-While this method performs well with regard to the false negative percentag, it is possible that is a relic of the NLA dataset and testing with additional data should be explore to further substatiate this result.
-
 There are numerous possible uses for the chlorophyll *a* and microcystin-LR advisory cut-off values.
 First, in the absence of microcystin-LR measurements, exceedence of the chlorophyll *a* concentrations could be a trigger for further actions.
-Given that there is uncertainity around these chlorophyll *a* cutoffs the best case scenario would be to monitor for chlorophyll *a* and in the event of exceeding a target concentration take water samples and have those samples tested in a lab for microcystin-LR.
+Given that there is uncertainity around these chlorophyll *a* cutoffs the best case scenario would be to monitor for chlorophyll *a* and in the event of exceeding a target concentration take water samples and have those samples tested for microcystin-LR.
 
 A second potential use is to identify past bloom events from historical data.
-As harmful algal blooms are made up of many species and have various mechanisms responsible for adverse impacts (e.g.
-toxins, hypoxia, odors), there is no single definition of a bloom.
-For cHABs one approach has been to identify an increase over a baseline concentration of phycocyanin [@miller2013spatiotemporal].
+As harmful algal blooms are made up of many species and have various mechanisms responsible for adverse impacts (e.g., toxins, hypoxia, odors), there is no single definition of a bloom.
+For cHABs, one approach has been to identify an increase over a baseline concentration of phycocyanin [@miller2013spatiotemporal].
 This is a useful approach for targeted studies, but phycocyanin is also not always available and measures the predominance of cyanobacterial pigments and not toxins.
@@
 Mention of trade names or commercial products does not constitute endorsement or recommendation for use.
-This contribution is identified by the tracking number ORD-XXXXX of the Atlantic Ecology Division, Office of Research and Development, National Health and Environmental Effects Research Laboratory, US Environmental Protection Agency.
+This contribution is identified by the tracking number ORD-015143 of the Atlantic Ecology Division, Office of Research and Development, National Health and Environmental Effects Research Laboratory, US Environmental Protection Agency.
 

### 2016-01-15 manuscript/manuscript.tex
Wayne and Tim comments/edits

--- 2015-12-03 manuscript/manuscript.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Original default rstudio/pandoc latex file
-%% upated by @jhollist 09/15/2014
-%% inspired by @cboetting https://github.com/cboettig/template and
-%% @rmflight blog posts:
-%% http://rmflight.github.io/posts/2014/07/analyses_as_packages.html 
-%% http://rmflight.github.io/posts/2014/07/vignetteAnalysis.html).  
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+---
+title: Associations between chlorophyll *a* and various microcystin-LR health advisory
+  Concentrations
+author:
+- affilnum: 1
+  email: hollister.jeff@epa.gov
+  name: Jeffrey W. Hollister
+- affilnum: 1
+  name: Betty J. Kreakie
+affiliation:
+- affil: US Environmental Protection Agency, Office of Research and Development, National
+    Health and Environmental Effects Research Laboratory, Atlantic Ecology Division,
+    27 Tarzwell Drive  Narragansett, RI, 02882, USA
+  affilnum: 1
+output:
+  pdf_document:
+    fig_caption: yes
+    keep_tex: yes
+    number_sections: yes
+    template: components/f1000research.latex
+  html_document: null
+capsize: normalsize
+csl: components/ecology.csl
+documentclass: article
+fontsize: 11pt
+linenumbers: yes
+bibliography: components/manuscript.bib
+spacing: doublespacing
+abstract: no
+---
 
-\documentclass[11pt,]{article}
-\usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-    \usepackage{xltxtra,xunicode}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-% use microtype if available
-\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\usepackage{longtable,booktabs}
-\usepackage{graphicx}
-% Redefine \includegraphics so that, unless explicit options are
-% given, the image width will not exceed the width of the page.
-% Images get their normal width if they fit onto the page, but
-% are scaled down if they would overflow the margins.
-\makeatletter
-\def\ScaleIfNeeded{%
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-\let\Oldincludegraphics\includegraphics
-{%
- \catcode`\@=11\relax%
- \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
-}%
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={Associations between Chlorophyll a and various Microcystin-LR Health Advisory Concentrations},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\setcounter{secnumdepth}{5}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%Changes borrowed from @cboettig, added by @jhollist 
-% A modified page layout 
-\textwidth 6.75in
-\oddsidemargin -0.15in
-\evensidemargin -0.15in
-\textheight 9in
-\topmargin -0.5in
-\usepackage{lineno} % add 
-  \linenumbers % turns line numbering on 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%Packages and layout changes by @jhollist 09/15/2014
-\usepackage{ragged2e}
-\usepackage[font=normalsize]{caption}
-  \usepackage[doublespacing]{setspace}
-\usepackage{parskip}
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\fancyhf{}
-\renewcommand{\headrulewidth}{0pt}
-\rfoot{\today}
-\lfoot{\thepage}
-%%Changed default abstract width and added lines
-\renewenvironment{abstract}{
-  \hfill\begin{minipage}{1\textwidth}
-  \rule{\textwidth}{1pt}\vspace{5pt}
-  \normalsize
-  \begin{justify}
-  \bfseries\abstractname\vspace{5pt}
-  \end{justify}}
-  {\par\noindent\rule{\textwidth}{1pt}\end{minipage}
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\title{Associations between Chlorophyll \emph{a} and various Microcystin-LR
-Health Advisory Concentrations}
-\author{
-Jeffrey W. Hollister
-Betty J. Kreakie
-}
-\date{}
-
-\begin{document}
-%%Edited by @jhollist 09/15/2014
-%%Adds title from YAML
-\begin{singlespace}
-\begin{center}
-\huge Associations between Chlorophyll \emph{a} and various Microcystin-LR
-Health Advisory Concentrations
-\end{center}
-%%Adds Author, correspond email asterisk, and affilnum from YAML
-\begin{center}
-\large
-Jeffrey W. Hollister \textsuperscript{*} \textsuperscript{1} 
-Betty J. Kreakie \textsuperscript{1} 
-\end{center}
-%%Adds affiliations from YAML
-\begin{justify}
-\footnotesize \emph{ 
-\\*
-\textsuperscript{1}US Environmental Protection Agency, Office of Research and Development,
-National Health and Environmental Effects Research Laboratory, Atlantic
-Ecology Division, 27 Tarzwell Drive Narragansett, RI, 02882, USA\\*
-}
-%%Adds corresponding author email(s) from YAML
-\newcounter{num}
-\setcounter{num}{1}
-\\[0.1cm]
-\footnotesize \emph{ 
-\ifnum\value{num}=1%
-\textsuperscript{*} corresponding author:
-\fi
-\href{mailto:hollister.jeff@epa.gov}{\nolinkurl{hollister.jeff@epa.gov}}
-\stepcounter{num}
-}
-\end{justify}
-%%Adds date from YAML
-\normalsize
-
-\end{singlespace}
+<!--
+%\VignetteEngine{knitr::rmarkdown}
+%\VignetteIndexEntry{Microsystin Chlorophyll Manusript}
+-->
 
@@
 
-\vspace{2mm}
+\vspace{2mm}\hrule
 
-\hrule
+<!-- Abstract is being wrapped in latex here so that all analysis can be run in the chunk above and the results reproducibly referenced in the abstract. -->
+Cyanobacteria harmful algal blooms (cHABs) are associated with a wide range of adverse health effects that stem mostly from the presence of cyanotoxins.
+To help protect against these impacts, several health advisory levels have been set for some toxins.
+In particular, one of the more common toxins, microcystin, has several advisory levels set for drinking water and recreational use.
+However, compared to other water quality measures, field measurements of microcystin are not commonly available due to cost and advanced understanding required to interpret results.
+Addressing these issues will take time and resources.
+Thus, there is utility in finding indicators of microcystin that are already widely available, can be estimated quickly and *in situ*, and used as a first defense against high levels of microcystin.
+Chlorophyll *a* is commonly measured, can be estimated *in situ*, and has been shown to be positively associated with microcystin.
+In this paper, we use this association to provide estimates of chlorophyll *a* concentrations that are indicative of a higher probability of exceeding select health advisory concentrations for microcystin-LR.
+Using the 2007 National Lakes Assessment and a conditional probability approach, we identify chlorophyll *a* concentrations that are more likely than not to be associated with an exceedance of a microcystin health advisory level.
+We look at the recent US EPA health advisories for drinking water as well as the World Health Organization levels for drinking water and recreational use and identify a range of chlorophyll *a* thresholds.
+A 50% chance of exceeding one of the specific advisory microcystin concentrations of 0.3, 1, 1.6, and 2 $\mu$g/L is associatied with chlorophyll *a* concentration thresholds of 22.72, 77.2, 84.96, and 89.71, respectively.
+When managing for these various microcystin levels, exceeding these reported chlorophyll *a* concentrations should be a trigger for further testing and possible management action.
 
-Cyanobacteria harmful algal blooms (cHABs) are associated with a wide
-range of adverse health effects that stem mostly from the presence of
-cyanotoxins. To help protect against these impacts, several health
-advisory levels have been set for some toxins. In particular, one of the
-more common toxins, microcystin, has several advisory levels set for
-drinking water and recreational use. However, compared to other water
-quality measures, field measurements of microcystin are not commonly
-available due to cost and advanced understanding required to interpret
-results. Addressing these issues will take time and resources. Thus,
-there is utility in finding indicators of microcystin that are already
-widely available, can be estimated quickly and \emph{in situ}, and used
-as a first defense against high levels of microcystin. In particular,
-chlorophyll \emph{a} is very commonly measured, can be estimated
-\emph{in situ}, and has been shown to be positively associated with
-microcystin. In this paper, we use this association to provide estimates
-of chlorophyll \emph{a} concentrations that if exceeded would be
-indicative of a higher probability of exceeding select health advisory
-concentrations for microcystin-LR. Using the 2007 National Lakes
-Assessment and a conditional probability approach, we identify
-chlorophyll \emph{a} concentrations that are more likely than not to be
-associated with an exceedance of a microcystin health advisory level. We
-look at the recent US EPA health advisories for drinking water as well
-as the World Health Organization levels for drinking water and
-recreational use. For the specific advisory microcystin concentrations
-of 0.3, 1, 1.6, and 2, we find chlorophyll \emph{a} concentrations of
-0.07, 0.07, 2.79, and 11.36, respectively. When managing for these
-various microcystin levels, exceeding these reported chlorophyll
-\emph{a} concentrations should be a trigger for further testing and
-possible management action.
-
-\vspace{3mm}
-
-\hrule
-
+\vspace{3mm}\hrule
 \doublespace
 
-\section{Introduction}\label{introduction}
+#Introduction
 
-Over the last decade or so numerous events and legislative activities
-have raised the public awareness of harmful algal blooms (Rinta-Kanto et
-al. 2009, HABHRCA 2014, Jetoo et al. 2015), and in response the US
-Environmental Protection Agency (USEPA) has recently released suggested
-microcystin-LR (one of the more common toxins) concentrations that would
-trigger health advisories (McElhiney and Lawton 2005, Zurawell et al.
-2005, USEPA 2015). Additionally, the World Health Organization (WHO) has
-had proposed advisory levels for drinking water and a range of
-recreational risk levels (Chorus and Bartram 1999, Organization 2003).
-While these levels and associated advisories are likely to help mitigate
-the impacts from harmful algal blooms, they are not without
-complications.
+Over the last decade, numerous events and legislative activities have raised the public awareness of harmful algal blooms [@jetoo2015toledo; @rinta2009lake; @HABHRCA2014].
+In response the US Environmental Protection Agency (USEPA) has recently released suggested microcystin-LR (one of the more common toxins) concentrations that would trigger health advisories [@mcelhiney2005detection; @zurawell2005hepatotoxic; @usepa2015drinking].
+Additionally, the World Health Organization (WHO) has proposed microcystin advisory levels for drinking water and a range of recreational risk levels [@world2003cyanobacterial; @chorus1999toxic].
+While these levels and associated advisories are likely to help mitigate the impacts from harmful algal blooms, they are not without complications.
 
-One of these complications is that they rely on available measurements
-of microcystin-LR. While laboratory testing remains the gold standard
-for quantifying microcystin-LR concentrations in water samples, several
-field test kits have been developed. Even though field tests provide a
-much needed means for rapid assessment, they are not yet widely used and
-are moderately expensive (approximately \$150-\$200 depending on
-specific kit) with a limited shelf life (typically one year) (James et
-al. 2011, Aranda-Rodriguez et al. 2015). Additionally, each technique
-requires nuanced understanding of the detection method (e.g., limit of
-detection, specific microcystin variants being measured, and sampling
-protocol).
+One of these complications is that they rely on available measurements of microcystin-LR.
+While laboratory testing (e.g., chromatography) remains the gold standard for quantifying microcystin-LR concentrations in water samples, several field test kits have been developed.
+Even though field tests provide a much needed means for rapid assessment, they are not yet widely used and are moderately expensive (approximately $150-$200 depending on specific kit) with a limited shelf life (typically one year) [@james2011environmental; @aranda2015evaluation].
+Additionally, each technique requires nuanced understanding of the detection method (e.g., limit of detection, specific microcystin variants being measured, and sampling protocol).
 
-Fortunately, microcystin-LR has been shown to be associated with several
-other, more commonly measured and well understood components of water
-quality that are readily assessed in the field. For instance, there are
-small or hand held fluorometers that measure chlorohpyll \emph{a}.
-Additionally, chlorophyll \emph{a} is a very commonly measured component
-of water quality that is also known to be positively associated with
-microsystin-LR concentrations (Pip and Bowman 2014, Yuan et al. 2014).
-Yuan et. al (2014) explore these associations in detail and control for
-other related variables. In their analysis they find that total nitrogen
-and chlorophyll \emph{a} show the strongest association with
-microcystin. Furthermore, they identify chlorophyll\emph{a} and total
-nitrogen concentrations that are associated with exceeding 1 \(\mu\)g/L
-of microcystin. Given these facts, it should be possible to identify
-chlorophyll \emph{a} concentrations that would be associated with the
-new USEPA Microcystin-LR health advisory levels for drinking water.
-Identifying these associations would provide another tool for water
-resource managers to help manage the threat to public health posed by
-cHABs and would be especially useful in the absence of microcystin-LR
-concentrations.
+Fortunately, microcystin-LR has been shown to be associated with several other, more commonly measured and well understood components of water quality that are readily assessed in the field.
+For instance, there are small or hand held fluorometers that measure chlorohpyll *a*.
+Additionally, chlorophyll *a* is a very commonly measured component of water quality that is also known to be positively associated with microsystin-LR concentrations [@pip2014microcystin; @yuan2014managing].
+Yuan et.
+al [-@yuan2014managing] explore these associations in detail and control for other related variables.
+In their analysis they find that total nitrogen and chlorophyll *a* show the strongest association with microcystin.
+Furthermore, they identify chlorophyll*a* and total nitrogen concentrations that are associated with exceeding 1 $\mu$g/L of microcystin.
+Given these facts, it should be possible to identify chlorophyll *a* concentrations that would be associated with the new USEPA microcystin-LR health advisory levels for drinking water.
+Identifying these associations would provide another tool for water resource managers to help manage the threat to public health posed by cHABs and would be especially useful in the absence of measured microcystin-LR concentrations.
 
-In this paper we build on past efforts and utilize the National Lakes
-Assessment (NLA) data and identify chlorophyll \emph{a} concentrations
-that are associated with higher probabilities of exceeding several
-microcystin-LR health advisory concentrations (Chorus and Bartram 1999,
-USEPA 2009, 2015). We add to past studies by exploring associations with
-newly announced advisory levels and by also applying a different method,
-conditional probability analysis. Utilizing different methods
-strengthens the evidence for suggested chlorophyll \emph{a} levels that
-are associated with increased risk of exceeding the health advisory
-levels as those levels are not predicated on a single analytical method.
-So that others may repeat or adjust this analysis, the data, code, and
-this manuscript are freely available via
-\href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USAPE/microcystinchla}}.
+In this paper we build on past efforts and utilize the National Lakes Assessment (NLA) data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin-LR health advisory concentrations [@usepa2009national; @usepa2015drinking; @chorus1999toxic].
+We add to past studies by exploring associations with newly announced advisory levels and by also applying a different method, conditional probability analysis.
+Utilizing different methods strengthens the evidence for suggested chlorophyll *a* levels that are associated with increased risk of exceeding the health advisory levels as those levels are not predicated on a single analytical method.
+So that others may repeat or adjust this analysis, the data, code, and this manuscript are freely available via [https://github.com/USEPA/microcystinchla](https://github.com/USAPE/microcystinchla).
 
-\section{Methods}\label{methods}
+#Methods
 
-\subsection{Data}\label{data}
+##Data
+We used the 2007 NLA chlorophyll *a* and microcystin-LR concentration data [@usepa2009national, [NLA](http://www2.epa.gov/national-aquatic-resource-surveys/national-lakes-assessment)].
+These data represent a snapshot of water quality from the summer of 2007 for the conterminous United States and were collected as part of an ongoing probabilistic monitoring program [@usepa2009national].
+Data on chlorophyll *a* and microcystin-LR concentrations are available for  lakes.
 
-We used the 2007 NLA water quality and microcystin-LR concentration data
-(USEPA 2009). These data represent a snapshot of water quality from the
-summer of 2007 for the conterminous United States. Data on chlorophyll
-\emph{a} and microcystin-LR concentrations are available for lakes.
+##Analytical Methods
+We used a conditional probability analysis (CPA) approach to explore associations between chlorophyll *a* concentrations and World Health Organization (WHO) and USEPA microcystin-LR health advisory levels [@paul2011probability].
+Many health advisory levels have been suggested (Table \ref{tab:microcystin_levels}), but lakes with higher microcystin-LR concentrations in the NLA were rare.
+Only 1.16% of lakes sampled had a concentration greater than 10 $\mu$g/L.
+Thus, for this analysis we focused on the microcystin concentrations that are better represented in the NLA data.
+These were the USEPA children's drinking water advisory level of  0.3 $\mu$g/L (USEPA Child), the WHO drinking water advisory level of 1 $\mu$g/L (WHO Drinking), the USEPA adult drinking water advisory level of 1.6 $\mu$g/L (USEPA Adult), and the WHO recreational, low probability of effect advisory level of 2 $\mu$g/L (WHO Recreational).
 
-\subsection{Conditional Probability
-Analysis}\label{conditional-probability-analysis}
+Conditional probability analysis provides information about the probability of observing one event given another event has also occured.
+For this analysis, we used CPA to examine how the conditional probability of exceeding one of the health advisories changes as chlorophyll *a* increases in a lake.
+We expect to find higher chlorohpyll *a* concentrations to be associated with higher probabilities of exceeding the microcystin-LR health advisory levels.
+We also calculated bootstrapped 95% confidence intervals (CI) using 1000 bootstrapped samples.
+Thus, to identify chlorophyll *a* concentrations of concern we identfiy the value of the upper 95% CI across a range of conditional probabilities of exceeding each health advisory level.
+Using the upper confidence limit to identify a threshold is justified as it ensures that a given threshold is unlikely to miss a microcystin exceedance.
 
-We used a conditional probability analysis (CPA) approach to explore
-associations between chlorophyll \emph{a} concentrations and World
-Health Organization (WHO) and USEPA microcystin-LR health advisory
-levels (Paul and Munns 2011). Many levels have been suggested (Table
-\ref{tab:microcystin_levels}), but lakes with higher microcystin-LR
-concentrations in the NLA were rare. Only 1.16 \% of lakes sampled had a
-concentration greater than 10. Thus, for this analysis we focus on the
-microcystin concentrations that are better represented in the NLA data.
-These were the USEPA children's drinking water advisory level of 0.3
-\(\mu\)g/L (USEPA Child), the WHO drinking water advisory level of 1
-\(\mu\)g/L (WHO Drinking), the USEPA adult drinking water advisory level
-of 1.6 \(\mu\)g/L (USEPA Adult), and the WHO recreational, low
-probability of effect advisory level of 2 \(\mu\)g/L (WHO Recreational).
+As both microcystin-LR and chlorophyll *a* values were highely skewed right, a log base 10 transformation was used.
+Additional details of the specific implementation are available at https://github.com/USEPA/microcystinchla.
+A more detailed discussion of CPA is beyond the scope of this paper, but see Paul et al.
+[-@paul2005development] and Hollister et al.
+[-@hollister2008cprob] for greater detail.
+Lastly, all analyses were conducted using R version 3.2.2 and code and data from this analysis are freely available as an R package at [https://github.com/USEPA/microcystinchla](https://github.com/USAPE/microcystinchla).
 
-Conditional probability analysis provides information about the
-probability of observing one event given another event has also occured.
-For this analysis, we used CPA to examine how the conditional
-probability of exceeding one of the health advisories changes as
-chlorophyll \emph{a} increases in a lake. We expect to find higher
-chlorohpyll \emph{a} concentrations to be associated with higher
-probabilities of exceeding the microcystin-LR health advisory levels. We
-also caclulated bootstrapped 95\% confidence intervals (CI) using 1000
-bootstrapped samples. Thus, to identify chlorophyll \emph{a}
-concentrations of concern we identfiy the value of the upper 95\% CI
-across a range of conditional probabilities of exceeding each health
-advisory level. As both microcystin-LR and chlorophyll \emph{a} values
-were highely skewed right, a log base 10 transformation was used.
-Additional details of the specific implementation are available at
-\url{https://github.com/USEPA/microcystinchla}. A more detailed
-discussion of CPA is beyond the scope of this paper, but see Paul et al.
-(2005) and Hollister et al. (2008) for greater detail.
+Lastly, we assess the ability of these chlorophyll *a* thresholds to predict microcystin exceedance.
+We use error matrices and calculate total accuracy as well as the proportion of false negatives.
+Total accuracy is the total number of correct predictions divided by total observations.
+The proportion of false negatives is the total number of lakes that were predicted to not exceed the microcystin guidelines but actually did, divided by the total number of observations.
 
-\section{Results}\label{results}
+#Results
+In the 2007 NLA, microcystin-LR concentrations ranged from 0.05 to 225 $\mu$g/L.
+Microcystin-LR concentrations of 0.05 $\mu$g/L represent the detection limits.
+Any value greater than that indicates the presence of microcystin-LR.
+Of those lakes with microcystin, the median concentration was 0.51$\mu$g/L and the mean was 3.17$\mu$g/L.
+Of all lakes sampled, 21% of lakes exceeded the USEPA Child level, 8.8% of lakes exceeded the USEPA Adult level, 11.7% of lakes exceeded the WHO Drinking level, and 7.3% of lakes exceeded the WHO Recreational level.
+For chlorophyll *a*, the range was 0.07 to 936 $\mu$g/L.
+All lakes had reported chlorophyll *a* concentrations that exceeded detection limits.
+The median concentration was 7.79 $\mu$g/L and the mean was 29.63 $\mu$g/L.
+The associations between chlorophyll *a* and the upper confidence interval across a range of conditional probability values are shown in Table \ref{tab:mc_chla_table}.
+Specific chlorophyll *a* that are associated with greater than even odds of exceeding the advisory levels were 0.07, 0.07, 2.91, and 12.53$\mu$g/L for 0.3, 1.0, 1.6, and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
 
-In the 2007 NLA, microcystin-LR concentrations ranged from 0.05 to 225
-\(\mu\)g/L. Microcystin-LR concentrations of 0.05 \(\mu\)g/L represent
-the detection limits. Any value greater than that indicates the presence
-of microcystin-LR. Of those lakes with microcystin, the median
-concentration was 0.51 and the mean was 3.17. Of all lakes sampled, 21\%
-of lakes exceeded the USEPA Child level, 8.8\% of lakes exceeded the
-USEPA Adult level,11.7\% of lakes exceeded the WHO Drinking level,and
-7.3\% of lakes exceeded the WHO Recreational level. For chlorophyll
-\emph{a}, the range was 0.07 to 936 \(\mu\)g/L. All lakes had reported
-chlorophyll \emph{a} concentrations that exceeded detection limits. The
-median concentration was 7.79 \(\mu\)g/L and the mean was 29.63
-\(\mu\)g/L. The associations between chlorophyll \emph{a} and the upper
-confidence interval across a range of conditional probability values is
-shown in Table \ref{tab:mc_chla_table}. Specific chlorophyll \emph{a}
-that are associated with greater than even odds of exceeding the
-advisory levels were 0.07, 0.07, 2.79, and 11.36 for 0.3, 1.0, 1.6 and
-2.0 \(\mu\)g/L advisory levels, respectively (Table
-\ref{tab:mc_chla_table} \& Figure \ref{fig:multi_cp_plot}).
+The chlorophyll *a* cutoffs may be used to predict whether or not a lake exceeds the microcystin-LR health advisories.
+Doing so allows us to compare the accuracy of the prediction as well as evaluate false negatives.
+Total accuracy of the four cutoffs predicting microcystin-LR exceedances were 19% for the USEPA children's advisory, 22% for the WHO drinking water advisory,  22% for the USEPA adult advisory, and 23% for the WHO regreational advisory (Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table}, \ref{tab:adult_conmat_table}, & \ref{tab:who_rec_conmat_table}).
+However, total accuracy is only one part of the prediction performace with which we are concerned.
 
-\section{Discussion}\label{discussion}
+When using the chlorophyll *a* cutoffs as an indicator of microcystin-LR exceedances, the error that should be avoided is predicting that no exceedance has occurred when in fact it has.
+In other words, we would like to avoid Type II errors and minimize the proportion of false negatives.
+For the four chlorophyll *a* cut-offs we had a proportion of false negatives of 2%, 2%, 1%, and 1% for the USEPA children's, the WHO drinking water, the USEPA adult, and the WHO recreational advisories, respectively.
+In each case we missed less than 10% of the lakes that in fact exceeded the microcystin-LR advisory.
+While this method performs well with regard to the false negative percentage, it is possible that is a relic of the NLA dataset and testing with additional data would allow us to confirm this result.
 
-The association between Log10 microcystin-LR and Log10 chlorophyll
-\emph{a} shows a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
-This indicates that higher concentrations of microcystin-LR almost
-always co-occur with higher concentrations of chlorophyll \emph{a} yet
-the inverse is not true. Higher chlorophyll \emph{a} is not necessarily
-predictive of higher microcystin-LR concentrations; however, chlorophyll
-\emph{a} may be predictive of the probability of exceeding a certain
-concentration.
+#Discussion
+The association between Log10 microcystin-LR and Log10 chlorophyll *a* shows a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
+This indicates that, in general, higher concentrations of microcystin-LR almost always co-occur with higher concentrations of chlorophyll *a* yet the inverse is not true.
+Higher chlorophyll *a* is not necessarily predictive of higher microcystin-LR concentrations; however, chlorophyll *a* may be predictive of the probability of exceeding a certain threshold.
 
-This is the case as the probability of exceeding each of the four tested
-health advisory levels increases as a function of chlorophyll \emph{a}
-concentration (Figure \ref{fig:multi_cp_plot}). We use this association
-to identify chlorophyll \emph{a} concentrations that are associated with
-greater than even odds of exceeding a given health advisory level (Table
-\ref{tab:mc_chla_table}). These represent 99.9\%, 99.9\%, 77\%, and
-43.8\% of sample lakes for the USEPA Child, the WHO Drinking, the USEPA
-Adult, and the WHO recreational levels, respecitvely.
+This is the case as the probability of exceeding each of the four tested health advisory levels increases as a function of chlorophyll *a* concentration (Figure \ref{fig:multi_cp_plot}).
+We used this association to identify chlorophyll *a* concentrations that are associated with a range of probabilities of exceeding a given health advisory level (Table \ref{tab:mc_chla_table}).
+For the purposes of this discussion we focus on a conditional probability of 50% or greater (i.e., greater than even odds to exceed a health advisory level).
+The 50% conditional probability chlorophyll *a* thresholds represents 28.7%, 9.8%,
+8.6%, and 8.3% of sample lakes for the USEPA Child, the WHO Drinking, the USEPA Adult, and the WHO recreational levels, respecitvely.
 
-Furthermore, the chlorophyll \emph{a} cutoffs may be used to predict
-whether or not a lake exceeds the microcystin-LR health advisories.
-Doing so allows us to compare the accuracy of the prediction as well as
-evaluate false negatives. Total accuracy of the four cutoffs predicting
-microcystin-LR exceedances were 21\% for the USEPA children's advisory,
-12\% for the WHO drinking water advisory, 31\% for the USEPA adult
-advisory and, 61\% for the WHO regreational advisory (Tables
-\ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table},
-\ref{tab:adult_conmat_table}, \& \ref{tab:who_rec_conmat_table}).
-However, total accuracy is only one part of the prediction performace
-with which we are concerned.
+There are numerous possible uses for the chlorophyll *a* and microcystin-LR advisory cut-off values.
+First, in the absence of microcystin-LR measurements, exceedence of the chlorophyll *a* concentrations could be a trigger for further actions.
+Given that there is uncertainity around these chlorophyll *a* cutoffs the best case scenario would be to monitor for chlorophyll *a* and in the event of exceeding a target concentration take water samples and have those samples tested for microcystin-LR.
 
-When using the chlorophyll \emph{a} cutoffs as an indicator of
-microcystin-LR exceedances, the error that should be avoided is
-predicting that no exceedance has occurred when in fact it has. In other
-words, we would like to avoid Type II errors and minimize the proportion
-of false negatives. For the four chlorophyll \emph{a} cut-offs we had a
-proportion of false negatives of 0\%, 0\%, 0\% and , 1\% for the U.S EPA
-childrens drinking water, the WHO drinking water, the USEPA adult
-drinking water, and the WHO recreational advisories, respectively. In
-each case we miss less than 10\% of the lakes that are in fact exceeding
-the microcystin-LR advisory.
+A second potential use is to identify past bloom events from historical data.
+As harmful algal blooms are made up of many species and have various mechanisms responsible for adverse impacts (e.g., toxins, hypoxia, odors), there is no single definition of a bloom.
+For cHABs, one approach has been to identify an increase over a baseline concentration of phycocyanin [@miller2013spatiotemporal].
+This is a useful approach for targeted studies, but phycocyanin is also not always available and measures the predominance of cyanobacterial pigments and not toxins.
+Using our chlorophyll *a* cutoffs provides a value that is more directly associated with microcystin-LR and can be used to classify lakes, from past surveys, as having bloomed.
 
-There are numerous possible uses for the chlorophyll \emph{a} and
-microcystin-LR advisory cut-off values. First, in the absence of
-microcystin-LR measurements, exceedence of the chlorophyll \emph{a}
-concentrations could be a trigger for further actions. Given that there
-is uncertainity around these chlorophyll \emph{a} cutoffs the best case
-scenario would be to monitor for chlorophyll \emph{a} and in the event
-of exceeding a target concentration take water samples and have those
-samples tested in a lab for microcystin-LR.
+Lastly, using chlorophyll *a* is not meant as a replacement for testing of microcystin-LR or other toxins.
+It should be used when other, direct measurements of cyanotoxins are not available.
+In those cases, which are likely to be common at least in the near future, using a more ubiquitous measurement, such as chlorophyll *a* will provide a reasonable proxy for the probability of exceeding a microcystin-LR health advisory level and provide better protection against adverse effects in both drinking and recreational use cases.
 
-A second potential use is to identify possible bloom events from
-historical data. As harmful algal blooms are made up of many species and
-have various mechanisms responsible for adverse impacts (e.g.~toxins,
-hypoxia, odors), there is no single definition of a bloom. For cHABs one
-approach has been to identify an increase over a baseline concentration
-of phycocyanin (Miller et al. 2013). This is a useful approach for
-targeted studies, but phycocyanin is also not always available and
-measures the predominance of cyanobacterial pigments and not toxins.
-Using our chlorophyll \emph{a} cutoffs provides a value that is more
-directly associated with microcystin-LR and can be used to classify
-lakes, from past surveys, as having bloomed.
-
-Lastly, using chlorophyll \emph{a} is not meant as a replacement for
-testing of microcystin-LR. It should be used when other, direct
-measurements of cyanotoxins are not available. In those cases, which are
-likely to be common at least in the near future, using a more ubiquitous
-measurement, such as chlorophyll \emph{a} will provide a reasonable
-proxy for the probability of exceeding a microcystin-LR health advisory
-level and provide better protection against adverse effects in both
-drinking and recreational use cases.
-
-\section{Acknowledgements}\label{acknowledgements}
+#Acknowledgements
+We would like to thank Anne Kuhn, Bryan Milstead, John Kiddon, Joe LiVolsi, Tim Gleason, and Wayne Munns for constructive reviews of this paper.
+This paper has not been subjected to Agency review.
+Therefore, it does not necessary reflect the views of the Agency.
+Mention of trade names or commercial products does not constitute endorsement or recommendation for use.
+This contribution is identified by the tracking number ORD-015143 of the Atlantic Ecology Division, Office of Research and Development, National Health and Environmental Effects Research Laboratory, US Environmental Protection Agency.
 
@@
 
-\section{Figures}\label{figures}
+#Figures
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/chla_micro_scatter-1.pdf}
-\caption{Scatterplot showing association betweeen chlorophyll \textit{a}
-and microcystin-LR. \label{fig:chla_micro_scatter}}
-\end{figure}
+![Scatterplot showing association betweeen chlorophyll \textit{a} and microcystin-LR. \label{fig:chla_micro_scatter}](figure/chla_micro_scatter-1.pdf)
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/epa_child_cp_plot-1.pdf}
-\caption{Conditional probability plots showing association between the
-probability of exceeding various microcystin-LR (MLR) health advisory
-Levels. A.) Plot for USEPA Child (0.3 \(\mu\)g/L). B.) Plot for WHO
-Drinking (1 \(\mu\)g/L). C.) Plot for USEPA Adult (1.6 \(\mu\)g/L). D.)
-Plot for WHO Recreational (2 \(\mu\)g/L). \label{fig:multi_cp_plot}}
-\end{figure}
+<!--
+
+```
+## No id variables; using all as measure variables
+```
+
+![Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels. \label{fig:chla_micro_cdf}](figure/chla_micro_cdf-1.pdf)
+
+\newpage-->
+
+```
+## Loading required package: grid
+```
+
+![Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
+A.) Plot for USEPA Child (0.3 $\mu$g/L).
+B.) Plot for WHO Drinking (1 $\mu$g/L).
+C.) Plot for USEPA Adult (1.6 $\mu$g/L).
+D.) Plot for WHO Recreational (2 $\mu$g/L).
+\label{fig:multi_cp_plot}](figure/epa_child_cp_plot-1.pdf)
 
@@
 
-\section{Tables}\label{tables}
+#Tables
 
-\begin{longtable}[c]{@{}lll@{}}
-\caption{Various suggested microcystin-LR health advisory
-concentrations. \label{tab:microcystin_levels}}\tabularnewline
-\toprule
-Source & Type & Concentration\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Source & Type & Concentration\tabularnewline
-\midrule
-\endhead
-USEPA & Adult Drinking Water Advisory & 1.6 \(\mu\)g/L\tabularnewline
-USEPA & Child Drinking Water Advisory & 0.3 \(\mu\)g/L\tabularnewline
-WHO & Drinking Water & 1 \(\mu\)g/L\tabularnewline
-WHO & Recreational: High Prob. of Effect & 20-2000
-\(\mu\)g/L\tabularnewline
-WHO & Recreational: Low Prob. of Effect & 2-4 \(\mu\)g/L\tabularnewline
-WHO & Recreational: Moderate Prob. of Effect & 10-20
-\(\mu\)g/L\tabularnewline
-WHO & Recreational: Very High Prob. of Effect & \textgreater{}2000
-\(\mu\)g/L\tabularnewline
-\bottomrule
-\end{longtable}
+|Source |Type                                    |Concentration    |
+|:------|:---------------------------------------|:----------------|
+|USEPA  |Adult Drinking Water Advisory           |1.6 $\mu$g/L     |
+|USEPA  |Child Drinking Water Advisory           |0.3 $\mu$g/L     |
+|WHO    |Drinking Water                          |1 $\mu$g/L       |
+|WHO    |Recreational: High Prob. of Effect      |20-2000 $\mu$g/L |
+|WHO    |Recreational: Low Prob. of Effect       |2-4 $\mu$g/L     |
+|WHO    |Recreational: Moderate Prob. of Effect  |10-20 $\mu$g/L   |
+|WHO    |Recreational: Very High Prob. of Effect |>2000 $\mu$g/L   |
+
+:Various suggested microcystin-LR health advisory concentrations. \label{tab:microcystin_levels}
 
@@
 
-\begin{longtable}[c]{@{}rrrrr@{}}
-\caption{Chlorophyll \textit{a} concentrations that are associated with
-a 50\% probability of exceeding a microcystin-LR health advisory
-concentration. \label{tab:mc_chla_table}}\tabularnewline
-\toprule
-Cond. Probability & USEPA Child & WHO Drink & USEPA Adult & WHO
-Recreational\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Cond. Probability & USEPA Child & WHO Drink & USEPA Adult & WHO
-Recreational\tabularnewline
-\midrule
-\endhead
-0.1 & 0.07 & 0.07 & 0.07 & 1.47\tabularnewline
-0.2 & 0.07 & 4.42 & 11.47 & 19.04\tabularnewline
-0.3 & 2.79 & 16.32 & 30.62 & 52.70\tabularnewline
-0.4 & 11.36 & 38.30 & 69.19 & 82.22\tabularnewline
-0.5 & 23.68 & 65.20 & 84.96 & 103.90\tabularnewline
-0.6 & 38.30 & 98.46 & 136.66 & 125.40\tabularnewline
-0.7 & 68.85 & 133.20 & 871.20 & 871.20\tabularnewline
-0.8 & 114.62 & 338.40 & 871.20 & 871.20\tabularnewline
-0.9 & 198.72 & 516.00 & 871.20 & 871.20\tabularnewline
-\bottomrule
-\end{longtable}
+| Cond. Probability| USEPA Child| WHO Drink| USEPA Adult| WHO Recreational|
+|-----------------:|-----------:|---------:|-----------:|----------------:|
+|               0.1|        0.07|      0.07|        0.07|             0.86|
+|               0.2|        0.07|      4.42|       11.01|            14.64|
+|               0.3|        2.91|     19.36|       31.59|            43.63|
+|               0.4|       12.53|     38.74|       66.96|            68.40|
+|               0.5|       22.72|     77.20|       84.96|            89.71|
+|               0.6|       41.60|    103.20|      125.40|           203.76|
+|               0.7|       65.60|    135.61|      318.24|           871.20|
+|               0.8|      113.14|    271.44|      871.20|           871.20|
+|               0.9|      167.04|    516.00|      871.20|           871.20|
+
+:Chlorophyll \textit{a} concentrations that are associated with a 50% probability of exceeding a microcystin-LR health advisory concentration. \label{tab:mc_chla_table}
 
@@
 
-\begin{longtable}[c]{@{}lrr@{}}
-\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
-exceedences (rows) versus real exceedances (columns) for the USEPA
-childrens drinking water advisory.
-\label{tab:child_conmat_table}}\tabularnewline
-\toprule
-& FALSE & TRUE\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-& FALSE & TRUE\tabularnewline
-\midrule
-\endhead
-FALSE & 1 & 0\tabularnewline
-TRUE & 810 & 217\tabularnewline
-\bottomrule
-\end{longtable}
+|              | Not Exceed| Exceed| Row Totals|
+|:-------------|----------:|------:|----------:|
+|Not Exceed    |        642|     95|        737|
+|Exceed        |        169|    122|        291|
+|Column Totals |        811|    217|       1028|
+
+:Confusion matrix comparing chlorophyll \textit{a} predicted exceedences (rows) versus real exceedances (columns) for the USEPA childrens drinking water advisory. \label{tab:child_conmat_table}
 
@@
 
-\begin{longtable}[c]{@{}lrr@{}}
-\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
-exceedences (rows) versus real exceedances (columns) for the WHO
-drinking water advisory.
-\label{tab:who_drink_conmat_table}}\tabularnewline
-\toprule
-& FALSE & TRUE\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-& FALSE & TRUE\tabularnewline
-\midrule
-\endhead
-FALSE & 1 & 0\tabularnewline
-TRUE & 906 & 121\tabularnewline
-\bottomrule
-\end{longtable}
+|              | Not Exceed| Exceed| Row Totals|
+|:-------------|----------:|------:|----------:|
+|Not Exceed    |        850|     82|        932|
+|Exceed        |         57|     39|         96|
+|Column Totals |        907|    121|       1028|
+
+:Confusion matrix comparing chlorophyll \textit{a} predicted exceedences (rows) versus real exceedances (columns) for the WHO drinking water advisory. \label{tab:who_drink_conmat_table}
 
@@
 
-\begin{longtable}[c]{@{}lrr@{}}
-\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
-exceedences (rows) versus real exceedances (columns) for the USEPA adult
-drinking water advisory. \label{tab:adult_conmat_table}}\tabularnewline
-\toprule
-& FALSE & TRUE\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-& FALSE & TRUE\tabularnewline
-\midrule
-\endhead
-FALSE & 234 & 4\tabularnewline
-TRUE & 703 & 87\tabularnewline
-\bottomrule
-\end{longtable}
+|              | Not Exceed| Exceed| Row Totals|
+|:-------------|----------:|------:|----------:|
+|Not Exceed    |        887|     57|        944|
+|Exceed        |         50|     34|         84|
+|Column Totals |        937|     91|       1028|
+
+:Confusion matrix comparing chlorophyll \textit{a} predicted exceedences (rows) versus real exceedances (columns) for the USEPA adult drinking water advisory. \label{tab:adult_conmat_table}
 
@@
 
-\begin{longtable}[c]{@{}lrr@{}}
-\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
-exceedences (rows) versus real exceedances (columns) for the WHO
-recreational water advisory.
-\label{tab:who_rec_conmat_table}}\tabularnewline
-\toprule
-& FALSE & TRUE\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-& FALSE & TRUE\tabularnewline
-\midrule
-\endhead
-FALSE & 566 & 15\tabularnewline
-TRUE & 387 & 60\tabularnewline
-\bottomrule
-\end{longtable}
+|              | Not Exceed| Exceed| Row Totals|
+|:-------------|----------:|------:|----------:|
+|Not Exceed    |        900|     47|        947|
+|Exceed        |         53|     28|         81|
+|Column Totals |        953|     75|       1028|
+
+:Confusion matrix comparing chlorophyll \textit{a} predicted exceedences (rows) versus real exceedances (columns) for the WHO recreational water advisory. \label{tab:who_rec_conmat_table}
 
@@
 
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
-
-Aranda-Rodriguez, R., Z. Jin, J. Harvie, and A. Cabecinha. 2015.
-Evaluation of three field test kits to detect microcystins from a public
-health perspective. Harmful Algae 42:34--42.
-
-Chorus, E. I., and J. Bartram. 1999. Toxic cyanobacteria in water: A
-guide to their public health consequences, monitoring and management.
-
-HABHRCA. 2014. Harmful algal bloom and hypoxia research and control
-amendments act of 2014.
-
-Hollister, J. W., H. A. Walker, and J. F. Paul. 2008. CProb: A
-computational tool for conducting conditional probability analysis.
-Journal of environmental quality 37:2392--2396.
-
-James, R., A. Gregg, A. Dindal, and J. McKernan. 2011. Environmental
-technology verification report: Abraxis microcystin test kits. Online
-document. Accessed online: June 22.
-
-Jetoo, S., V. I. Grover, and G. Krantzberg. 2015. The toledo drinking
-water advisory: Suggested application of the water safety planning
-approach. Sustainability 7:9787--9808.
-
-McElhiney, J., and L. A. Lawton. 2005. Detection of the cyanobacterial
-hepatotoxins microcystins. Toxicology and Applied Pharmacology
-203:219--230.
-
-Miller, T. R., L. Beversdorf, S. D. Chaston, and K. D. McMahon. 2013.
-Spatiotemporal molecular analysis of cyanobacteria blooms reveals
-microcystis-aphanizomenon interactions. PloS one 8:e74933.
-
-Organization, W. H. 2003. Cyanobacterial toxins: Microcystin-lR in
-drinking-water. background document for development of wHO guidelines
-for drinking-water quality. geneva, switzerland. World Health
-Organization, 2nd ed. Geneva.
-
-Paul, J. F., and M. E. McDonald. 2005. Development of empirical,
-geographically specific water quality criteria: A conditional
-probability analysis approach 41:1211--1223.
-
-Paul, J. F., and W. R. Munns. 2011. Probability surveys, conditional
-probability, and ecological risk assessment. Environmental Toxicology
-and Chemistry 30:1488--1495.
-
-Pip, E., and L. Bowman. 2014. Microcystin and algal chlorophyll in
-relation to nearshore nutrient concentrations in lake winnipeg, canada.
-Environment and Pollution 3:p36.
-
-Rinta-Kanto, J. M., E. A. Konopko, J. M. DeBruyn, R. A. Bourbonniere, G.
-L. Boyer, and S. W. Wilhelm. 2009. Lake erie microcystis: Relationship
-between microcystin production, dynamics of genotypes and environmental
-parameters in a large lake. Harmful Algae 8:665--673.
-
-USEPA. 2009. National lakes assessment: A collaborative survey of the
-nation's lakes. ePA 841-r-09-001. Office of Water; Office of Research;
-Development, US Environmental Protection Agency Washington, DC.
-
-USEPA. 2015. Drinking water health advisory for the cyanobacterial
-microcystin toxins. ePA-820R15100. Office of Water, US Environmental
-Protection Agency Washington, DC.
-
-Yuan, L. L., A. I. Pollard, S. Pather, J. L. Oliver, and L. D'Anglada.
-2014. Managing microcystin: Identifying national-scale thresholds for
-total nitrogen and chlorophyll a. Freshwater Biology 59:1970--1981.
-
-Zurawell, R. W., H. Chen, J. M. Burke, and E. E. Prepas. 2005.
-Hepatotoxic cyanobacteria: A review of the biological importance of
-microcystins in freshwater environments. Journal of Toxicology and
-Environmental Health, Part B 8:1--37.
-
-\end{document}
+#References

### 2016-01-17 manuscript/manuscript.Rmd
working on f1000 format

@@
 capsize: normalsize
-csl: components/ecology.csl
+csl: components/plos.csl
 documentclass: article

### 2016-01-17 manuscript/manuscript.tex
working on f1000 format

--- 2015-12-04 manuscript/manuscript.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Original default rstudio/pandoc latex file
-%% upated by @jhollist 09/15/2014
-%% inspired by @cboetting https://github.com/cboettig/template and
-%% @rmflight blog posts:
-%% http://rmflight.github.io/posts/2014/07/analyses_as_packages.html 
-%% http://rmflight.github.io/posts/2014/07/vignetteAnalysis.html).  
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% For more detailed article preparation guidelines, please see:
+% http://f1000research.com/author-guidelines
 
-\documentclass[11pt,]{article}
-\usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-    \usepackage{xltxtra,xunicode}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-% use microtype if available
-\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\usepackage{longtable,booktabs}
-\usepackage{graphicx}
-% Redefine \includegraphics so that, unless explicit options are
-% given, the image width will not exceed the width of the page.
-% Images get their normal width if they fit onto the page, but
-% are scaled down if they would overflow the margins.
-\makeatletter
-\def\ScaleIfNeeded{%
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-\let\Oldincludegraphics\includegraphics
-{%
- \catcode`\@=11\relax%
- \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
-}%
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={Associations between Chlorophyll a and various Microcystin-LR Health Advisory Concentrations},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\setcounter{secnumdepth}{5}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%Changes borrowed from @cboettig, added by @jhollist 
-% A modified page layout 
-\textwidth 6.75in
-\oddsidemargin -0.15in
-\evensidemargin -0.15in
-\textheight 9in
-\topmargin -0.5in
-\usepackage{lineno} % add 
-  \linenumbers % turns line numbering on 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%Packages and layout changes by @jhollist 09/15/2014
-\usepackage{ragged2e}
-\usepackage[font=normalsize]{caption}
-  \usepackage[doublespacing]{setspace}
-\usepackage{parskip}
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\fancyhf{}
-\renewcommand{\headrulewidth}{0pt}
-\rfoot{\today}
-\lfoot{\thepage}
-%%Changed default abstract width and added lines
-\renewenvironment{abstract}{
-  \hfill\begin{minipage}{1\textwidth}
-  \rule{\textwidth}{1pt}\vspace{5pt}
-  \normalsize
-  \begin{justify}
-  \bfseries\abstractname\vspace{5pt}
-  \end{justify}}
-  {\par\noindent\rule{\textwidth}{1pt}\end{minipage}
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\title{Associations between Chlorophyll \emph{a} and various Microcystin-LR
-Health Advisory Concentrations}
-\author{
-Jeffrey W. Hollister
-Betty J. Kreakie
-}
-\date{}
+\documentclass[10pt,a4paper,twocolumn]{article}
+\usepackage{f1000_styles}
+\bibliography{components/manuscript}
 
 \begin{document}
-%%Edited by @jhollist 09/15/2014
-%%Adds title from YAML
-\begin{singlespace}
-\begin{center}
-\huge Associations between Chlorophyll \emph{a} and various Microcystin-LR
-Health Advisory Concentrations
-\end{center}
-%%Adds Author, correspond email asterisk, and affilnum from YAML
-\begin{center}
-\large
-Jeffrey W. Hollister \textsuperscript{*} \textsuperscript{1} 
-Betty J. Kreakie \textsuperscript{1} 
-\end{center}
-%%Adds affiliations from YAML
-\begin{justify}
-\footnotesize \emph{ 
-\\*
-\textsuperscript{1}US Environmental Protection Agency, Office of Research and Development,
-National Health and Environmental Effects Research Laboratory, Atlantic
-Ecology Division, 27 Tarzwell Drive Narragansett, RI, 02882, USA\\*
-}
-%%Adds corresponding author email(s) from YAML
-\newcounter{num}
-\setcounter{num}{1}
-\\[0.1cm]
-\footnotesize \emph{ 
-\ifnum\value{num}=1%
-\textsuperscript{*} corresponding author:
-\fi
-\href{mailto:hollister.jeff@epa.gov}{\nolinkurl{hollister.jeff@epa.gov}}
-\stepcounter{num}
-}
-\end{justify}
-%%Adds date from YAML
-\normalsize
 
-\end{singlespace}
+%Still need to figure out how to get corresponding author on here.
+\title{\textit{F1000Research} Associations between chlorophyll \emph{a} and various microcystin-LR
+health advisory Concentrations}
+\author[1]{Jeffrey W. Hollister}
+\author[1]{Betty J. Kreakie}
+\affil[1]{}
+
+\maketitle
+\thispagestyle{fancy}
+
+\begin{abstract}
+
+\end{abstract}
+
+\clearpage
 
@@
 microcystin concentrations of 0.3, 1, 1.6, and 2 \(\mu\)g/L is
-associatied with chlorophyll \emph{a} concentration thresholds of 20.88,
-69.6, 84.24, and 97.49, respectively. When managing for these various
+associatied with chlorophyll \emph{a} concentration thresholds of 21.92,
+77.2, 84.96, and 105.84, respectively. When managing for these various
 microcystin levels, exceeding these reported chlorophyll \emph{a}
@@
 Over the last decade, numerous events and legislative activities have
-raised the public awareness of harmful algal blooms (Rinta-Kanto et al.
-2009, HABHRCA 2014, Jetoo et al. 2015). In response the US Environmental
-Protection Agency (USEPA) has recently released suggested microcystin-LR
-(one of the more common toxins) concentrations that would trigger health
-advisories (McElhiney and Lawton 2005, Zurawell et al. 2005, USEPA
-2015). Additionally, the World Health Organization (WHO) has proposed
+raised the public awareness of harmful algal blooms {[}1--3{]}. In
+response the US Environmental Protection Agency (USEPA) has recently
+released suggested microcystin-LR (one of the more common toxins)
+concentrations that would trigger health advisories {[}4--6{]}.
+Additionally, the World Health Organization (WHO) has proposed
 microcystin advisory levels for drinking water and a range of
-recreational risk levels (Chorus and Bartram 1999, Organization 2003).
-While these levels and associated advisories are likely to help mitigate
-the impacts from harmful algal blooms, they are not without
-complications.
+recreational risk levels {[}7,8{]}. While these levels and associated
+advisories are likely to help mitigate the impacts from harmful algal
+blooms, they are not without complications.
 
@@
 \$150-\$200 depending on specific kit) with a limited shelf life
-(typically one year) (James et al. 2011, Aranda-Rodriguez et al. 2015).
-Additionally, each technique requires nuanced understanding of the
-detection method (e.g., limit of detection, specific microcystin
-variants being measured, and sampling protocol).
+(typically one year) {[}9,10{]}. Additionally, each technique requires
+nuanced understanding of the detection method (e.g., limit of detection,
+specific microcystin variants being measured, and sampling protocol).
 
@@
 of water quality that is also known to be positively associated with
-microsystin-LR concentrations (Pip and Bowman 2014, Yuan et al. 2014).
-Yuan et. al (2014) explore these associations in detail and control for
-other related variables. In their analysis they find that total nitrogen
-and chlorophyll \emph{a} show the strongest association with
-microcystin. Furthermore, they identify chlorophyll\emph{a} and total
-nitrogen concentrations that are associated with exceeding 1 \(\mu\)g/L
-of microcystin. Given these facts, it should be possible to identify
-chlorophyll \emph{a} concentrations that would be associated with the
-new USEPA Microcystin-LR health advisory levels for drinking water.
-Identifying these associations would provide another tool for water
-resource managers to help manage the threat to public health posed by
-cHABs and would be especially useful in the absence of microcystin-LR
+microsystin-LR concentrations {[}11,12{]}. Yuan et. al {[}12{]} explore
+these associations in detail and control for other related variables. In
+their analysis they find that total nitrogen and chlorophyll \emph{a}
+show the strongest association with microcystin. Furthermore, they
+identify chlorophyll\emph{a} and total nitrogen concentrations that are
+associated with exceeding 1 \(\mu\)g/L of microcystin. Given these
+facts, it should be possible to identify chlorophyll \emph{a}
+concentrations that would be associated with the new USEPA
+microcystin-LR health advisory levels for drinking water. Identifying
+these associations would provide another tool for water resource
+managers to help manage the threat to public health posed by cHABs and
+would be especially useful in the absence of measured microcystin-LR
 concentrations.
@@
 that are associated with higher probabilities of exceeding several
-microcystin-LR health advisory concentrations (Chorus and Bartram 1999,
-USEPA 2009, 2015). We add to past studies by exploring associations with
-newly announced advisory levels and by also applying a different method,
-conditional probability analysis. Utilizing different methods
-strengthens the evidence for suggested chlorophyll \emph{a} levels that
-are associated with increased risk of exceeding the health advisory
-levels as those levels are not predicated on a single analytical method.
-So that others may repeat or adjust this analysis, the data, code, and
-this manuscript are freely available via
-\href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USAPE/microcystinchla}}.
+microcystin-LR health advisory concentrations {[}6,8,13{]}. We add to
+past studies by exploring associations with newly announced advisory
+levels and by also applying a different method, conditional probability
+analysis. Utilizing different methods strengthens the evidence for
+suggested chlorophyll \emph{a} levels that are associated with increased
+risk of exceeding the health advisory levels as those levels are not
+predicated on a single analytical method. So that others may repeat or
+adjust this analysis, the data, code, and this manuscript are freely
+available via
+\href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USEPA/microcystinchla}}.
 
@@
 We used the 2007 NLA chlorophyll \emph{a} and microcystin-LR
-concentration data (USEPA 2009). These data represent a snapshot of
-water quality from the summer of 2007 for the conterminous United States
-and were collected as part of an ongoing probabilistic monitoring
-program (USEPA 2009). Data on chlorophyll \emph{a} and microcystin-LR
-concentrations are available for lakes.
+concentration data {[}13{]}. These data represent a snapshot of water
+quality from the summer of 2007 for the conterminous United States and
+were collected as part of an ongoing probabilistic monitoring program
+{[}13{]}. Data on chlorophyll \emph{a} and microcystin-LR concentrations
+are available for lakes.
 
-\subsection{Conditional Probability
-Analysis}\label{conditional-probability-analysis}
+\subsection{Analytical Methods}\label{analytical-methods}
 
@@
 Health Organization (WHO) and USEPA microcystin-LR health advisory
-levels (Paul and Munns 2011). Many levels have been suggested (Table
+levels {[}14{]}. Many health advisory levels have been suggested (Table
 \ref{tab:microcystin_levels}), but lakes with higher microcystin-LR
-concentrations in the NLA were rare. Only 1.16 \% of lakes sampled had a
-concentration greater than 10. Thus, for this analysis we focus on the
-microcystin concentrations that are better represented in the NLA data.
-These were the USEPA children's drinking water advisory level of 0.3
-\(\mu\)g/L (USEPA Child), the WHO drinking water advisory level of 1
-\(\mu\)g/L (WHO Drinking), the USEPA adult drinking water advisory level
-of 1.6 \(\mu\)g/L (USEPA Adult), and the WHO recreational, low
-probability of effect advisory level of 2 \(\mu\)g/L (WHO Recreational).
+concentrations in the NLA were rare. Only 1.16\% of lakes sampled had a
+concentration greater than 10 \(\mu\)g/L. Thus, for this analysis we
+focused on the microcystin concentrations that are better represented in
+the NLA data. These were the USEPA children's drinking water advisory
+level of 0.3 \(\mu\)g/L (USEPA Child), the WHO drinking water advisory
+level of 1 \(\mu\)g/L (WHO Drinking), the USEPA adult drinking water
+advisory level of 1.6 \(\mu\)g/L (USEPA Adult), and the WHO
+recreational, low probability of effect advisory level of 2 \(\mu\)g/L
+(WHO Recreational).
 
@@
 probabilities of exceeding the microcystin-LR health advisory levels. We
-also caclulated bootstrapped 95\% confidence intervals (CI) using 1000
+also calculated bootstrapped 95\% confidence intervals (CI) using 1000
 bootstrapped samples. Thus, to identify chlorophyll \emph{a}
@@
 across a range of conditional probabilities of exceeding each health
-advisory level. As both microcystin-LR and chlorophyll \emph{a} values
-were highely skewed right, a log base 10 transformation was used.
-Additional details of the specific implementation are available at
+advisory level. Using the upper confidence limit to identify a threshold
+is justified as it ensures that a given threshold is unlikely to miss a
+microcystin exceedance.
+
+As both microcystin-LR and chlorophyll \emph{a} values were highely
+skewed right, a log base 10 transformation was used. Additional details
+of the specific implementation are available at
 \url{https://github.com/USEPA/microcystinchla}. A more detailed
 discussion of CPA is beyond the scope of this paper, but see Paul et al.
-(2005) and Hollister et al. (2008) for greater detail. Lastly, all
+{[}15{]} and Hollister et al. {[}16{]} for greater detail. Lastly, all
 analyses were conducted using R version 3.2.2 and code and data from
 this analysis are freely available as an R package at
-\href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USAPE/microcystinchla}}.
+\href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USEPA/microcystinchla}}.
+
+Lastly, we assess the ability of these chlorophyll \emph{a} thresholds
+to predict microcystin exceedance. We use error matrices and calculate
+total accuracy as well as the proportion of false negatives. Total
+accuracy is the total number of correct predictions divided by total
+observations. The proportion of false negatives is the total number of
+lakes that were predicted to not exceed the microcystin guidelines but
+actually did, divided by the total number of observations.
 
@@
 of microcystin-LR. Of those lakes with microcystin, the median
-concentration was 0.51 and the mean was 3.17. Of all lakes sampled, 21\%
-of lakes exceeded the USEPA Child level, 8.8\% of lakes exceeded the
-USEPA Adult level,11.7\% of lakes exceeded the WHO Drinking level,and
-7.3\% of lakes exceeded the WHO Recreational level. For chlorophyll
-\emph{a}, the range was 0.07 to 936 \(\mu\)g/L. All lakes had reported
-chlorophyll \emph{a} concentrations that exceeded detection limits. The
-median concentration was 7.79 \(\mu\)g/L and the mean was 29.63
-\(\mu\)g/L. The associations between chlorophyll \emph{a} and the upper
-confidence interval across a range of conditional probability values is
-shown in Table \ref{tab:mc_chla_table}. Specific chlorophyll \emph{a}
-that are associated with greater than even odds of exceeding the
-advisory levels were 0.07, 0.07, 2.89, and 10.27 for 0.3, 1.0, 1.6 and
-2.0 \(\mu\)g/L advisory levels, respectively (Table
-\ref{tab:mc_chla_table} \& Figure \ref{fig:multi_cp_plot}).
+concentration was 0.51\(\mu\)g/L and the mean was 3.17\(\mu\)g/L. Of all
+lakes sampled, 21\% of lakes exceeded the USEPA Child level, 8.8\% of
+lakes exceeded the USEPA Adult level, 11.7\% of lakes exceeded the WHO
+Drinking level, and 7.3\% of lakes exceeded the WHO Recreational level.
+For chlorophyll \emph{a}, the range was 0.07 to 936 \(\mu\)g/L. All
+lakes had reported chlorophyll \emph{a} concentrations that exceeded
+detection limits. The median concentration was 7.79 \(\mu\)g/L and the
+mean was 29.63 \(\mu\)g/L. The associations between chlorophyll \emph{a}
+and the upper confidence interval across a range of conditional
+probability values are shown in Table \ref{tab:mc_chla_table}. Specific
+chlorophyll \emph{a} that are associated with greater than even odds of
+exceeding the advisory levels were 0.07, 0.07, 2.78, and 9.97\(\mu\)g/L
+for 0.3, 1.0, 1.6, and 2.0 \(\mu\)g/L advisory levels, respectively
+(Table \ref{tab:mc_chla_table} \& Figure \ref{fig:multi_cp_plot}).
+
+The chlorophyll \emph{a} cutoffs may be used to predict whether or not a
+lake exceeds the microcystin-LR health advisories. Doing so allows us to
+compare the accuracy of the prediction as well as evaluate false
+negatives. Total accuracy of the four cutoffs predicting microcystin-LR
+exceedances were 19\% for the USEPA children's advisory, 22\% for the
+WHO drinking water advisory, 22\% for the USEPA adult advisory, and 23\%
+for the WHO regreational advisory (Tables \ref{tab:child_conmat_table},
+\ref{tab:who_drink_conmat_table}, \ref{tab:adult_conmat_table}, \&
+\ref{tab:who_rec_conmat_table}). However, total accuracy is only one
+part of the prediction performace with which we are concerned.
+
+When using the chlorophyll \emph{a} cutoffs as an indicator of
+microcystin-LR exceedances, the error that should be avoided is
+predicting that no exceedance has occurred when in fact it has. In other
+words, we would like to avoid Type II errors and minimize the proportion
+of false negatives. For the four chlorophyll \emph{a} cut-offs we had a
+proportion of false negatives of 2\%, 2\%, 1\%, and 1\% for the USEPA
+children's, the WHO drinking water, the USEPA adult, and the WHO
+recreational advisories, respectively. In each case we missed less than
+10\% of the lakes that in fact exceeded the microcystin-LR advisory.
+While this method performs well with regard to the false negative
+percentage, it is possible that is a relic of the NLA dataset and
+testing with additional data would allow us to confirm this result.
 
@@
 health advisory levels increases as a function of chlorophyll \emph{a}
-concentration (Figure \ref{fig:multi_cp_plot}). We use this association
+concentration (Figure \ref{fig:multi_cp_plot}). We used this association
 to identify chlorophyll \emph{a} concentrations that are associated with
@@
 (Table \ref{tab:mc_chla_table}). For the purposes of this discussion we
-focus on a conditional probability of 50\% or greater (i.e.~greater than
-even odds to exceed a health advisory level). The 50\% conditional
-probability chlorophyll \emph{a} thresholds represents 30.1\%, 10.6\%,
-8.8\%, and 7.9\% of sample lakes for the USEPA Child, the WHO Drinking,
+focus on a conditional probability of 50\% or greater (i.e., greater
+than even odds to exceed a health advisory level). The 50\% conditional
+probability chlorophyll \emph{a} thresholds represents 29.2\%, 9.8\%,
+8.6\%, and 6.9\% of sample lakes for the USEPA Child, the WHO Drinking,
 the USEPA Adult, and the WHO recreational levels, respecitvely.
-
-Furthermore, the chlorophyll \emph{a} cutoffs may be used to predict
-whether or not a lake exceeds the microcystin-LR health advisories.
-Doing so allows us to compare the accuracy of the prediction as well as
-evaluate false negatives. Total accuracy of the four cutoffs predicting
-microcystin-LR exceedances were 74\% for the USEPA children's advisory,
-86\% for the WHO drinking water advisory, 89\% for the USEPA adult
-advisory and, 90\% for the WHO regreational advisory (Tables
-\ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table},
-\ref{tab:adult_conmat_table}, \& \ref{tab:who_rec_conmat_table}).
-However, total accuracy is only one part of the prediction performace
-with which we are concerned.
-
-When using the chlorophyll \emph{a} cutoffs as an indicator of
-microcystin-LR exceedances, the error that should be avoided is
-predicting that no exceedance has occurred when in fact it has. In other
-words, we would like to avoid Type II errors and minimize the proportion
-of false negatives. For the four chlorophyll \emph{a} cut-offs we had a
-proportion of false negatives of 9\%, 8\%, 6\% and , 5\% for the U.S EPA
-childrens drinking water, the WHO drinking water, the USEPA adult
-drinking water, and the WHO recreational advisories, respectively. In
-each case we miss less than 10\% of the lakes that are in fact exceeding
-the microcystin-LR advisory.
 
@@
 of exceeding a target concentration take water samples and have those
-samples tested in a lab for microcystin-LR.
+samples tested for microcystin-LR.
 
@@
 data. As harmful algal blooms are made up of many species and have
-various mechanisms responsible for adverse impacts (e.g.~toxins,
-hypoxia, odors), there is no single definition of a bloom. For cHABs one
-approach has been to identify an increase over a baseline concentration
-of phycocyanin (Miller et al. 2013). This is a useful approach for
+various mechanisms responsible for adverse impacts (e.g., toxins,
+hypoxia, odors), there is no single definition of a bloom. For cHABs,
+one approach has been to identify an increase over a baseline
+concentration of phycocyanin {[}17{]}. This is a useful approach for
 targeted studies, but phycocyanin is also not always available and
@@
 \section{Acknowledgements}\label{acknowledgements}
+
+We would like to thank Anne Kuhn, Bryan Milstead, John Kiddon, Joe
+LiVolsi, Tim Gleason, and Wayne Munns for constructive reviews of this
+paper. This paper has not been subjected to Agency review. Therefore, it
+does not necessary reflect the views of the Agency. Mention of trade
+names or commercial products does not constitute endorsement or
+recommendation for use. This contribution is identified by the tracking
+number ORD-015143 of the Atlantic Ecology Division, Office of Research
+and Development, National Health and Environmental Effects Research
+Laboratory, US Environmental Protection Agency.
 
@@
 0.1 & 0.07 & 0.07 & 0.07 & 1.17\tabularnewline
-0.2 & 0.07 & 4.38 & 11.84 & 16.80\tabularnewline
-0.3 & 2.89 & 17.49 & 32.11 & 45.22\tabularnewline
-0.4 & 10.27 & 38.30 & 66.96 & 76.80\tabularnewline
-0.5 & 20.88 & 69.60 & 84.24 & 97.49\tabularnewline
-0.6 & 38.20 & 100.51 & 114.62 & 125.40\tabularnewline
-0.7 & 65.60 & 128.74 & 871.20 & 871.20\tabularnewline
-0.8 & 114.62 & 271.44 & 871.20 & 871.20\tabularnewline
-0.9 & 166.63 & 338.40 & 871.20 & 871.20\tabularnewline
+0.2 & 0.07 & 4.48 & 10.67 & 18.24\tabularnewline
+0.3 & 2.78 & 17.49 & 31.59 & 50.26\tabularnewline
+0.4 & 9.97 & 38.73 & 69.19 & 76.80\tabularnewline
+0.5 & 21.92 & 77.20 & 84.96 & 105.84\tabularnewline
+0.6 & 38.73 & 103.20 & 133.20 & 228.24\tabularnewline
+0.7 & 63.94 & 138.24 & 244.80 & 871.20\tabularnewline
+0.8 & 113.14 & 326.88 & 326.88 & 871.20\tabularnewline
+0.9 & 166.63 & 516.00 & 871.20 & 871.20\tabularnewline
 \bottomrule
@@
 
-\begin{longtable}[c]{@{}lrr@{}}
+\begin{longtable}[c]{@{}lrrr@{}}
 \caption{Confusion matrix comparing chlorophyll \textit{a} predicted
@@
 \toprule
-& FALSE & TRUE\tabularnewline
+& Not Exceed & Exceed & Row Totals\tabularnewline
 \midrule
@@
 \toprule
-& FALSE & TRUE\tabularnewline
+& Not Exceed & Exceed & Row Totals\tabularnewline
 \midrule
 \endhead
-FALSE & 632 & 90\tabularnewline
-TRUE & 179 & 127\tabularnewline
+Not Exceed & 638 & 93 & 731\tabularnewline
+Exceed & 173 & 124 & 297\tabularnewline
+Column Totals & 811 & 217 & 1028\tabularnewline
 \bottomrule
@@
 \toprule
-& FALSE & TRUE\tabularnewline
+& Not Exceed & Exceed & Row Totals\tabularnewline
 \midrule
 \endhead
-FALSE & 844 & 80\tabularnewline
-TRUE & 63 & 41\tabularnewline
+Not Exceed & 850 & 82 & 932\tabularnewline
+Exceed & 57 & 39 & 96\tabularnewline
+Column Totals & 907 & 121 & 1028\tabularnewline
 \bottomrule
@@
 \toprule
-& FALSE & TRUE\tabularnewline
+& Not Exceed & Exceed & Row Totals\tabularnewline
 \midrule
 \endhead
-FALSE & 885 & 57\tabularnewline
-TRUE & 52 & 34\tabularnewline
+Not Exceed & 887 & 57 & 944\tabularnewline
+Exceed & 50 & 34 & 84\tabularnewline
+Column Totals & 937 & 91 & 1028\tabularnewline
 \bottomrule
@@
 \toprule
-& FALSE & TRUE\tabularnewline
+& Not Exceed & Exceed & Row Totals\tabularnewline
 \midrule
 \endhead
-FALSE & 903 & 48\tabularnewline
-TRUE & 50 & 27\tabularnewline
+Not Exceed & 910 & 52 & 962\tabularnewline
+Exceed & 43 & 23 & 66\tabularnewline
+Column Totals & 953 & 75 & 1028\tabularnewline
 \bottomrule
@@
 
-Aranda-Rodriguez, R., Z. Jin, J. Harvie, and A. Cabecinha. 2015.
-Evaluation of three field test kits to detect microcystins from a public
-health perspective. Harmful Algae 42:34--42.
+1. Jetoo S, Grover VI, Krantzberg G (2015) The toledo drinking water
+advisory: Suggested application of the water safety planning approach.
+Sustainability 7: 9787--9808.
 
-Chorus, E. I., and J. Bartram. 1999. Toxic cyanobacteria in water: A
-guide to their public health consequences, monitoring and management.
+2. Rinta-Kanto JM, Konopko EA, DeBruyn JM, Bourbonniere RA, Boyer GL, et
+al. (2009) Lake erie microcystis: Relationship between microcystin
+production, dynamics of genotypes and environmental parameters in a
+large lake. Harmful Algae 8: 665--673.
 
-HABHRCA. 2014. Harmful algal bloom and hypoxia research and control
-amendments act of 2014.
+3. HABHRCA (2014) Harmful algal bloom and hypoxia research and control
+amendments act of 2014. Available:
+\url{https://www.congress.gov/113/bills/s1254/BILLS-113s1254enr.pdf}.
 
-Hollister, J. W., H. A. Walker, and J. F. Paul. 2008. CProb: A
-computational tool for conducting conditional probability analysis.
-Journal of environmental quality 37:2392--2396.
+4. McElhiney J, Lawton LA (2005) Detection of the cyanobacterial
+hepatotoxins microcystins. Toxicology and Applied Pharmacology 203:
+219--230.
 
-James, R., A. Gregg, A. Dindal, and J. McKernan. 2011. Environmental
-technology verification report: Abraxis microcystin test kits. Online
-document. Accessed online: June 22.
+5. Zurawell RW, Chen H, Burke JM, Prepas EE (2005) Hepatotoxic
+cyanobacteria: A review of the biological importance of microcystins in
+freshwater environments. Journal of Toxicology and Environmental Health,
+Part B 8: 1--37.
 
-Jetoo, S., V. I. Grover, and G. Krantzberg. 2015. The toledo drinking
-water advisory: Suggested application of the water safety planning
-approach. Sustainability 7:9787--9808.
+6. USEPA (2015) Drinking water health advisory for the cyanobacterial
+microcystin toxins. ePA-820R15100. Available:
+\url{http://www2.epa.gov/sites/production/files/2015-06/documents/microcystins-report-2015.pdf}.
 
-McElhiney, J., and L. A. Lawton. 2005. Detection of the cyanobacterial
-hepatotoxins microcystins. Toxicology and Applied Pharmacology
-203:219--230.
-
-Miller, T. R., L. Beversdorf, S. D. Chaston, and K. D. McMahon. 2013.
-Spatiotemporal molecular analysis of cyanobacteria blooms reveals
-microcystis-aphanizomenon interactions. PloS one 8:e74933.
-
-Organization, W. H. 2003. Cyanobacterial toxins: Microcystin-lR in
+7. Organization WH (2003) Cyanobacterial toxins: Microcystin-lR in
 drinking-water. background document for development of wHO guidelines
@@
 
-Paul, J. F., and M. E. McDonald. 2005. Development of empirical,
-geographically specific water quality criteria: A conditional
-probability analysis approach 41:1211--1223.
+8. Chorus EI, Bartram J (1999) Toxic cyanobacteria in water: A guide to
+their public health consequences, monitoring and management.
 
-Paul, J. F., and W. R. Munns. 2011. Probability surveys, conditional
+9. James R, Gregg A, Dindal A, McKernan J (2011) Environmental
+technology verification report: Abraxis microcystin test kits. Online
+document. Accessed online: June 22.
+
+10. Aranda-Rodriguez R, Jin Z, Harvie J, Cabecinha A (2015) Evaluation
+of three field test kits to detect microcystins from a public health
+perspective. Harmful Algae 42: 34--42.
+
+11. Pip E, Bowman L (2014) Microcystin and algal chlorophyll in relation
+to nearshore nutrient concentrations in lake winnipeg, canada.
+Environment and Pollution 3: p36.
+
+12. Yuan LL, Pollard AI, Pather S, Oliver JL, D'Anglada L (2014)
+Managing microcystin: Identifying national-scale thresholds for total
+nitrogen and chlorophyll a. Freshwater Biology 59: 1970--1981.
+
+13. USEPA (2009) National lakes assessment: A collaborative survey of
+the nation's lakes. ePA 841-r-09-001.
+
+14. Paul JF, Munns WR (2011) Probability surveys, conditional
 probability, and ecological risk assessment. Environmental Toxicology
-and Chemistry 30:1488--1495.
+and Chemistry 30: 1488--1495.
 
-Pip, E., and L. Bowman. 2014. Microcystin and algal chlorophyll in
-relation to nearshore nutrient concentrations in lake winnipeg, canada.
-Environment and Pollution 3:p36.
+15. Paul JF, McDonald ME (2005) Development of empirical, geographically
+specific water quality criteria: A conditional probability analysis
+approach. 41: 1211--1223.
 
-Rinta-Kanto, J. M., E. A. Konopko, J. M. DeBruyn, R. A. Bourbonniere, G.
-L. Boyer, and S. W. Wilhelm. 2009. Lake erie microcystis: Relationship
-between microcystin production, dynamics of genotypes and environmental
-parameters in a large lake. Harmful Algae 8:665--673.
+16. Hollister JW, Walker HA, Paul JF (2008) CProb: A computational tool
+for conducting conditional probability analysis. Journal of
+environmental quality 37: 2392--2396.
 
-USEPA. 2009. National lakes assessment: A collaborative survey of the
-nation's lakes. ePA 841-r-09-001. Office of Water; Office of Research;
-Development, US Environmental Protection Agency Washington, DC.
-
-USEPA. 2015. Drinking water health advisory for the cyanobacterial
-microcystin toxins. ePA-820R15100. Office of Water, US Environmental
-Protection Agency Washington, DC.
-
-Yuan, L. L., A. I. Pollard, S. Pather, J. L. Oliver, and L. D'Anglada.
-2014. Managing microcystin: Identifying national-scale thresholds for
-total nitrogen and chlorophyll a. Freshwater Biology 59:1970--1981.
-
-Zurawell, R. W., H. Chen, J. M. Burke, and E. E. Prepas. 2005.
-Hepatotoxic cyanobacteria: A review of the biological importance of
-microcystins in freshwater environments. Journal of Toxicology and
-Environmental Health, Part B 8:1--37.
+17. Miller TR, Beversdorf L, Chaston SD, McMahon KD (2013)
+Spatiotemporal molecular analysis of cyanobacteria blooms reveals
+microcystis-aphanizomenon interactions. PloS one 8: e74933.
 

### 2016-01-17 manuscript/manuscript.Rmd
f1000 foramt almost working.  Need to figure out tables and figures

@@
 spacing: doublespacing
-abstract: no
+abstract: "Cyanobacteria harmful algal blooms (cHABs) are associated with a wide range of adverse health effects that stem mostly from the presence of cyanotoxins.
+To help protect against these impacts, several health advisory levels have been set for some toxins.
+In particular, one of the more common toxins, microcystin, has several advisory levels set for drinking water and recreational use.
+However, compared to other water quality measures, field measurements of microcystin are not commonly available due to cost and advanced understanding required to interpret results.
+Addressing these issues will take time and resources.
+Thus, there is utility in finding indicators of microcystin that are already widely available, can be estimated quickly and in situ, and used as a first defense against high levels of microcystin.
+Chlorophyll *a* is commonly measured, can be estimated *in situ*, and has been shown to be positively associated with microcystin.
+In this paper, we use this association to provide estimates of chlorophyll a concentrations that are indicative of a higher probability of exceeding select health advisory concentrations for microcystin-LR.
+Using the 2007 National Lakes Assessment and a conditional probability approach, we identify chlorophyll *a* concentrations that are more likely than not to be associated with an exceedance of a microcystin health advisory level.
+We look at the recent US EPA health advisories for drinking water as well as the World Health Organization levels for drinking water and recreational use and identify a range of chlorophyll *a* thresholds.
+A 50% chance of exceeding one of the specific advisory microcystin concentrations of 0.3, 1, 1.6, and 2 g/L is associatied with chlorophyll *a* concentration thresholds of 24.64, 65.6, 89.71, and 94.93, respectively.
+When managing for these various microcystin levels, exceeding these reported chlorophyll *a* concentrations should be a trigger for further testing and possible management action."
 ---
@@
 
+<!--
 \singlespace
@@
 
-<!-- Abstract is being wrapped in latex here so that all analysis can be run in the chunk above and the results reproducibly referenced in the abstract. -->
+<!-- Abstract is being wrapped in latex here so that all analysis can be run in the chunk above and the results reproducibly referenced in the abstract. 
 Cyanobacteria harmful algal blooms (cHABs) are associated with a wide range of adverse health effects that stem mostly from the presence of cyanotoxins.
@@
 \doublespace
-
+-->
 #Introduction
@@
 #pandoc.table(mc_conc_table,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
-knitr::kable(mc_conc_table)
+knitr::kable(mc_conc_table,format="latex")
 ```
@@
 #pandoc.table(mc_chla_tbl,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
-knitr::kable(mc_chla_tbl)
+knitr::kable(mc_chla_tbl,format="latex")
 ```
@@
 \\label{tab:child_conmat_table}"}
-knitr::kable(child_conmat)
+knitr::kable(child_conmat,format="latex")
 ```
@@
 \\label{tab:who_drink_conmat_table}"}
-knitr::kable(who_drink_conmat)
+knitr::kable(who_drink_conmat,format="latex")
 ```
@@
 \\label{tab:adult_conmat_table}"}
-knitr::kable(adult_conmat)
+knitr::kable(adult_conmat,format="latex")
 ```
@@
 \\label{tab:who_rec_conmat_table}"}
-knitr::kable(who_rec_conmat)
+knitr::kable(who_rec_conmat,format="latex")
 ```

### 2016-01-17 manuscript/manuscript.tex
f1000 foramt almost working.  Need to figure out tables and figures

@@
 \documentclass[10pt,a4paper,twocolumn]{article}
-\usepackage{f1000_styles}
+\usepackage{components/f1000_styles}
+\usepackage{hyperref}
+\usepackage{longtable}
 \bibliography{components/manuscript}
@@
 
-\end{abstract}
-
-\clearpage
-
-\singlespace
-
-\vspace{2mm}
-
-\hrule
-
 Cyanobacteria harmful algal blooms (cHABs) are associated with a wide
@@
 there is utility in finding indicators of microcystin that are already
-widely available, can be estimated quickly and \emph{in situ}, and used
-as a first defense against high levels of microcystin. Chlorophyll
-\emph{a} is commonly measured, can be estimated \emph{in situ}, and has
-been shown to be positively associated with microcystin. In this paper,
-we use this association to provide estimates of chlorophyll \emph{a}
+widely available, can be estimated quickly and in situ, and used as a
+first defense against high levels of microcystin. Chlorophyll \emph{a}
+is commonly measured, can be estimated \emph{in situ}, and has been
+shown to be positively associated with microcystin. In this paper, we
+use this association to provide estimates of chlorophyll a
 concentrations that are indicative of a higher probability of exceeding
@@
 thresholds. A 50\% chance of exceeding one of the specific advisory
-microcystin concentrations of 0.3, 1, 1.6, and 2 \(\mu\)g/L is
-associatied with chlorophyll \emph{a} concentration thresholds of 21.92,
-77.2, 84.96, and 105.84, respectively. When managing for these various
-microcystin levels, exceeding these reported chlorophyll \emph{a}
-concentrations should be a trigger for further testing and possible
-management action.
+microcystin concentrations of 0.3, 1, 1.6, and 2 g/L is associatied with
+chlorophyll \emph{a} concentration thresholds of 24.64, 65.6, 89.71, and
+94.93, respectively. When managing for these various microcystin levels,
+exceeding these reported chlorophyll \emph{a} concentrations should be a
+trigger for further testing and possible management action.
 
-\vspace{3mm}
+\end{abstract}
 
-\hrule
-
-\doublespace
+\clearpage
 
@@
 
-\begin{longtable}[c]{@{}lll@{}}
-\caption{Various suggested microcystin-LR health advisory
-concentrations. \label{tab:microcystin_levels}}\tabularnewline
-\toprule
-Source & Type & Concentration\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Source & Type & Concentration\tabularnewline
-\midrule
-\endhead
-USEPA & Adult Drinking Water Advisory & 1.6 \(\mu\)g/L\tabularnewline
-USEPA & Child Drinking Water Advisory & 0.3 \(\mu\)g/L\tabularnewline
-WHO & Drinking Water & 1 \(\mu\)g/L\tabularnewline
-WHO & Recreational: High Prob. of Effect & 20-2000
-\(\mu\)g/L\tabularnewline
-WHO & Recreational: Low Prob. of Effect & 2-4 \(\mu\)g/L\tabularnewline
-WHO & Recreational: Moderate Prob. of Effect & 10-20
-\(\mu\)g/L\tabularnewline
-WHO & Recreational: Very High Prob. of Effect & \textgreater{}2000
-\(\mu\)g/L\tabularnewline
-\bottomrule
-\end{longtable}
+\begin{tabular}{l|l|l}
+\hline
+Source & Type & Concentration\\
+\hline
+USEPA & Adult Drinking Water Advisory & 1.6 \$\textbackslash{}mu\$g/L\\
+\hline
+USEPA & Child Drinking Water Advisory & 0.3 \$\textbackslash{}mu\$g/L\\
+\hline
+WHO & Drinking Water & 1 \$\textbackslash{}mu\$g/L\\
+\hline
+WHO & Recreational: High Prob. of Effect & 20-2000 \$\textbackslash{}mu\$g/L\\
+\hline
+WHO & Recreational: Low Prob. of Effect & 2-4 \$\textbackslash{}mu\$g/L\\
+\hline
+WHO & Recreational: Moderate Prob. of Effect & 10-20 \$\textbackslash{}mu\$g/L\\
+\hline
+WHO & Recreational: Very High Prob. of Effect & >2000 \$\textbackslash{}mu\$g/L\\
+\hline
+\end{tabular}
+
+:Various suggested microcystin-LR health advisory concentrations.
+\label{tab:microcystin_levels}
 
@@
 
-\begin{longtable}[c]{@{}rrrrr@{}}
-\caption{Chlorophyll \textit{a} concentrations that are associated with
-a 50\% probability of exceeding a microcystin-LR health advisory
-concentration. \label{tab:mc_chla_table}}\tabularnewline
-\toprule
-Cond. Probability & USEPA Child & WHO Drink & USEPA Adult & WHO
-Recreational\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Cond. Probability & USEPA Child & WHO Drink & USEPA Adult & WHO
-Recreational\tabularnewline
-\midrule
-\endhead
-0.1 & 0.07 & 0.07 & 0.07 & 1.17\tabularnewline
-0.2 & 0.07 & 4.48 & 10.67 & 18.24\tabularnewline
-0.3 & 2.78 & 17.49 & 31.59 & 50.26\tabularnewline
-0.4 & 9.97 & 38.73 & 69.19 & 76.80\tabularnewline
-0.5 & 21.92 & 77.20 & 84.96 & 105.84\tabularnewline
-0.6 & 38.73 & 103.20 & 133.20 & 228.24\tabularnewline
-0.7 & 63.94 & 138.24 & 244.80 & 871.20\tabularnewline
-0.8 & 113.14 & 326.88 & 326.88 & 871.20\tabularnewline
-0.9 & 166.63 & 516.00 & 871.20 & 871.20\tabularnewline
-\bottomrule
-\end{longtable}
+\begin{tabular}{r|r|r|r|r}
+\hline
+Cond. Probability & USEPA Child & WHO Drink & USEPA Adult & WHO Recreational\\
+\hline
+0.1 & 0.07 & 0.07 & 0.07 & 1.17\\
+\hline
+0.2 & 0.07 & 4.48 & 10.67 & 18.24\\
+\hline
+0.3 & 2.78 & 17.49 & 31.59 & 50.26\\
+\hline
+0.4 & 9.97 & 38.73 & 69.19 & 76.80\\
+\hline
+0.5 & 21.92 & 77.20 & 84.96 & 105.84\\
+\hline
+0.6 & 38.73 & 103.20 & 133.20 & 228.24\\
+\hline
+0.7 & 63.94 & 138.24 & 244.80 & 871.20\\
+\hline
+0.8 & 113.14 & 326.88 & 326.88 & 871.20\\
+\hline
+0.9 & 166.63 & 516.00 & 871.20 & 871.20\\
+\hline
+\end{tabular}
+
+:Chlorophyll \textit{a} concentrations that are associated with a 50\%
+probability of exceeding a microcystin-LR health advisory concentration.
+\label{tab:mc_chla_table}
 
@@
 
-\begin{longtable}[c]{@{}lrrr@{}}
-\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
-exceedences (rows) versus real exceedances (columns) for the USEPA
-childrens drinking water advisory.
-\label{tab:child_conmat_table}}\tabularnewline
-\toprule
-& Not Exceed & Exceed & Row Totals\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-& Not Exceed & Exceed & Row Totals\tabularnewline
-\midrule
-\endhead
-Not Exceed & 638 & 93 & 731\tabularnewline
-Exceed & 173 & 124 & 297\tabularnewline
-Column Totals & 811 & 217 & 1028\tabularnewline
-\bottomrule
-\end{longtable}
+\begin{tabular}{l|r|r|r}
+\hline
+  & Not Exceed & Exceed & Row Totals\\
+\hline
+Not Exceed & 638 & 93 & 731\\
+\hline
+Exceed & 173 & 124 & 297\\
+\hline
+Column Totals & 811 & 217 & 1028\\
+\hline
+\end{tabular}
+
+:Confusion matrix comparing chlorophyll \textit{a} predicted exceedences
+(rows) versus real exceedances (columns) for the USEPA childrens
+drinking water advisory. \label{tab:child_conmat_table}
 
@@
 
-\begin{longtable}[c]{@{}lrrr@{}}
-\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
-exceedences (rows) versus real exceedances (columns) for the WHO
-drinking water advisory.
-\label{tab:who_drink_conmat_table}}\tabularnewline
-\toprule
-& Not Exceed & Exceed & Row Totals\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-& Not Exceed & Exceed & Row Totals\tabularnewline
-\midrule
-\endhead
-Not Exceed & 850 & 82 & 932\tabularnewline
-Exceed & 57 & 39 & 96\tabularnewline
-Column Totals & 907 & 121 & 1028\tabularnewline
-\bottomrule
-\end{longtable}
+\begin{tabular}{l|r|r|r}
+\hline
+  & Not Exceed & Exceed & Row Totals\\
+\hline
+Not Exceed & 850 & 82 & 932\\
+\hline
+Exceed & 57 & 39 & 96\\
+\hline
+Column Totals & 907 & 121 & 1028\\
+\hline
+\end{tabular}
+
+:Confusion matrix comparing chlorophyll \textit{a} predicted exceedences
+(rows) versus real exceedances (columns) for the WHO drinking water
+advisory. \label{tab:who_drink_conmat_table}
 
@@
 
-\begin{longtable}[c]{@{}lrrr@{}}
-\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
-exceedences (rows) versus real exceedances (columns) for the USEPA adult
-drinking water advisory. \label{tab:adult_conmat_table}}\tabularnewline
-\toprule
-& Not Exceed & Exceed & Row Totals\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-& Not Exceed & Exceed & Row Totals\tabularnewline
-\midrule
-\endhead
-Not Exceed & 887 & 57 & 944\tabularnewline
-Exceed & 50 & 34 & 84\tabularnewline
-Column Totals & 937 & 91 & 1028\tabularnewline
-\bottomrule
-\end{longtable}
+\begin{tabular}{l|r|r|r}
+\hline
+  & Not Exceed & Exceed & Row Totals\\
+\hline
+Not Exceed & 887 & 57 & 944\\
+\hline
+Exceed & 50 & 34 & 84\\
+\hline
+Column Totals & 937 & 91 & 1028\\
+\hline
+\end{tabular}
+
+:Confusion matrix comparing chlorophyll \textit{a} predicted exceedences
+(rows) versus real exceedances (columns) for the USEPA adult drinking
+water advisory. \label{tab:adult_conmat_table}
 
@@
 
-\begin{longtable}[c]{@{}lrrr@{}}
-\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
-exceedences (rows) versus real exceedances (columns) for the WHO
-recreational water advisory.
-\label{tab:who_rec_conmat_table}}\tabularnewline
-\toprule
-& Not Exceed & Exceed & Row Totals\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-& Not Exceed & Exceed & Row Totals\tabularnewline
-\midrule
-\endhead
-Not Exceed & 910 & 52 & 962\tabularnewline
-Exceed & 43 & 23 & 66\tabularnewline
-Column Totals & 953 & 75 & 1028\tabularnewline
-\bottomrule
-\end{longtable}
+\begin{tabular}{l|r|r|r}
+\hline
+  & Not Exceed & Exceed & Row Totals\\
+\hline
+Not Exceed & 910 & 52 & 962\\
+\hline
+Exceed & 43 & 23 & 66\\
+\hline
+Column Totals & 953 & 75 & 1028\\
+\hline
+\end{tabular}
+
+:Confusion matrix comparing chlorophyll \textit{a} predicted exceedences
+(rows) versus real exceedances (columns) for the WHO recreational water
+advisory. \label{tab:who_rec_conmat_table}
 

### 2016-01-18 manuscript/f1000_submission/manuscript.Rmd
f1000 submission started

--- 2016-01-17 manuscript/manuscript.Rmd
@@
 title: Associations between chlorophyll *a* and various microcystin-LR health advisory
-  Concentrations
+  concentrations
 author:

### 2016-01-18 manuscript/manuscript.tex
f1000 submission started

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% For more detailed article preparation guidelines, please see:
-% http://f1000research.com/author-guidelines
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Original default rstudio/pandoc latex file
+%% upated by @jhollist 09/15/2014
+%% inspired by @cboetting https://github.com/cboettig/template and
+%% @rmflight blog posts:
+%% http://rmflight.github.io/posts/2014/07/analyses_as_packages.html 
+%% http://rmflight.github.io/posts/2014/07/vignetteAnalysis.html).  
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\documentclass[10pt,a4paper,twocolumn]{article}
-\usepackage{components/f1000_styles}
-\usepackage{hyperref}
-\usepackage{longtable}
-\bibliography{components/manuscript}
+\documentclass[11pt,]{article}
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+    \usepackage{xltxtra,xunicode}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage{graphicx}
+% Redefine \includegraphics so that, unless explicit options are
+% given, the image width will not exceed the width of the page.
+% Images get their normal width if they fit onto the page, but
+% are scaled down if they would overflow the margins.
+\makeatletter
+\def\ScaleIfNeeded{%
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+\let\Oldincludegraphics\includegraphics
+{%
+ \catcode`\@=11\relax%
+ \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
+}%
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={Associations between chlorophyll a and various microcystin-LR health advisory Concentrations},
+            colorlinks=true,
+            citecolor=blue,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\setcounter{secnumdepth}{5}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%Changes borrowed from @cboettig, added by @jhollist 
+% A modified page layout 
+\textwidth 6.75in
+\oddsidemargin -0.15in
+\evensidemargin -0.15in
+\textheight 9in
+\topmargin -0.5in
+\usepackage{lineno} % add 
+  \linenumbers % turns line numbering on 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%Packages and layout changes by @jhollist 09/15/2014
+\usepackage{ragged2e}
+\usepackage[font=normalsize]{caption}
+  \usepackage[doublespacing]{setspace}
+\usepackage{parskip}
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+\fancyhf{}
+\renewcommand{\headrulewidth}{0pt}
+\rfoot{\today}
+\lfoot{\thepage}
+%%Changed default abstract width and added lines
+\renewenvironment{abstract}{
+  \hfill\begin{minipage}{1\textwidth}
+  \rule{\textwidth}{1pt}\vspace{5pt}
+  \normalsize
+  \begin{justify}
+  \bfseries\abstractname\vspace{5pt}
+  \end{justify}}
+  {\par\noindent\rule{\textwidth}{1pt}\end{minipage}
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\title{Associations between chlorophyll \emph{a} and various microcystin-LR
+health advisory Concentrations}
+\author{
+Jeffrey W. Hollister
+Betty J. Kreakie
+}
+\date{}
 
 \begin{document}
+%%Edited by @jhollist 09/15/2014
+%%Adds title from YAML
+\begin{singlespace}
+\begin{center}
+\huge Associations between chlorophyll \emph{a} and various microcystin-LR
+health advisory Concentrations
+\end{center}
+%%Adds Author, correspond email asterisk, and affilnum from YAML
+\begin{center}
+\large
+Jeffrey W. Hollister \textsuperscript{*} \textsuperscript{1} 
+Betty J. Kreakie \textsuperscript{1} 
+\end{center}
+%%Adds affiliations from YAML
+\begin{justify}
+\footnotesize \emph{ 
+\\*
+\textsuperscript{1}US Environmental Protection Agency, Office of Research and Development,
+National Health and Environmental Effects Research Laboratory, Atlantic
+Ecology Division, 27 Tarzwell Drive Narragansett, RI, 02882, USA\\*
+}
+%%Adds corresponding author email(s) from YAML
+\newcounter{num}
+\setcounter{num}{1}
+\\[0.1cm]
+\footnotesize \emph{ 
+\ifnum\value{num}=1%
+\textsuperscript{*} corresponding author:
+\fi
+\href{mailto:hollister.jeff@epa.gov}{\nolinkurl{hollister.jeff@epa.gov}}
+\stepcounter{num}
+}
+\end{justify}
+%%Adds date from YAML
+\normalsize
 
-%Still need to figure out how to get corresponding author on here.
-\title{\textit{F1000Research} Associations between chlorophyll \emph{a} and various microcystin-LR
-health advisory Concentrations}
-\author[1]{Jeffrey W. Hollister}
-\author[1]{Betty J. Kreakie}
-\affil[1]{}
+\end{singlespace}
 
-\maketitle
-\thispagestyle{fancy}
+\singlespace
 
-\begin{abstract}
+\vspace{2mm}
+
+\hrule
 
@@
 there is utility in finding indicators of microcystin that are already
-widely available, can be estimated quickly and in situ, and used as a
-first defense against high levels of microcystin. Chlorophyll \emph{a}
-is commonly measured, can be estimated \emph{in situ}, and has been
-shown to be positively associated with microcystin. In this paper, we
-use this association to provide estimates of chlorophyll a
+widely available, can be estimated quickly and \emph{in situ}, and used
+as a first defense against high levels of microcystin. Chlorophyll
+\emph{a} is commonly measured, can be estimated \emph{in situ}, and has
+been shown to be positively associated with microcystin. In this paper,
+we use this association to provide estimates of chlorophyll \emph{a}
 concentrations that are indicative of a higher probability of exceeding
@@
 thresholds. A 50\% chance of exceeding one of the specific advisory
-microcystin concentrations of 0.3, 1, 1.6, and 2 g/L is associatied with
-chlorophyll \emph{a} concentration thresholds of 24.64, 65.6, 89.71, and
-94.93, respectively. When managing for these various microcystin levels,
-exceeding these reported chlorophyll \emph{a} concentrations should be a
-trigger for further testing and possible management action.
+microcystin concentrations of 0.3, 1, 1.6, and 2 \(\mu\)g/L is
+associatied with chlorophyll \emph{a} concentration thresholds of 21.92,
+77.2, 84.96, and 105.84, respectively. When managing for these various
+microcystin levels, exceeding these reported chlorophyll \emph{a}
+concentrations should be a trigger for further testing and possible
+management action.
 
-\end{abstract}
+\vspace{3mm}
 
-\clearpage
+\hrule
+
+\doublespace
 

### 2016-01-22 manuscript/f1000_submission/All.tex
closer to submission ready for f1000

%latex.default(All(mc_chla_tbl))%
\begin{table}[!tbp]
\begin{center}
\begin{tabular}{l}
\hline\hline
\multicolumn{1}{c}{}\tabularnewline
\hline
+\tabularnewline
Heading(`Cond. Probability`) * c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9) + Heading(`USEPA Child`) * c(0.07, 0.07, 2.98, 10.67, 23.36, 38.3, 65.6, 117.5, 167.04) + Heading(`WHO Drink`) * c(0.07, 4.43, 20.43, 42.24, 66.96, 103.2, 166.63, 338.4, 516) + Heading(`USEPA Adult`) * c(0.07, 11.8, 31.4, 66.96, 83.52, 133.2, 871.2, 871.2, 871.2)\tabularnewline
Heading(`WHO Recreational`) * c(1.22, 20, 53.64, 82.22, 105.84, 155.52, 216, 871.2, 871.2)\tabularnewline
\hline
\end{tabular}\end{center}

\end{table}

### 2016-01-22 manuscript/f1000_submission/manuscript.Rmd
closer to submission ready for f1000

@@
 affiliation:
-- affil: US Environmental Protection Agency, Office of Research and Development, National
+- affil: "US Environmental Protection Agency, Office of Research and Development, National
     Health and Environmental Effects Research Laboratory, Atlantic Ecology Division,
-    27 Tarzwell Drive  Narragansett, RI, 02882, USA
+    27 Tarzwell Drive  Narragansett, RI, 02882, USA"
   affilnum: 1
@@
     number_sections: yes
-    template: components/f1000research.latex
+    template: f1000research.latex
   html_document: null
 capsize: normalsize
-csl: components/plos.csl
+csl: plos.csl
 documentclass: article
@@
 linenumbers: yes
-bibliography: components/manuscript.bib
+bibliography: manuscript.bib
 spacing: doublespacing
@@
 }
+if(!require(microcystinchla)){
+  devtools::install_github("jhollist/autocrop",
+                           auth_token = getOption("Github_Access"))
+}
 if(!require(viridis)){
@@
 }
+library("autocrop")
 library("microcystinchla")
@@
 library("e1071")
-opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=5)
+#opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=5)
 
@@
 ##Data
+
 We used the 2007 NLA chlorophyll *a* and microcystin-LR concentration data [@usepa2009national, [NLA](http://www2.epa.gov/national-aquatic-resource-surveys/national-lakes-assessment)].
@@
 ##Analytical Methods
+
 We used a conditional probability analysis (CPA) approach to explore associations between chlorophyll *a* concentrations and World Health Organization (WHO) and USEPA microcystin-LR health advisory levels [@paul2011probability].
@@
 #Results
+
 In the 2007 NLA, microcystin-LR concentrations ranged from `r min(nla_dat$MCYST_TL_UGL,na.rm=T)` to `r max(nla_dat$MCYST_TL_UGL, na.rm=T)` $\mu$g/L.
@@
 Specific chlorophyll *a* that are associated with greater than even odds of exceeding the advisory levels were `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]`$\mu$g/L for 0.3, 1.0, 1.6, and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
+
+```{r epa_child_cp_plot, echo=FALSE, fig.width=6, fig.cap="",warning=FALSE,message=FALSE}
+jpeg(filename = "hollister_microcystin_fig2.jpg",
+       res = 450,
+       width = 7.82,
+       height = 6.8,
+       units = "in")
+multiplot(
+  plot_cp(epa_child_cp,x = "Log 10 Chlorophyll", 
+          y = "Conditional Probability",
+          title="A."),
+  plot_cp(who_drink_cp, x = "Log 10 Chlorophyll", 
+          y = "Conditional Probability", 
+          title="B."),
+  plot_cp(epa_adult_cp, x = "Log 10 Chlorophyll", 
+          y = "Conditional Probability", 
+          title="C."),
+  plot_cp(who_rec_low1_cp, x = "Log 10 Chlorophyll", 
+          y = "Conditional Probability", 
+          title="D."),
+  cols=2
+)
+x<-dev.off()
+x<-autocrop("hollister_microcystin_fig2.jpg",border = 25, 
+            outfile = "hollister_microcystin_fig2.jpg",
+            width = 6.1,
+            res = 450)
+```
+
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{hollister_microcystin_fig2.jpg}
+\caption{\label{fig:multi_cp_plot}Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
+A.) Plot for USEPA Child (0.3 $\mu$g/L).
+B.) Plot for WHO Drinking (1 $\mu$g/L).
+C.) Plot for USEPA Adult (1.6 $\mu$g/L).
+D.) Plot for WHO Recreational (2 $\mu$g/L).
+}
+\end{figure}
 
@@
 #Discussion
+
 The association between Log10 microcystin-LR and Log10 chlorophyll *a* shows a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
@@
 Higher chlorophyll *a* is not necessarily predictive of higher microcystin-LR concentrations; however, chlorophyll *a* may be predictive of the probability of exceeding a certain threshold.
+
+```{r chla_micro_scatter, echo=FALSE, fig.width=6, fig.cap="", warning=FALSE, message=FALSE}
+ggplot2::ggsave("hollister_microcystin_fig2.jpg",
+       plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",pt_col = "black",
+             x=expression(paste('Log10 Chl ', italic("a"),' (',mu,'g/L)')),
+             y=expression(paste('Log10 microcystin-LR',' (',mu,'g/L)'))),
+       dpi = 450,
+       height = 6.8,
+       units = "in")
+x<-autocrop("hollister_microcystin_fig2.jpg",border = 25, 
+            outfile = "hollister_microcystin_fig2.jpg",
+            width = 6.1,
+            res = 450)
+```
+
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{hollister_microcystin_fig2.jpg}
+\caption{\label{fig:chla_micro_scatter}Scatterplot showing association betweeen chlorophyll \textit{a} and microcystin-LR. }
+\end{figure}
 
@@
 #Acknowledgements
+
 We would like to thank Anne Kuhn, Bryan Milstead, John Kiddon, Joe LiVolsi, Tim Gleason, and Wayne Munns for constructive reviews of this paper.
@@
 
-\newpage
-
-#Figures
-
-```{r chla_micro_scatter, echo=FALSE, fig.width=6, fig.cap="Scatterplot showing association betweeen chlorophyll \\textit{a} and microcystin-LR. \\label{fig:chla_micro_scatter}"}
-plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",pt_col = "black",
-             x=expression(paste('Log10 Chl ', italic("a"),' (',mu,'g/L)')),
-             y=expression(paste('Log10 microcystin-LR',' (',mu,'g/L)')))
-```
-
-\newpage
-
-<!--
-```{r chla_micro_cdf, echo=FALSE, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
-\\label{fig:chla_micro_cdf}"}
-plot_cdf(nla_dat[,c("log_chla","log_micro")],"")
-```
-
-\newpage-->
-
-```{r epa_child_cp_plot, echo=FALSE, fig.width=6, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
-A.) Plot for USEPA Child (0.3 $\\mu$g/L).
-B.) Plot for WHO Drinking (1 $\\mu$g/L).
-C.) Plot for USEPA Adult (1.6 $\\mu$g/L).
-D.) Plot for WHO Recreational (2 $\\mu$g/L).
-\\label{fig:multi_cp_plot}"}
-multiplot(
-  plot_cp(epa_child_cp,x = "Log 10 Chlorophyll", 
-          y = "Conditional Probability",
-          title="A."),
-  plot_cp(who_drink_cp, x = "Log 10 Chlorophyll", 
-          y = "Conditional Probability", 
-          title="B."),
-  plot_cp(epa_adult_cp, x = "Log 10 Chlorophyll", 
-          y = "Conditional Probability", 
-          title="C."),
-  plot_cp(who_rec_low1_cp, x = "Log 10 Chlorophyll", 
-          y = "Conditional Probability", 
-          title="D."),
-  cols=2
-)
-```
-
-\newpage
+\onecolumn
 
@@
 
-```{r microcystin_levels,echo=FALSE,results='asis', tab.cap="Various suggested microcystin-LR health advisory concentrations. \\label{tab:microcystin_levels}"}
+```{r microcystin_levels,echo=FALSE,results='asis'}
 source_of_stnd <- c("WHO",
@@
                   "Adult Drinking Water Advisory")
-concentration <- c("2-4 $\\mu$g/L",
-                   "10-20 $\\mu$g/L",
-                   "20-2000 $\\mu$g/L",
-                   ">2000 $\\mu$g/L",
-                   "1 $\\mu$g/L",
-                   "0.3 $\\mu$g/L",
-                   "1.6 $\\mu$g/L")
+concentration <- c("2-4 $mu$g/L",
+                   "10-20 $mu$g/L",
+                   "20-2000 $mu$g/L",
+                   ">2000 $mu$g/L",
+                   "1 $mu$g/L",
+                   "0.3 $mu$g/L",
+                   "1.6 $mu$g/L")
 #note <- c("Low probability of adverse health effects",
@@
 #pandoc.table(mc_conc_table,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
-knitr::kable(mc_conc_table,format="latex")
+#knitr::kable(mc_conc_table,format="latex")
+print(xtable::xtable(mc_conc_table,
+                     label="microcystin_levels",
+                     caption="Various suggested microcystin-LR health advisory 
+                     concentrations."), 
+      include.rownames=FALSE, comment = FALSE,caption.placement = "top")
 ```
 
-\newpage
-
-```{r mc_chla_table, results='asis', echo=FALSE, tab.cap="Chlorophyll \\textit{a} concentrations that are associated with a 50% probability of exceeding a microcystin-LR health advisory concentration.
-\\label{tab:mc_chla_table}"}
+```{r mc_chla_table, results='asis', echo=FALSE}
 mc_chla_tbl <- data.frame(seq(0.1,0.9,0.1),mc_chla)
@@
 #pandoc.table(mc_chla_tbl,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
-knitr::kable(mc_chla_tbl,format="latex")
+#knitr::kable(mc_chla_tbl,format="latex")
+print(xtable::xtable(mc_chla_tbl,
+                     label="mc_chla_table",
+                     caption="Chlorophyll \\textit{a} concentrations that are 
+                     associated with a 50% probability of exceeding a 
+                     microcystin-LR health advisory concentration."),
+      include.rownames = FALSE, comment = FALSE,caption.placement = "top")
+```
+
+```{r child_conmat_table, results='asis', echo=FALSE}
+#knitr::kable(child_conmat,format="latex")
+print(xtable::xtable(child_conmat,
+                     label="child_conmat_table",
+                     caption="Confusion matrix comparing chlorophyll \\textit{a} 
+                     predicted exceedences (rows) versus real exceedances 
+                     (columns) for the USEPA childrens drinking water advisory."),
+      include.rownames = FALSE, comment = FALSE,caption.placement = "top")
+```
+
+```{r who_drink_conmat_table, results='asis', echo=FALSE}
+#knitr::kable(who_drink_conmat,format="latex")
+print(xtable::xtable(who_drink_conmat,
+                     label="who_drink_conmat_table",
+                     caption="Confusion matrix comparing chlorophyll \\textit{a} 
+                     predicted exceedences (rows) versus real exceedances 
+                     (columns) for the WHO drinking water advisory."),
+      include.rownames = FALSE, comment = FALSE,caption.placement = "top")
+```
+
+```{r adult_conmat_table, results='asis', echo=FALSE}
+#knitr::kable(adult_conmat,format="latex")
+print(xtable::xtable(adult_conmat,
+                     label="adult_conmat_table",
+                     caption="Confusion matrix comparing chlorophyll \\textit{a} 
+                     predicted exceedences (rows) versus real exceedances 
+                     (columns) for the USEPA adult drinking water advisory."),
+      include.rownames = FALSE, comment = FALSE,caption.placement = "top")
+```
+
+```{r who_rec_conmat_table, results='asis', echo=FALSE}
+#knitr::kable(who_rec_conmat,format="latex")
+print(xtable::xtable(who_rec_conmat,
+                     label="who_rec_conmat_table",
+                     caption="Confusion matrix comparing chlorophyll \\textit{a} 
+                     predicted exceedences (rows) versus real exceedances 
+                     (columns) for the WHO recreational water advisory."),
+      include.rownames = FALSE, comment = FALSE,caption.placement = "top")
 ```
@@
 \newpage
-
-```{r child_conmat_table, results='asis', echo=FALSE, tab.cap="Confusion matrix comparing chlorophyll \\textit{a} predicted exceedences (rows) versus real exceedances (columns) for the USEPA childrens drinking water advisory.
-\\label{tab:child_conmat_table}"}
-knitr::kable(child_conmat,format="latex")
-```
-
-\newpage
-
-```{r who_drink_conmat_table, results='asis', echo=FALSE, tab.cap="Confusion matrix comparing chlorophyll \\textit{a} predicted exceedences (rows) versus real exceedances (columns) for the WHO drinking water advisory.
-\\label{tab:who_drink_conmat_table}"}
-knitr::kable(who_drink_conmat,format="latex")
-```
-
-\newpage
-
-```{r adult_conmat_table, results='asis', echo=FALSE, tab.cap="Confusion matrix comparing chlorophyll \\textit{a} predicted exceedences (rows) versus real exceedances (columns) for the USEPA adult drinking water advisory.
-\\label{tab:adult_conmat_table}"}
-knitr::kable(adult_conmat,format="latex")
-```
-
-\newpage
-
-```{r who_rec_conmat_table, results='asis', echo=FALSE, tab.cap="Confusion matrix comparing chlorophyll \\textit{a} predicted exceedences (rows) versus real exceedances (columns) for the WHO recreational water advisory.
-\\label{tab:who_rec_conmat_table}"}
-knitr::kable(who_rec_conmat,format="latex")
-```
-
-\newpage
+\twocolumn
 

### 2016-01-22 manuscript/f1000_submission/manuscript.tex
closer to submission ready for f1000

--- 2016-01-17 manuscript/manuscript.tex
@@
 \documentclass[10pt,a4paper,twocolumn]{article}
-\usepackage{components/f1000_styles}
+\usepackage{f1000_styles}
 \usepackage{hyperref}
 \usepackage{longtable}
-\bibliography{components/manuscript}
+\bibliography{manuscript}
 
@@
 %Still need to figure out how to get corresponding author on here.
-\title{\textit{F1000Research} Associations between chlorophyll \emph{a} and various microcystin-LR
-health advisory Concentrations}
+\title{\textit{Associations between chlorophyll \emph{a} and various microcystin-LR
+health advisory concentrations} }
 \author[1]{Jeffrey W. Hollister}
 \author[1]{Betty J. Kreakie}
-\affil[1]{}
+\affil[1]{US Environmental Protection Agency, Office of Research and Development,
+National Health and Environmental Effects Research Laboratory, Atlantic
+Ecology Division, 27 Tarzwell Drive Narragansett, RI, 02882, USA}
 
@@
 chlorophyll \emph{a} that are associated with greater than even odds of
-exceeding the advisory levels were 0.07, 0.07, 2.78, and 9.97\(\mu\)g/L
+exceeding the advisory levels were 0.07, 0.07, 2.98, and 10.67\(\mu\)g/L
 for 0.3, 1.0, 1.6, and 2.0 \(\mu\)g/L advisory levels, respectively
 (Table \ref{tab:mc_chla_table} \& Figure \ref{fig:multi_cp_plot}).
+
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{hollister_microcystin_fig2.jpg}
+\caption{\label{fig:multi_cp_plot}Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
+A.) Plot for USEPA Child (0.3 $\mu$g/L).
+B.) Plot for WHO Drinking (1 $\mu$g/L).
+C.) Plot for USEPA Adult (1.6 $\mu$g/L).
+D.) Plot for WHO Recreational (2 $\mu$g/L).
+}
+\end{figure}
 
@@
 negatives. Total accuracy of the four cutoffs predicting microcystin-LR
-exceedances were 19\% for the USEPA children's advisory, 22\% for the
+exceedances were 19\% for the USEPA children's advisory, 21\% for the
 WHO drinking water advisory, 22\% for the USEPA adult advisory, and 23\%
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{hollister_microcystin_fig2.jpg}
+\caption{\label{fig:chla_micro_scatter}Scatterplot showing association betweeen chlorophyll \textit{a} and microcystin-LR. }
+\end{figure}
+
 This is the case as the probability of exceeding each of the four tested
@@
 than even odds to exceed a health advisory level). The 50\% conditional
-probability chlorophyll \emph{a} thresholds represents 29.2\%, 9.8\%,
-8.6\%, and 6.9\% of sample lakes for the USEPA Child, the WHO Drinking,
+probability chlorophyll \emph{a} thresholds represents 27.9\%, 11.3\%,
+9\%, and 6.9\% of sample lakes for the USEPA Child, the WHO Drinking,
 the USEPA Adult, and the WHO recreational levels, respecitvely.
@@
 
-\newpage
-
-\section{Figures}\label{figures}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/chla_micro_scatter-1.pdf}
-\caption{Scatterplot showing association betweeen chlorophyll \textit{a}
-and microcystin-LR. \label{fig:chla_micro_scatter}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/epa_child_cp_plot-1.pdf}
-\caption{Conditional probability plots showing association between the
-probability of exceeding various microcystin-LR (MLR) health advisory
-Levels. A.) Plot for USEPA Child (0.3 \(\mu\)g/L). B.) Plot for WHO
-Drinking (1 \(\mu\)g/L). C.) Plot for USEPA Adult (1.6 \(\mu\)g/L). D.)
-Plot for WHO Recreational (2 \(\mu\)g/L). \label{fig:multi_cp_plot}}
-\end{figure}
-
-\newpage
+\onecolumn
 
@@
 
-\begin{tabular}{l|l|l}
+\begin{table}[ht]
+\centering
+\caption{Various suggested microcystin-LR health advisory 
+                     concentrations.} 
+\label{microcystin_levels}
+\begin{tabular}{lll}
   \hline
-Source & Type & Concentration\\
+Source & Type & Concentration \\ 
   \hline
-USEPA & Adult Drinking Water Advisory & 1.6 \$\textbackslash{}mu\$g/L\\
-\hline
-USEPA & Child Drinking Water Advisory & 0.3 \$\textbackslash{}mu\$g/L\\
-\hline
-WHO & Drinking Water & 1 \$\textbackslash{}mu\$g/L\\
-\hline
-WHO & Recreational: High Prob. of Effect & 20-2000 \$\textbackslash{}mu\$g/L\\
-\hline
-WHO & Recreational: Low Prob. of Effect & 2-4 \$\textbackslash{}mu\$g/L\\
-\hline
-WHO & Recreational: Moderate Prob. of Effect & 10-20 \$\textbackslash{}mu\$g/L\\
-\hline
-WHO & Recreational: Very High Prob. of Effect & >2000 \$\textbackslash{}mu\$g/L\\
+USEPA & Adult Drinking Water Advisory & 1.6 \$mu\$g/L \\ 
+  USEPA & Child Drinking Water Advisory & 0.3 \$mu\$g/L \\ 
+  WHO & Drinking Water & 1 \$mu\$g/L \\ 
+  WHO & Recreational: High Prob. of Effect & 20-2000 \$mu\$g/L \\ 
+  WHO & Recreational: Low Prob. of Effect & 2-4 \$mu\$g/L \\ 
+  WHO & Recreational: Moderate Prob. of Effect & 10-20 \$mu\$g/L \\ 
+  WHO & Recreational: Very High Prob. of Effect & $>$2000 \$mu\$g/L \\ 
    \hline
 \end{tabular}
+\end{table}
 
-:Various suggested microcystin-LR health advisory concentrations.
-\label{tab:microcystin_levels}
+\begin{table}[ht]
+\centering
+\caption{Chlorophyll \textit{a} concentrations that are 
+                     associated with a 50% probability of exceeding a 
+                     microcystin-LR health advisory concentration.} 
+\label{mc_chla_table}
+\begin{tabular}{rrrrr}
+  \hline
+Cond. Probability & USEPA Child & WHO Drink & USEPA Adult & WHO Recreational \\ 
+  \hline
+0.10 & 0.07 & 0.07 & 0.07 & 1.22 \\ 
+  0.20 & 0.07 & 4.43 & 11.80 & 20.00 \\ 
+  0.30 & 2.98 & 20.43 & 31.40 & 53.64 \\ 
+  0.40 & 10.67 & 42.24 & 66.96 & 82.22 \\ 
+  0.50 & 23.36 & 66.96 & 83.52 & 105.84 \\ 
+  0.60 & 38.30 & 103.20 & 133.20 & 155.52 \\ 
+  0.70 & 65.60 & 166.63 & 871.20 & 216.00 \\ 
+  0.80 & 117.50 & 338.40 & 871.20 & 871.20 \\ 
+  0.90 & 167.04 & 516.00 & 871.20 & 871.20 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Confusion matrix comparing chlorophyll \textit{a} 
+                     predicted exceedences (rows) versus real exceedances 
+                     (columns) for the USEPA childrens drinking water advisory.} 
+\label{child_conmat_table}
+\begin{tabular}{rrr}
+  \hline
+Not Exceed & Exceed & Row Totals \\ 
+  \hline
+649 &  96 & 745 \\ 
+  162 & 121 & 283 \\ 
+  811 & 217 & 1028 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Confusion matrix comparing chlorophyll \textit{a} 
+                     predicted exceedences (rows) versus real exceedances 
+                     (columns) for the WHO drinking water advisory.} 
+\label{who_drink_conmat_table}
+\begin{tabular}{rrr}
+  \hline
+Not Exceed & Exceed & Row Totals \\ 
+  \hline
+839 &  77 & 916 \\ 
+   68 &  44 & 112 \\ 
+  907 & 121 & 1028 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Confusion matrix comparing chlorophyll \textit{a} 
+                     predicted exceedences (rows) versus real exceedances 
+                     (columns) for the USEPA adult drinking water advisory.} 
+\label{adult_conmat_table}
+\begin{tabular}{rrr}
+  \hline
+Not Exceed & Exceed & Row Totals \\ 
+  \hline
+883 &  57 & 940 \\ 
+   54 &  34 &  88 \\ 
+  937 &  91 & 1028 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\caption{Confusion matrix comparing chlorophyll \textit{a} 
+                     predicted exceedences (rows) versus real exceedances 
+                     (columns) for the WHO recreational water advisory.} 
+\label{who_rec_conmat_table}
+\begin{tabular}{rrr}
+  \hline
+Not Exceed & Exceed & Row Totals \\ 
+  \hline
+910 &  52 & 962 \\ 
+   43 &  23 &  66 \\ 
+  953 &  75 & 1028 \\ 
+   \hline
+\end{tabular}
+\end{table}
 
 \newpage
-
-\begin{tabular}{r|r|r|r|r}
-\hline
-Cond. Probability & USEPA Child & WHO Drink & USEPA Adult & WHO Recreational\\
-\hline
-0.1 & 0.07 & 0.07 & 0.07 & 1.17\\
-\hline
-0.2 & 0.07 & 4.48 & 10.67 & 18.24\\
-\hline
-0.3 & 2.78 & 17.49 & 31.59 & 50.26\\
-\hline
-0.4 & 9.97 & 38.73 & 69.19 & 76.80\\
-\hline
-0.5 & 21.92 & 77.20 & 84.96 & 105.84\\
-\hline
-0.6 & 38.73 & 103.20 & 133.20 & 228.24\\
-\hline
-0.7 & 63.94 & 138.24 & 244.80 & 871.20\\
-\hline
-0.8 & 113.14 & 326.88 & 326.88 & 871.20\\
-\hline
-0.9 & 166.63 & 516.00 & 871.20 & 871.20\\
-\hline
-\end{tabular}
-
-:Chlorophyll \textit{a} concentrations that are associated with a 50\%
-probability of exceeding a microcystin-LR health advisory concentration.
-\label{tab:mc_chla_table}
-
-\newpage
-
-\begin{tabular}{l|r|r|r}
-\hline
-  & Not Exceed & Exceed & Row Totals\\
-\hline
-Not Exceed & 638 & 93 & 731\\
-\hline
-Exceed & 173 & 124 & 297\\
-\hline
-Column Totals & 811 & 217 & 1028\\
-\hline
-\end{tabular}
-
-:Confusion matrix comparing chlorophyll \textit{a} predicted exceedences
-(rows) versus real exceedances (columns) for the USEPA childrens
-drinking water advisory. \label{tab:child_conmat_table}
-
-\newpage
-
-\begin{tabular}{l|r|r|r}
-\hline
-  & Not Exceed & Exceed & Row Totals\\
-\hline
-Not Exceed & 850 & 82 & 932\\
-\hline
-Exceed & 57 & 39 & 96\\
-\hline
-Column Totals & 907 & 121 & 1028\\
-\hline
-\end{tabular}
-
-:Confusion matrix comparing chlorophyll \textit{a} predicted exceedences
-(rows) versus real exceedances (columns) for the WHO drinking water
-advisory. \label{tab:who_drink_conmat_table}
-
-\newpage
-
-\begin{tabular}{l|r|r|r}
-\hline
-  & Not Exceed & Exceed & Row Totals\\
-\hline
-Not Exceed & 887 & 57 & 944\\
-\hline
-Exceed & 50 & 34 & 84\\
-\hline
-Column Totals & 937 & 91 & 1028\\
-\hline
-\end{tabular}
-
-:Confusion matrix comparing chlorophyll \textit{a} predicted exceedences
-(rows) versus real exceedances (columns) for the USEPA adult drinking
-water advisory. \label{tab:adult_conmat_table}
-
-\newpage
-
-\begin{tabular}{l|r|r|r}
-\hline
-  & Not Exceed & Exceed & Row Totals\\
-\hline
-Not Exceed & 910 & 52 & 962\\
-\hline
-Exceed & 43 & 23 & 66\\
-\hline
-Column Totals & 953 & 75 & 1028\\
-\hline
-\end{tabular}
-
-:Confusion matrix comparing chlorophyll \textit{a} predicted exceedences
-(rows) versus real exceedances (columns) for the WHO recreational water
-advisory. \label{tab:who_rec_conmat_table}
-
-\newpage
+\twocolumn
 

### 2016-01-25 manuscript/f1000_submission/manuscript.Rmd
version (almost) ready to submit to f1000

@@
 These data represent a snapshot of water quality from the summer of 2007 for the conterminous United States and were collected as part of an ongoing probabilistic monitoring program [@usepa2009national].
-Data on chlorophyll *a* and microcystin-LR concentrations are available for `r nrow(na.omit(nla_dat$CHLA,nla_dat$MCYST_TL_UGL))` lakes.
+Data on chlorophyll *a* and microcystin-LR concentrations are available for `r nrow(na.omit(data.frame(nla_dat$CHLA,nla_dat$MCYST_TL_UGL)))` lakes.
 
@@
 ```{r epa_child_cp_plot, echo=FALSE, fig.width=6, fig.cap="",warning=FALSE,message=FALSE}
-jpeg(filename = "hollister_microcystin_fig2.jpg",
+jpeg(filename = "hollister_microcystin_fig1.jpg",
        res = 450,
@@
 x<-dev.off()
-x<-autocrop("hollister_microcystin_fig2.jpg",border = 25, 
-            outfile = "hollister_microcystin_fig2.jpg",
+x<-autocrop("hollister_microcystin_fig1.jpg",border = 25, 
+            outfile = "hollister_microcystin_fig1.jpg",
             width = 6.1,
@@
 \centering
-\includegraphics[width=0.4\textwidth]{hollister_microcystin_fig2.jpg}
+\includegraphics[width=0.4\textwidth]{hollister_microcystin_fig1.jpg}
 \caption{\label{fig:multi_cp_plot}Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
@@
                   "Adult Drinking Water Advisory")
-concentration <- c("2-4 $mu$g/L",
-                   "10-20 $mu$g/L",
-                   "20-2000 $mu$g/L",
-                   ">2000 $mu$g/L",
-                   "1 $mu$g/L",
-                   "0.3 $mu$g/L",
-                   "1.6 $mu$g/L")
+concentration <- c("2-4 $\\mu$g/L",
+                   "10-20 $\\mu$g/L",
+                   "20-2000 $\\mu$g/L",
+                   ">2000 $\\mu$g/L",
+                   "1 $\\mu$g/L",
+                   "0.3 $\\mu$g/L",
+                   "1.6 $\\mu$g/L")
 #note <- c("Low probability of adverse health effects",
@@
                      caption="Various suggested microcystin-LR health advisory 
-                     concentrations."), 
-      include.rownames=FALSE, comment = FALSE,caption.placement = "top")
+                     concentrations from the USEPA and World Health Organization."), 
+      include.rownames=FALSE, comment = FALSE,caption.placement = "top",
+      latex.environments = NULL,sanitize.text.function=identity)
 ```
@@
                      microcystin-LR health advisory concentration."),
-      include.rownames = FALSE, comment = FALSE,caption.placement = "top")
+      include.rownames = FALSE, comment = FALSE,caption.placement = "top",latex.environments = NULL)
 ```
@@
                      (columns) for the USEPA childrens drinking water advisory."),
-      include.rownames = FALSE, comment = FALSE,caption.placement = "top")
+      include.rownames = T, comment = FALSE,caption.placement = "top",latex.environments = NULL)
 ```
@@
                      (columns) for the WHO drinking water advisory."),
-      include.rownames = FALSE, comment = FALSE,caption.placement = "top")
+      include.rownames = T, comment = FALSE,caption.placement = "top",latex.environments = NULL)
 ```
@@
                      (columns) for the USEPA adult drinking water advisory."),
-      include.rownames = FALSE, comment = FALSE,caption.placement = "top")
+      include.rownames = T, comment = FALSE,caption.placement = "top",latex.environments = NULL)
 ```
@@
                      (columns) for the WHO recreational water advisory."),
-      include.rownames = FALSE, comment = FALSE,caption.placement = "top")
+      include.rownames = T, comment = FALSE,caption.placement = "top",latex.environments = NULL)
 ```
 
-\newpage
 \twocolumn

### 2016-01-25 manuscript/f1000_submission/manuscript.tex
version (almost) ready to submit to f1000

@@
 {[}13{]}. Data on chlorophyll \emph{a} and microcystin-LR concentrations
-are available for lakes.
+are available for 1028 lakes.
 
@@
 \begin{table}[ht]
-\centering
 \caption{Various suggested microcystin-LR health advisory 
-                     concentrations.} 
+                     concentrations from the USEPA and World Health Organization.} 
 \label{microcystin_levels}
@@
   \hline
-USEPA & Adult Drinking Water Advisory & 1.6 \$mu\$g/L \\ 
-  USEPA & Child Drinking Water Advisory & 0.3 \$mu\$g/L \\ 
-  WHO & Drinking Water & 1 \$mu\$g/L \\ 
-  WHO & Recreational: High Prob. of Effect & 20-2000 \$mu\$g/L \\ 
-  WHO & Recreational: Low Prob. of Effect & 2-4 \$mu\$g/L \\ 
-  WHO & Recreational: Moderate Prob. of Effect & 10-20 \$mu\$g/L \\ 
-  WHO & Recreational: Very High Prob. of Effect & $>$2000 \$mu\$g/L \\ 
+USEPA & Adult Drinking Water Advisory & 1.6 $\mu$g/L \\ 
+  USEPA & Child Drinking Water Advisory & 0.3 $\mu$g/L \\ 
+  WHO & Drinking Water & 1 $\mu$g/L \\ 
+  WHO & Recreational: High Prob. of Effect & 20-2000 $\mu$g/L \\ 
+  WHO & Recreational: Low Prob. of Effect & 2-4 $\mu$g/L \\ 
+  WHO & Recreational: Moderate Prob. of Effect & 10-20 $\mu$g/L \\ 
+  WHO & Recreational: Very High Prob. of Effect & >2000 $\mu$g/L \\ 
    \hline
@@
 \begin{table}[ht]
-\centering
 \caption{Chlorophyll \textit{a} concentrations that are 
@@
 \begin{table}[ht]
-\centering
 \caption{Confusion matrix comparing chlorophyll \textit{a} 
@@
 \label{child_conmat_table}
-\begin{tabular}{rrr}
+\begin{tabular}{rrrr}
   \hline
-Not Exceed & Exceed & Row Totals \\ 
+ & Not Exceed & Exceed & Row Totals \\ 
   \hline
-649 &  96 & 745 \\ 
-  162 & 121 & 283 \\ 
-  811 & 217 & 1028 \\ 
+Not Exceed & 649 &  96 & 745 \\ 
+  Exceed & 162 & 121 & 283 \\ 
+  Column Totals & 811 & 217 & 1028 \\ 
    \hline
@@
 \label{who_drink_conmat_table}
-\begin{tabular}{rrr}
+\begin{tabular}{rrrr}
   \hline
-Not Exceed & Exceed & Row Totals \\ 
+ & Not Exceed & Exceed & Row Totals \\ 
   \hline
-839 &  77 & 916 \\ 
-   68 &  44 & 112 \\ 
-  907 & 121 & 1028 \\ 
+Not Exceed & 839 &  77 & 916 \\ 
+  Exceed &  68 &  44 & 112 \\ 
+  Column Totals & 907 & 121 & 1028 \\ 
    \hline
@@
 \label{adult_conmat_table}
-\begin{tabular}{rrr}
+\begin{tabular}{rrrr}
   \hline
-Not Exceed & Exceed & Row Totals \\ 
+ & Not Exceed & Exceed & Row Totals \\ 
   \hline
-883 &  57 & 940 \\ 
-   54 &  34 &  88 \\ 
-  937 &  91 & 1028 \\ 
+Not Exceed & 883 &  57 & 940 \\ 
+  Exceed &  54 &  34 &  88 \\ 
+  Column Totals & 937 &  91 & 1028 \\ 
    \hline
@@
 \label{who_rec_conmat_table}
-\begin{tabular}{rrr}
+\begin{tabular}{rrrr}
   \hline
-Not Exceed & Exceed & Row Totals \\ 
+ & Not Exceed & Exceed & Row Totals \\ 
   \hline
-910 &  52 & 962 \\ 
-   43 &  23 &  66 \\ 
-  953 &  75 & 1028 \\ 
+Not Exceed & 910 &  52 & 962 \\ 
+  Exceed &  43 &  23 &  66 \\ 
+  Column Totals & 953 &  75 & 1028 \\ 
    \hline
@@
 
-\newpage
 \twocolumn
@@
 
-3. HABHRCA (2014) Harmful algal bloom and hypoxia research and control
-amendments act of 2014. Available:
+3. Bloom HA, Research H, Act CA (2014) Harmful Algal Bloom and Hypoxia
+Research and Control Amendments Act of 2014. (S. 1254). Available:
 \url{https://www.congress.gov/113/bills/s1254/BILLS-113s1254enr.pdf}.
@@
 6. USEPA (2015) Drinking water health advisory for the cyanobacterial
-microcystin toxins. ePA-820R15100. Available:
+microcystin toxins. EPA-820-R-15100. Available:
 \url{http://www2.epa.gov/sites/production/files/2015-06/documents/microcystins-report-2015.pdf}.
 
-7. Organization WH (2003) Cyanobacterial toxins: Microcystin-lR in
-drinking-water. background document for development of wHO guidelines
-for drinking-water quality. geneva, switzerland. World Health
-Organization, 2nd ed. Geneva.
+7. World Health Organization (2003) Cyanobacterial toxins:
+Microcystin-LR in drinking-water. Background document for development of
+WHO guidelines for drinking-water quality. Geneva, Switzerland. World
+Health Organization, 2nd ed. Geneva.
 
 8. Chorus EI, Bartram J (1999) Toxic cyanobacteria in water: A guide to
-their public health consequences, monitoring and management.
+their public health consequences, monitoring and management. World
+Health Organization.
 
@@
 11. Pip E, Bowman L (2014) Microcystin and algal chlorophyll in relation
-to nearshore nutrient concentrations in lake winnipeg, canada.
+to nearshore nutrient concentrations in Lake Winnipeg, Canada.
 Environment and Pollution 3: p36.
@@
 13. USEPA (2009) National lakes assessment: A collaborative survey of
-the nation's lakes. ePA 841-r-09-001.
+the nation's lakes. EPA 841-R-09-001.
 
@@
 specific water quality criteria: A conditional probability analysis
-approach. 41: 1211--1223.
+approach. Journal of the American Water Resources Association 41:
+1211--1223.
 
@@
 for conducting conditional probability analysis. Journal of
-environmental quality 37: 2392--2396.
+Environmental Quality 37: 2392--2396.
 

### 2016-01-25 manuscript/manuscript.Rmd
version (almost) ready to submit to f1000

--- 2016-01-17 manuscript/manuscript.Rmd
@@
     number_sections: yes
-    template: components/f1000research.latex
+    template: components/manuscript.latex
   html_document: null
@@
 #pandoc.table(mc_conc_table,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
-knitr::kable(mc_conc_table)
+knitr::kable(mc_conc_table)#,format="latex")
 ```
@@
 #pandoc.table(mc_chla_tbl,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
-knitr::kable(mc_chla_tbl)
+knitr::kable(mc_chla_tbl)#,format="latex")
 ```
@@
 \\label{tab:child_conmat_table}"}
-knitr::kable(child_conmat)
+knitr::kable(child_conmat)#,format="latex")
 ```
@@
 \\label{tab:who_drink_conmat_table}"}
-knitr::kable(who_drink_conmat)
+knitr::kable(who_drink_conmat)#,format="latex")
 ```
@@
 \\label{tab:adult_conmat_table}"}
-knitr::kable(adult_conmat)
+knitr::kable(adult_conmat)#,format="latex")
 ```
@@
 \\label{tab:who_rec_conmat_table}"}
-knitr::kable(who_rec_conmat)
+knitr::kable(who_rec_conmat)#,format="latex")
 ```

### 2016-01-25 manuscript/f1000_submission/manuscript.Rmd
fixed table labels

@@
 print(xtable::xtable(mc_conc_table,
-                     label="microcystin_levels",
+                     label="tab:microcystin_levels",
                      caption="Various suggested microcystin-LR health advisory 
@@
 print(xtable::xtable(mc_chla_tbl,
-                     label="mc_chla_table",
+                     label="tab:mc_chla_table",
                      caption="Chlorophyll \\textit{a} concentrations that are 
@@
 print(xtable::xtable(child_conmat,
-                     label="child_conmat_table",
+                     label="tab:child_conmat_table",
                      caption="Confusion matrix comparing chlorophyll \\textit{a} 
@@
 print(xtable::xtable(who_drink_conmat,
-                     label="who_drink_conmat_table",
+                     label="tab:who_drink_conmat_table",
                      caption="Confusion matrix comparing chlorophyll \\textit{a} 
@@
 print(xtable::xtable(adult_conmat,
-                     label="adult_conmat_table",
+                     label="tab:adult_conmat_table",
                      caption="Confusion matrix comparing chlorophyll \\textit{a} 
@@
 print(xtable::xtable(who_rec_conmat,
-                     label="who_rec_conmat_table",
+                     label="tab:who_rec_conmat_table",
                      caption="Confusion matrix comparing chlorophyll \\textit{a} 

### 2016-01-25 manuscript/f1000_submission/cover_letter.Rmd
updated letter

---
output: pdf_document
fontsize: 12pt
geometry: 
  - top=0.05in
  - right=0.5in
  - left=0.5in
  - bottom=0.1in
---

![letterhead](epa_letterhead.jpg)

January 25, 2016 \newline

Editor, *F1000Research*\newline
London, England\newline

To whom it may concern,

Please accept, for your review, "Associations between chlorophyll *a* and various microcystin-LR health advisory concentrations" to be published in *F1000Research*.  In this article we detail a conditional probability analysis of the USEPA's 2007 National Lakes Assessment in which we identify chlorophyll *a* concentrations that are associated with several microcystin-LR health advisories.  Making these associations is important because it allows identification of a large number of lakes that may be posing a public health hazard because of high concentrations of cyanotoxins.  Direct measurements of microcystin-LR are becoming more commonplace, but at the current time they are still relatively rare.  Chlorophyll *a* measurements are much more common and can be used to help identify potential problems.  We are also pleased to submit this to *F1000Research* as we are proponents of practicing open science.  Towards that end all of the data and source code for this manuscript are available as an R package via GitHub (https://github.com/USEPA/microcystinchla).  Any questions or concerns about this submission, please do not hesitate to contact us.

Sincerely,

Jeffrey W. Hollister
401 782 9655
hollister.jeff@epa.gov

Betty J. Kreakie
kreakie.betty@epa.gov

### 2016-01-25 manuscript/f1000_submission/manuscript.Rmd
updated letter

@@
 To help protect against these impacts, several health advisory levels have been set for some toxins.
-In particular, one of the more common toxins, microcystin, has several advisory levels set for drinking water and recreational use.
-However, compared to other water quality measures, field measurements of microcystin are not commonly available due to cost and advanced understanding required to interpret results.
+In particular, one of the more common toxins, microcystin-LR, has several advisory levels set for drinking water and recreational use.
+However, compared to other water quality measures, field measurements of microcystin-LR are not commonly available due to cost and advanced understanding required to interpret results.
 Addressing these issues will take time and resources.
-Thus, there is utility in finding indicators of microcystin that are already widely available, can be estimated quickly and in situ, and used as a first defense against high levels of microcystin.
-Chlorophyll *a* is commonly measured, can be estimated *in situ*, and has been shown to be positively associated with microcystin.
-In this paper, we use this association to provide estimates of chlorophyll a concentrations that are indicative of a higher probability of exceeding select health advisory concentrations for microcystin-LR.
-Using the 2007 National Lakes Assessment and a conditional probability approach, we identify chlorophyll *a* concentrations that are more likely than not to be associated with an exceedance of a microcystin health advisory level.
+Thus, there is utility in finding indicators of microcystin-LR that are already widely available, can be estimated quickly and in situ, and used as a first defense against high concentrations of microcystin-LR.
+Chlorophyll *a* is commonly measured, can be estimated *in situ*, and has been shown to be positively associated with microcystin-LR.
+In this paper, we use this association to provide estimates of chlorophyll *a* concentrations that are indicative of a higher probability of exceeding select health advisory concentrations for microcystin-LR.
+Using the 2007 National Lakes Assessment and a conditional probability approach, we identify chlorophyll *a* concentrations that are more likely than not to be associated with an exceedance of a microcystin-LR health advisory level.
 We look at the recent US EPA health advisories for drinking water as well as the World Health Organization levels for drinking water and recreational use and identify a range of chlorophyll *a* thresholds.
-A 50% chance of exceeding one of the specific advisory microcystin concentrations of 0.3, 1, 1.6, and 2 g/L is associatied with chlorophyll *a* concentration thresholds of 24.64, 65.6, 89.71, and 94.93, respectively.
-When managing for these various microcystin levels, exceeding these reported chlorophyll *a* concentrations should be a trigger for further testing and possible management action."
+A 50% chance of exceeding one of the specific advisory microcystin-LR concentrations of 0.3, 1, 1.6, and 2 g/L is associatied with chlorophyll *a* concentration thresholds of 24.64, 65.6, 89.71, and 94.93, respectively.
+When managing for these various microcystin-LR levels, exceeding these reported chlorophyll *a* concentrations should be a trigger for further testing and possible management action."
 ---

### 2016-01-25 manuscript/f1000_submission/manuscript.tex
fixed table labels
updated letter

@@
 advisory levels have been set for some toxins. In particular, one of the
-more common toxins, microcystin, has several advisory levels set for
+more common toxins, microcystin-LR, has several advisory levels set for
 drinking water and recreational use. However, compared to other water
-quality measures, field measurements of microcystin are not commonly
+quality measures, field measurements of microcystin-LR are not commonly
 available due to cost and advanced understanding required to interpret
 results. Addressing these issues will take time and resources. Thus,
-there is utility in finding indicators of microcystin that are already
-widely available, can be estimated quickly and in situ, and used as a
-first defense against high levels of microcystin. Chlorophyll \emph{a}
-is commonly measured, can be estimated \emph{in situ}, and has been
-shown to be positively associated with microcystin. In this paper, we
-use this association to provide estimates of chlorophyll a
-concentrations that are indicative of a higher probability of exceeding
-select health advisory concentrations for microcystin-LR. Using the 2007
-National Lakes Assessment and a conditional probability approach, we
-identify chlorophyll \emph{a} concentrations that are more likely than
-not to be associated with an exceedance of a microcystin health advisory
-level. We look at the recent US EPA health advisories for drinking water
-as well as the World Health Organization levels for drinking water and
-recreational use and identify a range of chlorophyll \emph{a}
-thresholds. A 50\% chance of exceeding one of the specific advisory
-microcystin concentrations of 0.3, 1, 1.6, and 2 g/L is associatied with
-chlorophyll \emph{a} concentration thresholds of 24.64, 65.6, 89.71, and
-94.93, respectively. When managing for these various microcystin levels,
-exceeding these reported chlorophyll \emph{a} concentrations should be a
-trigger for further testing and possible management action.
+there is utility in finding indicators of microcystin-LR that are
+already widely available, can be estimated quickly and in situ, and used
+as a first defense against high concentrations of microcystin-LR.
+Chlorophyll \emph{a} is commonly measured, can be estimated \emph{in
+situ}, and has been shown to be positively associated with
+microcystin-LR. In this paper, we use this association to provide
+estimates of chlorophyll \emph{a} concentrations that are indicative of
+a higher probability of exceeding select health advisory concentrations
+for microcystin-LR. Using the 2007 National Lakes Assessment and a
+conditional probability approach, we identify chlorophyll \emph{a}
+concentrations that are more likely than not to be associated with an
+exceedance of a microcystin-LR health advisory level. We look at the
+recent US EPA health advisories for drinking water as well as the World
+Health Organization levels for drinking water and recreational use and
+identify a range of chlorophyll \emph{a} thresholds. A 50\% chance of
+exceeding one of the specific advisory microcystin-LR concentrations of
+0.3, 1, 1.6, and 2 g/L is associatied with chlorophyll \emph{a}
+concentration thresholds of 24.64, 65.6, 89.71, and 94.93, respectively.
+When managing for these various microcystin-LR levels, exceeding these
+reported chlorophyll \emph{a} concentrations should be a trigger for
+further testing and possible management action.
 
@@
                      concentrations from the USEPA and World Health Organization.} 
-\label{microcystin_levels}
+\label{tab:microcystin_levels}
 \begin{tabular}{lll}
@@
                      microcystin-LR health advisory concentration.} 
-\label{mc_chla_table}
+\label{tab:mc_chla_table}
 \begin{tabular}{rrrrr}
@@
                      (columns) for the USEPA childrens drinking water advisory.} 
-\label{child_conmat_table}
+\label{tab:child_conmat_table}
 \begin{tabular}{rrrr}
@@
                      (columns) for the WHO drinking water advisory.} 
-\label{who_drink_conmat_table}
+\label{tab:who_drink_conmat_table}
 \begin{tabular}{rrrr}
@@
                      (columns) for the USEPA adult drinking water advisory.} 
-\label{adult_conmat_table}
+\label{tab:adult_conmat_table}
 \begin{tabular}{rrrr}
@@
                      (columns) for the WHO recreational water advisory.} 
-\label{who_rec_conmat_table}
+\label{tab:who_rec_conmat_table}
 \begin{tabular}{rrrr}

### 2016-01-29 manuscript/f1000_submission/manuscript.Rmd
added keywords and other f1000 editorial comments

@@
 spacing: doublespacing
+keywords:
+-  keyword: Harmful Algal Blooms
+-  keyword: Cyanotoxins
+-  keyword: National Lakes Assessment
+-  keyword: Conditional Probability Analysis
+-  keyword: Cyanobacteria
 abstract: "Cyanobacteria harmful algal blooms (cHABs) are associated with a wide range of adverse health effects that stem mostly from the presence of cyanotoxins.
@@
 Data on chlorophyll *a* and microcystin-LR concentrations are available for `r nrow(na.omit(data.frame(nla_dat$CHLA,nla_dat$MCYST_TL_UGL)))` lakes.
+These data are available for download from the [National Lake Assessment Data Site](http://www.epa.gov/national-aquatic-resource-surveys/data-national-aquatic-resource-surveys)
 
@@
 Thus, for this analysis we focused on the microcystin concentrations that are better represented in the NLA data.
-These were the USEPA children's drinking water advisory level of  0.3 $\mu$g/L (USEPA Child), the WHO drinking water advisory level of 1 $\mu$g/L (WHO Drinking), the USEPA adult drinking water advisory level of 1.6 $\mu$g/L (USEPA Adult), and the WHO recreational, low probability of effect advisory level of 2 $\mu$g/L (WHO Recreational).
+These were the USEPA children's drinking water advisory level of  0.3 $\mu$g/L (USEPA Child), the WHO drinking water advisory level of 1 $\mu$g/L (WHO Drinking), the USEPA adult drinking water advisory level of 1.6 $\mu$g/L (USEPA Adult), and the WHO recreational, low probability of effect advisory level of 2 $\mu$g/L (WHO Recreational) [ @usepa2015drinking; @world2003cyanobacterial; @chorus1999toxic].
 
@@
 The associations between chlorophyll *a* and the upper confidence interval across a range of conditional probability values are shown in Table \ref{tab:mc_chla_table}.
-Specific chlorophyll *a* that are associated with greater than even odds of exceeding the advisory levels were `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]`$\mu$g/L for 0.3, 1.0, 1.6, and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
+Specific chlorophyll *a* concentrations that are associated with greater than even odds of exceeding the advisory levels were `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]`$\mu$g/L for 0.3, 1.0, 1.6, and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
 
@@
 In those cases, which are likely to be common at least in the near future, using a more ubiquitous measurement, such as chlorophyll *a* will provide a reasonable proxy for the probability of exceeding a microcystin-LR health advisory level and provide better protection against adverse effects in both drinking and recreational use cases.
+
+#Software and Data Availability
+Data and Latest Source Code: [https://github.com/USEPA/microcystinchla](https://github.com/USEPA/microcystinchla)
+
+Archived data and source code at time of publication: []()
+
+License: Creative Commons Zero 1.0: [http://creativecommons.org/publicdomain/zero/1.0/](http://creativecommons.org/publicdomain/zero/1.0/)
+
+#Author Contributions
+JH and BK conceived of and conducted the analysis.  Both authors reviewed the results and contributed to writing the manuscript.
+
+#Competing Interests
+No competing interests were disclosed.
+
+#Grant Information
+The author(s) declared that no grants were involved in supporting this work.
 

### 2016-01-29 manuscript/f1000_submission/manuscript.tex
added keywords and other f1000 editorial comments

@@
 \end{abstract}
+
+\textbf{Keywords:}
 
@@
 {[}13{]}. Data on chlorophyll \emph{a} and microcystin-LR concentrations
-are available for 1028 lakes.
+are available for 1028 lakes. These data are available for download from
+the
+\href{http://www.epa.gov/national-aquatic-resource-surveys/data-national-aquatic-resource-surveys}{National
+Lake Assessment Data Site}
 
@@
 recreational, low probability of effect advisory level of 2 \(\mu\)g/L
-(WHO Recreational).
+(WHO Recreational) {[}6--8{]}.
 
@@
 probability values are shown in Table \ref{tab:mc_chla_table}. Specific
-chlorophyll \emph{a} that are associated with greater than even odds of
-exceeding the advisory levels were 0.07, 0.07, 2.98, and 10.67\(\mu\)g/L
-for 0.3, 1.0, 1.6, and 2.0 \(\mu\)g/L advisory levels, respectively
-(Table \ref{tab:mc_chla_table} \& Figure \ref{fig:multi_cp_plot}).
+chlorophyll \emph{a} concentrations that are associated with greater
+than even odds of exceeding the advisory levels were 0.07, 0.07, 2.98,
+and 10.67\(\mu\)g/L for 0.3, 1.0, 1.6, and 2.0 \(\mu\)g/L advisory
+levels, respectively (Table \ref{tab:mc_chla_table} \& Figure
+\ref{fig:multi_cp_plot}).
 
@@
 effects in both drinking and recreational use cases.
+
+\section{Software and Data
+Availability}\label{software-and-data-availability}
+
+Data and Latest Source Code:
+\href{https://github.com/USEPA/microcystinchla}{\url{https://github.com/USEPA/microcystinchla}}
+
+Archived data and source code at time of publication: \href{}{}
+
+License: Creative Commons Zero 1.0:
+\href{http://creativecommons.org/publicdomain/zero/1.0/}{\url{http://creativecommons.org/publicdomain/zero/1.0/}}
+
+\section{Author Contributions}\label{author-contributions}
+
+JH and BK conceived of and conducted the analysis. Both authors reviewed
+the results and contributed to writing the manuscript.
+
+\section{Competing Interests}\label{competing-interests}
+
+No competing interests were disclosed.
+
+\section{Grant Information}\label{grant-information}
+
+The author(s) declared that no grants were involved in supporting this
+work.
 
@@
 
-3. Bloom HA, Research H, Act CA (2014) Harmful Algal Bloom and Hypoxia
-Research and Control Amendments Act of 2014. (S. 1254). Available:
+3. Harmful Algal Bloom and Hypoxia Research and Control Amendments Act
+(2014) Harmful Algal Bloom and Hypoxia Research and Control Amendments
+Act of 2014. (S. 1254). Available:
 \url{https://www.congress.gov/113/bills/s1254/BILLS-113s1254enr.pdf}.

### 2016-01-29 manuscript/f1000_submission/manuscript.Rmd
submission with archived DOI

@@
 #Software and Data Availability
-Data and Latest Source Code: [https://github.com/USEPA/microcystinchla](https://github.com/USEPA/microcystinchla)
 
-Archived data and source code at time of publication: []()
+Data and Latest Source Code:
+[https://github.com/USEPA/microcystinchla](https://github.com/USEPA/microcystinchla)
 
-License: Creative Commons Zero 1.0: [http://creativecommons.org/publicdomain/zero/1.0/](http://creativecommons.org/publicdomain/zero/1.0/)
+Archived data and source code at time of publication: [http://dx.doi.org/10.5281/zenodo.45317](http://dx.doi.org/10.5281/zenodo.45317)
+
+License: 
+Creative Commons Zero 1.0: [http://creativecommons.org/publicdomain/zero/1.0/](http://creativecommons.org/publicdomain/zero/1.0/)
 
 #Author Contributions
+
 JH and BK conceived of and conducted the analysis.  Both authors reviewed the results and contributed to writing the manuscript.
@@
 #Competing Interests
+
 No competing interests were disclosed.
@@
 #Grant Information
+
 The author(s) declared that no grants were involved in supporting this work.

### 2016-01-29 manuscript/f1000_submission/manuscript.tex
submission release for DOI minting.
submission with archived DOI

@@
 
-\textbf{Keywords:}
+\textbf{Keywords: }
+Harmful Algal Blooms;
+Cyanotoxins;
+National Lakes Assessment;
+Conditional Probability Analysis;
+Cyanobacteria;
 
@@
 
-Archived data and source code at time of publication: \href{}{}
+Archived data and source code at time of publication:
+\href{http://dx.doi.org/10.5281/zenodo.45317}{\url{http://dx.doi.org/10.5281/zenodo.45317}}
 

### 2016-05-25 R/plot_scatter.R
updated plot_scatter

@@ -7,7 +7,7 @@
 
 plot_scatter<-function(df,xvar,yvar,cat=NULL,pt_col=1,...){
   options(scipen=5)
-  xdf<-data.frame(xvar=df[[xvar]],yvar=df[[yvar]],pt_col=1)
+  xdf<-data.frame(xvar=df[[xvar]],yvar=df[[yvar]],pt_col=pt_col)
   if(!is.null(cat)){
     xdf<-data.frame(xdf,pt_col=df[[cat]])
   }

### 2016-05-25 manuscript/manuscript.Rmd
edits coming along

@@
 ---
-title: Associations between chlorophyll *a* and various microcystin-LR health advisory
+title: Associations between chlorophyll *a* and various microcystin health advisory
   Concentrations
@@
 Chlorophyll *a* is commonly measured, can be estimated *in situ*, and has been shown to be positively associated with microcystin.
-In this paper, we use this association to provide estimates of chlorophyll *a* concentrations that are indicative of a higher probability of exceeding select health advisory concentrations for microcystin-LR.
+In this paper, we use this association to provide estimates of chlorophyll *a* concentrations that are indicative of a higher probability of exceeding select health advisory concentrations for microcystin.
 Using the 2007 National Lakes Assessment and a conditional probability approach, we identify chlorophyll *a* concentrations that are more likely than not to be associated with an exceedance of a microcystin health advisory level.
@@
 Over the last decade, numerous events and legislative activities have raised the public awareness of harmful algal blooms [@jetoo2015toledo; @rinta2009lake; @HABHRCA2014].
-In response the US Environmental Protection Agency (USEPA) has recently released suggested microcystin-LR (one of the more common toxins) concentrations that would trigger health advisories [@mcelhiney2005detection; @zurawell2005hepatotoxic; @usepa2015drinking].
+In response the US Environmental Protection Agency (USEPA) has recently released suggested microcystin (one of the more common toxins) concentrations that would trigger health advisories [@mcelhiney2005detection; @zurawell2005hepatotoxic; @usepa2015drinking].
 Additionally, the World Health Organization (WHO) has proposed microcystin advisory levels for drinking water and a range of recreational risk levels [@world2003cyanobacterial; @chorus1999toxic].
@@
 
-One of these complications is that they rely on available measurements of microcystin-LR.
-While laboratory testing (e.g., chromatography) remains the gold standard for quantifying microcystin-LR concentrations in water samples, several field test kits have been developed.
+One of these complications is that they rely on available measurements of microcystin.
+While laboratory testing (e.g., chromatography) remains the gold standard for quantifying microcystin concentrations in water samples, several field test kits have been developed.
 Even though field tests provide a much needed means for rapid assessment, they are not yet widely used and are moderately expensive (approximately $150-$200 depending on specific kit) with a limited shelf life (typically one year) [@james2011environmental; @aranda2015evaluation].
@@
 Furthermore, they identify chlorophyll*a* and total nitrogen concentrations that are associated with exceeding 1 $\mu$g/L of microcystin.
-Given these facts, it should be possible to identify chlorophyll *a* concentrations that would be associated with the new USEPA microcystin-LR health advisory levels for drinking water.
-Identifying these associations would provide another tool for water resource managers to help manage the threat to public health posed by cHABs and would be especially useful in the absence of measured microcystin-LR concentrations.
+Given these facts, it should be possible to identify chlorophyll *a* concentrations that would be associated with the new USEPA microcystin health advisory levels for drinking water.
+Identifying these associations would provide another tool for water resource managers to help manage the threat to public health posed by cHABs and would be especially useful in the absence of measured microcystin concentrations.
 
-In this paper we build on past efforts and utilize the National Lakes Assessment (NLA) data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin-LR health advisory concentrations [@usepa2009national; @usepa2015drinking; @chorus1999toxic].
+In this paper we build on past efforts and utilize the National Lakes Assessment (NLA) data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin health advisory concentrations [@usepa2009national; @usepa2015drinking; @chorus1999toxic].
 We add to past studies by exploring associations with newly announced advisory levels and by also applying a different method, conditional probability analysis.
@@
 These data represent a snapshot of water quality from the summer of 2007 for the conterminous United States and were collected as part of an ongoing probabilistic monitoring program [@usepa2009national].
+Water quality data, including chlorophyll *a* and microcystin-LR were obtained via an intergated sample taken from the surface of the lake down to 2 meters.
+Samples were taken at the same time from the index site (e.g.
+near the centroid of the lake) and these provide the source for both chlorophyll *a* and microcystin-LR [@usepa2009national, @nla_fieldops].
+
+For our analysis we only used samples that were part of the probability sampling design (i.e.
+no reference samples) and from the first visit to the lake (e.g.
+some lakes were sampled mutliple times).
+The detection limit for microcysin-LR was 0.05 $\mu$g/L.
+Approximately `r round(sum(nla_dat$MCYST_TL_UGL==0.05)/nrow(nla_dat),2)*100`% of lakes reported microcystin-LR at the detection limit.
+For this analysis we retained these values as removing them would erroneously reduce the confidence intervals around the conditional probabilities.
 Data on chlorophyll *a* and microcystin-LR concentrations are available for `r nrow(na.omit(data.frame(nla_dat$CHLA,nla_dat$MCYST_TL_UGL)))` lakes.
@@
 ##Analytical Methods
-We used a conditional probability analysis (CPA) approach to explore associations between chlorophyll *a* concentrations and World Health Organization (WHO) and USEPA microcystin-LR health advisory levels [@paul2011probability].
+We used a conditional probability analysis (CPA) approach to explore associations between chlorophyll *a* concentrations and World Health Organization (WHO) and USEPA microcystin health advisory levels [@paul2011probability].
 Many health advisory levels have been suggested (Table \ref{tab:microcystin_levels}), but lakes with higher microcystin-LR concentrations in the NLA were rare.
@@
 Thus, for this analysis we focused on the microcystin concentrations that are better represented in the NLA data.
-These were the USEPA children's drinking water advisory level of  0.3 $\mu$g/L (USEPA Child), the WHO drinking water advisory level of 1 $\mu$g/L (WHO Drinking), the USEPA adult drinking water advisory level of 1.6 $\mu$g/L (USEPA Adult), and the WHO recreational, low probability of effect advisory level of 2 $\mu$g/L (WHO Recreational).
+These were the USEPA children's (i.e.
+bottle fed infants to pre-school age children) drinking water advisory level of  0.3 $\mu$g/L (USEPA Child), the WHO drinking water advisory level of 1 $\mu$g/L (WHO Drinking), the USEPA adult (i.e.
+beyond pre-school aged individuals) drinking water advisory level of 1.6 $\mu$g/L (USEPA Adult), and the WHO recreational, low probability of effect advisory level of 2 $\mu$g/L (WHO Recreational).
 
@@
 For this analysis, we used CPA to examine how the conditional probability of exceeding one of the health advisories changes as chlorophyll *a* increases in a lake.
-We expect to find higher chlorohpyll *a* concentrations to be associated with higher probabilities of exceeding the microcystin-LR health advisory levels.
+We expect to find higher chlorohpyll *a* concentrations to be associated with higher probabilities of exceeding the microcystin health advisory levels.
 We also calculated bootstrapped 95% confidence intervals (CI) using 1000 bootstrapped samples.
@@
 
-The chlorophyll *a* cutoffs may be used to predict whether or not a lake exceeds the microcystin-LR health advisories.
+The chlorophyll *a* cutoffs may be used to predict whether or not a lake exceeds the microcystin health advisories.
 Doing so allows us to compare the accuracy of the prediction as well as evaluate false negatives.
-Total accuracy of the four cutoffs predicting microcystin-LR exceedances were `r total_accuracy$epa_child*100`% for the USEPA children's advisory, `r total_accuracy$who_drink*100`% for the WHO drinking water advisory,  `r total_accuracy$epa_adult*100`% for the USEPA adult advisory, and `r total_accuracy$who_rec*100`% for the WHO regreational advisory (Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table}, \ref{tab:adult_conmat_table}, & \ref{tab:who_rec_conmat_table}).
+Total accuracy of the four cutoffs predicting microcystin exceedances were `r total_accuracy$epa_child*100`% for the USEPA children's advisory, `r total_accuracy$who_drink*100`% for the WHO drinking water advisory,  `r total_accuracy$epa_adult*100`% for the USEPA adult advisory, and `r total_accuracy$who_rec*100`% for the WHO regreational advisory (Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table}, \ref{tab:adult_conmat_table}, & \ref{tab:who_rec_conmat_table}).
 However, total accuracy is only one part of the prediction performace with which we are concerned.
 
-When using the chlorophyll *a* cutoffs as an indicator of microcystin-LR exceedances, the error that should be avoided is predicting that no exceedance has occurred when in fact it has.
+When using the chlorophyll *a* cutoffs as an indicator of microcystin exceedances, the error that should be avoided is predicting that no exceedance has occurred when in fact it has.
 In other words, we would like to avoid Type II errors and minimize the proportion of false negatives.
 For the four chlorophyll *a* cut-offs we had a proportion of false negatives of `r fn_percent$epa_child*100`%, `r fn_percent$who_drink*100`%, `r fn_percent$epa_adult*100`%, and `r fn_percent$who_rec*100`% for the USEPA children's, the WHO drinking water, the USEPA adult, and the WHO recreational advisories, respectively.
-In each case we missed less than 10% of the lakes that in fact exceeded the microcystin-LR advisory.
+In each case we missed less than 10% of the lakes that in fact exceeded the microcystin advisory.
 While this method performs well with regard to the false negative percentage, it is possible that is a relic of the NLA dataset and testing with additional data would allow us to confirm this result.
@@
 #Discussion
-The association between Log10 microcystin-LR and Log10 chlorophyll *a* shows a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
+The association between log microcystin-LR and log chlorophyll *a* shows a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
 This indicates that, in general, higher concentrations of microcystin-LR almost always co-occur with higher concentrations of chlorophyll *a* yet the inverse is not true.
@@
 
-There are numerous possible uses for the chlorophyll *a* and microcystin-LR advisory cut-off values.
+There are numerous possible uses for the chlorophyll *a* and microcystin advisory cut-off values.
 First, in the absence of microcystin-LR measurements, exceedence of the chlorophyll *a* concentrations could be a trigger for further actions.
@@
 
+The values we propose are national and may miss regional variation in water quality, including, chlorophyll *a* and microcystin-LR [@beaver2014land, @wagner2011landscape].
+A set of regional conditional probabilities would be interesting; however, limiting the analysis to the data available per region would make interpretaion difficult.
+The sample size for each of the conditional probabilities would be reduced (it ranges from `r  min(table(nla_dat$WSA_ECO9))` to `r  max(table(nla_dat$WSA_ECO9))`) and the number of lakes in each region that exceed the microcystin values is also reduced.
+The result is that our confidence in the conditional probabilites would be less (i.e.
+greatly increased confidence intervals) and the relationships less pronounced as we have fewer lakes on which to base the probabilites.
+Thus, this dataset is best for making national scale recommendations.
+
 Lastly, using chlorophyll *a* is not meant as a replacement for testing of microcystin-LR or other toxins.
 It should be used when other, direct measurements of cyanotoxins are not available.
-In those cases, which are likely to be common at least in the near future, using a more ubiquitous measurement, such as chlorophyll *a* will provide a reasonable proxy for the probability of exceeding a microcystin-LR health advisory level and provide better protection against adverse effects in both drinking and recreational use cases.
+In those cases, which are likely to be common at least in the near future, using a more ubiquitous measurement, such as chlorophyll *a* will provide a reasonable proxy for the probability of exceeding a microcystin health advisory level and provide better protection against adverse effects in both drinking and recreational use cases.
 
 #Acknowledgements
-We would like to thank Anne Kuhn, Bryan Milstead, John Kiddon, Joe LiVolsi, Tim Gleason, and Wayne Munns for constructive reviews of this paper.
+We would like to thank Anne Kuhn, Bryan Milstead, John Kiddon, Joe LiVolsi, Tim Gleason, Wayne Munns, and Leslie D’Anglada for constructive reviews of this paper.
+Special thanks to Jason Marion, Alan Wilson, and Zofia Taranu for reviews of the submitted manuscript.
 This paper has not been subjected to Agency review.
@@
 ```{r chla_micro_scatter, echo=FALSE, fig.width=6, fig.cap="Scatterplot showing association betweeen chlorophyll \\textit{a} and microcystin-LR. \\label{fig:chla_micro_scatter}"}
-plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",pt_col = "black",
+plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",pt_col = nla_dat$WSA_ECO9,
              x=expression(paste('Log10 Chl ', italic("a"),' (',mu,'g/L)')),
@@
 <!--
-```{r chla_micro_cdf, echo=FALSE, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
+```{r chla_micro_cdf, echo=FALSE, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin health advisory Levels.
 \\label{fig:chla_micro_cdf}"}
@@
 
-```{r epa_child_cp_plot, echo=FALSE, fig.width=6, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
+```{r epa_child_cp_plot, echo=FALSE, fig.width=6, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin health advisory Levels.
 A.) Plot for USEPA Child (0.3 $\\mu$g/L).
@@
 
-```{r microcystin_levels,echo=FALSE,results='asis', tab.cap="Various suggested microcystin-LR health advisory concentrations. \\label{tab:microcystin_levels}"}
+```{r microcystin_levels,echo=FALSE,results='asis', tab.cap="Various suggested microcystin health advisory concentrations. \\label{tab:microcystin_levels}"}
 source_of_stnd <- c("WHO",
@@
 
-```{r mc_chla_table, results='asis', echo=FALSE, tab.cap="Chlorophyll \\textit{a} concentrations that are associated with a 50% probability of exceeding a microcystin-LR health advisory concentration.
+```{r mc_chla_table, results='asis', echo=FALSE, tab.cap="Chlorophyll \\textit{a} concentrations that are associated with a 50% probability of exceeding a microcystin health advisory concentration.
 \\label{tab:mc_chla_table}"}

### 2016-05-25 R/get_nla.R

@@ -9,7 +9,9 @@
 #' @examples
 #' nla_dat <- get_nla("http://water.epa.gov/type/lakes/assessmonitor/lakessurvey/upload/NLA2007_Recreational_ConditionEstimates_20091123.csv")
 get_nla<-function(data_url){
-  dat <- read.csv(data_url) %>%
+  x <- tempfile()
+  download.file(data_url,x)
+  dat <- read.csv(x) %>%
     filter(VISIT_NO == 1) %>%
     filter(SITE_TYPE == "PROB_Lake") %>%
     select(SITE_ID,LAT_DD,LON_DD,ST,EPA_REG,WSA_ECO3,WSA_ECO9,LAKE_ORIGIN,CHLA,

### 2016-05-25 R/plot_scatter.R

@@ -13,7 +13,7 @@ plot_scatter<-function(df,xvar,yvar,cat=NULL,pt_col=1,...){
   }
   xdf<-xdf[complete.cases(xdf),]
   x <- ggplot(xdf,aes(xvar,yvar))+
-    geom_point(size=3) +
+    geom_point(size=3,aes(color = pt_col)) +
     theme_bw()+
     theme(text = element_text(family="sans"),
           axis.title.x = element_text(family="sans",vjust = -0.5, size = 14),

### 2016-05-25 manuscript/manuscript.Rmd
more edits!

@@
 
-<!--
+<!--marion2012vivo
 %\VignetteEngine{knitr::rmarkdown}
@@
                   }
-                })
+                })https://github.com/USEPA/Microcystinchla/blob/master/R/get_nla.R
 default_output_hook = knit_hooks$get("output")
@@
 mc_chla<-cbind(epa_child,who_drink,epa_adult,who_rec)
-
+mc_chla_rnd <- mc_chla[mc_chla>0.1]
 #Agreement Tables
@@
 total_accuracy<-NULL
-total_accuracy$epa_child<-round((child_conmat[1,1]+child_conmat[2,2])/
+total_accuracy$epa_child<-https://github.com/USEPA/Microcystinchla/blob/master/R/get_nla.R((child_conmat[1,1]+child_conmat[2,2])/
   sum(child_conmat),2)
@@
 
-Fortunately, microcystin-LR has been shown to be associated with several other, more commonly measured and well understood components of water quality that are readily assessed in the field.
+Fortunately, cyanobacteria and microcystin-LR has been shown to be associated with several other, more commonly measured and well understood components of water quality that are readily assessed in the field [e.g.
+@yuan2015deriving].
 For instance, there are small or hand held fluorometers that measure chlorohpyll *a*.
@@
 For chlorophyll *a*, the range was `r min(nla_dat$CHLA,na.rm=T)` to `r max(nla_dat$CHLA, na.rm=T)` $\mu$g/L.
-All lakes had reported chlorophyll *a* concentrations that exceeded detection limits.
+All lakes had detectable levels of chlorophyll *a*.
 The median concentration was `r median(nla_dat$CHLA,na.rm=T)` $\mu$g/L and the mean was `r round(mean(nla_dat$CHLA,na.rm=T),2)` $\mu$g/L.
@@
 Doing so allows us to compare the accuracy of the prediction as well as evaluate false negatives.
-Total accuracy of the four cutoffs predicting microcystin exceedances were `r total_accuracy$epa_child*100`% for the USEPA children's advisory, `r total_accuracy$who_drink*100`% for the WHO drinking water advisory,  `r total_accuracy$epa_adult*100`% for the USEPA adult advisory, and `r total_accuracy$who_rec*100`% for the WHO regreational advisory (Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table}, \ref{tab:adult_conmat_table}, & \ref{tab:who_rec_conmat_table}).
+Total accuracy of the four cutoffs predicting microcystin exceedances were `r total_accuracy$epa_child*100`% for the USEPA children's drinking water advisory, `r total_accuracy$who_drink*100`% for the WHO drinking water advisory,  `r total_accuracy$epa_adult*100`% for the USEPA adult drinking water advisory, and `r total_accuracy$who_rec*100`% for the WHO regreational advisory (Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table}, \ref{tab:adult_conmat_table}, & \ref{tab:who_rec_conmat_table}).
 However, total accuracy is only one part of the prediction performace with which we are concerned.
@@
 #Discussion
-The association between log microcystin-LR and log chlorophyll *a* shows a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
-This indicates that, in general, higher concentrations of microcystin-LR almost always co-occur with higher concentrations of chlorophyll *a* yet the inverse is not true.
+The log-log association between microcystin-LR and chlorophyll *a* indicates that, in general, higher concentrations of microcystin-LR almost always co-occur with higher concentrations of chlorophyll *a* yet the inverse is not true (Figure \ref{fig:chla_micro_scatter}).
 Higher chlorophyll *a* is not necessarily predictive of higher microcystin-LR concentrations; however, chlorophyll *a* may be predictive of the probability of exceeding a certain threshold.
@@
 As harmful algal blooms are made up of many species and have various mechanisms responsible for adverse impacts (e.g., toxins, hypoxia, odors), there is no single definition of a bloom.
-For cHABs, one approach has been to identify an increase over a baseline concentration of phycocyanin [@miller2013spatiotemporal].
-This is a useful approach for targeted studies, but phycocyanin is also not always available and measures the predominance of cyanobacterial pigments and not toxins.
-Using our chlorophyll *a* cutoffs provides a value that is more directly associated with microcystin-LR and can be used to classify lakes, from past surveys, as having bloomed.
+For cHABs, one approach has been to utilize phycocyanin to screen for or identify bloom events [@miller2013spatiotemporal, ahn2007alternative, @marion2012vivo].
+This is a useful approach, but phycocyanin is not always available, thus limiting its utility especially for examining historical data.
+Using our chlorophyll *a* cutoffs provides a value that is also associated with microcystin-LR and can be used to classify lakes, from past surveys, as having bloomed.
 
@@
 A set of regional conditional probabilities would be interesting; however, limiting the analysis to the data available per region would make interpretaion difficult.
-The sample size for each of the conditional probabilities would be reduced (it ranges from `r  min(table(nla_dat$WSA_ECO9))` to `r  max(table(nla_dat$WSA_ECO9))`) and the number of lakes in each region that exceed the microcystin values is also reduced.
+The sample size for each of the conditional probabilities would be reduced (it ranges from `r  min(table(nla_dat$WSA_EChttps://github.com/USEPA/Microcystinchla/blob/master/R/get_nla.RO9))` to `r  max(table(nla_dat$WSA_ECO9))`) and the number of lakes in each region that exceed the microcystin values is also reduced.
 The result is that our confidence in the conditional probabilites would be less (i.e.
@@
 Thus, this dataset is best for making national scale recommendations.
+
+Two other limitations with the 2007 National Lakes Assessment data are that they represent a single sample from that lake and do not capture temporal dynamics and without subsetting the data do not provide us the ability to validate the presence of microcystin-LR.
+As of this writing, the 2012 National Lakes Assessment data are not public.
+When these data are released, a validation of this approach can be completed then.
 

### 2016-05-26 manuscript/manuscript.Rmd
more edits and data

@@
                   }
-                })https://github.com/USEPA/Microcystinchla/blob/master/R/get_nla.R
+                })
 default_output_hook = knit_hooks$get("output")
@@
   epa_child[i] <- get_chla(epa_child_cp,cps[i])[1]
+  if(epa_child[i]>0.1){epa_child[i]<-round(epa_child[i])}
   who_drink[i] <- get_chla(who_drink_cp,cps[i])[1]
+  if(who_drink[i]>0.1){who_drink[i]<-round(who_drink[i])}
   epa_adult[i] <- get_chla(epa_adult_cp,cps[i])[1]
+  if(epa_adult[i]>0.1){epa_adult[i]<-round(epa_adult[i])}
   who_rec[i] <- get_chla(who_rec_low1_cp,cps[i])[1]
+  if(who_rec[i]>0.1){who_rec[i]<-round(who_rec[i])}
 }
-mc_chla<-cbind(epa_child,who_drink,epa_adult,who_rec)
-mc_chla_rnd <- mc_chla[mc_chla>0.1]
+
+mc_chla<-cbind(as.character(epa_child),as.character(who_drink),
+               as.character(epa_adult),as.character(who_rec))
+
 #Agreement Tables
@@
                    "1.6 $\\mu$g/L")
+order <- c(4,5,6,7,2,1,3)
 #note <- c("Low probability of adverse health effects",
@@
 #          "Individuals School age and older")
-mc_conc_table <- data.frame(source_of_stnd, type_of_stnd, concentration) %>%
-  arrange(type_of_stnd)
+mc_conc_table <- data.frame(source_of_stnd, type_of_stnd, concentration,order) %>%
+  arrange(order) %>%
+  select(source_of_stnd,type_of_stnd,concentration)
 names(mc_conc_table)<-c("Source","Type","Concentration")
@@
 mc_chla_tbl <- data.frame(seq(0.1,0.9,0.1),mc_chla)
-names(mc_chla_tbl)<-c("Cond. Probability","USEPA Child", "WHO Drink", 
-                      "USEPA Adult", "WHO Recreational")
+names(mc_chla_tbl)<-c("Cond. Probability","USEPA Child (0.3 $\\mu$g/L)", "WHO Drink (1 $\\mu$g/L)", "USEPA Adult (1.6 $\\mu$g/L)", "WHO Recreational (2 $\\mu$g/L) ")
 #pandoc.table(mc_chla_tbl,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))

### 2016-05-26 manuscript/manuscript.Rmd
edits done.  Need to transfer to overleaf now...

@@
 total_accuracy<-NULL
-total_accuracy$epa_child<-https://github.com/USEPA/Microcystinchla/blob/master/R/get_nla.R((child_conmat[1,1]+child_conmat[2,2])/
+total_accuracy$epa_child<-round((child_conmat[1,1]+child_conmat[2,2])/
   sum(child_conmat),2)
@@
 Additionally, chlorophyll *a* is a very commonly measured component of water quality that is also known to be positively associated with microsystin-LR concentrations [@pip2014microcystin; @yuan2014managing].
-Yuan et.
-al [-@yuan2014managing] explore these associations in detail and control for other related variables.
+Recently, Yuan et.
+al [-@yuan2014managing] explored these associations in detail and controlled for other related variables.
 In their analysis they find that total nitrogen and chlorophyll *a* show the strongest association with microcystin.
 Furthermore, they identify chlorophyll*a* and total nitrogen concentrations that are associated with exceeding 1 $\mu$g/L of microcystin.
-Given these facts, it should be possible to identify chlorophyll *a* concentrations that would be associated with the new USEPA microcystin health advisory levels for drinking water.
-Identifying these associations would provide another tool for water resource managers to help manage the threat to public health posed by cHABs and would be especially useful in the absence of measured microcystin concentrations.
+These findings suggest that chlorophyll a concentrations could also track the new USEPA microcystin health advisory levels for drinking water.
+Identifying this association would provide an important tool for water resource managers to help manage the threat to public health posed by cHABs and would be especially useful in the absence of measured microcystin concentrations.
 
 In this paper we build on past efforts and utilize the National Lakes Assessment (NLA) data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin health advisory concentrations [@usepa2009national; @usepa2015drinking; @chorus1999toxic].
-We add to past studies by exploring associations with newly announced advisory levels and by also applying a different method, conditional probability analysis.
+We build on past studies by exploring associations with the newly announced advisory levels and by also applying a different method, conditional probability analysis.
 Utilizing different methods strengthens the evidence for suggested chlorophyll *a* levels that are associated with increased risk of exceeding the health advisory levels as those levels are not predicated on a single analytical method.
@@
 We also calculated bootstrapped 95% confidence intervals (CI) using 1000 bootstrapped samples.
-Thus, to identify chlorophyll *a* concentrations of concern we identfiy the value of the upper 95% CI across a range of conditional probabilities of exceeding each health advisory level.
+Thus, to identify chlorophyll *a* concentrations of concern we identified the value of the upper 95% CI across a range of conditional probabilities of exceeding each health advisory level.
 Using the upper confidence limit to identify a threshold is justified as it ensures that a given threshold is unlikely to miss a microcystin exceedance.
 
-As both microcystin-LR and chlorophyll *a* values were highely skewed right, a log base 10 transformation was used.
+As both microcystin-LR and chlorophyll *a* values were highly right skewed, a log base 10 transformation was used.
 Additional details of the specific implementation are available at https://github.com/USEPA/microcystinchla.
@@
 
-Lastly, we assess the ability of these chlorophyll *a* thresholds to predict microcystin exceedance.
-We use error matrices and calculate total accuracy as well as the proportion of false negatives.
+Lastly, we assessed the ability of these chlorophyll *a* thresholds to predict microcystin exceedance.
+We used error matrices and calculate total accuracy as well as the proportion of false negatives.
 Total accuracy is the total number of correct predictions divided by total observations.
@@
 Of all lakes sampled, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=0.3])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the USEPA Child level, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1.6])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the USEPA Adult level, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO Drinking level, and `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=2])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO Recreational level.
-For chlorophyll *a*, the range was `r min(nla_dat$CHLA,na.rm=T)` to `r max(nla_dat$CHLA, na.rm=T)` $\mu$g/L.
+Chlorophyll *a*, ranged from `r min(nla_dat$CHLA,na.rm=T)` to `r max(nla_dat$CHLA, na.rm=T)` $\mu$g/L and this captures the range of trrophic states from oligotrophic to hypereutrophic.
 All lakes had detectable levels of chlorophyll *a*.
 The median concentration was `r median(nla_dat$CHLA,na.rm=T)` $\mu$g/L and the mean was `r round(mean(nla_dat$CHLA,na.rm=T),2)` $\mu$g/L.
-The associations between chlorophyll *a* and the upper confidence interval across a range of conditional probability values are shown in Table \ref{tab:mc_chla_table}.
+The association between chlorophyll *a* and the upper confidence interval across a range of conditional probability values are shown in Table \ref{tab:mc_chla_table}.
 Specific chlorophyll *a* that are associated with greater than even odds of exceeding the advisory levels were `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]`$\mu$g/L for 0.3, 1.0, 1.6, and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
@@
 
-This is the case as the probability of exceeding each of the four tested health advisory levels increases as a function of chlorophyll *a* concentration (Figure \ref{fig:multi_cp_plot}).
-We used this association to identify chlorophyll *a* concentrations that are associated with a range of probabilities of exceeding a given health advisory level (Table \ref{tab:mc_chla_table}).
+Indeed, the probability of exceeding each of the four tested health advisory levels increased as a function of chlorophyll *a* concentration (Figure \ref{fig:multi_cp_plot}).
+We used this association to identify chlorophyll *a* concentrations that were associated with a range of probabilities of exceeding a given health advisory level (Table \ref{tab:mc_chla_table}).
 For the purposes of this discussion we focus on a conditional probability of 50% or greater (i.e., greater than even odds to exceed a health advisory level).
@@
 A set of regional conditional probabilities would be interesting; however, limiting the analysis to the data available per region would make interpretaion difficult.
-The sample size for each of the conditional probabilities would be reduced (it ranges from `r  min(table(nla_dat$WSA_EChttps://github.com/USEPA/Microcystinchla/blob/master/R/get_nla.RO9))` to `r  max(table(nla_dat$WSA_ECO9))`) and the number of lakes in each region that exceed the microcystin values is also reduced.
+The sample size for each of the conditional probabilities would be reduced (it ranges from `r  min(table(nla_dat$WSA_ECO9))` to `r  max(table(nla_dat$WSA_ECO9))`) and the number of lakes in each region that exceed the microcystin values is also reduced.
 The result is that our confidence in the conditional probabilites would be less (i.e.

### 2016-05-31 manuscript/f1000_submission/manuscript.Rmd
update

@@
 })
+
+set.seed(2042016)
 ```
@@
 total_accuracy$epa_child<-round((child_conmat[1,1]+child_conmat[2,2])/
-  sum(child_conmat),2)
+  sum(child_conmat[1:2,1:2]),2)
 total_accuracy$who_drink<-round((who_drink_conmat[1,1]+who_drink_conmat[2,2])/
-  sum(who_drink_conmat),2)
+  sum(who_drink_conmat[1:2,1:2]),2)
 total_accuracy$epa_adult<-round((adult_conmat[1,1]+adult_conmat[2,2])/
-  sum(adult_conmat),2)
+  sum(adult_conmat[1:2,1:2]),2)
 total_accuracy$who_rec<-round((who_rec_conmat[1,1]+who_rec_conmat[2,2])/
-  sum(who_rec_conmat),2)
+  sum(who_rec_conmat[1:2,1:2]),2)
 
@@
 fn_percent$epa_child<-round((child_conmat[1,2])/
-  sum(child_conmat),2)
+  sum(child_conmat[1:2,1:2]),2)
 fn_percent$who_drink<-round((who_drink_conmat[1,2])/
-  sum(who_drink_conmat),2)
+  sum(who_drink_conmat[1:2,1:2]),2)
 fn_percent$epa_adult<-round((adult_conmat[1,2])/
-  sum(adult_conmat),2)
+  sum(adult_conmat[1:2,1:2]),2)
 fn_percent$who_rec<-round((who_rec_conmat[1,2])/
-  sum(who_rec_conmat),2)
+  sum(who_rec_conmat[1:2,1:2]),2)
 
@@
 The associations between chlorophyll *a* and the upper confidence interval across a range of conditional probability values are shown in Table \ref{tab:mc_chla_table}.
-Specific chlorophyll *a* concentrations that are associated with greater than even odds of exceeding the advisory levels were `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]`$\mu$g/L for 0.3, 1.0, 1.6, and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
+Specific chlorophyll *a* concentrations that are associated with greater than even odds of exceeding the advisory levels were `r mc_chla[5,1]`, `r mc_chla[5,2]`, `r mc_chla[5,3]`, and `r mc_chla[5,4]`$\mu$g/L for 0.3, 1.0, 1.6, and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
 

### 2016-05-31 manuscript/f1000_submission/manuscript.tex
update

@@
 chlorophyll \emph{a} concentrations that are associated with greater
-than even odds of exceeding the advisory levels were 0.07, 0.07, 2.98,
-and 10.67\(\mu\)g/L for 0.3, 1.0, 1.6, and 2.0 \(\mu\)g/L advisory
-levels, respectively (Table \ref{tab:mc_chla_table} \& Figure
+than even odds of exceeding the advisory levels were 23.36, 62.64,
+83.52, and 103.9\(\mu\)g/L for 0.3, 1.0, 1.6, and 2.0 \(\mu\)g/L
+advisory levels, respectively (Table \ref{tab:mc_chla_table} \& Figure
 \ref{fig:multi_cp_plot}).
@@
 negatives. Total accuracy of the four cutoffs predicting microcystin-LR
-exceedances were 19\% for the USEPA children's advisory, 21\% for the
-WHO drinking water advisory, 22\% for the USEPA adult advisory, and 23\%
+exceedances were 75\% for the USEPA children's advisory, 86\% for the
+WHO drinking water advisory, 89\% for the USEPA adult advisory, and 91\%
 for the WHO regreational advisory (Tables \ref{tab:child_conmat_table},
@@
 of false negatives. For the four chlorophyll \emph{a} cut-offs we had a
-proportion of false negatives of 2\%, 2\%, 1\%, and 1\% for the USEPA
+proportion of false negatives of 9\%, 7\%, 6\%, and 5\% for the USEPA
 children's, the WHO drinking water, the USEPA adult, and the WHO
@@
 than even odds to exceed a health advisory level). The 50\% conditional
-probability chlorophyll \emph{a} thresholds represents 27.9\%, 11.3\%,
-9\%, and 6.9\% of sample lakes for the USEPA Child, the WHO Drinking,
+probability chlorophyll \emph{a} thresholds represents 27.9\%, 12.1\%,
+9\%, and 7.3\% of sample lakes for the USEPA Child, the WHO Drinking,
 the USEPA Adult, and the WHO recreational levels, respecitvely.
@@
   \hline
-0.10 & 0.07 & 0.07 & 0.07 & 1.22 \\ 
-  0.20 & 0.07 & 4.43 & 11.80 & 20.00 \\ 
-  0.30 & 2.98 & 20.43 & 31.40 & 53.64 \\ 
-  0.40 & 10.67 & 42.24 & 66.96 & 82.22 \\ 
-  0.50 & 23.36 & 66.96 & 83.52 & 105.84 \\ 
-  0.60 & 38.30 & 103.20 & 133.20 & 155.52 \\ 
-  0.70 & 65.60 & 166.63 & 871.20 & 216.00 \\ 
-  0.80 & 117.50 & 338.40 & 871.20 & 871.20 \\ 
-  0.90 & 167.04 & 516.00 & 871.20 & 871.20 \\ 
+0.10 & 0.07 & 0.07 & 0.07 & 1.65 \\ 
+  0.20 & 0.07 & 4.30 & 9.52 & 20.52 \\ 
+  0.30 & 2.91 & 18.72 & 25.76 & 53.64 \\ 
+  0.40 & 10.88 & 38.02 & 65.60 & 82.22 \\ 
+  0.50 & 23.36 & 62.64 & 83.52 & 103.90 \\ 
+  0.60 & 39.80 & 94.93 & 114.62 & 155.52 \\ 
+  0.70 & 65.60 & 125.40 & 273.60 & 871.20 \\ 
+  0.80 & 125.28 & 314.61 & 871.20 & 871.20 \\ 
+  0.90 & 197.00 & 516.00 & 871.20 & 871.20 \\ 
    \hline
@@
   \hline
-Not Exceed & 839 &  77 & 916 \\ 
-  Exceed &  68 &  44 & 112 \\ 
+Not Exceed & 833 &  75 & 908 \\ 
+  Exceed &  74 &  46 & 120 \\ 
   Column Totals & 907 & 121 & 1028 \\ 
@@
   \hline
-Not Exceed & 910 &  52 & 962 \\ 
-  Exceed &  43 &  23 &  66 \\ 
+Not Exceed & 908 &  50 & 958 \\ 
+  Exceed &  45 &  25 &  70 \\ 
   Column Totals & 953 &  75 & 1028 \\ 

### 2016-05-31 manuscript/manuscript.tex
update

--- 2016-01-17 manuscript/manuscript.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% For more detailed article preparation guidelines, please see:
-% http://f1000research.com/author-guidelines
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Original default rstudio/pandoc latex file
+%% upated by @jhollist 09/15/2014
+%% inspired by @cboetting https://github.com/cboettig/template and
+%% @rmflight blog posts:
+%% http://rmflight.github.io/posts/2014/07/analyses_as_packages.html 
+%% http://rmflight.github.io/posts/2014/07/vignetteAnalysis.html).  
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\documentclass[10pt,a4paper,twocolumn]{article}
-\usepackage{f1000_styles}
-\bibliography{components/manuscript}
+\documentclass[11pt,]{article}
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+    \usepackage{xltxtra,xunicode}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage{longtable,booktabs}
+\usepackage{graphicx}
+% Redefine \includegraphics so that, unless explicit options are
+% given, the image width will not exceed the width of the page.
+% Images get their normal width if they fit onto the page, but
+% are scaled down if they would overflow the margins.
+\makeatletter
+\def\ScaleIfNeeded{%
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+\let\Oldincludegraphics\includegraphics
+{%
+ \catcode`\@=11\relax%
+ \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
+}%
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={Associations between chlorophyll a and various microcystin-LR health advisory Concentrations},
+            colorlinks=true,
+            citecolor=blue,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\setcounter{secnumdepth}{5}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%Changes borrowed from @cboettig, added by @jhollist 
+% A modified page layout 
+\textwidth 6.75in
+\oddsidemargin -0.15in
+\evensidemargin -0.15in
+\textheight 9in
+\topmargin -0.5in
+\usepackage{lineno} % add 
+  \linenumbers % turns line numbering on 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%Packages and layout changes by @jhollist 09/15/2014
+\usepackage{ragged2e}
+\usepackage[font=normalsize]{caption}
+  \usepackage[doublespacing]{setspace}
+\usepackage{parskip}
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+\fancyhf{}
+\renewcommand{\headrulewidth}{0pt}
+\rfoot{\today}
+\lfoot{\thepage}
+%%Changed default abstract width and added lines
+\renewenvironment{abstract}{
+  \hfill\begin{minipage}{1\textwidth}
+  \rule{\textwidth}{1pt}\vspace{5pt}
+  \normalsize
+  \begin{justify}
+  \bfseries\abstractname\vspace{5pt}
+  \end{justify}}
+  {\par\noindent\rule{\textwidth}{1pt}\end{minipage}
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\title{Associations between chlorophyll \emph{a} and various microcystin-LR
+health advisory Concentrations}
+\author{
+Jeffrey W. Hollister
+Betty J. Kreakie
+}
+\date{}
 
 \begin{document}
+%%Edited by @jhollist 09/15/2014
+%%Adds title from YAML
+\begin{singlespace}
+\begin{center}
+\huge Associations between chlorophyll \emph{a} and various microcystin-LR
+health advisory Concentrations
+\end{center}
+%%Adds Author, correspond email asterisk, and affilnum from YAML
+\begin{center}
+\large
+Jeffrey W. Hollister \textsuperscript{*} \textsuperscript{1} 
+Betty J. Kreakie \textsuperscript{1} 
+\end{center}
+%%Adds affiliations from YAML
+\begin{justify}
+\footnotesize \emph{ 
+\\*
+\textsuperscript{1}US Environmental Protection Agency, Office of Research and Development,
+National Health and Environmental Effects Research Laboratory, Atlantic
+Ecology Division, 27 Tarzwell Drive Narragansett, RI, 02882, USA\\*
+}
+%%Adds corresponding author email(s) from YAML
+\newcounter{num}
+\setcounter{num}{1}
+\\[0.1cm]
+\footnotesize \emph{ 
+\ifnum\value{num}=1%
+\textsuperscript{*} corresponding author:
+\fi
+\href{mailto:hollister.jeff@epa.gov}{\nolinkurl{hollister.jeff@epa.gov}}
+\stepcounter{num}
+}
+\end{justify}
+%%Adds date from YAML
+\normalsize
 
-%Still need to figure out how to get corresponding author on here.
-\title{\textit{F1000Research} Associations between chlorophyll \emph{a} and various microcystin-LR
-health advisory Concentrations}
-\author[1]{Jeffrey W. Hollister}
-\author[1]{Betty J. Kreakie}
-\affil[1]{}
-
-\maketitle
-\thispagestyle{fancy}
-
-\begin{abstract}
-
-\end{abstract}
-
-\clearpage
+\end{singlespace}
 

### 2016-05-31 numbers.Rmd
update

---
title: "NLA and Microcystin"
author: "Jeff W. Hollister"
date: "3/1/2016"
output: pdf_document
---

Review of some numbers from the 2007 NLA data on microcystin.

```{r get_data}
nla_url <- "http://www2.epa.gov/sites/production/files/2014-10/nla2007_recreational_conditionestimates_20091123.csv"
nla_dat <- read.csv(nla_url)
```

The raw dataset from the `nla_url` has `r nrow(nla_dat)` rows and `r ncol(nla_dat)`.

Some filtering of this data is required.  For our analyses we prefer to use the probabilistic draw in case we want to use the weighted values.  Do this we need to only use PROB_Lakes and the first visit.

```{r filter_prob}
library(dplyr)
nla_dat <- nla_dat %>%
  filter(SITE_TYPE == "PROB_Lake") %>%
  filter(VISIT_NO == 1) %>%
  select(SITE_ID, CHLA, MCYST_TL_UGL)
```

Filtering this out results in `r nrow(nla_dat)` rows and `r ncol(nla_dat)`.

From this we can see how many lakes had microcystin present.

```{r counts}
sum(nla_dat$MCYST_TL_UGL > 0.05)
sum(nla_dat$MCYST_TL_UGL > 0.05)/nrow(nla_dat)
```

### 2016-05-31 manuscript/components/manuscript.tex
local commit

--- 2016-01-17 manuscript/manuscript.tex
@@
 \documentclass[10pt,a4paper,twocolumn]{article}
-\usepackage{components/f1000_styles}
+\usepackage{f1000_styles}
 \usepackage{hyperref}
 \usepackage{longtable}
-\bibliography{components/manuscript}
+\bibliography{../components/manuscript}
 
@@
 %Still need to figure out how to get corresponding author on here.
-\title{\textit{F1000Research} Associations between chlorophyll \emph{a} and various microcystin-LR
-health advisory Concentrations}
+\title{\textit{Associations between chlorophyll \emph{a} and various microcystin-LR
+health advisory concentrations} }
 \author[1]{Jeffrey W. Hollister}
@@
 chlorophyll \emph{a} that are associated with greater than even odds of
-exceeding the advisory levels were 0.07, 0.07, 2.78, and 9.97\(\mu\)g/L
+exceeding the advisory levels were 0.07, 0.07, 2.98, and 10.67\(\mu\)g/L
 for 0.3, 1.0, 1.6, and 2.0 \(\mu\)g/L advisory levels, respectively
@@
 
-\section{Figures}\label{figures}
+\section{Figure Legends}\label{figure-legends}
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/chla_micro_scatter-1.pdf}
-\caption{Scatterplot showing association betweeen chlorophyll \textit{a}
-and microcystin-LR. \label{fig:chla_micro_scatter}}
-\end{figure}
+Figure 1. Scatterplot showing association betweeen chlorophyll
+\textbackslash{}textit\{a\} and microcystin-LR.
+\textbackslash{}label\{fig:chla\_micro\_scatter\}
+
+Figure 2. Conditional probability plots showing association between the
+probability of exceeding various microcystin-LR (MLR) health advisory
+Levels. A.) Plot for USEPA Child (0.3 \(\\mu\)g/L). B.) Plot for WHO
+Drinking (1 \(\\mu\)g/L). C.) Plot for USEPA Adult (1.6 \(\\mu\)g/L).
+D.) Plot for WHO Recreational (2 \(\\mu\)g/L).
+\textbackslash{}label\{fig:multi\_cp\_plot\}
+
+\begin{verbatim}
+## Saving 6 x 6.8 in image
+\end{verbatim}
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/epa_child_cp_plot-1.pdf}
-\caption{Conditional probability plots showing association between the
-probability of exceeding various microcystin-LR (MLR) health advisory
-Levels. A.) Plot for USEPA Child (0.3 \(\mu\)g/L). B.) Plot for WHO
-Drinking (1 \(\mu\)g/L). C.) Plot for USEPA Adult (1.6 \(\mu\)g/L). D.)
-Plot for WHO Recreational (2 \(\mu\)g/L). \label{fig:multi_cp_plot}}
-\end{figure}
+\begin{verbatim}
+## pdf 
+##   2
+\end{verbatim}
 
@@
 \hline
-0.1 & 0.07 & 0.07 & 0.07 & 1.17\\
+0.1 & 0.07 & 0.07 & 0.07 & 1.22\\
 \hline
-0.2 & 0.07 & 4.48 & 10.67 & 18.24\\
+0.2 & 0.07 & 4.43 & 11.80 & 20.00\\
 \hline
-0.3 & 2.78 & 17.49 & 31.59 & 50.26\\
+0.3 & 2.98 & 20.43 & 31.40 & 53.64\\
 \hline
-0.4 & 9.97 & 38.73 & 69.19 & 76.80\\
+0.4 & 10.67 & 42.24 & 66.96 & 82.22\\
 \hline
-0.5 & 21.92 & 77.20 & 84.96 & 105.84\\
+0.5 & 23.36 & 66.96 & 83.52 & 105.84\\
 \hline
-0.6 & 38.73 & 103.20 & 133.20 & 228.24\\
+0.6 & 38.30 & 103.20 & 133.20 & 155.52\\
 \hline
-0.7 & 63.94 & 138.24 & 244.80 & 871.20\\
+0.7 & 65.60 & 166.63 & 871.20 & 216.00\\
 \hline
-0.8 & 113.14 & 326.88 & 326.88 & 871.20\\
+0.8 & 117.50 & 338.40 & 871.20 & 871.20\\
 \hline
-0.9 & 166.63 & 516.00 & 871.20 & 871.20\\
+0.9 & 167.04 & 516.00 & 871.20 & 871.20\\
 \hline
@@
 \hline
-Not Exceed & 638 & 93 & 731\\
+Not Exceed & 649 & 96 & 745\\
 \hline
-Exceed & 173 & 124 & 297\\
+Exceed & 162 & 121 & 283\\
 \hline
@@
 \hline
-Not Exceed & 850 & 82 & 932\\
+Not Exceed & 839 & 77 & 916\\
 \hline
-Exceed & 57 & 39 & 96\\
+Exceed & 68 & 44 & 112\\
 \hline
@@
 \hline
-Not Exceed & 887 & 57 & 944\\
+Not Exceed & 883 & 57 & 940\\
 \hline
-Exceed & 50 & 34 & 84\\
+Exceed & 54 & 34 & 88\\
 \hline

### 2016-05-31 manuscript/manuscript.tex
local commit

@@
             pdfauthor={},
-            pdftitle={Associations between chlorophyll a and various microcystin-LR health advisory Concentrations},
+            pdftitle={Associations between chlorophyll a and various microcystin health advisory Concentrations},
             colorlinks=true,
@@
 
-\title{Associations between chlorophyll \emph{a} and various microcystin-LR
-health advisory Concentrations}
+\title{Associations between chlorophyll \emph{a} and various microcystin health
+advisory Concentrations}
 \author{
@@
 \begin{center}
-\huge Associations between chlorophyll \emph{a} and various microcystin-LR
-health advisory Concentrations
+\huge Associations between chlorophyll \emph{a} and various microcystin health
+advisory Concentrations
 \end{center}
@@
 concentrations that are indicative of a higher probability of exceeding
-select health advisory concentrations for microcystin-LR. Using the 2007
+select health advisory concentrations for microcystin. Using the 2007
 National Lakes Assessment and a conditional probability approach, we
@@
 microcystin concentrations of 0.3, 1, 1.6, and 2 \(\mu\)g/L is
-associatied with chlorophyll \emph{a} concentration thresholds of 21.92,
-77.2, 84.96, and 105.84, respectively. When managing for these various
+associatied with chlorophyll \emph{a} concentration thresholds of 23,
+64, 84, and 103, respectively. When managing for these various
 microcystin levels, exceeding these reported chlorophyll \emph{a}
@@
 response the US Environmental Protection Agency (USEPA) has recently
-released suggested microcystin-LR (one of the more common toxins)
+released suggested microcystin (one of the more common toxins)
 concentrations that would trigger health advisories {[}4--6{]}.
@@
 One of these complications is that they rely on available measurements
-of microcystin-LR. While laboratory testing (e.g., chromatography)
-remains the gold standard for quantifying microcystin-LR concentrations
-in water samples, several field test kits have been developed. Even
-though field tests provide a much needed means for rapid assessment,
-they are not yet widely used and are moderately expensive (approximately
-\$150-\$200 depending on specific kit) with a limited shelf life
-(typically one year) {[}9,10{]}. Additionally, each technique requires
-nuanced understanding of the detection method (e.g., limit of detection,
+of microcystin. While laboratory testing (e.g., chromatography) remains
+the gold standard for quantifying microcystin concentrations in water
+samples, several field test kits have been developed. Even though field
+tests provide a much needed means for rapid assessment, they are not yet
+widely used and are moderately expensive (approximately \$150-\$200
+depending on specific kit) with a limited shelf life (typically one
+year) {[}9,10{]}. Additionally, each technique requires nuanced
+understanding of the detection method (e.g., limit of detection,
 specific microcystin variants being measured, and sampling protocol).
 
-Fortunately, microcystin-LR has been shown to be associated with several
-other, more commonly measured and well understood components of water
-quality that are readily assessed in the field. For instance, there are
-small or hand held fluorometers that measure chlorohpyll \emph{a}.
-Additionally, chlorophyll \emph{a} is a very commonly measured component
-of water quality that is also known to be positively associated with
-microsystin-LR concentrations {[}11,12{]}. Yuan et. al {[}12{]} explore
-these associations in detail and control for other related variables. In
-their analysis they find that total nitrogen and chlorophyll \emph{a}
-show the strongest association with microcystin. Furthermore, they
-identify chlorophyll\emph{a} and total nitrogen concentrations that are
-associated with exceeding 1 \(\mu\)g/L of microcystin. Given these
-facts, it should be possible to identify chlorophyll \emph{a}
-concentrations that would be associated with the new USEPA
-microcystin-LR health advisory levels for drinking water. Identifying
-these associations would provide another tool for water resource
-managers to help manage the threat to public health posed by cHABs and
-would be especially useful in the absence of measured microcystin-LR
-concentrations.
+Fortunately, cyanobacteria and microcystin-LR has been shown to be
+associated with several other, more commonly measured and well
+understood components of water quality that are readily assessed in the
+field {[}11{]}. For instance, there are small or hand held fluorometers
+that measure chlorohpyll \emph{a}. Additionally, chlorophyll \emph{a} is
+a very commonly measured component of water quality that is also known
+to be positively associated with microsystin-LR concentrations
+{[}12,13{]}. Recently, Yuan et. al {[}13{]} explored these associations
+in detail and controlled for other related variables. In their analysis
+they find that total nitrogen and chlorophyll \emph{a} show the
+strongest association with microcystin. Furthermore, they identify
+chlorophyll\emph{a} and total nitrogen concentrations that are
+associated with exceeding 1 \(\mu\)g/L of microcystin. These findings
+suggest that chlorophyll a concentrations could also track the new USEPA
+microcystin health advisory levels for drinking water. Identifying this
+association would provide an important tool for water resource managers
+to help manage the threat to public health posed by cHABs and would be
+especially useful in the absence of measured microcystin concentrations.
 
@@
 that are associated with higher probabilities of exceeding several
-microcystin-LR health advisory concentrations {[}6,8,13{]}. We add to
-past studies by exploring associations with newly announced advisory
+microcystin health advisory concentrations {[}6,8,14{]}. We build on
+past studies by exploring associations with the newly announced advisory
 levels and by also applying a different method, conditional probability
@@
 available via
-\href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USEPA/microcystinchla}}.
+\href{https://github.com/USAPE/microcystinchla}{https://github.com/USEPA/microcystinchla}.
 
@@
 We used the 2007 NLA chlorophyll \emph{a} and microcystin-LR
-concentration data {[}13{]}. These data represent a snapshot of water
+concentration data {[}14{]}. These data represent a snapshot of water
 quality from the summer of 2007 for the conterminous United States and
 were collected as part of an ongoing probabilistic monitoring program
-{[}13{]}. Data on chlorophyll \emph{a} and microcystin-LR concentrations
-are available for 1028 lakes.
+{[}14{]}. Water quality data, including chlorophyll \emph{a} and
+microcystin-LR were obtained via an intergated sample taken from the
+surface of the lake down to 2 meters. Samples were taken at the same
+time from the index site (e.g.~near the centroid of the lake) and these
+provide the source for both chlorophyll \emph{a} and microcystin-LR
+{[}14{]}.
+
+For our analysis we only used samples that were part of the probability
+sampling design (i.e.~no reference samples) and from the first visit to
+the lake (e.g.~some lakes were sampled mutliple times). The detection
+limit for microcysin-LR was 0.05 \(\mu\)g/L. Approximately 67\% of lakes
+reported microcystin-LR at the detection limit. For this analysis we
+retained these values as removing them would erroneously reduce the
+confidence intervals around the conditional probabilities. Data on
+chlorophyll \emph{a} and microcystin-LR concentrations are available for
+1028 lakes.
 
@@
 associations between chlorophyll \emph{a} concentrations and World
-Health Organization (WHO) and USEPA microcystin-LR health advisory
-levels {[}14{]}. Many health advisory levels have been suggested (Table
+Health Organization (WHO) and USEPA microcystin health advisory levels
+{[}16{]}. Many health advisory levels have been suggested (Table
 \ref{tab:microcystin_levels}), but lakes with higher microcystin-LR
@@
 focused on the microcystin concentrations that are better represented in
-the NLA data. These were the USEPA children's drinking water advisory
-level of 0.3 \(\mu\)g/L (USEPA Child), the WHO drinking water advisory
-level of 1 \(\mu\)g/L (WHO Drinking), the USEPA adult drinking water
-advisory level of 1.6 \(\mu\)g/L (USEPA Adult), and the WHO
-recreational, low probability of effect advisory level of 2 \(\mu\)g/L
-(WHO Recreational).
+the NLA data. These were the USEPA children's (i.e.~bottle fed infants
+to pre-school age children) drinking water advisory level of 0.3
+\(\mu\)g/L (USEPA Child), the WHO drinking water advisory level of 1
+\(\mu\)g/L (WHO Drinking), the USEPA adult (i.e.~beyond pre-school aged
+individuals) drinking water advisory level of 1.6 \(\mu\)g/L (USEPA
+Adult), and the WHO recreational, low probability of effect advisory
+level of 2 \(\mu\)g/L (WHO Recreational).
 
@@
 chlorohpyll \emph{a} concentrations to be associated with higher
-probabilities of exceeding the microcystin-LR health advisory levels. We
+probabilities of exceeding the microcystin health advisory levels. We
 also calculated bootstrapped 95\% confidence intervals (CI) using 1000
 bootstrapped samples. Thus, to identify chlorophyll \emph{a}
-concentrations of concern we identfiy the value of the upper 95\% CI
+concentrations of concern we identified the value of the upper 95\% CI
 across a range of conditional probabilities of exceeding each health
@@
 
-As both microcystin-LR and chlorophyll \emph{a} values were highely
-skewed right, a log base 10 transformation was used. Additional details
-of the specific implementation are available at
+As both microcystin-LR and chlorophyll \emph{a} values were highly right
+skewed, a log base 10 transformation was used. Additional details of the
+specific implementation are available at
 \url{https://github.com/USEPA/microcystinchla}. A more detailed
 discussion of CPA is beyond the scope of this paper, but see Paul et al.
-{[}15{]} and Hollister et al. {[}16{]} for greater detail. Lastly, all
+{[}17{]} and Hollister et al. {[}18{]} for greater detail. Lastly, all
 analyses were conducted using R version 3.2.2 and code and data from
 this analysis are freely available as an R package at
-\href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USEPA/microcystinchla}}.
+\href{https://github.com/USAPE/microcystinchla}{https://github.com/USEPA/microcystinchla}.
 
-Lastly, we assess the ability of these chlorophyll \emph{a} thresholds
-to predict microcystin exceedance. We use error matrices and calculate
+Lastly, we assessed the ability of these chlorophyll \emph{a} thresholds
+to predict microcystin exceedance. We used error matrices and calculate
 total accuracy as well as the proportion of false negatives. Total
@@
 Drinking level, and 7.3\% of lakes exceeded the WHO Recreational level.
-For chlorophyll \emph{a}, the range was 0.07 to 936 \(\mu\)g/L. All
-lakes had reported chlorophyll \emph{a} concentrations that exceeded
-detection limits. The median concentration was 7.79 \(\mu\)g/L and the
-mean was 29.63 \(\mu\)g/L. The associations between chlorophyll \emph{a}
-and the upper confidence interval across a range of conditional
-probability values are shown in Table \ref{tab:mc_chla_table}. Specific
-chlorophyll \emph{a} that are associated with greater than even odds of
-exceeding the advisory levels were 0.07, 0.07, 2.78, and 9.97\(\mu\)g/L
-for 0.3, 1.0, 1.6, and 2.0 \(\mu\)g/L advisory levels, respectively
-(Table \ref{tab:mc_chla_table} \& Figure \ref{fig:multi_cp_plot}).
+Chlorophyll \emph{a}, ranged from 0.07 to 936 \(\mu\)g/L and this
+captures the range of trrophic states from oligotrophic to
+hypereutrophic. All lakes had detectable levels of chlorophyll \emph{a}.
+The median concentration was 7.79 \(\mu\)g/L and the mean was 29.63
+\(\mu\)g/L. The association between chlorophyll \emph{a} and the upper
+confidence interval across a range of conditional probability values are
+shown in Table \ref{tab:mc_chla_table}. Specific chlorophyll \emph{a}
+that are associated with greater than even odds of exceeding the
+advisory levels were 0.07, 0.07, 3, and 11\(\mu\)g/L for 0.3, 1.0, 1.6,
+and 2.0 \(\mu\)g/L advisory levels, respectively (Table
+\ref{tab:mc_chla_table} \& Figure \ref{fig:multi_cp_plot}).
 
 The chlorophyll \emph{a} cutoffs may be used to predict whether or not a
-lake exceeds the microcystin-LR health advisories. Doing so allows us to
+lake exceeds the microcystin health advisories. Doing so allows us to
 compare the accuracy of the prediction as well as evaluate false
-negatives. Total accuracy of the four cutoffs predicting microcystin-LR
-exceedances were 19\% for the USEPA children's advisory, 22\% for the
-WHO drinking water advisory, 22\% for the USEPA adult advisory, and 23\%
-for the WHO regreational advisory (Tables \ref{tab:child_conmat_table},
-\ref{tab:who_drink_conmat_table}, \ref{tab:adult_conmat_table}, \&
-\ref{tab:who_rec_conmat_table}). However, total accuracy is only one
-part of the prediction performace with which we are concerned.
+negatives. Total accuracy of the four cutoffs predicting microcystin
+exceedances were 12\% for the USEPA children's drinking water advisory,
+20\% for the WHO drinking water advisory, 22\% for the USEPA adult
+drinking water advisory, and 3\% for the WHO regreational advisory
+(Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table},
+\ref{tab:adult_conmat_table}, \& \ref{tab:who_rec_conmat_table}).
+However, total accuracy is only one part of the prediction performace
+with which we are concerned.
 
 When using the chlorophyll \emph{a} cutoffs as an indicator of
-microcystin-LR exceedances, the error that should be avoided is
-predicting that no exceedance has occurred when in fact it has. In other
-words, we would like to avoid Type II errors and minimize the proportion
-of false negatives. For the four chlorophyll \emph{a} cut-offs we had a
-proportion of false negatives of 2\%, 2\%, 1\%, and 1\% for the USEPA
+microcystin exceedances, the error that should be avoided is predicting
+that no exceedance has occurred when in fact it has. In other words, we
+would like to avoid Type II errors and minimize the proportion of false
+negatives. For the four chlorophyll \emph{a} cut-offs we had a
+proportion of false negatives of 2\%, 2\%, 2\%, and 0\% for the USEPA
 children's, the WHO drinking water, the USEPA adult, and the WHO
 recreational advisories, respectively. In each case we missed less than
-10\% of the lakes that in fact exceeded the microcystin-LR advisory.
-While this method performs well with regard to the false negative
-percentage, it is possible that is a relic of the NLA dataset and
-testing with additional data would allow us to confirm this result.
+10\% of the lakes that in fact exceeded the microcystin advisory. While
+this method performs well with regard to the false negative percentage,
+it is possible that is a relic of the NLA dataset and testing with
+additional data would allow us to confirm this result.
 
@@
 
-The association between Log10 microcystin-LR and Log10 chlorophyll
-\emph{a} shows a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
-This indicates that, in general, higher concentrations of microcystin-LR
+The log-log association between microcystin-LR and chlorophyll \emph{a}
+indicates that, in general, higher concentrations of microcystin-LR
 almost always co-occur with higher concentrations of chlorophyll
-\emph{a} yet the inverse is not true. Higher chlorophyll \emph{a} is not
+\emph{a} yet the inverse is not true (Figure
+\ref{fig:chla_micro_scatter}). Higher chlorophyll \emph{a} is not
 necessarily predictive of higher microcystin-LR concentrations; however,
@@
 
-This is the case as the probability of exceeding each of the four tested
-health advisory levels increases as a function of chlorophyll \emph{a}
+Indeed, the probability of exceeding each of the four tested health
+advisory levels increased as a function of chlorophyll \emph{a}
 concentration (Figure \ref{fig:multi_cp_plot}). We used this association
-to identify chlorophyll \emph{a} concentrations that are associated with
-a range of probabilities of exceeding a given health advisory level
+to identify chlorophyll \emph{a} concentrations that were associated
+with a range of probabilities of exceeding a given health advisory level
 (Table \ref{tab:mc_chla_table}). For the purposes of this discussion we
@@
 than even odds to exceed a health advisory level). The 50\% conditional
-probability chlorophyll \emph{a} thresholds represents 29.2\%, 9.8\%,
-8.6\%, and 6.9\% of sample lakes for the USEPA Child, the WHO Drinking,
+probability chlorophyll \emph{a} thresholds represents 59.1\%, 14.3\%,
+5.2\%, and 93.9\% of sample lakes for the USEPA Child, the WHO Drinking,
 the USEPA Adult, and the WHO recreational levels, respecitvely.
@@
 There are numerous possible uses for the chlorophyll \emph{a} and
-microcystin-LR advisory cut-off values. First, in the absence of
+microcystin advisory cut-off values. First, in the absence of
 microcystin-LR measurements, exceedence of the chlorophyll \emph{a}
@@
 hypoxia, odors), there is no single definition of a bloom. For cHABs,
-one approach has been to identify an increase over a baseline
-concentration of phycocyanin {[}17{]}. This is a useful approach for
-targeted studies, but phycocyanin is also not always available and
-measures the predominance of cyanobacterial pigments and not toxins.
-Using our chlorophyll \emph{a} cutoffs provides a value that is more
-directly associated with microcystin-LR and can be used to classify
+one approach has been to utilize phycocyanin to screen for or identify
+bloom events {[}19{]}. This is a useful approach, but phycocyanin is not
+always available, thus limiting its utility especially for examining
+historical data. Using our chlorophyll \emph{a} cutoffs provides a value
+that is also associated with microcystin-LR and can be used to classify
 lakes, from past surveys, as having bloomed.
+
+The values we propose are national and may miss regional variation in
+water quality, including, chlorophyll \emph{a} and microcystin-LR
+{[}21{]}. A set of regional conditional probabilities would be
+interesting; however, limiting the analysis to the data available per
+region would make interpretaion difficult. The sample size for each of
+the conditional probabilities would be reduced (it ranges from 67 to
+155) and the number of lakes in each region that exceed the microcystin
+values is also reduced. The result is that our confidence in the
+conditional probabilites would be less (i.e.~greatly increased
+confidence intervals) and the relationships less pronounced as we have
+fewer lakes on which to base the probabilites. Thus, this dataset is
+best for making national scale recommendations.
+
+Two other limitations with the 2007 National Lakes Assessment data are
+that they represent a single sample from that lake and do not capture
+temporal dynamics and without subsetting the data do not provide us the
+ability to validate the presence of microcystin-LR. As of this writing,
+the 2012 National Lakes Assessment data are not public. When these data
+are released, a validation of this approach can be completed then.
 
@@
 ubiquitous measurement, such as chlorophyll \emph{a} will provide a
-reasonable proxy for the probability of exceeding a microcystin-LR
-health advisory level and provide better protection against adverse
-effects in both drinking and recreational use cases.
+reasonable proxy for the probability of exceeding a microcystin health
+advisory level and provide better protection against adverse effects in
+both drinking and recreational use cases.
 
@@
 We would like to thank Anne Kuhn, Bryan Milstead, John Kiddon, Joe
-LiVolsi, Tim Gleason, and Wayne Munns for constructive reviews of this
-paper. This paper has not been subjected to Agency review. Therefore, it
-does not necessary reflect the views of the Agency. Mention of trade
-names or commercial products does not constitute endorsement or
-recommendation for use. This contribution is identified by the tracking
-number ORD-015143 of the Atlantic Ecology Division, Office of Research
-and Development, National Health and Environmental Effects Research
-Laboratory, US Environmental Protection Agency.
+LiVolsi, Tim Gleason, Wayne Munns, and Leslie D'Anglada for constructive
+reviews of this paper. Special thanks to Jason Marion, Alan Wilson, and
+Zofia Taranu for reviews of the submitted manuscript. This paper has not
+been subjected to Agency review. Therefore, it does not necessary
+reflect the views of the Agency. Mention of trade names or commercial
+products does not constitute endorsement or recommendation for use. This
+contribution is identified by the tracking number ORD-015143 of the
+Atlantic Ecology Division, Office of Research and Development, National
+Health and Environmental Effects Research Laboratory, US Environmental
+Protection Agency.
 
@@
 \caption{Conditional probability plots showing association between the
-probability of exceeding various microcystin-LR (MLR) health advisory
-Levels. A.) Plot for USEPA Child (0.3 \(\mu\)g/L). B.) Plot for WHO
-Drinking (1 \(\mu\)g/L). C.) Plot for USEPA Adult (1.6 \(\mu\)g/L). D.)
-Plot for WHO Recreational (2 \(\mu\)g/L). \label{fig:multi_cp_plot}}
+probability of exceeding various microcystin health advisory Levels. A.)
+Plot for USEPA Child (0.3 \(\mu\)g/L). B.) Plot for WHO Drinking (1
+\(\mu\)g/L). C.) Plot for USEPA Adult (1.6 \(\mu\)g/L). D.) Plot for WHO
+Recreational (2 \(\mu\)g/L). \label{fig:multi_cp_plot}}
 \end{figure}
@@
 \begin{longtable}[c]{@{}lll@{}}
-\caption{Various suggested microcystin-LR health advisory
-concentrations. \label{tab:microcystin_levels}}\tabularnewline
+\caption{Various suggested microcystin health advisory concentrations.
+\label{tab:microcystin_levels}}\tabularnewline
 \toprule
@@
 \endhead
-USEPA & Adult Drinking Water Advisory & 1.6 \(\mu\)g/L\tabularnewline
 USEPA & Child Drinking Water Advisory & 0.3 \(\mu\)g/L\tabularnewline
 WHO & Drinking Water & 1 \(\mu\)g/L\tabularnewline
-WHO & Recreational: High Prob. of Effect & 20-2000
-\(\mu\)g/L\tabularnewline
+USEPA & Adult Drinking Water Advisory & 1.6 \(\mu\)g/L\tabularnewline
 WHO & Recreational: Low Prob. of Effect & 2-4 \(\mu\)g/L\tabularnewline
 WHO & Recreational: Moderate Prob. of Effect & 10-20
+\(\mu\)g/L\tabularnewline
+WHO & Recreational: High Prob. of Effect & 20-2000
 \(\mu\)g/L\tabularnewline
@@
 
-\begin{longtable}[c]{@{}rrrrr@{}}
+\begin{longtable}[c]{@{}lllll@{}}
 \caption{Chlorophyll \textit{a} concentrations that are associated with
-a 50\% probability of exceeding a microcystin-LR health advisory
+a 50\% probability of exceeding a microcystin health advisory
 concentration. \label{tab:mc_chla_table}}\tabularnewline
 \toprule
-Cond. Probability & USEPA Child & WHO Drink & USEPA Adult & WHO
-Recreational\tabularnewline
+Cond. Probability U & SEPA Child (0.3 \(\mu\)g/L) W & HO Drink (1
+\(\mu\)g/L) U & SEPA Adult (1.6 \(\mu\)g/L) W & HO Recreational (2
+\(\mu\)g/L)\tabularnewline
 \midrule
@@
 \toprule
-Cond. Probability & USEPA Child & WHO Drink & USEPA Adult & WHO
-Recreational\tabularnewline
+Cond. Probability U & SEPA Child (0.3 \(\mu\)g/L) W & HO Drink (1
+\(\mu\)g/L) U & SEPA Adult (1.6 \(\mu\)g/L) W & HO Recreational (2
+\(\mu\)g/L)\tabularnewline
 \midrule
 \endhead
-0.1 & 0.07 & 0.07 & 0.07 & 1.17\tabularnewline
-0.2 & 0.07 & 4.48 & 10.67 & 18.24\tabularnewline
-0.3 & 2.78 & 17.49 & 31.59 & 50.26\tabularnewline
-0.4 & 9.97 & 38.73 & 69.19 & 76.80\tabularnewline
-0.5 & 21.92 & 77.20 & 84.96 & 105.84\tabularnewline
-0.6 & 38.73 & 103.20 & 133.20 & 228.24\tabularnewline
-0.7 & 63.94 & 138.24 & 244.80 & 871.20\tabularnewline
-0.8 & 113.14 & 326.88 & 326.88 & 871.20\tabularnewline
-0.9 & 166.63 & 516.00 & 871.20 & 871.20\tabularnewline
+0.1 & 0.07 & 0.07 & 0.07 & 1\tabularnewline
+0.2 & 0.07 & 5 & 11 & 15\tabularnewline
+0.3 & 3 & 18 & 32 & 39\tabularnewline
+0.4 & 11 & 39 & 67 & 78\tabularnewline
+0.5 & 23 & 64 & 84 & 103\tabularnewline
+0.6 & 39 & 92 & 115 & 167\tabularnewline
+0.7 & 65 & 116 & 274 & 274\tabularnewline
+0.8 & 115 & 256 & 871 & 871\tabularnewline
+0.9 & 138 & 318 & 871 & 871\tabularnewline
 \bottomrule
@@
 
-\begin{longtable}[c]{@{}lrrr@{}}
+\begin{longtable}[c]{@{}llll@{}}
 \caption{Confusion matrix comparing chlorophyll \textit{a} predicted
@@
 \toprule
-& Not Exceed & Exceed & Row Totals\tabularnewline
+Not Exceed Exc & eed Row Tot & als &\tabularnewline
 \midrule
@@
 \toprule
-& Not Exceed & Exceed & Row Totals\tabularnewline
+Not Exceed Exc & eed Row Tot & als &\tabularnewline
 \midrule
 \endhead
-Not Exceed & 638 & 93 & 731\tabularnewline
-Exceed & 173 & 124 & 297\tabularnewline
+Not Exceed & 344 & 78 & 422\tabularnewline
+Exceed & 467 & 139 & 606\tabularnewline
 Column Totals & 811 & 217 & 1028\tabularnewline
@@
 \toprule
-& Not Exceed & Exceed & Row Totals\tabularnewline
+Not Exceed Exc & eed Row Tot & als &\tabularnewline
 \midrule
 \endhead
-Not Exceed & 850 & 82 & 932\tabularnewline
-Exceed & 57 & 39 & 96\tabularnewline
+Not Exceed & 787 & 98 & 885\tabularnewline
+Exceed & 120 & 23 & 143\tabularnewline
 Column Totals & 907 & 121 & 1028\tabularnewline
@@
 \toprule
-& Not Exceed & Exceed & Row Totals\tabularnewline
+Not Exceed Exc & eed Row Tot & als &\tabularnewline
 \midrule
 \endhead
-Not Exceed & 887 & 57 & 944\tabularnewline
-Exceed & 50 & 34 & 84\tabularnewline
+Not Exceed & 897 & 82 & 979\tabularnewline
+Exceed & 40 & 9 & 49\tabularnewline
 Column Totals & 937 & 91 & 1028\tabularnewline
@@
 \toprule
-& Not Exceed & Exceed & Row Totals\tabularnewline
+Not Exceed Exc & eed Row Tot & als &\tabularnewline
 \midrule
 \endhead
-Not Exceed & 910 & 52 & 962\tabularnewline
-Exceed & 43 & 23 & 66\tabularnewline
+Not Exceed & 61 & 2 & 63\tabularnewline
+Exceed & 892 & 73 & 965\tabularnewline
 Column Totals & 953 & 75 & 1028\tabularnewline
@@
 
+\hypertarget{refs}{}
+\hypertarget{ref-jetoo2015toledo}{}
 1. Jetoo S, Grover VI, Krantzberg G (2015) The toledo drinking water
@@
 
+\hypertarget{ref-rinta2009lake}{}
 2. Rinta-Kanto JM, Konopko EA, DeBruyn JM, Bourbonniere RA, Boyer GL, et
@@
 
+\hypertarget{ref-HABHRCA2014}{}
 3. Bloom HA, Research H, Act CA (2014) Harmful Algal Bloom and Hypoxia
@@
 
+\hypertarget{ref-mcelhiney2005detection}{}
 4. McElhiney J, Lawton LA (2005) Detection of the cyanobacterial
@@
 
+\hypertarget{ref-zurawell2005hepatotoxic}{}
 5. Zurawell RW, Chen H, Burke JM, Prepas EE (2005) Hepatotoxic
@@
 
+\hypertarget{ref-usepa2015drinking}{}
 6. USEPA (2015) Drinking water health advisory for the cyanobacterial
@@
 
+\hypertarget{ref-world2003cyanobacterial}{}
 7. World Health Organization (2003) Cyanobacterial toxins:
@@
 
+\hypertarget{ref-chorus1999toxic}{}
 8. Chorus EI, Bartram J (1999) Toxic cyanobacteria in water: A guide to
@@
 
+\hypertarget{ref-james2011environmental}{}
 9. James R, Gregg A, Dindal A, McKernan J (2011) Environmental
@@
 
+\hypertarget{ref-aranda2015evaluation}{}
 10. Aranda-Rodriguez R, Jin Z, Harvie J, Cabecinha A (2015) Evaluation
@@
 
-11. Pip E, Bowman L (2014) Microcystin and algal chlorophyll in relation
+\hypertarget{ref-yuan2015deriving}{}
+11. Yuan LL, Pollard AI (2015) Deriving nutrient targets to prevent
+excessive cyanobacterial densities in uS lakes and reservoirs.
+Freshwater Biology 60: 1901--1916.
+
+\hypertarget{ref-pip2014microcystin}{}
+12. Pip E, Bowman L (2014) Microcystin and algal chlorophyll in relation
 to nearshore nutrient concentrations in Lake Winnipeg, Canada.
@@
 
-12. Yuan LL, Pollard AI, Pather S, Oliver JL, D'Anglada L (2014)
+\hypertarget{ref-yuan2014managing}{}
+13. Yuan LL, Pollard AI, Pather S, Oliver JL, D'Anglada L (2014)
 Managing microcystin: Identifying national-scale thresholds for total
@@
 
-13. USEPA (2009) National lakes assessment: A collaborative survey of
+\hypertarget{ref-usepa2009national}{}
+14. USEPA (2009) National lakes assessment: A collaborative survey of
 the nation's lakes. EPA 841-R-09-001.
 
-14. Paul JF, Munns WR (2011) Probability surveys, conditional
+\hypertarget{ref-nlaux5ffieldops}{}
+15. USEPA (2007) Survey of the nation's lakes field operations manual
+EPA841-B-07-004.
+
+\hypertarget{ref-paul2011probability}{}
+16. Paul JF, Munns WR (2011) Probability surveys, conditional
 probability, and ecological risk assessment. Environmental Toxicology
@@
 
-15. Paul JF, McDonald ME (2005) Development of empirical, geographically
+\hypertarget{ref-paul2005development}{}
+17. Paul JF, McDonald ME (2005) Development of empirical, geographically
 specific water quality criteria: A conditional probability analysis
@@
 
-16. Hollister JW, Walker HA, Paul JF (2008) CProb: A computational tool
+\hypertarget{ref-hollister2008cprob}{}
+18. Hollister JW, Walker HA, Paul JF (2008) CProb: A computational tool
 for conducting conditional probability analysis. Journal of
@@
 
-17. Miller TR, Beversdorf L, Chaston SD, McMahon KD (2013)
+\hypertarget{ref-miller2013spatiotemporal}{}
+19. Miller TR, Beversdorf L, Chaston SD, McMahon KD (2013)
 Spatiotemporal molecular analysis of cyanobacteria blooms reveals
@@
 
+\hypertarget{ref-marion2012vivo}{}
+20. Marion JW, Lee J, Wilkins III J, Lemeshow S, Lee C, et al. (2012) In
+vivo phycocyanin flourometry as a potential rapid screening tool for
+predicting elevated microcystin concentrations at eutrophic lakes.
+Environmental science \& technology 46: 4523--4531.
+
+\hypertarget{ref-beaver2014land}{}
+21. Beaver JR, Manis EE, Loftin KA, Graham JL, Pollard AI, et al. (2014)
+Land use patterns, ecoregion, and microcystin relationships in uS lakes
+and reservoirs: A preliminary evaluation. Harmful Algae 36: 57--62.
+
+\hypertarget{ref-wagner2011landscape}{}
+22. Wagner T, Soranno PA, Webster KE, Cheruvelil KS (2011) Landscape
+drivers of regional variation in the relationship between total
+phosphorus and chlorophyll in lakes. Freshwater Biology 56: 1811--1824.
+
 \end{document}

### 2016-06-01 manuscript/manuscript.Rmd
revisions entered into overleaf

@@
 }
+if(!require(autocrop)){
+  devtools::install_github("jhollist/autocrop") 
+}
 library("microcystinchla")
@@
 library("e1071")
-opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=5)
+library("autocrop")
+#opts_chunk$set(dev = 'jpg', fig.width=6, fig.height=5)
+
+if(file.exists("manuscript_cache")){
+  for(i in gsub(".rdb","",list.files("manuscript_cache/latex",".rdb",full.names=T))){
+    lazyLoad(i)
+  }
+}
 
@@
 
-epa_child_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log10(mc_guides[1]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100, xW=wgt)
+#set.seed(19720923)
+#epa_child_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
+#                     xImpair = log10(mc_guides[1]), ProbComp = "gt", Exceed = "gte",
+#                     ci = TRUE, R = 10000, xW=wgt)
 
 #mc_chla[1] <- get_chla(epa_child_cp,0.5)
-
-who_drink_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log10(mc_guides[2]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100, xW=wgt)
+#set.seed(19720923)
+#who_drink_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
+#                     xImpair = log10(mc_guides[2]), ProbComp = "gt", Exceed = "gte",
+#                     ci = TRUE, R = 10000, xW=wgt)
 
 #mc_chla[2] <- get_chla(who_drink_cp,0.5)
-
-epa_adult_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log10(mc_guides[3]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100, xW=wgt)
+#set.seed(19720923)
+#epa_adult_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
+#                     xImpair = log10(mc_guides[3]), ProbComp = "gt", Exceed = "gte",
+#                     ci = TRUE, R = 10000, xW=wgt)
 
 #mc_chla[3] <- get_chla(epa_adult_cp,0.5)
-
-who_rec_low1_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
-                     xImpair = log10(mc_guides[4]), ProbComp = "gt", Exceed = "gte",
-                     ci = TRUE, R = 100, xW=wgt)
+#set.seed(19720923)
+#who_rec_low1_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
+#                     xImpair = log10(mc_guides[4]), ProbComp = "gt", Exceed = "gte",
+#                     ci = TRUE, R = 10000, xW=wgt)
 
@@
 
-mc_chla<-cbind(as.character(epa_child),as.character(who_drink),
-               as.character(epa_adult),as.character(who_rec))
+mc_chla<-cbind(epa_child,who_drink,
+               epa_adult,who_rec)
 
@@
 We look at the recent US EPA health advisories for drinking water as well as the World Health Organization levels for drinking water and recreational use and identify a range of chlorophyll *a* thresholds.
-A 50% chance of exceeding one of the specific advisory microcystin concentrations of 0.3, 1, 1.6, and 2 $\mu$g/L is associatied with chlorophyll *a* concentration thresholds of `r mc_chla[5,1]`, `r mc_chla[5,2]`, `r mc_chla[5,3]`, and `r mc_chla[5,4]`, respectively.
+A 50% chance of exceeding one of the specific advisory microcystin concentrations of 0.3, 1, 1.6, and 2 $\mu$g/L is associatied with chlorophyll *a* concentration thresholds of `r mc_chla[5,1]`, `r mc_chla[5,2]`, `r mc_chla[5,3]`, and `r mc_chla[5,4]` $\mu$g/L, respectively.
 When managing for these various microcystin levels, exceeding these reported chlorophyll *a* concentrations should be a trigger for further testing and possible management action.
@@
 The association between chlorophyll *a* and the upper confidence interval across a range of conditional probability values are shown in Table \ref{tab:mc_chla_table}.
-Specific chlorophyll *a* that are associated with greater than even odds of exceeding the advisory levels were `r mc_chla[1]`, `r mc_chla[2]`, `r mc_chla[3]`, and `r mc_chla[4]`$\mu$g/L for 0.3, 1.0, 1.6, and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
+Specific chlorophyll *a* that are associated with greater than even odds of exceeding the advisory levels were `r mc_chla[5,1]`, `r mc_chla[5,2]`, `r mc_chla[5,3]`, and `r mc_chla[5,4]`$\mu$g/L for 0.3, 1.0, 1.6, and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
 
@@
 #Acknowledgements
-We would like to thank Anne Kuhn, Bryan Milstead, John Kiddon, Joe LiVolsi, Tim Gleason, Wayne Munns, and Leslie D’Anglada for constructive reviews of this paper.
+We would like to thank Anne Kuhn, Bryan Milstead, John Kiddon, Joe LiVolsi, Tim Gleason, Wayne Munns, and Leslie D'Anglada for constructive reviews of this paper.
 Special thanks to Jason Marion, Alan Wilson, and Zofia Taranu for reviews of the submitted manuscript.
@@
 ```{r chla_micro_scatter, echo=FALSE, fig.width=6, fig.cap="Scatterplot showing association betweeen chlorophyll \\textit{a} and microcystin-LR. \\label{fig:chla_micro_scatter}"}
-plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",pt_col = nla_dat$WSA_ECO9,
+ggplot2::ggsave("hollister_microcystin_fig2.jpg",
+       plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",pt_col = "black",
              x=expression(paste('Log10 Chl ', italic("a"),' (',mu,'g/L)')),
-             y=expression(paste('Log10 microcystin-LR',' (',mu,'g/L)')))
+             y=expression(paste('Log10 microcystin-LR',' (',mu,'g/L)'))),
+       dpi = 450,
+       height = 6.8,
+       units = "in")
+x<-autocrop("hollister_microcystin_fig2.jpg",border = 25, 
+            outfile = "hollister_microcystin_fig2.jpg",
+            width = 6.1,
+            res = 450)
 ```
@@
 \\label{fig:multi_cp_plot}"}
+jpeg(filename = "hollister_microcystin_fig1.jpg",
+       res = 450,
+       width = 7.82,
+       height = 6.8,
+       units = "in")
 multiplot(
@@
 )
+x<-dev.off()
+x<-autocrop("hollister_microcystin_fig1.jpg",border = 25, 
+            outfile = "hollister_microcystin_fig1.jpg",
+            width = 6.1,
+            res = 450)
 ```

### 2016-06-01 manuscript/manuscript.tex
revisions entered into overleaf

@@
 \usepackage{longtable,booktabs}
-\usepackage{graphicx}
-% Redefine \includegraphics so that, unless explicit options are
-% given, the image width will not exceed the width of the page.
-% Images get their normal width if they fit onto the page, but
-% are scaled down if they would overflow the margins.
-\makeatletter
-\def\ScaleIfNeeded{%
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-\let\Oldincludegraphics\includegraphics
-{%
- \catcode`\@=11\relax%
- \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
-}%
 \ifxetex
@@
 associatied with chlorophyll \emph{a} concentration thresholds of 23,
-64, 84, and 103, respectively. When managing for these various
-microcystin levels, exceeding these reported chlorophyll \emph{a}
-concentrations should be a trigger for further testing and possible
-management action.
+68, 84, and 104 \(\mu\)g/L, respectively. When managing for these
+various microcystin levels, exceeding these reported chlorophyll
+\emph{a} concentrations should be a trigger for further testing and
+possible management action.
 
@@
 that are associated with greater than even odds of exceeding the
-advisory levels were 0.07, 0.07, 3, and 11\(\mu\)g/L for 0.3, 1.0, 1.6,
+advisory levels were 23, 68, 84, and 104\(\mu\)g/L for 0.3, 1.0, 1.6,
 and 2.0 \(\mu\)g/L advisory levels, respectively (Table
@@
 negatives. Total accuracy of the four cutoffs predicting microcystin
-exceedances were 12\% for the USEPA children's drinking water advisory,
-20\% for the WHO drinking water advisory, 22\% for the USEPA adult
-drinking water advisory, and 3\% for the WHO regreational advisory
+exceedances were 74\% for the USEPA children's drinking water advisory,
+86\% for the WHO drinking water advisory, 89\% for the USEPA adult
+drinking water advisory, and 91\% for the WHO regreational advisory
 (Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table},
@@
 negatives. For the four chlorophyll \emph{a} cut-offs we had a
-proportion of false negatives of 2\%, 2\%, 2\%, and 0\% for the USEPA
+proportion of false negatives of 9\%, 8\%, 6\%, and 5\% for the USEPA
 children's, the WHO drinking water, the USEPA adult, and the WHO
@@
 than even odds to exceed a health advisory level). The 50\% conditional
-probability chlorophyll \emph{a} thresholds represents 59.1\%, 14.3\%,
-5.2\%, and 93.9\% of sample lakes for the USEPA Child, the WHO Drinking,
+probability chlorophyll \emph{a} thresholds represents 28.6\%, 11\%,
+8.9\%, and 7.2\% of sample lakes for the USEPA Child, the WHO Drinking,
 the USEPA Adult, and the WHO recreational levels, respecitvely.
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/chla_micro_scatter-1.pdf}
-\caption{Scatterplot showing association betweeen chlorophyll \textit{a}
-and microcystin-LR. \label{fig:chla_micro_scatter}}
-\end{figure}
+\begin{verbatim}
+## Saving 6 x 6.8 in image
+\end{verbatim}
 
 \newpage
-
-\begin{verbatim}
-## Loading required package: grid
-\end{verbatim}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{manuscript_files/figure-latex/epa_child_cp_plot-1.pdf}
-\caption{Conditional probability plots showing association between the
-probability of exceeding various microcystin health advisory Levels. A.)
-Plot for USEPA Child (0.3 \(\mu\)g/L). B.) Plot for WHO Drinking (1
-\(\mu\)g/L). C.) Plot for USEPA Adult (1.6 \(\mu\)g/L). D.) Plot for WHO
-Recreational (2 \(\mu\)g/L). \label{fig:multi_cp_plot}}
-\end{figure}
 
@@
 \toprule
-Cond. Probability U & SEPA Child (0.3 \(\mu\)g/L) W & HO Drink (1
-\(\mu\)g/L) U & SEPA Adult (1.6 \(\mu\)g/L) W & HO Recreational (2
+Cond. Probability & USEPA Child (0.3 \(\mu\)g/L) & WHO Drink (1
+\(\mu\)g/L) & USEPA Adult (1.6 \(\mu\)g/L) & WHO Recreational (2
 \(\mu\)g/L)\tabularnewline
@@
 0.1 & 0.07 & 0.07 & 0.07 & 1\tabularnewline
-0.2 & 0.07 & 5 & 11 & 15\tabularnewline
-0.3 & 3 & 18 & 32 & 39\tabularnewline
-0.4 & 11 & 39 & 67 & 78\tabularnewline
-0.5 & 23 & 64 & 84 & 103\tabularnewline
-0.6 & 39 & 92 & 115 & 167\tabularnewline
-0.7 & 65 & 116 & 274 & 274\tabularnewline
-0.8 & 115 & 256 & 871 & 871\tabularnewline
-0.9 & 138 & 318 & 871 & 871\tabularnewline
+0.2 & 0.07 & 4.00 & 12.00 & 17\tabularnewline
+0.3 & 3.00 & 17.00 & 32.00 & 45\tabularnewline
+0.4 & 11.00 & 37.00 & 68.00 & 77\tabularnewline
+0.5 & 23.00 & 68.00 & 84.00 & 104\tabularnewline
+0.6 & 39.00 & 97.00 & 115.00 & 185\tabularnewline
+0.7 & 66.00 & 126.00 & 871.00 & 871\tabularnewline
+0.8 & 116.00 & 271.00 & 871.00 & 871\tabularnewline
+0.9 & 170.00 & 516.00 & 871.00 & 871\tabularnewline
 \bottomrule
@@
 \endhead
-Not Exceed & 344 & 78 & 422\tabularnewline
-Exceed & 467 & 139 & 606\tabularnewline
+Not Exceed & 643 & 95 & 738\tabularnewline
+Exceed & 168 & 122 & 290\tabularnewline
 Column Totals & 811 & 217 & 1028\tabularnewline
@@
 \endhead
-Not Exceed & 787 & 98 & 885\tabularnewline
-Exceed & 120 & 23 & 143\tabularnewline
+Not Exceed & 841 & 78 & 919\tabularnewline
+Exceed & 66 & 43 & 109\tabularnewline
 Column Totals & 907 & 121 & 1028\tabularnewline
@@
 \endhead
-Not Exceed & 897 & 82 & 979\tabularnewline
-Exceed & 40 & 9 & 49\tabularnewline
+Not Exceed & 884 & 57 & 941\tabularnewline
+Exceed & 53 & 34 & 87\tabularnewline
 Column Totals & 937 & 91 & 1028\tabularnewline
@@
 \endhead
-Not Exceed & 61 & 2 & 63\tabularnewline
-Exceed & 892 & 73 & 965\tabularnewline
+Not Exceed & 908 & 51 & 959\tabularnewline
+Exceed & 45 & 24 & 69\tabularnewline
 Column Totals & 953 & 75 & 1028\tabularnewline

### 2016-06-07 manuscript/f1000_revisions/manuscript.Rmd
updates, including proofs

--- 2016-06-01 manuscript/manuscript.Rmd
@@
 
-```{r analysis , include=FALSE, echo=FALSE, cache=TRUE}
+```{r analysis , include=FALSE, echo=FALSE, cache=FALSE}
 #All analysis in here, that way all bits of the paper have access to the final objects
@@
 In response the US Environmental Protection Agency (USEPA) has recently released suggested microcystin (one of the more common toxins) concentrations that would trigger health advisories [@mcelhiney2005detection; @zurawell2005hepatotoxic; @usepa2015drinking].
-Additionally, the World Health Organization (WHO) has proposed microcystin advisory levels for drinking water and a range of recreational risk levels [@world2003cyanobacterial; @chorus1999toxic].
+Additionally, the World Health Organization (WHO) has microcystin advisory levels for drinking water and for a range of recreational risk levels [@world2003cyanobacterial; @chorus1999toxic].
 While these levels and associated advisories are likely to help mitigate the impacts from harmful algal blooms, they are not without complications.
@@
 In their analysis they find that total nitrogen and chlorophyll *a* show the strongest association with microcystin.
-Furthermore, they identify chlorophyll*a* and total nitrogen concentrations that are associated with exceeding 1 $\mu$g/L of microcystin.
-These findings suggest that chlorophyll a concentrations could also track the new USEPA microcystin health advisory levels for drinking water.
+Furthermore, they identify chlorophyll *a* and total nitrogen concentrations that are associated with exceeding 1 $\mu$g/L of microcystin.
+These findings suggest that chlorophyll *a* concentrations could also track the new USEPA microcystin health advisory levels for drinking water.
 Identifying this association would provide an important tool for water resource managers to help manage the threat to public health posed by cHABs and would be especially useful in the absence of measured microcystin concentrations.
 
-In this paper we build on past efforts and utilize the National Lakes Assessment (NLA) data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin health advisory concentrations [@usepa2009national; @usepa2015drinking; @chorus1999toxic].
+In fact, this is a similar tact to the World Health Organization who, in addition to advisory levels for microcystin, have also proposed related advisory levels for cyanobacteria abundance and chlorophyll *a* [@world2003cyanobacterial; @chorus1999toxic].
+The chlorophyll *a* concentrations proposed by the WHO are for low (< 10 $\mu$g/L), moderate (between 10 and 50 $\mu$g/L, high (between 50 and 5000 $\mu$g/L), and very high risk (>5000 $\mu$g/L) [@chorus1999toxic].
+While these advisories have proven to be useful tools they do suffer from being coarse, broad, and have been found to overestimate actual risk [@loftin2016cyanotoxins].
+
+In this paper we build on these past efforts and utilize the National Lakes Assessment (NLA) data and identify chlorophyll *a* concentrations that are associated with higher probabilities of exceeding several microcystin health advisory concentrations [@usepa2009national; @usepa2015drinking; @chorus1999toxic].
 We build on past studies by exploring associations with the newly announced advisory levels and by also applying a different method, conditional probability analysis.
@@
 We expect to find higher chlorohpyll *a* concentrations to be associated with higher probabilities of exceeding the microcystin health advisory levels.
-We also calculated bootstrapped 95% confidence intervals (CI) using 1000 bootstrapped samples.
+We also calculated bootstrapped 95% confidence intervals (CI) using 10,000 bootstrapped samples.
 Thus, to identify chlorophyll *a* concentrations of concern we identified the value of the upper 95% CI across a range of conditional probabilities of exceeding each health advisory level.
@@
 [-@hollister2008cprob] for greater detail.
-Lastly, all analyses were conducted using R version 3.2.2 and code and data from this analysis are freely available as an R package at [https://github.com/USEPA/microcystinchla](https://github.com/USAPE/microcystinchla).
+All analyses were conducted using R version 3.2.3 and code and data from this analysis are freely available as an R package at https://github.com/USEPA/microcystinchla.
 
@@
 Any value greater than that indicates the presence of microcystin-LR.
-Of those lakes with microcystin, the median concentration was `r round(median(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T), 2)`$\mu$g/L and the mean was `r round(mean(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T), 2)`$\mu$g/L.
+Of those lakes with microcystin-LR, the median concentration was `r round(median(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T), 2)` $\mu$g/L and the mean was `r round(mean(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>0.05],na.rm=T), 2)` $\mu$g/L.
 Of all lakes sampled, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=0.3])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the USEPA Child level, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1.6])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the USEPA Adult level, `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=1])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO Drinking level, and `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=2])/length(nla_dat$MCYST_TL_UGL)),1)`% of lakes exceeded the WHO Recreational level.
@@
 The association between chlorophyll *a* and the upper confidence interval across a range of conditional probability values are shown in Table \ref{tab:mc_chla_table}.
-Specific chlorophyll *a* that are associated with greater than even odds of exceeding the advisory levels were `r mc_chla[5,1]`, `r mc_chla[5,2]`, `r mc_chla[5,3]`, and `r mc_chla[5,4]`$\mu$g/L for 0.3, 1.0, 1.6, and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
+Specific chlorophyll *a* concentrations that are associated with greater than even odds of exceeding the advisory levels were `r mc_chla[5,1]`, `r mc_chla[5,2]`, `r mc_chla[5,3]`, and `r mc_chla[5,4]` $\mu$g/L for 0.3, 1.0, 1.6, and 2.0 $\mu$g/L advisory levels, respectively (Table \ref{tab:mc_chla_table} & Figure \ref{fig:multi_cp_plot}).
 
-The chlorophyll *a* cutoffs may be used to predict whether or not a lake exceeds the microcystin health advisories.
-Doing so allows us to compare the accuracy of the prediction as well as evaluate false negatives.
-Total accuracy of the four cutoffs predicting microcystin exceedances were `r total_accuracy$epa_child*100`% for the USEPA children's drinking water advisory, `r total_accuracy$who_drink*100`% for the WHO drinking water advisory,  `r total_accuracy$epa_adult*100`% for the USEPA adult drinking water advisory, and `r total_accuracy$who_rec*100`% for the WHO regreational advisory (Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table}, \ref{tab:adult_conmat_table}, & \ref{tab:who_rec_conmat_table}).
-However, total accuracy is only one part of the prediction performace with which we are concerned.
-
-When using the chlorophyll *a* cutoffs as an indicator of microcystin exceedances, the error that should be avoided is predicting that no exceedance has occurred when in fact it has.
-In other words, we would like to avoid Type II errors and minimize the proportion of false negatives.
-For the four chlorophyll *a* cut-offs we had a proportion of false negatives of `r fn_percent$epa_child*100`%, `r fn_percent$who_drink*100`%, `r fn_percent$epa_adult*100`%, and `r fn_percent$who_rec*100`% for the USEPA children's, the WHO drinking water, the USEPA adult, and the WHO recreational advisories, respectively.
-In each case we missed less than 10% of the lakes that in fact exceeded the microcystin advisory.
-While this method performs well with regard to the false negative percentage, it is possible that is a relic of the NLA dataset and testing with additional data would allow us to confirm this result.
-
-#Discussion
-The log-log association between microcystin-LR and chlorophyll *a* indicates that, in general, higher concentrations of microcystin-LR almost always co-occur with higher concentrations of chlorophyll *a* yet the inverse is not true (Figure \ref{fig:chla_micro_scatter}).
-Higher chlorophyll *a* is not necessarily predictive of higher microcystin-LR concentrations; however, chlorophyll *a* may be predictive of the probability of exceeding a certain threshold.
-
-Indeed, the probability of exceeding each of the four tested health advisory levels increased as a function of chlorophyll *a* concentration (Figure \ref{fig:multi_cp_plot}).
-We used this association to identify chlorophyll *a* concentrations that were associated with a range of probabilities of exceeding a given health advisory level (Table \ref{tab:mc_chla_table}).
-For the purposes of this discussion we focus on a conditional probability of 50% or greater (i.e., greater than even odds to exceed a health advisory level).
-The 50% conditional probability chlorophyll *a* thresholds represents `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[5,1]])/length(nla_dat$CHLA)),1)`%, `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[5,2]])/length(nla_dat$CHLA)),1)`%,
-`r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[5,3]])/length(nla_dat$CHLA)),1)`%, and `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[5,4]])/length(nla_dat$CHLA)),1)`% of sample lakes for the USEPA Child, the WHO Drinking, the USEPA Adult, and the WHO recreational levels, respecitvely.
-
-There are numerous possible uses for the chlorophyll *a* and microcystin advisory cut-off values.
-First, in the absence of microcystin-LR measurements, exceedence of the chlorophyll *a* concentrations could be a trigger for further actions.
-Given that there is uncertainity around these chlorophyll *a* cutoffs the best case scenario would be to monitor for chlorophyll *a* and in the event of exceeding a target concentration take water samples and have those samples tested for microcystin-LR.
-
-A second potential use is to identify past bloom events from historical data.
-As harmful algal blooms are made up of many species and have various mechanisms responsible for adverse impacts (e.g., toxins, hypoxia, odors), there is no single definition of a bloom.
-For cHABs, one approach has been to utilize phycocyanin to screen for or identify bloom events [@miller2013spatiotemporal, ahn2007alternative, @marion2012vivo].
-This is a useful approach, but phycocyanin is not always available, thus limiting its utility especially for examining historical data.
-Using our chlorophyll *a* cutoffs provides a value that is also associated with microcystin-LR and can be used to classify lakes, from past surveys, as having bloomed.
-
-The values we propose are national and may miss regional variation in water quality, including, chlorophyll *a* and microcystin-LR [@beaver2014land, @wagner2011landscape].
-A set of regional conditional probabilities would be interesting; however, limiting the analysis to the data available per region would make interpretaion difficult.
-The sample size for each of the conditional probabilities would be reduced (it ranges from `r  min(table(nla_dat$WSA_ECO9))` to `r  max(table(nla_dat$WSA_ECO9))`) and the number of lakes in each region that exceed the microcystin values is also reduced.
-The result is that our confidence in the conditional probabilites would be less (i.e.
-greatly increased confidence intervals) and the relationships less pronounced as we have fewer lakes on which to base the probabilites.
-Thus, this dataset is best for making national scale recommendations.
-
-Two other limitations with the 2007 National Lakes Assessment data are that they represent a single sample from that lake and do not capture temporal dynamics and without subsetting the data do not provide us the ability to validate the presence of microcystin-LR.
-As of this writing, the 2012 National Lakes Assessment data are not public.
-When these data are released, a validation of this approach can be completed then.
-
-Lastly, using chlorophyll *a* is not meant as a replacement for testing of microcystin-LR or other toxins.
-It should be used when other, direct measurements of cyanotoxins are not available.
-In those cases, which are likely to be common at least in the near future, using a more ubiquitous measurement, such as chlorophyll *a* will provide a reasonable proxy for the probability of exceeding a microcystin health advisory level and provide better protection against adverse effects in both drinking and recreational use cases.
-
-#Acknowledgements
-We would like to thank Anne Kuhn, Bryan Milstead, John Kiddon, Joe LiVolsi, Tim Gleason, Wayne Munns, and Leslie D'Anglada for constructive reviews of this paper.
-Special thanks to Jason Marion, Alan Wilson, and Zofia Taranu for reviews of the submitted manuscript.
-This paper has not been subjected to Agency review.
-Therefore, it does not necessary reflect the views of the Agency.
-Mention of trade names or commercial products does not constitute endorsement or recommendation for use.
-This contribution is identified by the tracking number ORD-015143 of the Atlantic Ecology Division, Office of Research and Development, National Health and Environmental Effects Research Laboratory, US Environmental Protection Agency.
-
-\newpage
-
-#Figures
-
-```{r chla_micro_scatter, echo=FALSE, fig.width=6, fig.cap="Scatterplot showing association betweeen chlorophyll \\textit{a} and microcystin-LR. \\label{fig:chla_micro_scatter}"}
-ggplot2::ggsave("hollister_microcystin_fig2.jpg",
-       plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",pt_col = "black",
-             x=expression(paste('Log10 Chl ', italic("a"),' (',mu,'g/L)')),
-             y=expression(paste('Log10 microcystin-LR',' (',mu,'g/L)'))),
-       dpi = 450,
-       height = 6.8,
-       units = "in")
-x<-autocrop("hollister_microcystin_fig2.jpg",border = 25, 
-            outfile = "hollister_microcystin_fig2.jpg",
-            width = 6.1,
-            res = 450)
-```
-
-\newpage
-
-<!--
-```{r chla_micro_cdf, echo=FALSE, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin health advisory Levels.
-\\label{fig:chla_micro_cdf}"}
-plot_cdf(nla_dat[,c("log_chla","log_micro")],"")
-```
-
-\newpage-->
-
-```{r epa_child_cp_plot, echo=FALSE, fig.width=6, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin health advisory Levels.
-A.) Plot for USEPA Child (0.3 $\\mu$g/L).
-B.) Plot for WHO Drinking (1 $\\mu$g/L).
-C.) Plot for USEPA Adult (1.6 $\\mu$g/L).
-D.) Plot for WHO Recreational (2 $\\mu$g/L).
-\\label{fig:multi_cp_plot}"}
+```{r epa_child_cp_plot, echo=FALSE, fig.width=6, fig.cap="",warning=FALSE,message=FALSE}
 jpeg(filename = "hollister_microcystin_fig1.jpg",
@@
 
+<!--
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{hollister_microcystin_fig1.jpg}
+\caption{\label{fig:multi_cp_plot}Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels.
+A.) Plot for USEPA Child (0.3 $\mu$g/L).
+B.) Plot for WHO Drinking (1 $\mu$g/L).
+C.) Plot for USEPA Adult (1.6 $\mu$g/L).
+D.) Plot for WHO Recreational (2 $\mu$g/L).
+}
+\end{figure}
+-->
+
+The chlorophyll *a* cutoffs may be used to predict whether or not a lake exceeds the microcystin health advisories.
+Doing so allows us to compare the accuracy of the prediction as well as evaluate false negatives.
+Total accuracy of the four cutoffs predicting microcystin exceedances were `r total_accuracy$epa_child*100`% for the USEPA children's drinking water advisory, `r total_accuracy$who_drink*100`% for the WHO drinking water advisory,  `r total_accuracy$epa_adult*100`% for the USEPA adult drinking water advisory, and `r total_accuracy$who_rec*100`% for the WHO recreational advisory (Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table}, \ref{tab:adult_conmat_table}, & \ref{tab:who_rec_conmat_table}).
+However, total accuracy is only one part of the prediction performace with which we are concerned.
+
+When using the chlorophyll *a* cutoffs as an indicator of microcystin exceedances, the error that should be avoided is predicting that no exceedance has occurred when in fact it has.
+In other words, we would like to avoid Type II errors and minimize the proportion of false negatives.
+For the four chlorophyll *a* cut-offs we had a proportion of false negatives of `r fn_percent$epa_child*100`%, `r fn_percent$who_drink*100`%, `r fn_percent$epa_adult*100`%, and `r fn_percent$who_rec*100`% for the USEPA children's, the WHO drinking water, the USEPA adult, and the WHO recreational advisories, respectively.
+In each case we missed less than 10% of the lakes that in fact exceeded the microcystin advisory.
+While this method performs well with regard to the false negative percentage, it is possible that is a relic of the NLA dataset and testing with additional data would allow us to confirm this result.
+
+#Discussion
+The log-log association between microcystin-LR and chlorophyll *a* indicates that, in general, higher concentrations of microcystin-LR almost always co-occur with higher concentrations of chlorophyll *a* yet the inverse is not true (Figure \ref{fig:chla_micro_scatter}).
+Higher chlorophyll *a* is not necessarily predictive of higher microcystin-LR concentrations; however, chlorophyll *a* may be predictive of the probability of exceeding a certain threshold.
+
+```{r chla_micro_scatter, echo=FALSE, fig.width=6, fig.cap="", warning=FALSE, message=FALSE}
+ggplot2::ggsave("hollister_microcystin_fig2.jpg",
+       plot_scatter(nla_dat,xvar="log_chla",yvar = "log_micro",pt_col = "black",
+             x=expression(paste('Log10 Chl ', italic("a"),' (',mu,'g/L)')),
+             y=expression(paste('Log10 microcystin-LR',' (',mu,'g/L)'))),
+       dpi = 450,
+       height = 6.8,
+       units = "in")
+x<-autocrop("hollister_microcystin_fig2.jpg",border = 25, 
+            outfile = "hollister_microcystin_fig2.jpg",
+            width = 6.1,
+            res = 450)
+```
+
+<!--
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{hollister_microcystin_fig2.jpg}
+\caption{\label{fig:chla_micro_scatter}Scatterplot showing association betweeen chlorophyll \textit{a} and microcystin-LR. }
+\end{figure}
+-->
+
+Indeed, the probability of exceeding each of the four tested health advisory levels increased as a function of chlorophyll *a* concentration (Figure \ref{fig:multi_cp_plot}).
+We used this association to identify chlorophyll *a* concentrations that were associated with a range of probabilities of exceeding a given health advisory level (Table \ref{tab:mc_chla_table}).
+For the purposes of this discussion we focus on a conditional probability of 50% or greater (i.e., greater than even odds to exceed a health advisory level).
+The 50% conditional probability chlorophyll *a* thresholds represents `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[5,1]])/length(nla_dat$CHLA)),1)`%, `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[5,2]])/length(nla_dat$CHLA)),1)`%,
+`r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[5,3]])/length(nla_dat$CHLA)),1)`%, and `r round(100*(length(nla_dat$CHLA[nla_dat$CHLA>=mc_chla[5,4]])/length(nla_dat$CHLA)),1)`% of sample lakes for the USEPA Child, the WHO Drinking, the USEPA Adult, and the WHO recreational levels, respecitvely.
+
+There are numerous possible uses for the chlorophyll *a* and microcystin advisory cut-off values.
+First, in the absence of microcystin-LR measurements, exceedence of the chlorophyll *a* concentrations could be a trigger for further actions.
+Given that there is uncertainity around these chlorophyll *a* cutoffs the best case scenario would be to monitor for chlorophyll *a* and in the event of exceeding a target concentration take water samples and have those samples tested for microcystin-LR.
+
+A second potential use is to identify past bloom events from historical data.
+As harmful algal blooms are made up of many species and have various mechanisms responsible for adverse impacts (e.g., toxins, hypoxia, odors), there is no single definition of a bloom.
+For cHABs, one approach has been to utilize phycocyanin to screen for or identify bloom events [@miller2013spatiotemporal, ahn2007alternative, @marion2012vivo].
+This is a useful approach, but phycocyanin is not always available, thus limiting its utility especially for examining historical data.
+Using our chlorophyll *a* cutoffs provides a value that is also associated with microcystin-LR and can be used to classify lakes, from past surveys, as having bloomed.
+
+The values we propose are national and may miss regional variation in water quality, including, chlorophyll *a* and microcystin-LR [@beaver2014land, @wagner2011landscape].
+A set of regional conditional probabilities would be interesting; however, limiting the analysis to the data available per region would make interpretation difficult.
+The sample size for each of the regional conditional probabilities would be reduced and the number of lakes in each region that exceed the microcystin values would also be reduced.
+Thus, our confidence in the conditional probabilites would be less (i.e.
+greatly increased confidence intervals) and the relationships less pronounced as we have fewer lakes on which to base the probabilites.
+Thus, this dataset is best for making national scale recommendations.
+
+There are two other limitations with the 2007 National Lakes Assessment dataset.
+First, it represents a single sample from a lake and does not capture temporal dynamics.
+Second, validation of the predictions with the 2007 data alone would be challenging as the data would need to be subset and this would only sever to increase the uncertainty of our conditional probabilities, reducing our ability to validate the presence of microcystin-LR.
+The 2017 National Lakes Assessment would be ideal for this task.
+However, as of this writing, the 2012 National Lakes Assessment data are not public.
+When these data are released, a validation of this approach can be completed then.
+
+Lastly, using chlorophyll *a* is not meant as a replacement for testing of microcystin-LR or other toxins.
+It should be used when other, direct measurements of cyanotoxins are not available.
+In those cases, which are likely to be common at least in the near future, using a more ubiquitous measurement such as chlorophyll *a* will provide a reasonable proxy for the probability of exceeding a microcystin health advisory level and provide better protection against adverse effects in both drinking and recreational use cases.
+
+#Acknowledgements
+We would like to thank Anne Kuhn, Bryan Milstead, John Kiddon, Joe LiVolsi, Tim Gleason, Wayne Munns, and Leslie D'Anglada for constructive reviews of this paper.
+Special thanks to Jason Marion, Alan Wilson, and Zofia Taranu for reviews of the submitted manuscript.
+This paper has not been subjected to Agency review.
+Therefore, it does not necessary reflect the views of the Agency.
+Mention of trade names or commercial products does not constitute endorsement or recommendation for use.
+This contribution is identified by the tracking number ORD-015143 of the Atlantic Ecology Division, Office of Research and Development, National Health and Environmental Effects Research Laboratory, US Environmental Protection Agency.
+
 \newpage
+
+<!--
+```{r chla_micro_cdf, echo=FALSE, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin health advisory Levels.
+\\label{fig:chla_micro_cdf}"}
+plot_cdf(nla_dat[,c("log_chla","log_micro")],"")
+```
+
+\newpage
+-->
 
@@
 
-```{r microcystin_levels,echo=FALSE,results='asis', tab.cap="Various suggested microcystin health advisory concentrations. \\label{tab:microcystin_levels}"}
+```{r microcystin_levels,echo=FALSE,results='asis', tab.cap="Various microcystin health advisory concentrations. \\label{tab:microcystin_levels}"}
 source_of_stnd <- c("WHO",
@@
 
-```{r mc_chla_table, results='asis', echo=FALSE, tab.cap="Chlorophyll \\textit{a} concentrations that are associated with a 50% probability of exceeding a microcystin health advisory concentration.
+```{r mc_chla_table, results='asis', echo=FALSE, tab.cap="Chlorophyll \\textit{a} concentrations that are associated with a 50\\% probability of exceeding a microcystin health advisory concentration.
 \\label{tab:mc_chla_table}"}

### 2016-06-07 manuscript/manuscript.tex
updates, including proofs

@@
 concentrations that would trigger health advisories {[}4--6{]}.
-Additionally, the World Health Organization (WHO) has proposed
-microcystin advisory levels for drinking water and a range of
-recreational risk levels {[}7,8{]}. While these levels and associated
-advisories are likely to help mitigate the impacts from harmful algal
-blooms, they are not without complications.
+Additionally, the World Health Organization (WHO) has microcystin
+advisory levels for drinking water and for a range of recreational risk
+levels {[}7,8{]}. While these levels and associated advisories are
+likely to help mitigate the impacts from harmful algal blooms, they are
+not without complications.
 
@@
 strongest association with microcystin. Furthermore, they identify
-chlorophyll\emph{a} and total nitrogen concentrations that are
+chlorophyll \emph{a} and total nitrogen concentrations that are
 associated with exceeding 1 \(\mu\)g/L of microcystin. These findings
-suggest that chlorophyll a concentrations could also track the new USEPA
-microcystin health advisory levels for drinking water. Identifying this
-association would provide an important tool for water resource managers
-to help manage the threat to public health posed by cHABs and would be
-especially useful in the absence of measured microcystin concentrations.
+suggest that chlorophyll \emph{a} concentrations could also track the
+new USEPA microcystin health advisory levels for drinking water.
+Identifying this association would provide an important tool for water
+resource managers to help manage the threat to public health posed by
+cHABs and would be especially useful in the absence of measured
+microcystin concentrations.
 
-In this paper we build on past efforts and utilize the National Lakes
-Assessment (NLA) data and identify chlorophyll \emph{a} concentrations
-that are associated with higher probabilities of exceeding several
-microcystin health advisory concentrations {[}6,8,14{]}. We build on
-past studies by exploring associations with the newly announced advisory
-levels and by also applying a different method, conditional probability
-analysis. Utilizing different methods strengthens the evidence for
-suggested chlorophyll \emph{a} levels that are associated with increased
-risk of exceeding the health advisory levels as those levels are not
-predicated on a single analytical method. So that others may repeat or
-adjust this analysis, the data, code, and this manuscript are freely
-available via
+In fact, this is a similar tact to the World Health Organization who, in
+addition to advisory levels for microcystin, have also proposed related
+advisory levels for cyanobacteria abundance and chlorophyll \emph{a}
+{[}7,8{]}. The chlorophyll \emph{a} concentrations proposed by the WHO
+are for low (\textless{} 10 \(\mu\)g/L), moderate (between 10 and 50
+\(\mu\)g/L, high (between 50 and 5000 \(\mu\)g/L), and very high risk
+(\textgreater{}5000 \(\mu\)g/L) {[}8{]}. While these advisories have
+proven to be useful tools they do suffer from being coarse, broad, and
+have been found to overestimate actual risk {[}14{]}.
+
+In this paper we build on these past efforts and utilize the National
+Lakes Assessment (NLA) data and identify chlorophyll \emph{a}
+concentrations that are associated with higher probabilities of
+exceeding several microcystin health advisory concentrations
+{[}6,8,15{]}. We build on past studies by exploring associations with
+the newly announced advisory levels and by also applying a different
+method, conditional probability analysis. Utilizing different methods
+strengthens the evidence for suggested chlorophyll \emph{a} levels that
+are associated with increased risk of exceeding the health advisory
+levels as those levels are not predicated on a single analytical method.
+So that others may repeat or adjust this analysis, the data, code, and
+this manuscript are freely available via
 \href{https://github.com/USAPE/microcystinchla}{https://github.com/USEPA/microcystinchla}.
@@
 We used the 2007 NLA chlorophyll \emph{a} and microcystin-LR
-concentration data {[}14{]}. These data represent a snapshot of water
+concentration data {[}15{]}. These data represent a snapshot of water
 quality from the summer of 2007 for the conterminous United States and
 were collected as part of an ongoing probabilistic monitoring program
-{[}14{]}. Water quality data, including chlorophyll \emph{a} and
+{[}15{]}. Water quality data, including chlorophyll \emph{a} and
 microcystin-LR were obtained via an intergated sample taken from the
@@
 provide the source for both chlorophyll \emph{a} and microcystin-LR
-{[}14{]}.
+{[}15{]}.
 
@@
 Health Organization (WHO) and USEPA microcystin health advisory levels
-{[}16{]}. Many health advisory levels have been suggested (Table
+{[}17{]}. Many health advisory levels have been suggested (Table
 \ref{tab:microcystin_levels}), but lakes with higher microcystin-LR
@@
 probabilities of exceeding the microcystin health advisory levels. We
-also calculated bootstrapped 95\% confidence intervals (CI) using 1000
+also calculated bootstrapped 95\% confidence intervals (CI) using 10,000
 bootstrapped samples. Thus, to identify chlorophyll \emph{a}
@@
 discussion of CPA is beyond the scope of this paper, but see Paul et al.
-{[}17{]} and Hollister et al. {[}18{]} for greater detail. Lastly, all
-analyses were conducted using R version 3.2.2 and code and data from
-this analysis are freely available as an R package at
-\href{https://github.com/USAPE/microcystinchla}{https://github.com/USEPA/microcystinchla}.
+{[}18{]} and Hollister et al. {[}19{]} for greater detail. All analyses
+were conducted using R version 3.2.3 and code and data from this
+analysis are freely available as an R package at
+\url{https://github.com/USEPA/microcystinchla}.
 
@@
 the detection limits. Any value greater than that indicates the presence
-of microcystin-LR. Of those lakes with microcystin, the median
-concentration was 0.51\(\mu\)g/L and the mean was 3.17\(\mu\)g/L. Of all
-lakes sampled, 21\% of lakes exceeded the USEPA Child level, 8.8\% of
-lakes exceeded the USEPA Adult level, 11.7\% of lakes exceeded the WHO
-Drinking level, and 7.3\% of lakes exceeded the WHO Recreational level.
-Chlorophyll \emph{a}, ranged from 0.07 to 936 \(\mu\)g/L and this
+of microcystin-LR. Of those lakes with microcystin-LR, the median
+concentration was 0.51 \(\mu\)g/L and the mean was 3.17 \(\mu\)g/L. Of
+all lakes sampled, 21\% of lakes exceeded the USEPA Child level, 8.8\%
+of lakes exceeded the USEPA Adult level, 11.7\% of lakes exceeded the
+WHO Drinking level, and 7.3\% of lakes exceeded the WHO Recreational
+level. Chlorophyll \emph{a}, ranged from 0.07 to 936 \(\mu\)g/L and this
 captures the range of trrophic states from oligotrophic to
@@
 shown in Table \ref{tab:mc_chla_table}. Specific chlorophyll \emph{a}
-that are associated with greater than even odds of exceeding the
-advisory levels were 23, 68, 84, and 104\(\mu\)g/L for 0.3, 1.0, 1.6,
-and 2.0 \(\mu\)g/L advisory levels, respectively (Table
+concentrations that are associated with greater than even odds of
+exceeding the advisory levels were 23, 68, 84, and 104 \(\mu\)g/L for
+0.3, 1.0, 1.6, and 2.0 \(\mu\)g/L advisory levels, respectively (Table
 \ref{tab:mc_chla_table} \& Figure \ref{fig:multi_cp_plot}).
@@
 86\% for the WHO drinking water advisory, 89\% for the USEPA adult
-drinking water advisory, and 91\% for the WHO regreational advisory
+drinking water advisory, and 91\% for the WHO recreational advisory
 (Tables \ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table},
@@
 one approach has been to utilize phycocyanin to screen for or identify
-bloom events {[}19{]}. This is a useful approach, but phycocyanin is not
+bloom events {[}20{]}. This is a useful approach, but phycocyanin is not
 always available, thus limiting its utility especially for examining
@@
 water quality, including, chlorophyll \emph{a} and microcystin-LR
-{[}21{]}. A set of regional conditional probabilities would be
+{[}22{]}. A set of regional conditional probabilities would be
 interesting; however, limiting the analysis to the data available per
-region would make interpretaion difficult. The sample size for each of
-the conditional probabilities would be reduced (it ranges from 67 to
-155) and the number of lakes in each region that exceed the microcystin
-values is also reduced. The result is that our confidence in the
-conditional probabilites would be less (i.e.~greatly increased
-confidence intervals) and the relationships less pronounced as we have
-fewer lakes on which to base the probabilites. Thus, this dataset is
-best for making national scale recommendations.
+region would make interpretation difficult. The sample size for each of
+the regional conditional probabilities would be reduced and the number
+of lakes in each region that exceed the microcystin values would also be
+reduced. Thus, our confidence in the conditional probabilites would be
+less (i.e.~greatly increased confidence intervals) and the relationships
+less pronounced as we have fewer lakes on which to base the
+probabilites. Thus, this dataset is best for making national scale
+recommendations.
 
-Two other limitations with the 2007 National Lakes Assessment data are
-that they represent a single sample from that lake and do not capture
-temporal dynamics and without subsetting the data do not provide us the
-ability to validate the presence of microcystin-LR. As of this writing,
-the 2012 National Lakes Assessment data are not public. When these data
-are released, a validation of this approach can be completed then.
+There are two other limitations with the 2007 National Lakes Assessment
+dataset. First, it represents a single sample from a lake and does not
+capture temporal dynamics. Second, validation of the predictions with
+the 2007 data alone would be challenging as the data would need to be
+subset and this would only sever to increase the uncertainty of our
+conditional probabilities, reducing our ability to validate the presence
+of microcystin-LR. The 2017 National Lakes Assessment would be ideal for
+this task. However, as of this writing, the 2012 National Lakes
+Assessment data are not public. When these data are released, a
+validation of this approach can be completed then.
 
@@
 which are likely to be common at least in the near future, using a more
-ubiquitous measurement, such as chlorophyll \emph{a} will provide a
+ubiquitous measurement such as chlorophyll \emph{a} will provide a
 reasonable proxy for the probability of exceeding a microcystin health
@@
 
-\section{Figures}\label{figures}
-
-\begin{verbatim}
-## Saving 6 x 6.8 in image
-\end{verbatim}
-
-\newpage
-
-\newpage
-
 \section{Tables}\label{tables}
@@
 \begin{longtable}[c]{@{}lll@{}}
-\caption{Various suggested microcystin health advisory concentrations.
+\caption{Various microcystin health advisory concentrations.
 \label{tab:microcystin_levels}}\tabularnewline
@@
 
+\hypertarget{ref-loftin2016cyanotoxins}{}
+14. Loftin KA, Graham JL, Hilborn ED, Lehmann SC, Meyer MT, et al.
+(2016) Cyanotoxins in inland lakes of the united states: Occurrence and
+potential recreational health risks in the ePA national lakes assessment
+2007. Harmful Algae 56: 77--90.
+
 \hypertarget{ref-usepa2009national}{}
-14. USEPA (2009) National lakes assessment: A collaborative survey of
+15. USEPA (2009) National lakes assessment: A collaborative survey of
 the nation's lakes. EPA 841-R-09-001.
@@
 \hypertarget{ref-nlaux5ffieldops}{}
-15. USEPA (2007) Survey of the nation's lakes field operations manual
+16. USEPA (2007) Survey of the nation's lakes field operations manual
 EPA841-B-07-004.
@@
 \hypertarget{ref-paul2011probability}{}
-16. Paul JF, Munns WR (2011) Probability surveys, conditional
+17. Paul JF, Munns WR (2011) Probability surveys, conditional
 probability, and ecological risk assessment. Environmental Toxicology
@@
 \hypertarget{ref-paul2005development}{}
-17. Paul JF, McDonald ME (2005) Development of empirical, geographically
+18. Paul JF, McDonald ME (2005) Development of empirical, geographically
 specific water quality criteria: A conditional probability analysis
@@
 \hypertarget{ref-hollister2008cprob}{}
-18. Hollister JW, Walker HA, Paul JF (2008) CProb: A computational tool
+19. Hollister JW, Walker HA, Paul JF (2008) CProb: A computational tool
 for conducting conditional probability analysis. Journal of
@@
 \hypertarget{ref-miller2013spatiotemporal}{}
-19. Miller TR, Beversdorf L, Chaston SD, McMahon KD (2013)
+20. Miller TR, Beversdorf L, Chaston SD, McMahon KD (2013)
 Spatiotemporal molecular analysis of cyanobacteria blooms reveals
@@
 \hypertarget{ref-marion2012vivo}{}
-20. Marion JW, Lee J, Wilkins III J, Lemeshow S, Lee C, et al. (2012) In
+21. Marion JW, Lee J, Wilkins III J, Lemeshow S, Lee C, et al. (2012) In
 vivo phycocyanin flourometry as a potential rapid screening tool for
@@
 \hypertarget{ref-beaver2014land}{}
-21. Beaver JR, Manis EE, Loftin KA, Graham JL, Pollard AI, et al. (2014)
+22. Beaver JR, Manis EE, Loftin KA, Graham JL, Pollard AI, et al. (2014)
 Land use patterns, ecoregion, and microcystin relationships in uS lakes
@@
 \hypertarget{ref-wagner2011landscape}{}
-22. Wagner T, Soranno PA, Webster KE, Cheruvelil KS (2011) Landscape
+23. Wagner T, Soranno PA, Webster KE, Cheruvelil KS (2011) Landscape
 drivers of regional variation in the relationship between total

### 2016-12-02 manuscript/manuscript.Rmd
fixed url

@@
 ################################################################################
-data_url<-"http://www2.epa.gov/sites/production/files/2014-10/nla2007_recreational_conditionestimates_20091123.csv"
+data_url<-"https://www2.epa.gov/sites/production/files/2014-10/nla2007_recreational_conditionestimates_20091123.csv"
 nla_dat <- get_nla(data_url)
